### 2021-09-14 notes-rkhs.tex
Start RKHS notes

\documentclass[11pt]{article}
\usepackage[margin=1.3in]{geometry}
%\usepackage[compact,small]{titlesec}
%--------------------------------------------------------
\usepackage{sectsty,todonotes}
\sectionfont{\normalsize\bfseries}
\subsectionfont{\normalsize\sf}
%----------------------------------------------------
\usepackage[mathcal]{euscript}

\usepackage{bm}                     %AMS Packages
\usepackage{amsfonts, framed}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}               %Graphics
%\usepackage{pstricks,pst-plot}
%\usepackage{wrapfig}
\usepackage{natbib}                 %Citation style
\usepackage{colortbl}               %Tables
\usepackage{booktabs}
\usepackage{hyperref}
\setlength{\bibsep}{1pt}
%\usepackage{hypernat}

% Definiciones

\definecolor{violet}{rgb}{0.7,0,0.6}
\newcommand{\explainindetail}[1]{\todo[color=violet]{#1}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\eps}{\varepsilon}
\newcommand{\ind}{\mathbb{I}}
\newcommand{\pr}{\mathbb{P}}
\newcommand{\B}{B}
\newcommand{\A}{\mathcal{A}}\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\newcommand{\E}{\mathcal{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\F}{\mathfrak{F}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\ds}{\displaystyle}
\newcommand{\D}{\mathcal{D}}
\newcommand{\deb}{\stackrel{\mathcal{L}}{\longrightarrow}}
\newcommand{\Vor}{\mbox{Vor}}
\newcommand{\ang}{\measuredangle}
\def\cals_+{{\cals_+}}
\def\cala{{\mathcal{A}}}
\def\calb{{\mathcal{B}}}
\def\calc{{\mathcal{C}}}
\def\cald{{\mathcal{D}}}
\def\calf{{\mathcal{F}}}
\def\calh{{\mathcal{H}}}
\def\call{{\mathcal{L}}}
\def\cals{{\mathcal{S}}}
\def\calv{{\mathcal{V}}}
\def\calg{{\mathcal{G}}}
\def\calp{{\mathcal{P}}}
\def\cale{{\mathcal{E}}}
\def\calo{{\mathcal{O}}}
\def\calj{{\mathcal{J}}}
\def\caln{{\mathcal{N}}}
\def\calm{{\mathcal{M}}}

\hypersetup{
  colorlinks=true,
  citecolor=violet
}

\begin{document}

\begin{center}
\large \bf  Preliminaries on RKHS's\rm
\end{center}

RKHS's are spaces of ``regular'' FUNCTIONS (not equivalence classes in $L^2[0,1]$) with a different norm to that inherited from $L^2[0,1]$.

\section*{Definition}

	Let $K(s,t)$ be a positive definite (the matrix $K(t_i,t_j)$ is always positive definite) real function on $[0,1]^2$.

	Let $\calh_0(K):=\{f\in L^2[0,1] \ : \ f(\cdot)=\sum_{i=1}^n a_i K(t_i,\cdot),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\}$, be
	the space of all finite linear combinations of evaluations of $K$.  This space is endowed with the inner product
	$\langle f,g\rangle_K=\sum_{i,j}\alpha_i\beta_j K(t_i,s_j)$,
	where $f(\cdot)=\sum_i\alpha_i K(t_i,\cdot)$ and $g(\cdot)=\sum_j\beta_j K(s_j,\cdot)$.

	Then, the RKHS associated with $K$ is defined as the completion of $\calh_0(K)$. In other words, $\calh(K)$ is made of all functions obtained as pointwise limits of Cauchy sequences in $\calh_0(K)$. The inner product is extended accordingly to the whole space $\calh(K)$.

\section*{Reproducing property}

	These spaces are named after the so-called \textit{reproducing property},
	$\langle f, K(s,\cdot)\rangle_K = f(s)$, for all  $f\in\calh(K), s\in[0,1]$, which is particularly important in the applications. On account of this property it is sometimes said that the RKHS are spaces of ``true functions'', in the sense that the pointwise values $f(s)$ at a given $s$ do matter, in contrast with $L^2[0,1]$ whose elements are in fact equivalence classes of functions.

  Note that in an RKHS the evaluation functional is continuous, so a small perturbation $f\mapsto f + \delta f$ results in a small change in $f(x)$ for all $x$.

\section*{Loève's isometry}

	A property of RKHS's especially useful in statistical applications is given by the following \textit{Lo\`eve's isometry}:  let $L^2(\Omega) $ be the Hilbert space of real random variables with finite second moment, endowed with the usual inner product whose associated norm is $\| U\|^2={\mathbb E}(U^2)$. Define
	$$\call_0(X) = \big\{U\in L^2(\Omega) \ : \ U=\sum_{i=1}^n a_i \big(X(t_i)-m(t_i)\big),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\big\},$$
	where $m(t)={\mathbb E}[X(t)]$, and let $\call(X)$ be the completion of $\call_0(X)$ in  $L^2(\Omega)$; hence, in other words, is a closed subspace of $L^2(\Omega)$ defined as the linear span of the centred one-dimensional marginals of the process $X$. It turns out that the transformation $\Psi_X$, from  $\call(X)$ to $\calh(K)$, defined by
	\begin{equation}\label{eq:isometry}
	\Psi_X(U)(s) = {\mathbb E}[ U(X(s)-m(s))] = \langle U, X(s)-m(s) \rangle \in \calh(K), \ \text{ for } U\in \call(X)
	\end{equation}
	is an isometry (sometimes called Lo\`eve's isometry) between $\call(X)$ and $\calh(K)$, that is, $\Psi_X(U)$ is bijective and preserves the inner product. As a consequence, the Hilbert spaces $\call(X)$ and $\calh(K)$ can be identified.

	\textbf{In more informal terms, this isometry is nothing but the completion} of
	$$
	\sum_ia_iK(t_i,\cdot)\longleftrightarrow \sum_ia_i(X(t_i)-m(t_i))
	$$

\section*{Alternative definitions}

	Assume that $K(s,t)$ is continuous. Denote by ${\mathcal K}$ the (linear continuous) covariance operator ${\mathcal K}:L^2[0,1]\rightarrow L^2[0,1]$ defined by

	$$
	{\mathcal K}(x)(t)=\int_0^1K(s,t)x(s)ds
	$$

	Assume that ${\mathcal K}$ is injective.
	We know, from Spectral Theorem, that
	$${\mathcal K}(x)=\sum_i\lambda_i\langle x,e_i\rangle e_i,$$

	$\lambda_i$ and $e_i$ being the eigenvalues and eigenfunctions of ${\mathcal K}$.

		One could also define the Reproducing Kernel Hilbert Space associated with $K$, by
$$
	{\mathcal H}(K)=\left\{x:\in L^2[0,1]:  \sum_{i=1}^\infty\frac{\langle x,e_i\rangle^2}{\lambda_i} <\infty \right\}
	$$
	with the inner product and norm given by
	$$
	\langle f,g\rangle_K=\sum_{i=1}^\infty \frac{\langle f,e_i\rangle \langle g,e_i\rangle}{\lambda_i},\ \Vert f\Vert_K^2=\sum_{i=1}^\infty \frac{\langle f,e_i\rangle^2 }{\lambda_i}
	$$
An equivalent definition:
	${\mathcal H}(K)={\mathcal K}^{1/2}(L^2[0,1])$,
	where, from Spectral Theorem, ${\mathcal K}^{1/2}(x)=\sum_i\lambda_i^{1/2}\langle x,e_i\rangle e_i$

\section*{General properties}

  An important thing to note is that in general the trajectories of the process $X$ \textbf{do not belong to the corresponding RKHS}. More precisely, $X(t)\notin \mathcal H(K)$ with probability $1$ if $\mathcal K$ has infinite non-zero eigenvalues. While a rigorous proof of this fact is given in \citet{hajek1962}, an heuristic argument presented in \citet{wahba1990} is as follows. Consider the Karhunen-Loève expansion of $X$, i.e.,
  \[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
  \]

  and the truncated version $X_N(t)$ up to the $N$-th term. On the one hand, $X_N(t)\to X(t)$ in the quadratic mean (for each fixed $t$) by the Karhunen-Loève theorem, but on the other hand we have
  \[
\mathbb E\|X_N(\cdot)\|^2_K = \mathbb E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
  \]

  Thus, functions in the RKHS are in a way ``smoother'' than the trajectories of the process themselves. For convenience, if $x=x(\omega)$ is a realization of $X$ (for a fixed $\omega$), we will denote $\langle x, f\rangle_K \equiv \left[\Psi_X^{-1}(f)\right](\omega)$. Some interesting properties stemming from this interpretation are the following (see \citet[p.~974]{parzen1961b}):

  \begin{enumerate}

    \item $\langle X, K(t, \cdot)\rangle_K = X(t) - m(t)$

    \item $\mathbb{E} \langle X, h\rangle_K = \langle m, h\rangle_K$

    \item $\operatorname{Cov}\left(\langle X, h \rangle_K , \langle X, g\rangle_K\right) = \langle h, g\rangle_K$

  \end{enumerate}

\section*{Convergence properties}

Note that
\[
		\sup_t \|\langle f_n,K(\cdot,t)\rangle_K\|=\sup_t|f_n(t)|\leq \Vert f_n\Vert_K \sup_t\Vert K(\cdot,t)\Vert_K
		=\Vert f_n\Vert_K \sup_t K(t,t),
\]
		so $\Vert f_n\Vert_K \to 0$ implies $f_n(t)\to 0$ uniformly on $t$, that is, convergence in the RKHS norm \textbf{entails uniform convergence}.

\end{itemize}

\begin{thebibliography}{9}
	\bibitem[\protect\citeauthoryear{Hsing \& Eubank}{2015}]{hsi15}
	Hsing, T., \& Eubank, R. (2015). \textit{Theoretical foundations of functional data analysis, with an introduction to linear operators} (Vol. 997). John Wiley \& Sons.

	\bibitem[\protect\citeauthoryear{Paulsen \& Raghupathi }{2015}]{pau16}
	Paulsen, V. I., \& Raghupathi, M. (2016). \textit{An introduction to the theory of reproducing kernel Hilbert spaces} (Vol. 152). Cambridge University press.
  \bibitem[\protect\citeauthoryear{Parzen}{1961}]{parzen1961b} Parzen, E. (1961). \textit{An Approach to Time Series Analysis}. The Annals of Mathematical Statistics 32(4), 951–989.

  \bibitem[\protect\citeauthoryear{Wahba}{1990}]{wahba1990} Wahba G. (1990). \textit{Spline models for observational data}. CBMS 59, SIAM, Philadelphia.

  \bibitem[\protect\citeauthoryear{Hajek}{1962}]{hajek1962} Hajek J. (1962). \textit{On linear Statistical problems in stochastic processes}. Chekoslovak Math. J., vol. 12, pp.404-443.

\end{thebibliography}

\end{document}

### 2021-09-14 notes.tex
Add global notes with kaobook style

\documentclass[
	fontsize=11pt, % Base font size
	twoside=false, % Use different layouts for even and odd pages (in particular, if twoside=true, the margin column will be always on the outside)
	%open=any, % If twoside=true, uncomment this to force new chapters to start on any page, not only on right (odd) pages
	%chapterprefix=true, % Uncomment to use the word "Chapter" before chapter numbers everywhere they appear
	%chapterentrydots=true, % Uncomment to output dots from the chapter name to the page number in the table of contents
	numbers=noenddot, % Comment to output dots after chapter numbers; the most common values for this option are: enddot, noenddot and auto (see the KOMAScript documentation for an in-depth explanation)
	%draft=true, % If uncommented, rulers will be added in the header and footer
	%overfullrule=true, % If uncommented, overly long lines will be marked by a black box; useful for correcting spacing problems
]{kaohandt}

% Choose the language
\usepackage[english]{babel} % Load characters and hyphenation
\usepackage[english=british]{csquotes}	% English quotes

% Load the bibliography package
\usepackage{kaobiblio}
\addbibresource{bibliography.bib} % Bibliography file

% Load the package for hyperreferences
\usepackage{kaorefs}

% Set the paths where to look for images
\usepackage{subcaption}
\graphicspath{{figures/}}

\begin{document}

\title{Notes}
\author[FM]{Antonio Coín Castro}
\date{\today}
\maketitle

\defbibnote{bibnote}{Here are the references in citation order.\par\bigskip} % Prepend this text to the bibliography
\renewcommand*{\bibfont}{\small}
\printbibliography[title=Bibliography]% , prenote=bibnote]

\end{document}

### 2021-10-06 notes-rkhs.tex
Add notes on theory & practice of Bayesian Linear Regression

@@
-\documentclass[11pt]{article}
-\usepackage[margin=1.3in]{geometry}
-%\usepackage[compact,small]{titlesec}
-%--------------------------------------------------------
-\usepackage{sectsty,todonotes}
-\sectionfont{\normalsize\bfseries}
-\subsectionfont{\normalsize\sf}
-%----------------------------------------------------
-\usepackage[mathcal]{euscript}
-
-\usepackage{bm}                     %AMS Packages
-\usepackage{amsfonts, framed}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{graphicx}               %Graphics
-%\usepackage{pstricks,pst-plot}
-%\usepackage{wrapfig}
-\usepackage{natbib}                 %Citation style
-\usepackage{colortbl}               %Tables
-\usepackage{booktabs}
-\usepackage{hyperref}
-\setlength{\bibsep}{1pt}
-%\usepackage{hypernat}
-
-% Definiciones
-
-\definecolor{violet}{rgb}{0.7,0,0.6}
-\newcommand{\explainindetail}[1]{\todo[color=violet]{#1}}
-\newcommand{\nat}{\mathbb{N}}
-\newcommand{\eps}{\varepsilon}
-\newcommand{\ind}{\mathbb{I}}
-\newcommand{\pr}{\mathbb{P}}
-\newcommand{\B}{B}
-\newcommand{\A}{\mathcal{A}}\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\newcommand{\E}{\mathcal{E}}
-\newcommand{\V}{\mathbb{V}}
-\newcommand{\F}{\mathfrak{F}}
-\newcommand{\G}{\mathbb{G}}
-\newcommand{\M}{\mathcal{M}}
-\newcommand{\X}{\mathcal{X}}
-\newcommand{\Z}{\mathcal{Z}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\Y}{\mathcal{Y}}
-\newcommand{\C}{\mathcal{C}}
-\newcommand{\ds}{\displaystyle}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\deb}{\stackrel{\mathcal{L}}{\longrightarrow}}
-\newcommand{\Vor}{\mbox{Vor}}
-\newcommand{\ang}{\measuredangle}
-\def\cals_+{{\cals_+}}
-\def\cala{{\mathcal{A}}}
-\def\calb{{\mathcal{B}}}
-\def\calc{{\mathcal{C}}}
-\def\cald{{\mathcal{D}}}
-\def\calf{{\mathcal{F}}}
-\def\calh{{\mathcal{H}}}
-\def\call{{\mathcal{L}}}
-\def\cals{{\mathcal{S}}}
-\def\calv{{\mathcal{V}}}
-\def\calg{{\mathcal{G}}}
-\def\calp{{\mathcal{P}}}
-\def\cale{{\mathcal{E}}}
-\def\calo{{\mathcal{O}}}
-\def\calj{{\mathcal{J}}}
-\def\caln{{\mathcal{N}}}
-\def\calm{{\mathcal{M}}}
-
-\hypersetup{
-  colorlinks=true,
-  citecolor=violet
-}
-
-\begin{document}
+% \documentclass[11pt]{article}
+% \usepackage[margin=1.3in]{geometry}
+% %\usepackage[compact,small]{titlesec}
+% %--------------------------------------------------------
+% \usepackage{sectsty,todonotes}
+% \sectionfont{\normalsize\bfseries}
+% \subsectionfont{\normalsize\sf}
+% %----------------------------------------------------
+% \usepackage[mathcal]{euscript}
+%
+% \usepackage{bm}                     %AMS Packages
+% \usepackage{amsfonts, framed}
+% \usepackage{amssymb}
+% \usepackage{amsmath}
+% \usepackage{amsthm}
+% \usepackage{graphicx}               %Graphics
+% %\usepackage{pstricks,pst-plot}
+% %\usepackage{wrapfig}
+% \usepackage{natbib}                 %Citation style
+% \usepackage{colortbl}               %Tables
+% \usepackage{booktabs}
+% \usepackage{hyperref}
+% \setlength{\bibsep}{1pt}
+% %\usepackage{hypernat}
+%
+% % Definiciones
+%
+% \definecolor{violet}{rgb}{0.7,0,0.6}
+% \newcommand{\explainindetail}[1]{\todo[color=violet]{#1}}
+% \newcommand{\nat}{\mathbb{N}}
+% \newcommand{\eps}{\varepsilon}
+% \newcommand{\ind}{\mathbb{I}}
+% \newcommand{\pr}{\mathbb{P}}
+% \newcommand{\B}{B}
+% \newcommand{\A}{\mathcal{A}}\usepackage{amsfonts}
+% \usepackage{amssymb}
+% \usepackage{amsmath}
+% \usepackage{amsthm}
+% \newcommand{\E}{\mathcal{E}}
+% \newcommand{\V}{\mathbb{V}}
+% \newcommand{\F}{\mathfrak{F}}
+% \newcommand{\G}{\mathbb{G}}
+% \newcommand{\M}{\mathcal{M}}
+% \newcommand{\X}{\mathcal{X}}
+% \newcommand{\Z}{\mathcal{Z}}
+% \newcommand{\T}{\mathcal{T}}
+% \newcommand{\Y}{\mathcal{Y}}
+% \newcommand{\C}{\mathcal{C}}
+% \newcommand{\ds}{\displaystyle}
+% \newcommand{\D}{\mathcal{D}}
+% \newcommand{\deb}{\stackrel{\mathcal{L}}{\longrightarrow}}
+% \newcommand{\Vor}{\mbox{Vor}}
+% \newcommand{\ang}{\measuredangle}
+% \def\cals_+{{\cals_+}}
+% \def\cala{{\mathcal{A}}}
+% \def\calb{{\mathcal{B}}}
+% \def\calc{{\mathcal{C}}}
+% \def\cald{{\mathcal{D}}}
+% \def\calf{{\mathcal{F}}}
+% \def\calh{{\mathcal{H}}}
+% \def\call{{\mathcal{L}}}
+% \def\cals{{\mathcal{S}}}
+% \def\calv{{\mathcal{V}}}
+% \def\calg{{\mathcal{G}}}
+% \def\calp{{\mathcal{P}}}
+% \def\cale{{\mathcal{E}}}
+% \def\calo{{\mathcal{O}}}
+% \def\calj{{\mathcal{J}}}
+% \def\caln{{\mathcal{N}}}
+% \def\calm{{\mathcal{M}}}
+%
+% \hypersetup{
+% colorlinks=true,
+% citecolor=violet
+% }
+%
+% \begin{document}
 
@@
 $$
-	{\mathcal H}(K)=\left\{x:\in L^2[0,1]:  \sum_{i=1}^\infty\frac{\langle x,e_i\rangle^2}{\lambda_i} <\infty \right\}
+	{\mathcal H}(K)=\left\{x\in L^2[0,1]:  \sum_{i=1}^\infty\frac{\langle x,e_i\rangle^2}{\lambda_i} <\infty \right\}
 	$$
@@
 	${\mathcal H}(K)={\mathcal K}^{1/2}(L^2[0,1])$,
-	where, from Spectral Theorem, ${\mathcal K}^{1/2}(x)=\sum_i\lambda_i^{1/2}\langle x,e_i\rangle e_i$
+	where, from Spectral Theorem, ${\mathcal K}^{1/2}(x)=\sum_i\lambda_i^{1/2}\langle x,e_i\rangle e_i$.
+It can be easily shown that ${\mathcal K}(L^2) \varsubsetneq {\mathcal K}^{1/2}(L^2)=\mathcal H(K)$ (the inclusion is straightforward, and $K(s, \cdot)\notin \mathcal K(L^2)$).
 
@@
 
-\end{document}
+%\end{document}

### 2021-10-06 notes.tex
Add copyright notice and latex class info
Add notes on theory & practice of Bayesian Linear Regression

@@
+%%%
+% Class 'kaobook' from https://github.com/fmarotta/kaobook
+%
+% Copyright (c) 2022 Antonio Coín Castro
+%%%
+
 \documentclass[
+  a4paper,
 	fontsize=11pt, % Base font size
@@
 	%chapterentrydots=true, % Uncomment to output dots from the chapter name to the page number in the table of contents
+  secnumdepth=2,
 	numbers=noenddot, % Comment to output dots after chapter numbers; the most common values for this option are: enddot, noenddot and auto (see the KOMAScript documentation for an in-depth explanation)
@@
 % Choose the language
+\ifxetexorluatex
+	\usepackage{polyglossia}
+	\setmainlanguage{english}
+\else
 	\usepackage[english]{babel} % Load characters and hyphenation
+\fi
 \usepackage[english=british]{csquotes} % English quotes
 
+% Change default fonts
+\setmathfont[range={\mathcal,\mathbfcal, \mathbb}]{NewCMMath-Regular.otf}
+
+% Change page style
+\pagestyle{scrheadings}
+
 % Load the bibliography package
-\usepackage{kaobiblio}
-\addbibresource{bibliography.bib} % Bibliography file
+\usepackage[
+  style=authoryear-comp
+]{kaobiblio}
+
+% Bibliography side cite style
+\renewcommand{\formatmargincitation}[1]{%
+  \parencite{#1}: \citetitle{#1}.%
+}
+
+% Bibliography file
+\addbibresource{bibliography.bib}
 
@@
 
+% Set reference colors
+\hypersetup{
+  pdfauthor = {Antonio Coín},
+  pdftitle = {Notes},
+  citecolor=teal,
+  urlcolor=blue!50!black,
+  %anchorcolor=OliveGreen,
+  linkcolor=teal
+}
+
+% Load mathematical packages for theorems and related environments
+\usepackage{kaotheorems}
+
 % Set the paths where to look for images
@@
 \graphicspath{{figures/}}
+
+% Only show section titles in margintoc
+\setcounter{margintocdepth}{\sectiontocdepth}
+
+% Load additional packages
+\usepackage{dsfont}  %\mathds
+
+% Some custom commands
+\newcommand{\R} {\ensuremath{\mathds{R}}}
+\newcommand{\E} {\ensuremath{\mathds{E}}}
+\newcommand{\eps}{\varepsilon}
+\newcommand{\bm}{\symbf}
+\DeclareMathOperator{\var} {Var}
+\DeclareMathOperator{\cov} {Cov}
+
+%%%%%%%%%%%%%%%
+% DOCUMENT
+%%%%%%%%%%%%%%%
 
@@
 
+\margintoc[*-2]
+
+\section{Questions}
+
+Here are some questions.
+
+\begin{enumerate}
+
+  \item Why does it not make sense to regress \(Y\) on a finite combination of projections of \(X\) whenever we consider the \(L^2\) model?
+
+  \item Why is it that \(k(\cdot, t) \notin \mathcal K(L^2)\)?
+
+  \item We are assuming that \(\beta, \tau\) and \(\sigma^2\) are independent? Independent given \(Y\) and/or \(X\)?
+
+  \item How many samples is it usually enough to successfully predict a relatively simple model? What about the execution time?
+
+  \item How to set a sensible value for \(g\) and \(b_0\)? Cross-validation?
+
+  \item Which MCMC sampling algorithm should we use?
+
+  \item Can we ignore the posterior on \(\sigma^2\) and treat it as a nuisance parameter?
+
+  \item \textbf{How to include the case in which} \(\bm{\mathds EX(t)\neq 0}\) \textbf{and/or} \(\bm{\mathds EY\neq 0}\)?
+Do we add a term \(\mu\) in the model and impose a prior on it as well?
+How would we treat the terms \(X_i(t_j) - m(t_j)\) resulting from \(\Psi^{-1}_{X_i}(k(\cdot, t_j))\)?
+Can we just subtract the sample mean to every sample trajectory and response and work with our current model?
+
+  \item How do we get an estimate of the parameters given the posterior distribution?
+
+  \item How do we try the algorithm in a more realistic setting? Which is the standard error measure for the latent parameters?
+
+\end{enumerate}
+
+\newpage
+
+\section{Bayesian Linear Regression}
+
+\marginnote{We assume the functional regressor \(X=X(t)\) to be an \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), a vanishing mean function and a continuous covariance function \(k=k(s, t)\).}
+
+We consider the RKHS-based functional linear regression model given by\sidenote[][*7]{This formulation comes from considering the (centered) model \(Y=\Psi_X^{-1}(\alpha) + \eps\), where \(\eps\) is an independent normal error term with \(\E\eps=0\) and \(\var(\eps)=\sigma^2\).}
+\begin{equation}
+  \label{eq:model_lin}
+  Y_i \mid \theta, X_i \stackrel{\text{i.i.d.}}{\sim} \mathcal N(\Psi_{X_i}^{-1}(\alpha), \sigma^2),
+\end{equation}
+where \(\Psi_X\) is \textit{Loeve's isometry} and \(\alpha \in \mathcal H_{k}\) is the regression coefficient function.
+The main idea is to restrict \(\alpha\) to the linear span of a finite subset of evaluations of the kernel \(\{k(\cdot, t_j)\}_j\), since the whole set of such evaluations is dense in \(\mathcal H_k\).
+Specifically, we only consider functions of the form
+\begin{equation}
+  \label{eq:alpha}
+  \alpha(t) = \sum_{j=1}^p \beta_j k(t, t_j),
+\end{equation}
+where \(p\) is a fixed hyperparameter. Under these assumptions, the parametric space is the set \(\Theta \subseteq \R^p \times [0, 1]^p \times \R^+\), with elements
+\[
+  \theta = (\beta_1, \dots, \beta_p, t_1,\dots, t_p, \sigma^2)' \equiv (\beta, \tau, \sigma^2)'.
+\]
+
+Following a Bayesian approach, we can set a prior distribution on the parameters, such as:
+\begin{align*}
+  f(\sigma^2)              & \propto 1/\sigma^2,                                                     \\
+  \tau                     & \sim \mathscr U([0, 1]^p),                                              \\
+  \beta\mid \tau, \sigma^2 & \sim \mathcal N(b_0, g\sigma^2(\mathcal X_\tau' \mathcal X_\tau)^{-1}),
+\end{align*}
+where\sidenote[][*1]{This is because of the correspondence between \(k(\cdot, t_j)\) and \(X(t_j)\), which becomes relevant when plugging the approximation~\eqref{eq:alpha} into the model~\eqref{eq:model_lin}.}
+\(\mathcal X_\tau(i, j)=X_i(t_j)\).
+The prior on \(\beta\) is known as Zellner's \(g\)-prior owing to the main hyperparameter \(g\in \R^+\), which controls the confidence we have in the initial estimate \(b_0 \in \R^p\) for the true value of the coefficients.
+The choice of a non-informative improper prior for \(\sigma^2\) is of course subject to change.
+All in all, the prior distribution for the parameters becomes
+\begin{equation}
+  \label{eq:prior}
+  \pi(\theta) = \pi(\beta, \tau, \sigma^2) = \pi(\beta \mid \tau, \sigma^2)\pi(\tau \mid \sigma^2)\pi(\sigma^2).
+\end{equation}
+
+For the inference step we rely on Bayes' theorem, which in the case of i.i.d. samples amounts to
+\[
+  \pi(\theta \mid Y_1, \dots, Y_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid \theta) \right]\pi(\theta),
+\]
+that is, the \textit{posterior} is proportional to the \textit{likelihood} times the \textit{prior}.
+For convenience, let \(G_\tau=\mathcal X_\tau' \mathcal X_\tau\) and \(Y = (Y_1,\dots, Y_n)'\).
+Then, substituting the expressions derived for our model above, we get:
+\begin{widepar}
+  \begin{align*}
+    \pi(\theta \mid Y) & \propto \left[ \prod_{i=1}^n  \frac{1}{\sqrt{2\pi}\sigma} \exp\left\{ -\frac{1}{2\sigma^2} \left(Y_i - \sum_{j=1}^{p}\beta_j X_i(t_j)\right)^2 \right\}\right]\frac{|G_\tau|^{1/2}}{(2\pi g)^{p/2}\sigma^{p+2}} \exp\left\{ -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0)\right\} \\
+                       & \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n+2}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right) \right\}.
+  \end{align*}
+\end{widepar}
+Note that the resulting expression is not a true distribution, since we have omitted the possibly intractable integral related to the normalizing constant.
+Nevertheless, we can sample from the posterior distribution\sidenote[][*0]{Provided that the \textit{marginal likelihood} \(\int_\Theta \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using several Markov chain Monte Carlo (MCMC) sampling algorithms, which are discussed in \refsubsec{subsec:mcmc}.
+
+Finally, to recover the marginal posterior distribution of each parameter, we ``only'' need to integrate out the desired quantities, e.g.:
+\[
+  \pi(\beta\mid Y) = \int_{\R^+}\int_{\R^p} \pi(\beta, \tau, \sigma^2\mid Y)\, d\tau\, d\sigma^2.
+\]
+However, since these integrals can be hard to compute, in most cases we can also exploit the MCMC sampling algorithms to approximate the marginal distributions directly, or even derive other Bayesian estimators such as the posterior expected value or a MAP\marginnote[*-1]{MAP stands for \textit{Maximum a Posteriori Probability}.} estimate for each parameter.
+\subsection{MCMC methods}
+\labsubsec{subsec:mcmc}
+
+In most MCMC methods it is preferable to work with the \textit{log-posterior} probability, both in terms of efficiency and stability. In our case, we have:
+\begin{widepar}
+  \[
+    \log \pi(\theta\mid Y) \propto \frac{1}{2}\log |G_\tau| - (p+n+2)\log \sigma -\frac{1}{2\sigma^2} \left(\|Y- \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right).
+  \]
+\end{widepar}
+
+\subsubsection{Gibbs sampling}
+
+Some notes:
+
+\begin{enumerate}
+  \item To compute initial values: random, or sample from prior.
+  \item It is more efficient to work with the natural logarithm of the objective function and then exponentiate when needed (to avoid overflow and underflow).
+  \item Gibbs sampling is better than standard Metropolis due to the curse of dimensionality and the concentration of density in small regions.
+  \item We can obtain the marginal posterior directly, considering only samples for each parameter.
+We can also compute other estimates such as expectation, or a MAP estimate, by binning samples and choosing the ones that occur the most.
+  \item \textbf{Always} allow for a burn-in initial period (how many samples?).
+  \item To compute expectation: average only every \(n\)-th value (because of correlation).
+\end{enumerate}
+
+Some possible modifications of the methods include:
+
+\begin{enumerate}
+  \item Use \textit{collapsed Gibbs} or \textit{blocked Gibbs} to sample from more than one variable at a time. For this it would be useful to derive a graphical Bayesian model for the parameters.
+  \item Use \textit{simulated annealing} to try to reduce the autocorrelation of samples.
+
+\end{enumerate}
+
+\subsubsection{Affine-invariant ensemble sampler}
+
+See \sidetextcite[*0]{foreman2013emcee}.
+
+\subsubsection{Resources for MCMC methods}
+
+Some resources on MCMC methods in Python are:
+
+\begin{enumerate}
+  \item \href{https://www.tweag.io/blog/2019-10-25-mcmc-intro1/}{A step-by-step tutorial}
+  \item \href{http://www.stat.columbia.edu/~gelman/research/published/nuts.pdf}{NUTS sampler}
+  \item \href{https://github.com/michaelnowotny/pyjags}{pyjags}
+  \item \href{https://www.cdslab.org/paramonte/index.html}{ParaMonte}
+  \item \href{https://github.com/prmiles/pymcmcstat}{pymcmcstat}
+  \item \href{https://docs.pymc.io/}{\textbf{PyMC3}}
+  \item \href{https://mc-stan.org/}{Stan}
+  \item \href{https://emcee.readthedocs.io/en/stable/tutorials/quickstart/#quickstart}{\textbf{emcee}}
+\end{enumerate}
+
+\subsection{Model extensions}
+
+Some immediate extensions to the model are the following:
+\begin{enumerate}
+
+  \item Include \(p\) in the set of model parameters and define a prior distribution on it as well.
+
+  \item Consider \textit{multiple} functional covariates \(\alpha_1(t), \dots, \alpha_m(t)\) and adapt the model (e.g.
+via Gibbs sampling) to this new setting, considering \(\theta=(\beta^{(1)}, \tau^{(1)}, \dots, \beta^{(m)}, \tau^{(m)}, \sigma^2)\).
+
+  \item Change the prior distribution of the parameters in various ways: impose an informative prior on \(\sigma^2\), change the prior on \(\tau\) to the Dirichlet distribution on the hypercube, or even introduce prior information on the kernel \(k\) when defining the prior on \(\beta\).
+
+  \item Substitute the matrix \(\mathcal X\) for \(\mathcal X + \eta I\) as a form of regularization, to avoid ill-conditioning problems and to increase the stability of the computations.
+
+  \item Derive Bayesian estimates of the support and/or the coefficient function by minimizing a loss function integrated with respect to the posterior distribution, as done in \sidetextcite[*-2]{grollemund2019bayesian}.
+In particular, it is worth exploring the \textit{prior and posterior predictive probabilities} of the parameters to gain insight into them (as explained \href{https://en.wikipedia.org/wiki/Bayesian_inference#Bayesian_prediction}{here}).
+
+\end{enumerate}
+
+\section{Bayesian Logistic Regression}
+
+-- Work in progress --
+
+%%%%%%%%%%%%%%%
+% BIBLIOGRAPHY
+%%%%%%%%%%%%%%%
+
 \defbibnote{bibnote}{Here are the references in citation order.\par\bigskip} % Prepend this text to the bibliography
 \renewcommand*{\bibfont}{\small}
-\printbibliography[title=Bibliography]% , prenote=bibnote]
+\printbibliography[title=Bibliography]%, prenote=bibnote]
 
+%%%%%%%%%%%%%%%
+% APPENDICES
+%%%%%%%%%%%%%%%
+
+\appendix
+
+\section{Notes on the theory of RKHS's}
+
+An important thing to note is that in general the trajectories of the process \(X\) \textbf{do not belong to the corresponding RKHS}.
+More precisely, \(X(t)\notin \mathcal H_k\) with probability \(1\) if \(\mathcal K\) has infinite non-zero eigenvalues.
+While a rigorous proof of this fact is given in \sidetextcite{hajek1962linear}, an heuristic argument presented in \sidetextcite{wahba1990spline} is as follows.
+Consider the Karhunen-Loève expansion of \(X\), i.e.,
+\[
+  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
+\]
+and the truncated version \(X_N(t)\) up to the \(N\)-th term.
+On the one hand, \(X_N(t)\to X(t)\) in the quadratic mean (for each fixed \(t\)) by the Karhunen-Loève theorem, but on the other hand we have
+\[
+  \mathds E\|X_N(\cdot)\|^2_k = \mathds E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
+\]
+
+Thus, functions in the RKHS are in a way \textit{smoother} than the trajectories of the process themselves.
+For convenience, if \(x=x(\omega)\) is a realization of \(X\) (for a fixed \(\omega\)), we will denote \(\langle x, f\rangle_k \equiv \left[\Psi_X^{-1}(f)\right](\omega)\).
+Some interesting properties stemming from this interpretation are the following (see \sidetextcite[][p.~974]{parzen1961approach}):
+
+\begin{enumerate}
+
+  \item \(\langle X, k(\cdot, t)\rangle_k = X(t) - m(t)\).
+
+  \item \(\mathds{E} \langle X, f\rangle_k = \langle m, f\rangle_k\).
+
+  \item \(\cov \left(\langle X, f \rangle_k , \langle X, g\rangle_k\right) = \langle f, g\rangle_k\).
+
+\end{enumerate}
+
+%\include{notes-rkhs.tex}
+
+\section{Further study}
+
+Here are some ideas and additional materials.
+
+\begin{enumerate}
+
+  \item \textbf{Asymptotic properties of standard Bayesian functional linear regression.} In the paper by Lian et al.
+(2016) they consider a Gaussian process prior on the slope coefficient and show some asymptotic properties of the posterior distribution (which is also Gaussian).
+
+  \item \textbf{Posterior concentration for a misspecified Bayesian regression model with functional covariates.} Theoretical paper by Abraham and Grollemund (2020).
+
+  \item \textbf{Bliss method in R.} \url{https://rdrr.io/cran/bliss/f/vignettes/BlissIntro.Rmd}
+
+\end{enumerate}
 \end{document}

### 2022-02-14 notes.tex
Start paper

@@
 
-\margintoc[*-2]
+\margintoc[*-7]
 
-\section{Questions}
-
-Here are some questions.
+\section{TODO}
 
@@
 
-  \item Why is it that \(k(\cdot, t) \notin \mathcal K(L^2)\)?
+\textbf{ANSWER}: Continuous linear functionals are only obtained by integrating a coefficient function against $x(t)$ (Riesz-Fisher), and it is not possible to select a $\beta(t)$ that produces finite linear combinations of the process after integrating.
 
-  \item We are assuming that \(\beta, \tau\) and \(\sigma^2\) are independent? Independent given \(Y\) and/or \(X\)?
+\item \(\alpha(t)=E [X(t)U]\) is the correlation/covariance of $X$ and $U$ (Berrendero et al. 2019, 2020), if we assume that $E[X(t) \varepsilon]=0$. Moreover, are we assuming $E[(X(t) - m(t))U)]=0$?
 
-  \item How many samples is it usually enough to successfully predict a relatively simple model? What about the execution time?
-
-  \item How to set a sensible value for \(g\) and \(b_0\)? Cross-validation?
-
-  \item Which MCMC sampling algorithm should we use?
-
-  \item Can we ignore the posterior on \(\sigma^2\) and treat it as a nuisance parameter?
-
-  \item \textbf{How to include the case in which} \(\bm{\mathds EX(t)\neq 0}\) \textbf{and/or} \(\bm{\mathds EY\neq 0}\)?
-Do we add a term \(\mu\) in the model and impose a prior on it as well?
-How would we treat the terms \(X_i(t_j) - m(t_j)\) resulting from \(\Psi^{-1}_{X_i}(k(\cdot, t_j))\)?
-Can we just subtract the sample mean to every sample trajectory and response and work with our current model?
-
-  \item How do we get an estimate of the parameters given the posterior distribution?
-
-  \item How do we try the algorithm in a more realistic setting? Which is the standard error measure for the latent parameters?
+\item Since the RMSE is measured on the same scale, with the same units as y, one can expect 68\% of the y values to be within 1 RMSE — given the data is normally distributed.
 
 \end{enumerate}
-
-\newpage
 
@@
 The prior on \(\beta\) is known as Zellner's \(g\)-prior owing to the main hyperparameter \(g\in \R^+\), which controls the confidence we have in the initial estimate \(b_0 \in \R^p\) for the true value of the coefficients.
-The choice of a non-informative improper prior for \(\sigma^2\) is of course subject to change.
+The choice of a non-informative improper prior for \(\sigma^2\) (known as Jeffreys prior) is of course subject to change.
 All in all, the prior distribution for the parameters becomes
@@
 
-\subsubsection{Resources for MCMC methods}
+\subsubsection{Probabilistic programming in Python: PyMC3}
 
-Some resources on MCMC methods in Python are:
-
-\begin{enumerate}
-  \item \href{https://www.tweag.io/blog/2019-10-25-mcmc-intro1/}{A step-by-step tutorial}
-  \item \href{http://www.stat.columbia.edu/~gelman/research/published/nuts.pdf}{NUTS sampler}
-  \item \href{https://github.com/michaelnowotny/pyjags}{pyjags}
-  \item \href{https://www.cdslab.org/paramonte/index.html}{ParaMonte}
-  \item \href{https://github.com/prmiles/pymcmcstat}{pymcmcstat}
-  \item \href{https://docs.pymc.io/}{\textbf{PyMC3}}
-  \item \href{https://mc-stan.org/}{Stan}
-  \item \href{https://emcee.readthedocs.io/en/stable/tutorials/quickstart/#quickstart}{\textbf{emcee}}
-\end{enumerate}
+See \sidetextcite[*0]{salvatier2016probabilistic}.
 
@@
   \item Derive Bayesian estimates of the support and/or the coefficient function by minimizing a loss function integrated with respect to the posterior distribution, as done in \sidetextcite[*-2]{grollemund2019bayesian}.
-In particular, it is worth exploring the \textit{prior and posterior predictive probabilities} of the parameters to gain insight into them (as explained \href{https://en.wikipedia.org/wiki/Bayesian_inference#Bayesian_prediction}{here}).
+In particular, it is worth exploring the \textit{prior and posterior predictive probabilities} of the parameters to gain insight into them (as explained \href{https://en.wikipedia.org/wiki/Bayesian_inference#Bayesian_prediction}{here} and illustrated \href{https://docs.pymc.io/en/stable/pymc-examples/examples/diagnostics_and_criticism/posterior_predictive.html}{here}).
 

### 2022-02-14 paper/paper.tex
Start paper

% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
%
\pubyear{2022}
\volume{TBA}
\issue{TBA}
\doi{0000}
\arxiv{2010.00000}
\firstpage{1}
\lastpage{1}

%
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}{Proposition}[section]

\renewcommand{\epsilon}{\varepsilon}

\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\endlocaldefs

%% Environment for colored comments
\newenvironment{comment}
{
\noindent \em \color{red}
}
{
\color{black}
}

%% Inline comments
\newcommand\incomment[1]{\color{red}[\textit{#1}]\color{black}}

\begin{document}

\begin{frontmatter}
\title{Bayesian RKHS-based methods in functional regression\support{Support information of the article.}}
\runtitle{Bayesian RKHS-based methods in functional regression }

\begin{aug}
\author{\fnms{First} \snm{Author}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{first@somewhere.com}},
\author{\fnms{Second} \snm{Author}\thanksref{addr1,t3,m1,m2}\ead[label=e2]{second@somewhere.com}}
\and
\author{\fnms{Third} \snm{Author}\thanksref{addr2,t1,m2}%
\ead[label=e3]{third@somewhere.com}%
\ead[label=u1,url]{http://www.foo.com}}

\runauthor{F. Author et al.}

\address[addr1]{Address of the First and Second authors
     Usually a few lines long
    \printead{e1} % print email address of "e1"
    \printead*{e2}
}

\address[addr2]{Address of the Third author
    Usually a few lines long
    Usually a few lines long
    \printead{e3}
    \printead{u1}
}

\thankstext{t1}{Some comment}
\thankstext{t2}{First supporter of the project}
\thankstext{t3}{Second supporter of the project}

\end{aug}

\begin{abstract}
In this work we propose a Bayesian approach for functional linear or logistic  regression models, based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's). These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional alternative to the classical functional regression paradigm. The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subset of the RKHS generated by the underlying covariance function. By imposing a suitable prior distribution on such RKHS, we can perform data-driven inference via standard Bayes methodology. The posterior distribution can be approximated from Markov Chain Monte Carlo (MCMC) methods. Several estimators  derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real datasets.
\end{abstract}

\begin{keyword}[class=MSC]
\kwd[Primary ]{62M20}
\kwd[; secondary ]{62F15}
\end{keyword}

\begin{keyword}
\kwd{functional data}
\kwd{linear regression}
\kwd{logistic regression}
\kwd{reproducing kernel Hilbert space}
\kwd{Bayesian inference}
\end{keyword}

\end{frontmatter}

\section{Introduction}

\subsection{A brief review of the theory of RKHS's}

The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces. In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and the covariance function \(K(s, t)= \mathbb E[(X(t) - m(t)(X(s) - m(s)))]\) are both continuous. To construct the RKHS associated with the covariance function, we start by defining the functional space \(H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\[
H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(\cdot, t_i), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\]
This space is endowed with the inner product
\[
\langle f, g\rangle_K = \sum_{i, j} a_i b_j K(s_i, t_j),
\]
given that \(f(\cdot)=\sum_i a_i K(\cdot, s_i) \) and \(g(\cdot)=\sum_j b_j K(\cdot, t_j)\). Lastly, we define \(H(K)\) to be the completion of \(H_0(K)\) under the norm induced by the scalar product defined above.

As we can see, the covariance function plays a crucial role in characterizing the RKHS. An operator closely related to this covariance function is the so-called covariance operator, i.e.:
\[
\mathcal Kf(\cdot) = \int_0^1 K(\cdot, t)f(t)\, dt.
\]
This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that \(X\) can be decomposed as
\[
X(t) = \sum_{j=1}^\infty \xi_j \phi_j(t),
\]
where \(\{\phi_j\}\) are the eigenfunctions of the covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\):

There are at least three alternative, equivalent ways of defining the RKHS associated with \(K\), and each of them gives a different insight into the relationship between the process itself and the underlying RKHS.

\begin{enumerate}
  \item Loeve's isometry
  \item Sum with lambda j
  \item Square root of K

\end{enumerate}3

\subsection{Overview of Markov chain Monte Carlo methods}

\section{Bayesian methodology for RKHS-based functional regression models}

Note that starting from a probability distribution \(\mathbb{P}_0\) on \(H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap H_0(K))\) for all Borel sets \(B\). The following result proves that any such extension is indeed unique.

\begin{prop} Let \(H\) be a separable Hilbert space and consider a dense subset \(H_0\subseteq H\). Then, any probability distribution on \((H, \mathcal{B}(H))\) is uniquely determined by its restriction to \((H_0, \mathcal B(H_0))\).
\end{prop}

\begin{comment}
  ¿Es necesario especificar que \(\mathcal B\) es la sigma álgebra de Borel?
\end{comment}

\begin{comment}
    ¿La prueba va en el cuerpo del artículo, o en un anexo/material suplementario?
\end{comment}

\begin{proof}

First, note that Theorem 7.1.2 in \citet[p.~177]{hsing2015theoretical} establishes that the distribution of any random variable \(X: (\Omega, \mathcal A, \mathbb{P})\to (H, \mathcal B(H))\) is uniquely determined by the distribution of its random projections \(\langle X, h\rangle_H\) for all \(h \in H\).
Consider two random variables \(\alpha\) and \(\beta\), taking values in \(H\) and equally distributed on \(H_0\), and let \(h \in H\). Since \(H_0\) is dense in \(H\), there is a sequence \(\{h_n\}\subseteq H_0\) such that \(\|h_n - h\|_H \to 0\), and necessarily \(\langle \alpha, h_n\rangle_H \equiv \langle \beta, h_n\rangle_H\) for all \(n\), where the equality is in distribution.

Observe now that \(\langle \alpha, h_n\rangle_H \overset{P}{\to} \langle \alpha, h\rangle_H\), since
\[
\forall \epsilon > 0 \quad \mathbb{P}\{|\langle \alpha, h_n - h\rangle_H| > \epsilon\} \leq \mathbb{P}\{\|\alpha\|_H \|h_n - h\|_H > \epsilon\} \to 0,
\]
and equivalently \(\langle \beta, h_n \rangle_H \overset{P}{\to} \langle \beta, h\rangle_H\), so that \(\langle \alpha, h \rangle_H \equiv \langle \beta,h\rangle_H\). As this is valid for every \(h \in H\), we can conclude that \(\alpha \equiv \beta\) on \(H\).
\end{proof}

\subsection{Functional linear regression}

\subsection{Functional logistic regression}

\section{Experimental results}

\subsection{Simulation studies}

\subsection{Application to real data}

\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supplementary Material, if any, should   %%
%% be provided in {supplement} environment  %%
%% with title and short description.        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{supplement}
\stitle{Title of Supplement A}
\sdescription{Short description of Supplement A.}
\end{supplement}

\bibliographystyle{ba}
\bibliography{sample}

\begin{acks}[Acknowledgments]
This is an acknowledgements section.
\end{acks}

\end{document}

### 2022-02-19 notes-rkhs.tex
Reorganize and minor corrections

@@
-% \documentclass[11pt]{article}
-% \usepackage[margin=1.3in]{geometry}
-% %\usepackage[compact,small]{titlesec}
-% %--------------------------------------------------------
-% \usepackage{sectsty,todonotes}
-% \sectionfont{\normalsize\bfseries}
-% \subsectionfont{\normalsize\sf}
-% %----------------------------------------------------
-% \usepackage[mathcal]{euscript}
-%
-% \usepackage{bm}                     %AMS Packages
-% \usepackage{amsfonts, framed}
-% \usepackage{amssymb}
-% \usepackage{amsmath}
-% \usepackage{amsthm}
-% \usepackage{graphicx}               %Graphics
-% %\usepackage{pstricks,pst-plot}
-% %\usepackage{wrapfig}
-% \usepackage{natbib}                 %Citation style
-% \usepackage{colortbl}               %Tables
-% \usepackage{booktabs}
-% \usepackage{hyperref}
-% \setlength{\bibsep}{1pt}
-% %\usepackage{hypernat}
-%
-% % Definiciones
-%
-% \definecolor{violet}{rgb}{0.7,0,0.6}
-% \newcommand{\explainindetail}[1]{\todo[color=violet]{#1}}
-% \newcommand{\nat}{\mathbb{N}}
-% \newcommand{\eps}{\varepsilon}
-% \newcommand{\ind}{\mathbb{I}}
-% \newcommand{\pr}{\mathbb{P}}
-% \newcommand{\B}{B}
-% \newcommand{\A}{\mathcal{A}}\usepackage{amsfonts}
-% \usepackage{amssymb}
-% \usepackage{amsmath}
-% \usepackage{amsthm}
-% \newcommand{\E}{\mathcal{E}}
-% \newcommand{\V}{\mathbb{V}}
-% \newcommand{\F}{\mathfrak{F}}
-% \newcommand{\G}{\mathbb{G}}
-% \newcommand{\M}{\mathcal{M}}
-% \newcommand{\X}{\mathcal{X}}
-% \newcommand{\Z}{\mathcal{Z}}
-% \newcommand{\T}{\mathcal{T}}
-% \newcommand{\Y}{\mathcal{Y}}
-% \newcommand{\C}{\mathcal{C}}
-% \newcommand{\ds}{\displaystyle}
-% \newcommand{\D}{\mathcal{D}}
-% \newcommand{\deb}{\stackrel{\mathcal{L}}{\longrightarrow}}
-% \newcommand{\Vor}{\mbox{Vor}}
-% \newcommand{\ang}{\measuredangle}
-% \def\cals_+{{\cals_+}}
-% \def\cala{{\mathcal{A}}}
-% \def\calb{{\mathcal{B}}}
-% \def\calc{{\mathcal{C}}}
-% \def\cald{{\mathcal{D}}}
-% \def\calf{{\mathcal{F}}}
-% \def\calh{{\mathcal{H}}}
-% \def\call{{\mathcal{L}}}
-% \def\cals{{\mathcal{S}}}
-% \def\calv{{\mathcal{V}}}
-% \def\calg{{\mathcal{G}}}
-% \def\calp{{\mathcal{P}}}
-% \def\cale{{\mathcal{E}}}
-% \def\calo{{\mathcal{O}}}
-% \def\calj{{\mathcal{J}}}
-% \def\caln{{\mathcal{N}}}
-% \def\calm{{\mathcal{M}}}
-%
-% \hypersetup{
-% colorlinks=true,
-% citecolor=violet
-% }
-%
-% \begin{document}
+RKHS's are spaces of ``regular'' FUNCTIONS (not equivalence classes in $L^2[0,1]$) with a different norm to that inherited from $L^2[0,1]$.
+See \sidetextcite{paulsen2016introduction} for a more detailed introduction.
 
-\begin{center}
-\large \bf  Preliminaries on RKHS's\rm
-\end{center}
-
-RKHS's are spaces of ``regular'' FUNCTIONS (not equivalence classes in $L^2[0,1]$) with a different norm to that inherited from $L^2[0,1]$.
-
-\section*{Definition}
+\subsection*{Definition}
 
@@
 
-	Let $\calh_0(K):=\{f\in L^2[0,1] \ : \ f(\cdot)=\sum_{i=1}^n a_i K(t_i,\cdot),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\}$, be
+	Let $\mathcal H_0(K):=\{f\in L^2[0,1] \ : \ f(\cdot)=\sum_{i=1}^n a_i K(t_i,\cdot),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\}$, be
 	the space of all finite linear combinations of evaluations of $K$. This space is endowed with the inner product
@@
 
-	Then, the RKHS associated with $K$ is defined as the completion of $\calh_0(K)$.
-In other words, $\calh(K)$ is made of all functions obtained as pointwise limits of Cauchy sequences in $\calh_0(K)$.
-The inner product is extended accordingly to the whole space $\calh(K)$.
+	Then, the RKHS associated with $K$ is defined as the completion of $\mathcal H_0(K)$.
+In other words, $\mathcal H(K)$ is made of all functions obtained as pointwise limits of Cauchy sequences in $\mathcal H_0(K)$.
+The inner product is extended accordingly to the whole space $\mathcal H(K)$.
 
-\section*{Reproducing property}
+\subsection*{Reproducing property}
 
 	These spaces are named after the so-called \textit{reproducing property},
-	$\langle f, K(s,\cdot)\rangle_K = f(s)$, for all  $f\in\calh(K), s\in[0,1]$, which is particularly important in the applications.
+	$\langle f, K(s,\cdot)\rangle_K = f(s)$, for all  $f\in\mathcal H(K), s\in[0,1]$, which is particularly important in the applications.
 On account of this property it is sometimes said that the RKHS are spaces of ``true functions'', in the sense that the pointwise values $f(s)$ at a given $s$ do matter, in contrast with $L^2[0,1]$ whose elements are in fact equivalence classes of functions.
@@
 
-\section*{Loève's isometry}
+\subsection*{Loève's isometry}
 
@@
 Define
-	$$\call_0(X) = \big\{U\in L^2(\Omega) \ : \ U=\sum_{i=1}^n a_i \big(X(t_i)-m(t_i)\big),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\big\},$$
-	where $m(t)={\mathbb E}[X(t)]$, and let $\call(X)$ be the completion of $\call_0(X)$ in  $L^2(\Omega)$; hence, in other words, is a closed subspace of $L^2(\Omega)$ defined as the linear span of the centred one-dimensional marginals of the process $X$.
-It turns out that the transformation $\Psi_X$, from  $\call(X)$ to $\calh(K)$, defined by
+	$$\mathcal L_0(X) = \big\{U\in L^2(\Omega) \ : \ U=\sum_{i=1}^n a_i \big(X(t_i)-m(t_i)\big),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\big\},$$
+	where $m(t)={\mathbb E}[X(t)]$, and let $\mathcal L(X)$ be the completion of $\mathcal L_0(X)$ in  $L^2(\Omega)$; hence, in other words, is a closed subspace of $L^2(\Omega)$ defined as the linear span of the centred one-dimensional marginals of the process $X$.
+It turns out that the transformation $\Psi_X$, from  $\mathcal L(X)$ to $\mathcal H(K)$, defined by
 	\begin{equation}\label{eq:isometry}
-	\Psi_X(U)(s) = {\mathbb E}[ U(X(s)-m(s))] = \langle U, X(s)-m(s) \rangle \in \calh(K), \ \text{ for } U\in \call(X)
+	\Psi_X(U)(s) = {\mathbb E}[ U(X(s)-m(s))] = \langle U, X(s)-m(s) \rangle \in \mathcal H(K), \ \text{ for } U\in \mathcal L(X)
 	\end{equation}
-	is an isometry (sometimes called Lo\`eve's isometry) between $\call(X)$ and $\calh(K)$, that is, $\Psi_X(U)$ is bijective and preserves the inner product.
-As a consequence, the Hilbert spaces $\call(X)$ and $\calh(K)$ can be identified.
+	is an isometry (sometimes called Lo\`eve's isometry) between $\mathcal L(X)$ and $\mathcal H(K)$, that is, $\Psi_X(U)$ is bijective and preserves the inner product.
+As a consequence, the Hilbert spaces $\mathcal L(X)$ and $\mathcal H(K)$ can be identified.
 
-	\textbf{In more informal terms, this isometry is nothing but the completion} of
+	In more informal terms, this isometry is nothing but the completion of
 	$$
@@
 
-\section*{Alternative definitions}
+\subsection*{Alternative definitions}
 
@@
 
-\section*{General properties}
+\subsection*{General properties}
 
-  An important thing to note is that in general the trajectories of the process $X$ \textbf{do not belong to the corresponding RKHS}.
-More precisely, $X(t)\notin \mathcal H(K)$ with probability $1$ if $\mathcal K$ has infinite non-zero eigenvalues.
-While a rigorous proof of this fact is given in \citet{hajek1962}, an heuristic argument presented in \citet{wahba1990} is as follows.
-Consider the Karhunen-Loève expansion of $X$, i.e.,
+An important thing to note is that in general the trajectories of the process \(X\) \textbf{do not belong to the corresponding RKHS}.
+More precisely, \(X(t)\notin \mathcal H_k\) with probability \(1\) if \(\mathcal K\) has infinite non-zero eigenvalues.
+While a rigorous proof of this fact is given in \sidetextcite{hajek1962linear}, an heuristic argument presented in \sidetextcite{wahba1990spline} is as follows.
+Consider the Karhunen-Loève expansion of \(X\), i.e.,
 \[
@@
 \]
-
-  and the truncated version $X_N(t)$ up to the $N$-th term. On the one hand, $X_N(t)\to X(t)$ in the quadratic mean (for each fixed $t$) by the Karhunen-Loève theorem, but on the other hand we have
+and the truncated version \(X_N(t)\) up to the \(N\)-th term.
+On the one hand, \(X_N(t)\to X(t)\) in the quadratic mean (for each fixed \(t\)) by the Karhunen-Loève theorem, but on the other hand we have
 \[
-\mathbb E\|X_N(\cdot)\|^2_K = \mathbb E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
+  \mathds E\|X_N(\cdot)\|^2_k = \mathds E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
 \]
 
-  Thus, functions in the RKHS are in a way ``smoother'' than the trajectories of the process themselves.
-For convenience, if $x=x(\omega)$ is a realization of $X$ (for a fixed $\omega$), we will denote $\langle x, f\rangle_K \equiv \left[\Psi_X^{-1}(f)\right](\omega)$.
-Some interesting properties stemming from this interpretation are the following (see \citet[p.~974]{parzen1961b}):
+Thus, functions in the RKHS are in a way \textit{smoother} than the trajectories of the process themselves.
+For convenience, if \(x=x(\omega)\) is a realization of \(X\) (for a fixed \(\omega\)), we will denote \(\langle x, f\rangle_k \equiv \left[\Psi_X^{-1}(f)\right](\omega)\).
+Some interesting properties stemming from this interpretation are the following (see \sidetextcite[][p.~974]{parzen1961approach}):
 
@@
 
-    \item $\langle X, K(t, \cdot)\rangle_K = X(t) - m(t)$
+  \item \(\langle X, k(\cdot, t)\rangle_k = X(t) - m(t)\).
 
-    \item $\mathbb{E} \langle X, h\rangle_K = \langle m, h\rangle_K$
+  \item \(\mathds{E} \langle X, f\rangle_k = \langle m, f\rangle_k\).
 
-    \item $\operatorname{Cov}\left(\langle X, h \rangle_K , \langle X, g\rangle_K\right) = \langle h, g\rangle_K$
+  \item \(\cov \left(\langle X, f \rangle_k , \langle X, g\rangle_k\right) = \langle f, g\rangle_k\).
 
@@
 
-\section*{Convergence properties}
+\subsection*{Convergence properties}
 
@@
 		so $\Vert f_n\Vert_K \to 0$ implies $f_n(t)\to 0$ uniformly on $t$, that is, convergence in the RKHS norm \textbf{entails uniform convergence}.
-
-\end{itemize}
-
-\begin{thebibliography}{9}
-	\bibitem[\protect\citeauthoryear{Hsing \& Eubank}{2015}]{hsi15}
-	Hsing, T., \& Eubank, R. (2015). \textit{Theoretical foundations of functional data analysis, with an introduction to linear operators} (Vol. 997). John Wiley \& Sons.
-
-	\bibitem[\protect\citeauthoryear{Paulsen \& Raghupathi }{2015}]{pau16}
-	Paulsen, V. I., \& Raghupathi, M. (2016). \textit{An introduction to the theory of reproducing kernel Hilbert spaces} (Vol. 152). Cambridge University press.
-  \bibitem[\protect\citeauthoryear{Parzen}{1961}]{parzen1961b} Parzen, E. (1961). \textit{An Approach to Time Series Analysis}. The Annals of Mathematical Statistics 32(4), 951–989.
-
-  \bibitem[\protect\citeauthoryear{Wahba}{1990}]{wahba1990} Wahba G. (1990). \textit{Spline models for observational data}. CBMS 59, SIAM, Philadelphia.
-
-  \bibitem[\protect\citeauthoryear{Hajek}{1962}]{hajek1962} Hajek J. (1962). \textit{On linear Statistical problems in stochastic processes}. Chekoslovak Math. J., vol. 12, pp.404-443.
-
-\end{thebibliography}
-
-%\end{document}

### 2022-02-19 notes.tex
Reorganize and minor corrections

@@
 
-\section{TODO}
+\section{Functional Data Analysis}
 
-\begin{enumerate}
-
-\item Why does it not make sense to regress \(Y\) on a finite combination of projections of \(X\) whenever we consider the \(L^2\) model?
-
-\textbf{ANSWER}: Continuous linear functionals are only obtained by integrating a coefficient function against $x(t)$ (Riesz-Fisher), and it is not possible to select a $\beta(t)$ that produces finite linear combinations of the process after integrating.
-
-\item \(\alpha(t)=E [X(t)U]\) is the correlation/covariance of $X$ and $U$ (Berrendero et al. 2019, 2020), if we assume that $E[X(t) \varepsilon]=0$. Moreover, are we assuming $E[(X(t) - m(t))U)]=0$?
-
-\item Since the RMSE is measured on the same scale, with the same units as y, one can expect 68\% of the y values to be within 1 RMSE — given the data is normally distributed.
-
-\end{enumerate}
+See \sidetextcite{hsing2015theoretical} for a complete discussion on the subject.
 
@@
 
-\marginnote{We assume the functional regressor \(X=X(t)\) to be an \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), a vanishing mean function and a continuous covariance function \(k=k(s, t)\).}
+\marginnote{We assume the functional regressor \(X=X(t)\) to be an \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), a vanishing mean function and a continuous covariance function \(K=K(t, s)\).}
 
@@
 \end{equation}
-where \(\Psi_X\) is \textit{Loeve's isometry} and \(\alpha \in \mathcal H_{k}\) is the regression coefficient function.
-The main idea is to restrict \(\alpha\) to the linear span of a finite subset of evaluations of the kernel \(\{k(\cdot, t_j)\}_j\), since the whole set of such evaluations is dense in \(\mathcal H_k\).
+where \(\Psi_X\) is \textit{Loeve's isometry} and \(\alpha \in \mathcal H(K)\) is the regression coefficient function.
+The main idea is to restrict \(\alpha\) to the linear span of a finite subset of evaluations of the kernel \(\{K(\cdot, t_j)\}_j\), since the whole set of such evaluations is dense in \(\mathcal H(K)\).
 Specifically, we only consider functions of the form
@@
   \label{eq:alpha}
-  \alpha(t) = \sum_{j=1}^p \beta_j k(t, t_j),
+  \alpha(t) = \sum_{j=1}^p \beta_j K(t, t_j),
 \end{equation}
@@
 \end{align*}
-where\sidenote[][*1]{This is because of the correspondence between \(k(\cdot, t_j)\) and \(X(t_j)\), which becomes relevant when plugging the approximation~\eqref{eq:alpha} into the model~\eqref{eq:model_lin}.}
+where\sidenote[][*1]{This is because of the correspondence between \(K(\cdot, t_j)\) and \(X(t_j)\), which becomes relevant when plugging the approximation~\eqref{eq:alpha} into the model~\eqref{eq:model_lin}.}
 \(\mathcal X_\tau(i, j)=X_i(t_j)\).
@@
 
+\section{Notes on the theory of RKHS's}
+
+\input{notes-rkhs.tex}
+
 %%%%%%%%%%%%%%%
@@
 \appendix
-
-\section{Notes on the theory of RKHS's}
-
-An important thing to note is that in general the trajectories of the process \(X\) \textbf{do not belong to the corresponding RKHS}.
-More precisely, \(X(t)\notin \mathcal H_k\) with probability \(1\) if \(\mathcal K\) has infinite non-zero eigenvalues.
-While a rigorous proof of this fact is given in \sidetextcite{hajek1962linear}, an heuristic argument presented in \sidetextcite{wahba1990spline} is as follows.
-Consider the Karhunen-Loève expansion of \(X\), i.e.,
-\[
-  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
-\]
-and the truncated version \(X_N(t)\) up to the \(N\)-th term.
-On the one hand, \(X_N(t)\to X(t)\) in the quadratic mean (for each fixed \(t\)) by the Karhunen-Loève theorem, but on the other hand we have
-\[
-  \mathds E\|X_N(\cdot)\|^2_k = \mathds E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
-\]
-
-Thus, functions in the RKHS are in a way \textit{smoother} than the trajectories of the process themselves.
-For convenience, if \(x=x(\omega)\) is a realization of \(X\) (for a fixed \(\omega\)), we will denote \(\langle x, f\rangle_k \equiv \left[\Psi_X^{-1}(f)\right](\omega)\).
-Some interesting properties stemming from this interpretation are the following (see \sidetextcite[][p.~974]{parzen1961approach}):
-
-\begin{enumerate}
-
-  \item \(\langle X, k(\cdot, t)\rangle_k = X(t) - m(t)\).
-
-  \item \(\mathds{E} \langle X, f\rangle_k = \langle m, f\rangle_k\).
-
-  \item \(\cov \left(\langle X, f \rangle_k , \langle X, g\rangle_k\right) = \langle f, g\rangle_k\).
-
-\end{enumerate}
-
-%\include{notes-rkhs.tex}
 

### 2022-02-19 paper/paper.tex
Reorganize and minor corrections

@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
+
+The problem of inferring a scalar response \(Y\) from a functional covariate \(X(t)\)
+
+\subsubsection{Organization of the paper}
+
+In Section~\ref{sec:intro} we review...
 
@@
 The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces.
-In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and the covariance function \(K(s, t)= \mathbb E[(X(t) - m(t)(X(s) - m(s)))]\) are both continuous.
-To construct the RKHS associated with the covariance function, we start by defining the functional space \(H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
+In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and its covariance function \(K(t, s)= \mathbb E[(X(t) - m(t))(X(s) - m(s))]\) are both continuous.
+To construct the RKHS associated with the covariance function, we start by defining the functional space \(\mathcal H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
 \[
-H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(\cdot, t_i), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
+\mathcal H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
 \]
@@
 \[
-\langle f, g\rangle_K = \sum_{i, j} a_i b_j K(s_i, t_j),
+\langle f, g\rangle_K = \sum_{i, j} a_i b_j K(t_i, s_j),
 \]
-given that \(f(\cdot)=\sum_i a_i K(\cdot, s_i) \) and \(g(\cdot)=\sum_j b_j K(\cdot, t_j)\).
-Lastly, we define \(H(K)\) to be the completion of \(H_0(K)\) under the norm induced by the scalar product defined above.
+given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
+Lastly, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.
 
-As we can see, the covariance function plays a crucial role in characterizing the RKHS. An operator closely related to this covariance function is the so-called covariance operator, i.e.:
+\begin{comment}
+  \(K\) is positive and symmetric. We will refer to it as a \textit{kernel}.
+\end{comment}
+
+\begin{comment}
+  Reproducing property? Convergence entails uniform convergence?
+\end{comment}
+
+As we can see, the covariance function plays a crucial role in characterizing the RKHS. A compact operator closely related to this covariance function is the so-called covariance operator, i.e.:
 \[
-\mathcal Kf(\cdot) = \int_0^1 K(\cdot, t)f(t)\, dt.
+\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1].
 \]
 This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that \(X\) can be decomposed as
-\[
+\begin{equation}\label{eq:karhunen-loeve}
 X(t) = \sum_{j=1}^\infty \xi_j \phi_j(t),
-\]
-where \(\{\phi_j\}\) are the eigenfunctions of the covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\):
+\end{equation}
+where \(\phi_j\) are the (orthonormal) eigenfunctions of the covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\).
+This expansion is the main justification of many FDA techniques, such as functional principal component analysis \incomment{cita}.
 
@@
 \begin{enumerate}
-  \item Loeve's isometry
-  \item Sum with lambda j
-  \item Square root of K
+  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process \(X - m\) in the space of all random variables with finite second moment, \(L^2(\Omega)\) \incomment{cita}.
+This isometry is essentially the completion of the correspondence
+  \begin{equation}\label{eq:loeve}
+  \sum_{i=1}^p a_i (X(t_i) - m(t_i)) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot).
+\end{equation}
+  \item The space \(\mathcal H(K)\) can also be defined as the image of the square root of the covariance operator, i.e.:
+  \begin{equation}\label{eq:rkhs-square-root}
+  \mathcal H(K) = \mathcal K^{1/2}(L^2[0, 1]).
+\end{equation}
 
-\end{enumerate}3
+  \item Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \incomment{cita}, we can also express
+  \begin{equation}\label{rkhs-sum-lambda}
+    \mathcal H(K) = \left\{f \in L^2[0, 1]: \sum_{j=1}^\infty \frac{\langle f, \phi_j\rangle}{\lambda_j} < \infty \right\},
+  \end{equation}
+  with the corresponding inner product
+  \[
+  \langle f, g\rangle_K = \sum_{j=1}^\infty \frac{\langle f, \phi_j\rangle\langle g, \phi_j\rangle}{\lambda_j}.
+  \]
+Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly \incomment{and?}.
+\end{enumerate}
 
-\subsection{Overview of Markov chain Monte Carlo methods}
+\begin{comment}
+  Trajectories not in RKHS
+\end{comment}
+
+\begin{comment}
+  Small RKHS norm implies small variations and hence a higher degree of smoothness.
+Specifically, \(\|f\|_K\) controls how fast the function varies over the input space with respect to the geometry defined by the kernel:
+  \[
+  |f(t) - f(s)| = |\langle f, K_t - K_s\rangle_K| \leq \|f\|_K \|K_t - K_s\|_K.
+  \]
+\end{comment}
 
@@
 
-Note that starting from a probability distribution \(\mathbb{P}_0\) on \(H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap H_0(K))\) for all Borel sets \(B\).
+The general idea explored in this work revolves around imposing a prior distribution on the functional space \(\mathcal H_0(K)\), which we know is dense in \(\mathcal H(K)\).
+Specifically, we consider a functional parameter of the form
+\begin{equation}\label{eq:parameter-rkhs}
+  \beta(t) = \sum_{i=1}^p \beta_i K(t_i, \cdot),
+\end{equation}
+and regard \(p, \beta_i\) and \(t_i\) as free parameters in the parameter space \(\Theta\).
+
+Note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B\).
 The following result proves that any such extension is indeed unique.
@@
 
+\begin{comment}
+  Note that in an RKHS the evaluation functional is continuous, so a small perturbation \(f\mapsto f + \delta f\) results in a small change in \(f(x)\) for all \(x\).
+It makes sense to choose the parameter in \(\mathcal H(K)\).
+Also smooth functions and linear: the model is \textit{linear} because of the correspondence between the RKHS and the linear span of the process (Loeve's isometry).
+\end{comment}
+
+\begin{comment}
+  Overview of MCMC methods.
+\end{comment}
+
 \subsection{Functional linear regression}
+
+Berrendero et al. (2020).
 
@@
 \bibliographystyle{ba}
-\bibliography{sample}
+\bibliography{bibliography}
 

### 2022-02-25 paper/paper.tex
Update paper

@@
 \newcommand{\E}{\mathbb{E}}
-\endlocaldefs
+
+%% Scalar product
+\newcommand\dotprod[2]{\left\langle #1,\, #2 \right\rangle}
 
@@
 \newcommand\incomment[1]{\color{red}[\textit{#1}]\color{black}}
+\endlocaldefs
 
@@
 
+\begin{comment}
+  - Grollemund et al.
+
+  - FDA (Cuevas et al. 2013)
+
+  - L2 model and finite model as a particular case of RKHS model. Define dot-product in L2.
+\end{comment}
+
 \subsubsection{Organization of the paper}
@@
 \subsection{A brief review of the theory of RKHS's}
+
+\begin{comment}
+  Es posible que esta sección sea demasiado larga...?
+\end{comment}
 
@@
 \[
-\langle f, g\rangle_K = \sum_{i, j} a_i b_j K(t_i, s_j),
+\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
 \]
@@
 Lastly, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.
+As it turns out, functions in this space verify the \textit{reproducing property}, thus giving it its name:
+\begin{equation}\label{eq:reproducing-property}
+  \dotprod{K(t, \cdot)}{f}_K = f(t), \quad \text{for all } f \in \mathcal H(K), \ t \in [0, 1].
+\end{equation}
 
-\begin{comment}
-  \(K\) is positive and symmetric. We will refer to it as a \textit{kernel}.
-\end{comment}
-
-\begin{comment}
-  Reproducing property? Convergence entails uniform convergence?
-\end{comment}
-
-As we can see, the covariance function plays a crucial role in characterizing the RKHS. A compact operator closely related to this covariance function is the so-called covariance operator, i.e.:
+As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
+A linear compact operator closely related to this covariance function is the so-called covariance operator, i.e.:
 \[
@@
 \]
-This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that \(X\) can be decomposed as
+This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that a centered process \(X\) can be decomposed as
 \begin{equation}\label{eq:karhunen-loeve}
@@
 \end{equation}
-where \(\phi_j\) are the (orthonormal) eigenfunctions of the covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\).
-This expansion is the main justification of many FDA techniques, such as functional principal component analysis \incomment{cita}.
+where \(\phi_j\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\).
+This expansion is the main justification of many FDA techniques, such as functional principal component analysis (FPCA) (see for example \incomment{cita}).
 
@@
 \begin{enumerate}
-  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process \(X - m\) in the space of all random variables with finite second moment, \(L^2(\Omega)\) \incomment{cita}.
+  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \incomment{cita}.
 This isometry is essentially the completion of the correspondence
-  \begin{equation}\label{eq:loeve}
-  \sum_{i=1}^p a_i (X(t_i) - m(t_i)) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot).
+  \[
+  \sum_{i=1}^p a_i (X(t_i) - m(t_i)) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
+\]
+and can be formally defined as (see \incomment{cita})
+\begin{equation}\label{eq:loeves-isometry}
+  \Psi_X(U)(t) = \E[U(X(t) - m(t))], \quad U \in \mathcal L(X).
 \end{equation}
@@
 \end{equation}
+In this case, the inner product can be expressed as
+\[
+\dotprod{f}{g}_K = \dotprod{\mathcal K^{1/2}(f)}{\mathcal K^{1/2}(g)}.
+\]
 
@@
   \begin{equation}\label{rkhs-sum-lambda}
-    \mathcal H(K) = \left\{f \in L^2[0, 1]: \sum_{j=1}^\infty \frac{\langle f, \phi_j\rangle}{\lambda_j} < \infty \right\},
+    \mathcal H(K) = \left\{f \in L^2[0, 1]: \sum_{j=1}^\infty \frac{\dotprod{f}{\phi_j}^2}{\lambda_j} < \infty \right\},
   \end{equation}
@@
   \[
-  \langle f, g\rangle_K = \sum_{j=1}^\infty \frac{\langle f, \phi_j\rangle\langle g, \phi_j\rangle}{\lambda_j}.
+  \dotprod{f}{g}_K = \sum_{j=1}^\infty \frac{\dotprod{f}
+  {\phi_j}\dotprod{g}{\phi_j}}{\lambda_j}.
   \]
-Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly \incomment{and?}.
+Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
 \end{enumerate}
 
-\begin{comment}
-  Trajectories not in RKHS
-\end{comment}
+Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since in general the trajectories do not belong to the RKHS with probability one (see \incomment{cita}).
+As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined for \(x\) a realization of \(X\).
+However, following Parzen's approach \incomment{cita}\ we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
+This notation often proves to be useful and convenient.
+
+\section{Bayesian methodology for RKHS-based functional regression models}
+
+The classical functional linear regression model, firstly introduced in \incomment{cita Ramsay \& Silverman}, can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
+\begin{equation}\label{eq:l2-model}
+Y = \dotprod{X}{\beta} + \epsilon = \int_0^1 X(t)\beta(t)\, dt + \epsilon.
+\end{equation}
+In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinte-dimensional space \(L^2[0, 1]\).
+Aside from the fact that functional (i.e.
+infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle inconveniences with this model.
+For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case (see \incomment{Berrendero et al 2020, on a general definition of the flm.}).
+
+It turns out that a natural alternative to the \(L^2\) mode is the RKHS model.....
+Not only simpler, but also smoother and more manageable functions.
+The model remains linear because of the correspondence between the RKHS and the linear span of the process (Loève’s isometry).
 
 \begin{comment}
-  Small RKHS norm implies small variations and hence a higher degree of smoothness.
-Specifically, \(\|f\|_K\) controls how fast the function varies over the input space with respect to the geometry defined by the kernel:
-  \[
-  |f(t) - f(s)| = |\langle f, K_t - K_s\rangle_K| \leq \|f\|_K \|K_t - K_s\|_K.
-  \]
+  - The RKHS model includes both L2 model and finite model. Cite Berrendero et al 2020.
+  - Parametric approach with \(\beta \in H_0(K)\) (further constrain the parameter to this subset of the RKHS, which is dense by construction).
+Then we can impose a prior on the finite-dimensional parameter space.
 \end{comment}
-
-\section{Bayesian methodology for RKHS-based functional regression models}
 
@@
 
-Note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B\).
+Note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
 The following result proves that any such extension is indeed unique.
@@
 \end{prop}
-
-\begin{comment}
-  ¿Es necesario especificar que \(\mathcal B\) es la sigma álgebra de Borel?
-\end{comment}
-
 \begin{comment}
@@
 
-First, note that Theorem 7.1.2 in \citet[p.~177]{hsing2015theoretical} establishes that the distribution of any random variable \(X: (\Omega, \mathcal A, \mathbb{P})\to (H, \mathcal B(H))\) is uniquely determined by the distribution of its random projections \(\langle X, h\rangle_H\) for all \(h \in H\).
+First, note that Theorem 7.1.2 in \citet[p.~177]{hsing2015theoretical} establishes that the distribution of any random variable \(X: (\Omega, \mathcal A, \mathbb{P})\to (H, \mathcal B(H))\) is uniquely determined by the distribution of its random projections \(\dotprod{X}{h}_H\) for all \(h \in H\).
 Consider two random variables \(\alpha\) and \(\beta\), taking values in \(H\) and equally distributed on \(H_0\), and let \(h \in H\).
-Since \(H_0\) is dense in \(H\), there is a sequence \(\{h_n\}\subseteq H_0\) such that \(\|h_n - h\|_H \to 0\), and necessarily \(\langle \alpha, h_n\rangle_H \equiv \langle \beta, h_n\rangle_H\) for all \(n\), where the equality is in distribution.
+Since \(H_0\) is dense in \(H\), there is a sequence \(\{h_n\}\subseteq H_0\) such that \(\|h_n - h\|_H \to 0\), and necessarily \(\dotprod{\alpha}{h_n}_H \equiv \dotprod{\beta}{h_n}_H\) for all \(n\), where the equality is in distribution.
 
-Observe now that \(\langle \alpha, h_n\rangle_H \overset{P}{\to} \langle \alpha, h\rangle_H\), since
+Observe now that \(\dotprod{\alpha}{h_n}_H \overset{P}{\to} \dotprod{\alpha}{h}_H\), since
 \[
-\forall \epsilon > 0 \quad \mathbb{P}\{|\langle \alpha, h_n - h\rangle_H| > \epsilon\} \leq \mathbb{P}\{\|\alpha\|_H \|h_n - h\|_H > \epsilon\} \to 0,
+\forall \epsilon > 0 \quad \mathbb{P}\{|\dotprod{\alpha}{h_n - h}_H| > \epsilon\} \leq \mathbb{P}\{\|\alpha\|_H \|h_n - h\|_H > \epsilon\} \to 0,
 \]
-and equivalently \(\langle \beta, h_n \rangle_H \overset{P}{\to} \langle \beta, h\rangle_H\), so that \(\langle \alpha, h \rangle_H \equiv \langle \beta,h\rangle_H\).
+and equivalently \(\dotprod{\beta}{h_n}_H \overset{P}{\to} \dotprod{\beta}{h}_H\), so that \(\dotprod{\alpha}{h}_H \equiv \dotprod{\beta}{h}_H\).
 As this is valid for every \(h \in H\), we can conclude that \(\alpha \equiv \beta\) on \(H\).
@@
 
-\begin{comment}
-  Note that in an RKHS the evaluation functional is continuous, so a small perturbation \(f\mapsto f + \delta f\) results in a small change in \(f(x)\) for all \(x\).
-It makes sense to choose the parameter in \(\mathcal H(K)\).
-Also smooth functions and linear: the model is \textit{linear} because of the correspondence between the RKHS and the linear span of the process (Loeve's isometry).
-\end{comment}
-
-\begin{comment}
-  Overview of MCMC methods.
-\end{comment}
+\subsubsection{Overview of MCMC methods}
 
@@
 
-Berrendero et al. (2020).
+\begin{comment}
+  - Estimators derived.
+\end{comment}
 
@@
 
+In this case the model/functional equation in $L^2$ is....... and the same problems apply, plus the MLE existence problem. An RKHS aproach is given by:.......
+
+ and this solves the problems (see Bueno-Larraz 2021)
+
+\begin{comment}
+  - Estimators derived.
+\end{comment}
+
+\subsection{Model choice and validation}
+
+\begin{comment}
+  - Choice of p, MLE, BIC, ...
+
+  - Posterior predictive checks, ¿Bayesian p-values?, ...
+\end{comment}
+
 \section{Experimental results}
+
+\begin{comment}
+  Not only MSE/Acc but also credible intervals, points-of-impact model --> variable selection included.
+\end{comment}
 
@@
 \stitle{Title of Supplement A}
-\sdescription{Short description of Supplement A.}
+\sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}
 \end{supplement}

### 2022-02-25 paper/paper.tex
Update paper

@@
 
+\begin{comment}
+  Algunos comentarios generales:
+\end{comment}
+
 \section{Introduction}\label{sec:intro}
 
-The problem of inferring a scalar response \(Y\) from a functional covariate \(X(t)\)
+The problem of inferring a scalar response \(Y\) from a functional covariate \(X=X(t)\) is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
+While in principle we could simply treat the functional data as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
+As a consequence, numerous proposals have arisen on how to suitably treat functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on continuous data.
+A recent survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or \citet{berrendero2019rkhs}.
 
-\begin{comment}
-  - Grollemund et al.
+In this work we are concerned with (supervised) functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
+Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges that emerge as a result of working in infinite dimensions.
+To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some hidden dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
+We will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i(t), Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
-  - FDA (Cuevas et al. 2013)
+The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fueron realmente los primeros?}.
+It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
+\begin{equation}\label{eq:l2-linear-model}
+Y = \dotprod{X}{\beta} + \epsilon = \int_0^1 X(t)\beta(t)\, dt + \epsilon,
+\end{equation}
+where \(\epsilon\) is an independent error term with certain simplifying assumptions.
+In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinte-dimensional space \(L^2[0, 1]\).
+Aside from the fact that functional (i.e.
+infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle drawbacks that arise in this \(L^2\)-framework.
+For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case (see \citet{berrendero2019rkhs}).
+Moreover, the non-invertibility of the covariance operator (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least-squares theory.
 
-  - L2 model and finite model as a particular case of RKHS model. Define dot-product in L2.
-\end{comment}
+A similar \(L^2\)-model can be derived for the binary classification problem via the logistic function:
+\begin{equation}\label{l2-logistic-model}
+  \mathbb P(Y \mid X(t)) = \frac{1}{}
+\end{equation}
+
+It turns out that a natural alternative to the \(L^2\) model is the RKHS model.....
+Not only simpler, but also smoother and more manageable functions.
+Sustituir producto escalar por el de \(H(K)\).
+This model has already been proposed in \incomment{cita Berrendero et al.
+2019}, and in \incomment{cita Torrecilla 2017} for the case of functional classification.
+
+Bayesian RKHS-based variable selection (cite Berrendero et als.).
 
@@
 
-\subsection{A brief review of the theory of RKHS's}
+\subsection{A brief review of the theory of RKHS's}\label{sec:rkhs}
 
@@
 
-The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces.
+The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
 In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and its covariance function \(K(t, s)= \mathbb E[(X(t) - m(t))(X(s) - m(s))]\) are both continuous.
@@
 given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Lastly, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.
+Then, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.
 As it turns out, functions in this space verify the \textit{reproducing property}, thus giving it its name:
@@
 
-As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-A linear compact operator closely related to this covariance function is the so-called covariance operator, i.e.:
+As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS \incomment{¿Es necesario destacar alguna suposición/propiedad extra sobre \(K\)?
+Por ejemplo \(\int\int K^2 <\infty\) o que es simétrica y (semi-) definida positiva}.
+A compact operator closely related to this covariance function is the so-called covariance operator, i.e.: \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear, ...}
 \[
@@
 
+\begin{comment}
+  Puede que las definiciones 2 y 3 arriba sean innecesarias.
+\end{comment}
+
 Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since in general the trajectories do not belong to the RKHS with probability one (see \incomment{cita}).
-As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined for \(x\) a realization of \(X\).
+As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined when \(x\) is a realization of \(X\).
 However, following Parzen's approach \incomment{cita}\ we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
@@
 
-The classical functional linear regression model, firstly introduced in \incomment{cita Ramsay \& Silverman}, can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
-\begin{equation}\label{eq:l2-model}
-Y = \dotprod{X}{\beta} + \epsilon = \int_0^1 X(t)\beta(t)\, dt + \epsilon.
-\end{equation}
-In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinte-dimensional space \(L^2[0, 1]\).
-Aside from the fact that functional (i.e.
-infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle inconveniences with this model.
-For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case (see \incomment{Berrendero et al 2020, on a general definition of the flm.}).
+In this section...
 
-It turns out that a natural alternative to the \(L^2\) mode is the RKHS model.....
-Not only simpler, but also smoother and more manageable functions.
-The model remains linear because of the correspondence between the RKHS and the linear span of the process (Loève’s isometry).
+We will assume? the notation from the previous section and suppose WLOG that \(X\) is centered.
 
 \begin{comment}
-  - The RKHS model includes both L2 model and finite model. Cite Berrendero et al 2020.
+
   - Parametric approach with \(\beta \in H_0(K)\) (further constrain the parameter to this subset of the RKHS, which is dense by construction).
@@
 \end{equation}
-and regard \(p, \beta_i\) and \(t_i\) as free parameters in the parameter space \(\Theta\).
+and regard \(p \in \N\), \(\beta_i \in \R\) and \(t_i \in [0, 1]\) as free parameters in the parameter space \(\Theta = \N \times \R^p \times [0, 1]^p\).
+
+As a matter of fact, to simplify things further we will consider \(\pi(p) = \delta_{\hat p}\), or equivalently, we will consider a fixed value \(\hat p\) for the dimension of the model.
 
@@
 \subsection{Functional linear regression}
+ The model remains linear because of the correspondence between the RKHS and the linear span of the process (Loève’s isometry).
+The RKHS model includes both L2 model and finite model.
+Cite Berrendero et al 2020.
 

### 2022-02-28 paper/paper.tex
Update paper

@@
 %% Scalar product
-\newcommand\dotprod[2]{\left\langle #1,\, #2 \right\rangle}
+\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}
 
@@
 The posterior distribution can be approximated from Markov Chain Monte Carlo (MCMC) methods.
-Several estimators  derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real datasets.
+Several estimators derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real datasets, including a Bayesian-motivated variable selection method.
 \end{abstract}
@@
 
-The problem of inferring a scalar response \(Y\) from a functional covariate \(X=X(t)\) is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
+The problem of inferring a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
 While in principle we could simply treat the functional data as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
 As a consequence, numerous proposals have arisen on how to suitably treat functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on continuous data.
-A recent survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or \citet{berrendero2019rkhs}.
+A recent survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} and \citet{horváth2012inference}.
 
@@
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some hidden dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i(t), Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+We will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
@@
 \end{equation}
-where \(\epsilon\) is an independent error term with certain simplifying assumptions.
+\incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\epsilon\) is an independent error term with certain simplifying assumptions.
 In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinte-dimensional space \(L^2[0, 1]\).
@@
 For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case (see \citet{berrendero2019rkhs}).
-Moreover, the non-invertibility of the covariance operator (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least-squares theory.
+Moreover, the non-invertibility of the covariance operator (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
 
-A similar \(L^2\)-model can be derived for the binary classification problem via the logistic function:
+A similar \(L^2\)-based logistic regression model can be derived for the binary classification problem via the logistic function:
 \begin{equation}\label{l2-logistic-model}
-  \mathbb P(Y \mid X(t)) = \frac{1}{}
+  \mathbb P(Y=1 \mid X=x) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{\beta}{x}\}},
 \end{equation}
+where \(\beta_0 \in \R\) and \(\beta \in L^2[0, 1]\).
+In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE).
+However, not only do the same complications as in the linear regression model apply in this situation, but also there is the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions (see~\citet[p.~10]{buenolarraz2021functional}).
 
-It turns out that a natural alternative to the \(L^2\) model is the RKHS model.....
-Not only simpler, but also smoother and more manageable functions.
-Sustituir producto escalar por el de \(H(K)\).
-This model has already been proposed in \incomment{cita Berrendero et al.
-2019}, and in \incomment{cita Torrecilla 2017} for the case of functional classification.
+It turns out that a natural alternative to the \(L^2\) model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
+As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
+This approach has already been proposed in \citet{berrendero2019rkhs} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
 
-Bayesian RKHS-based variable selection (cite Berrendero et als.).
+Having provided a general perspective on the problem we are trying to solve, our main contribution is a novel Bayesian approach to linear and logistic regression with functional data, building on the newly introduced RKHS models for these tasks, and deviating from the usual methods of basis expansion and finite-dimensional least squares for parameter estimation.
+In essence, we suggest to impose a prior distribution on the RKHS associated with the functional regressor, and perform inference on this parametric space via standard Bayesian methodology.
+Luckily, the posterior distribution can be numerically approximated with Markov Chain Monte Carlo (MCMC) methods (see Section \ref{sec:methodology}), so that a complete end-to-end estimator can be derived.
+Moreover, the fact that we have a posterior distribution to work with allows us to perform diverse kinds of analyses to validate and assess the model.
+Similar ideas were presented in \citet{grollemund2019bayesian}, but the authors did not consider an RKHS-based model.
+
+Another set of techniques widely studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion.
+As it happens, some variable selection methods have already been proposed, both in the RKHS framework and in more general settings (see for example \citet{berrendero2019rkhs}, \citet{berrendero2016mrmr} or \citet{berrendero2016variable}), but in general they have their own dedicated algorithms and procedures.
+Given the nature of our proposed Bayesian model, we can easily isolate the marginal posterior distribution corresponding to a finite set of points of impact \(\{t_i\}\), thus providing a variable selection method along with the other estimators that naturally arise within our model.
+In this way, we can not only make predictions about our data, but also evaluate exactly which marginals of the process contain the most information.
 
@@
 As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS \incomment{¿Es necesario destacar alguna suposición/propiedad extra sobre \(K\)?
-Por ejemplo \(\int\int K^2 <\infty\) o que es simétrica y (semi-) definida positiva}.
-A compact operator closely related to this covariance function is the so-called covariance operator, i.e.: \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear, ...}
+Por ejemplo \(\int\int K^2 <\infty\) o que es simétrico y (semi-) definido positivo}.
+A linear operator closely related to this covariance function is the so-called covariance operator, i.e.: \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}
 \[
@@
 \end{equation}
-  \item The space \(\mathcal H(K)\) can also be defined as the image of the square root of the covariance operator, i.e.:
+  \item The space \(\mathcal H(K)\) can also be identified with the image of the square root of the covariance operator, i.e.:
   \begin{equation}\label{eq:rkhs-square-root}
@@
 
-\section{Bayesian methodology for RKHS-based functional regression models}
+\section{Bayesian methodology for RKHS-based functional regression models}\label{sec:methodology}
 
@@
 Then we can impose a prior on the finite-dimensional parameter space.
+
+  - Substitute scalar product of L2 for that of H(K).
 \end{comment}
@@
 
+\begin{comment}
+Bayesian variable selection: select marginal and then apply multivariate methods.
+\end{comment}
+
 \subsubsection{Overview of MCMC methods}
@@
  The model remains linear because of the correspondence between the RKHS and the linear span of the process (Loève’s isometry).
-The RKHS model includes both L2 model and finite model.
+The RKHS model includes both L2 model and finite model (most problems solved).
 Cite Berrendero et al 2020.
@@
 
-In this case the model/functional equation in $L^2$ is....... and the same problems apply, plus the MLE existence problem. An RKHS aproach is given by:.......
+An RKHS aproach is given by:.......
 
- and this solves the problems (see Bueno-Larraz 2021)
+As well as the other problems, The problem of MLE are solved, and to avoid the problem in finite dimension, we could use the strategy in Bueno Larraz 2021 (Firth + 2 phase algorithm for t and beta).
 
@@
   Not only MSE/Acc but also credible intervals, points-of-impact model --> variable selection included.
+
+  - Comparar con Grollemund et al. ?
 \end{comment}

### 2022-03-01 paper/paper.tex
Continue sections 1 and 2

@@
 
-\begin{comment}
-  Algunos comentarios generales:
-\end{comment}
-
 \section{Introduction}\label{sec:intro}
@@
 \begin{equation}\label{eq:l2-linear-model}
-Y = \dotprod{X}{\beta} + \epsilon = \int_0^1 X(t)\beta(t)\, dt + \epsilon,
+Y = \beta_0 + \dotprod{X}{\beta} + \epsilon = \beta_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
 \end{equation}
-\incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\epsilon\) is an independent error term with certain simplifying assumptions.
+\incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\beta_0\in \R\) and \(\epsilon\) is an independent error term with certain simplifying assumptions.
 In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinte-dimensional space \(L^2[0, 1]\).
@@
 A similar \(L^2\)-based logistic regression model can be derived for the binary classification problem via the logistic function:
-\begin{equation}\label{l2-logistic-model}
+\begin{equation}\label{eq:l2-logistic-model}
   \mathbb P(Y=1 \mid X=x) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{\beta}{x}\}},
@@
 In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but also there is the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions (see~\citet[p.~10]{buenolarraz2021functional}).
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions (see~\citet[p.~10]{buenolarraz2021functional}).
 
@@
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-This approach has already been proposed in \citet{berrendero2019rkhs} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
+This approach has already been explored in \citet{berrendero2019rkhs} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
 
@@
 Luckily, the posterior distribution can be numerically approximated with Markov Chain Monte Carlo (MCMC) methods (see Section \ref{sec:methodology}), so that a complete end-to-end estimator can be derived.
-Moreover, the fact that we have a posterior distribution to work with allows us to perform diverse kinds of analyses to validate and assess the model.
+Moreover, the fact that we have a posterior distribution to work with allows us to perform diverse kinds of posterior checks to validate and assess the model.
 Similar ideas were presented in \citet{grollemund2019bayesian}, but the authors did not consider an RKHS-based model.
@@
 Another set of techniques widely studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion.
-As it happens, some variable selection methods have already been proposed, both in the RKHS framework and in more general settings (see for example \citet{berrendero2019rkhs}, \citet{berrendero2016mrmr} or \citet{berrendero2016variable}), but in general they have their own dedicated algorithms and procedures.
-Given the nature of our proposed Bayesian model, we can easily isolate the marginal posterior distribution corresponding to a finite set of points of impact \(\{t_i\}\), thus providing a variable selection method along with the other estimators that naturally arise within our model.
-In this way, we can not only make predictions about our data, but also evaluate exactly which marginals of the process contain the most information.
+As it happens, some variable selection methods have already been proposed in the RKHS framework (see for example \citet{berrendero2019rkhs}), but in general they have their own dedicated algorithms and procedures.
+Given the nature of our suggested Bayesian model, we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model.
+In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
+These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent} or \citet{ferraty2010most} by way of illustration.
 
@@
 
-In Section~\ref{sec:intro} we review...
+In Section~\ref{sec:rkhs} we carry out a quick review of the theory of RKHS's from a probabilistic point of view.
+Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models proposed.
+The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of implementation details.
+Lastly, the conclusions drawn from this work are presented in Section~\ref{sec:conclusion}.
 
@@
 As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS \incomment{¿Es necesario destacar alguna suposición/propiedad extra sobre \(K\)?
-Por ejemplo \(\int\int K^2 <\infty\) o que es simétrico y (semi-) definido positivo}.
-A linear operator closely related to this covariance function is the so-called covariance operator, i.e.: \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}
+Por ejemplo \(\int\int K^2 <\infty\) o que es simétrica y (semi-) definida positiva}.
+A linear operator \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}\  closely related to this covariance function is the so-called covariance operator, namely
 \[
@@
 \]
-This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that a centered process \(X\) can be decomposed as
+This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita?}, which asserts that a centered process \(X\) can be decomposed as
 \begin{equation}\label{eq:karhunen-loeve}
@@
 \begin{enumerate}
-  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \incomment{cita}.
+  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see \incomment{cita}).
 This isometry is essentially the completion of the correspondence
@@
 \]
-and can be formally defined as (see \incomment{cita})
+and can be formally defined as
 \begin{equation}\label{eq:loeves-isometry}
@@
 
-  \item Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \incomment{cita}, we can also express
+  \item Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \incomment{cita?}, we can also express
   \begin{equation}\label{rkhs-sum-lambda}
@@
   \]
-Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
+Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita?}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
 \end{enumerate}
@@
 As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined when \(x\) is a realization of \(X\).
-However, following Parzen's approach \incomment{cita}\ we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
+However, following Parzen's approach \incomment{cita}\ we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
 This notation often proves to be useful and convenient.
@@
 
-In this section...
+In this section present the precise models and Bayesian methodologies explored in this work.
+We will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\).
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\beta \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\beta}_K\) in the \(L^2\) models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+However, to further simplify things we will follow a parametric approach and suppose that \(\beta\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
+\begin{equation}\label{eq:beta-parameter-rkhs}
+  \beta(\cdot) = \sum_{i=1}^p \beta_i K(t_i, \cdot).
+\end{equation}
 
-We will assume? the notation from the previous section and suppose WLOG that \(X\) is centered.
+The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_i \in \R\) and \(t_i \in [0, 1]\) as free parameters.
+In view of expression \eqref{eq:beta-parameter-rkhs}, to set a prior distribution on the unknown function \(\beta\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_i\}\) and the times \(\{t_i\}\) separately.
+Moreover, as we said before, we will take the expression \(\dotprod{x}{\beta}_K\) to mean \(\Psi_x^{-1}(\beta)\), where \(\Psi\) is \textit{Loève's isometry}.
+Hence, taking into account that \(\Psi_x^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
+\[
+\dotprod{x}{\beta}_K \equiv \sum_{i=1}^p \beta_i x(t_i).
+\]
 
-\begin{comment}
-
-  - Parametric approach with \(\beta \in H_0(K)\) (further constrain the parameter to this subset of the RKHS, which is dense by construction).
-Then we can impose a prior on the finite-dimensional parameter space.
-
-  - Substitute scalar product of L2 for that of H(K).
-\end{comment}
-
-The general idea explored in this work revolves around imposing a prior distribution on the functional space \(\mathcal H_0(K)\), which we know is dense in \(\mathcal H(K)\).
-Specifically, we consider a functional parameter of the form
-\begin{equation}\label{eq:parameter-rkhs}
-  \beta(t) = \sum_{i=1}^p \beta_i K(t_i, \cdot),
-\end{equation}
-and regard \(p \in \N\), \(\beta_i \in \R\) and \(t_i \in [0, 1]\) as free parameters in the parameter space \(\Theta = \N \times \R^p \times [0, 1]^p\).
-
-As a matter of fact, to simplify things further we will consider \(\pi(p) = \delta_{\hat p}\), or equivalently, we will consider a fixed value \(\hat p\) for the dimension of the model.
-
-Note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
+On a separate note, it is worth mentioning that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
+This is why our simplifying assumption on \(\beta\) is not actually very restricting, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
 The following result proves that any such extension is indeed unique.
@@
 
-\begin{comment}
-Bayesian variable selection: select marginal and then apply multivariate methods.
-\end{comment}
+In any case, after having selected a suitable prior distribution \(\pi(\theta)\) on the chosen finite-dimensional parameter space \(\Theta_p\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
+samples amounts to \incomment{es necesario escribirlo?}
+\begin{equation}\label{eq:bayes-theorem}
+  \pi(\theta \mid Y_1, \dots, Y_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid \theta) \right]\pi(\theta).
+\end{equation}
+In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(Y\mid \theta\) considered in the case of functional linear regression and functional logistic regression, respectively.
 
-\subsubsection{Overview of MCMC methods}
+A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant.
+Nevertheless, we can sample from the posterior distribution\footnote{Provided that the \textit{marginal likelihood} \(\int_{\Theta_p} \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of MCMC sampling methods.
+These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known.
+Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the function given.
+In short, they build stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution.
+The Metropolis-Hastings algorithm (e.g.~\cite{chib1995understanding}) is arguably the most well-known method in this regard.
 
-\subsection{Functional linear regression}
+\subsection{Functional linear regression}\label{sec:rkhs-linear-model}
  The model remains linear because of the correspondence between the RKHS and the linear span of the process (Loève’s isometry).
@@
 \begin{comment}
+  - Parameter space \(\Theta_p = \R^p \times [0, 1]^p\ldots\).
+  - Prior distributions (cite "A study of variable selection using
+g-prior distribution with ridge parameter")+ log-posterior
   - Estimators derived.
+  - Bayesian variable selection: select marginal and then apply multivariate methods.
 \end{comment}
 
-\subsection{Functional logistic regression}
+\subsection{Functional logistic regression}\label{sec:rkhs-logistic-model}
 
@@
 \begin{comment}
+  - Parameter space \(\Theta_p = \R^p \times [0, 1]^p\ldots\).
+  - Prior distributions + log-posterior
   - Estimators derived.
+  - Bayesian variable selection: similar to the linear regression case.
 \end{comment}
 
-\subsection{Model choice and validation}
+\subsection{Model choice and validation}\label{sec:model-choice}
 
 \begin{comment}
-  - Choice of p, MLE, BIC, ...
-
+  - The parameter p could be included; but problems: label switching, difficulty in implementation, etc.
+  - Choice of p: equivalent to Dirac's delta prior; chosen via MLE and BIC, ...
   - Posterior predictive checks, ¿Bayesian p-values?, ...
@@
 \begin{comment}
-  Not only MSE/Acc but also credible intervals, points-of-impact model --> variable selection included.
+  - comentar el método MCMC usado y por qué (vs. Metropolis and other gradient-based such as NUTS).
+  - Not only MSE/Acc but also credible intervals, points-of-impact model --> variable selection included.
 

### 2022-03-04 paper/paper.tex
FLinReg and FLogReg

@@
 \usepackage{natbib}
-\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
+\usepackage[
+  colorlinks,
+  citecolor=blue,
+  urlcolor=blue,
+  filecolor=blue,
+  backref=page
+]{hyperref}
 \usepackage{graphicx}
@@
 \newcommand{\E}{\mathbb{E}}
+\newcommand{\Var}{\ensuremath\operatorname{Var}}
 
@@
 The posterior distribution can be approximated from Markov Chain Monte Carlo (MCMC) methods.
-Several estimators derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real datasets, including a Bayesian-motivated variable selection method.
+Several estimators derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection method.
 \end{abstract}
@@
 
-The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fueron realmente los primeros?}.
+The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fueron realmente así?}.
 It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
@@
 infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle drawbacks that arise in this \(L^2\)-framework.
-For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case (see \citet{berrendero2019rkhs}).
+For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case \citep[see][]{berrendero2019rkhs}.
 Moreover, the non-invertibility of the covariance operator (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
@@
 In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions (see~\citet[p.~10]{buenolarraz2021functional}).
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][p.~10]{buenolarraz2021functional}.
 
@@
 
-Having provided a general perspective on the problem we are trying to solve, our main contribution is a novel Bayesian approach to linear and logistic regression with functional data, building on the newly introduced RKHS models for these tasks, and deviating from the usual methods of basis expansion and finite-dimensional least squares for parameter estimation.
-In essence, we suggest to impose a prior distribution on the RKHS associated with the functional regressor, and perform inference on this parametric space via standard Bayesian methodology.
-Luckily, the posterior distribution can be numerically approximated with Markov Chain Monte Carlo (MCMC) methods (see Section \ref{sec:methodology}), so that a complete end-to-end estimator can be derived.
-Moreover, the fact that we have a posterior distribution to work with allows us to perform diverse kinds of posterior checks to validate and assess the model.
-Similar ideas were presented in \citet{grollemund2019bayesian}, but the authors did not consider an RKHS-based model.
-
 Another set of techniques widely studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion.
-As it happens, some variable selection methods have already been proposed in the RKHS framework (see for example \citet{berrendero2019rkhs}), but in general they have their own dedicated algorithms and procedures.
+As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
 Given the nature of our suggested Bayesian model, we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model.
 In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
-These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent} or \citet{ferraty2010most} by way of illustration.
+These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
 
@@
 
+\begin{comment}
+¿Mejor poner esto después de la sección 1.1?
+\end{comment}
+
 In Section~\ref{sec:rkhs} we carry out a quick review of the theory of RKHS's from a probabilistic point of view.
 Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models proposed.
-The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of implementation details.
+The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of computational details.
 Lastly, the conclusions drawn from this work are presented in Section~\ref{sec:conclusion}.
@@
 \begin{equation}\label{eq:karhunen-loeve}
-X(t) = \sum_{j=1}^\infty \xi_j \phi_j(t),
+X(t) = \sum_{i=1}^\infty \xi_i \phi_i(t),
 \end{equation}
-where \(\phi_j\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\).
+where \(\phi_i\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_i = \int X\phi_i\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_i\), the eigenvalue associated with \(\phi_i\).
 This expansion is the main justification of many FDA techniques, such as functional principal component analysis (FPCA) (see for example \incomment{cita}).
@@
 
-In this section present the precise models and Bayesian methodologies explored in this work.
-We will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\).
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\beta \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\beta}_K\) in the \(L^2\) models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\beta\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
-\begin{equation}\label{eq:beta-parameter-rkhs}
-  \beta(\cdot) = \sum_{i=1}^p \beta_i K(t_i, \cdot).
+In this section we present the precise models and Bayesian methodologies explored in this work.
+In the sequel, we will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\).
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\) models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
+\begin{equation}\label{eq:alpha-parameter-rkhs}
+  \alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot) \in \mathcal H_0(K).
 \end{equation}
 
-The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_i \in \R\) and \(t_i \in [0, 1]\) as free parameters.
-In view of expression \eqref{eq:beta-parameter-rkhs}, to set a prior distribution on the unknown function \(\beta\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_i\}\) and the times \(\{t_i\}\) separately.
-Moreover, as we said before, we will take the expression \(\dotprod{x}{\beta}_K\) to mean \(\Psi_x^{-1}(\beta)\), where \(\Psi\) is \textit{Loève's isometry}.
+The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j \in \R\) and \(t_j \in [0, 1]\) as free parameters.
+The general idea will be to impose a prior distribution on these parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data.
+In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_i\}\) separately.
+Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry.
 Hence, taking into account that \(\Psi_x^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
 \[
-\dotprod{x}{\beta}_K \equiv \sum_{i=1}^p \beta_i x(t_i).
+\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j).
 \]
@@
 On a separate note, it is worth mentioning that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
-This is why our simplifying assumption on \(\beta\) is not actually very restricting, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
+This is why our simplifying assumption on \(\alpha\) is not actually very restricting, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
 The following result proves that any such extension is indeed unique.
@@
 \end{prop}
-\begin{comment}
-    ¿La prueba va en el cuerpo del artículo, o en un anexo/material suplementario?
-\end{comment}
 
@@
 
-In any case, after having selected a suitable prior distribution \(\pi(\theta)\) on the chosen finite-dimensional parameter space \(\Theta_p\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
-samples amounts to \incomment{es necesario escribirlo?}
+In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the parameter vector \(\theta\) on the chosen finite-dimensional parameter space, we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
+samples amounts to
 \begin{equation}\label{eq:bayes-theorem}
-  \pi(\theta \mid Y_1, \dots, Y_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid \theta) \right]\pi(\theta).
+  \pi(\theta \mid \mathcal D_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right]\pi(\theta).
 \end{equation}
-In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(Y\mid \theta\) considered in the case of functional linear regression and functional logistic regression, respectively.
+In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid \theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
 
 A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant.
-Nevertheless, we can sample from the posterior distribution\footnote{Provided that the \textit{marginal likelihood} \(\int_{\Theta_p} \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of MCMC sampling methods.
+Nevertheless, we can sample from the posterior distribution\footnote{Provided that the marginal likelihood \(\int \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of MCMC sampling methods.
 These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known.
@@
 \subsection{Functional linear regression}\label{sec:rkhs-linear-model}
- The model remains linear because of the correspondence between the RKHS and the linear span of the process (Loève’s isometry).
-The RKHS model includes both L2 model and finite model (most problems solved).
-Cite Berrendero et al 2020.
 
-\begin{comment}
-  - Parameter space \(\Theta_p = \R^p \times [0, 1]^p\ldots\).
-  - Prior distributions (cite "A study of variable selection using
-g-prior distribution with ridge parameter")+ log-posterior
-  - Estimators derived.
-  - Bayesian variable selection: select marginal and then apply multivariate methods.
-\end{comment}
+In the case of functional linear regression, the RKHS model considered (with \(\alpha\) as in \eqref{eq:alpha-parameter-rkhs}) is
+\begin{equation}\label{eq:rkhs-model-linear}
+  Y = \alpha_0 + \Psi^{-1}_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
+\end{equation}
+where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term verifying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multivariate linear model with the \(X(t_j)\) as scalar covariates.
+Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information on \(\mathcal D_n\).
+For a generic element \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (\beta, \tau, \alpha_0, \sigma^2) \in \Theta_p\), the reinterpreted model assumes the form
+\begin{equation}\label{eq:rkhs-model-linear-2}
+  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
+\end{equation}
+
+It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
+Furthermore, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model \eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
+In addition, as far as generalizations go the model is easily extensible, in the sense that several functional covariates could be added.
+In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
+
+The prior distribution \(\pi(\theta)\) suggested for the parameter \(\theta \in \Theta_p\) is given by
+\begin{align}\label{eq:prior-linear}
+  \begin{split}
+  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
+  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
+  \beta\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
+\end{split}
+\end{align}
+where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \eta \in \R^+\) and \(g\in \R\) are hyperparameters of the model.
+On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
+In any event, the estimation of \(\alpha_0\) is straightforward, since \(\alpha_0 = \E[Y]\) by~\eqref{eq:rkhs-model-linear-2}, so we could have left this parameter out of the model altogether.
+On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
+
+When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get a sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
+Assuming now a previously unseen test set \(\mathcal D_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples:
+
+\begin{description}
+  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\}_m)\) and predict the responses in the usual way by
+  \[
+    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
+  \]
+  The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function.
+On the other hand, the mode is also a good candidate because it represents the point of highest probability.
+At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
+  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution \(Y^*\) of the ``sampled responses''
+  \[
+  \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n', \ m=1,\dots,M\right\}
+  \]
+  following model~\eqref{eq:rkhs-model-linear-2}. Then, we can take the mean of all such samples as a proxy for the response:
+  \[
+  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
+  \]
+  \item[Variable selection estimates] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\).
+Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\}_m)\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\).
+After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known multivariate algorithms suited for this situation.
+\end{description}
+
+Note that these estimators can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
+As a consequence, what we have in practice is a single algorithm that can produce multiple estimators at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
+Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available estimators is performed to produce a final result.
 
@@
 
-An RKHS aproach is given by:.......
+In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable, and consider an RKHS approach given by the following functional logistic equation:
 
-As well as the other problems, The problem of MLE are solved, and to avoid the problem in finite dimension, we could use the strategy in Bueno Larraz 2021 (Firth + 2 phase algorithm for t and beta).
+\begin{equation}\label{eq:rkhs-model-logistic}
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi^{-1}_X(\alpha)\}}, \quad \alpha \in \mathcal H_0(K).
+\end{equation}
+After taking into account our assumptions on \(\alpha\) (namely the choice of a fixed value of \(p\)) and incorporating the sample information, we can rewrite the classification model \eqref{eq:rkhs-model-logistic} as
+\begin{equation}\label{eq:rkhs-model-logistic-2}
+  \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}},
+\end{equation}
+with \(\beta_j\in\R\), \(t_j\in[0, 1]\) and \(\alpha_0\in\R\).
 
-\begin{comment}
-  - Parameter space \(\Theta_p = \R^p \times [0, 1]^p\ldots\).
-  - Prior distributions + log-posterior
-  - Estimators derived.
-  - Bayesian variable selection: similar to the linear regression case.
-\end{comment}
+In much the same way as the functional linear regression model explained in the previous section, this RKHS-based logistic regression model partially solves some of the issues with the \(L^2\)-model~\eqref{eq:l2-logistic-model} stated in Section~\ref{sec:intro}.
+Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-perfect separation in which the MLE does not exist.
+However, this problem can be easily circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.}.
+Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.
+1]{buenolarraz2021functional}.
+
+As far as prior distributions go, we propose to use the same ones as we did in \eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner}.
+However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.
+\incomment{estimators derived are the same except the distributional case, threshold of 0.5, ...}
 
@@
 \begin{comment}
+  - The estimated variance can be used?
   - The parameter p could be included; but problems: label switching, difficulty in implementation, etc.
+  - g=5 good choice.
     - Choice of p: equivalent to Dirac's delta prior; chosen via MLE and BIC, ...
-  - Posterior predictive checks, ¿Bayesian p-values?, ...
+  - Posterior predictive checks, credible intervals (especially for variable selection), ¿Bayesian p-values?, ...
 \end{comment}
 
-\section{Experimental results}
+\section{Experimental results}\label{sec:results}
 
 \begin{comment}
-  - comentar el método MCMC usado y por qué (vs. Metropolis and other gradient-based such as NUTS).
+  - comentar el método MCMC usado y por qué (vs. Metropolis and other gradient-based such as NUTS). Comentar que hay varias cadenas en paralelo y luego se hace la media, para aumentar estabilidad.
+  - Algoritmos de comparación
   - Not only MSE/Acc but also credible intervals, points-of-impact model --> variable selection included.
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
@@
 \sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}
+
+\begin{comment}
+  - Details of the log-posterior distribution formula, the use of the \(\log\sigma\) for \(\sigma^2\), the use of the maximum eigenvalue of the Gram matrix in the Zellner prior, etc..
+  - Some implementation decisions/possibilities: burn-in, thinning when computing the distributional estimates, ....
+  - Graphs of traces, posteriors, etc.
+\end{comment}
+
 \end{supplement}

### 2022-03-04 paper/paper.tex
Finish section 2

@@
 
-The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fueron realmente así?}.
+The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}.
 It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
@@
 \incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\beta_0\in \R\) and \(\epsilon\) is an independent error term with certain simplifying assumptions.
-In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinte-dimensional space \(L^2[0, 1]\).
+In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinite-dimensional space \(L^2[0, 1]\).
 Aside from the fact that functional (i.e.
@@
 \begin{equation}\label{eq:l2-logistic-model}
-  \mathbb P(Y=1 \mid X=x) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{\beta}{x}\}},
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{X}{\beta}\}},
 \end{equation}
@@
 In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][p.~10]{buenolarraz2021functional}.
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{buenolarraz2021functional}.
 
@@
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-This approach has already been explored in \citet{berrendero2019rkhs} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
+This approach has already been explored in \citet{berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
 
@@
 \end{equation}
+An important detail to note is that \(\mathcal H(K)\) is a space of genuine functions, and not of equivalence classes of functions.
+Hence, the values of these functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
 
@@
 \]
-This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita?}, which asserts that a centered process \(X\) can be decomposed as
+This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear} which asserts that a centered process \(X\) can be decomposed as
 \begin{equation}\label{eq:karhunen-loeve}
@@
 where \(\phi_i\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_i = \int X\phi_i\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_i\), the eigenvalue associated with \(\phi_i\).
-This expansion is the main justification of many FDA techniques, such as functional principal component analysis (FPCA) (see for example \incomment{cita}).
+This expansion is the main justification of many FDA techniques, such as functional principal component analysis (FPCA)~\cite[see for example][Ch.~8]{ramsay2005functional}.
 
@@
 
-\begin{enumerate}
-  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see \incomment{cita}).
+\begin{description}
+  \item[From an isometry] Via \textit{Loève's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see Lemma 1.1 in \citet{lukic2001stochastic}).
 This isometry is essentially the completion of the correspondence
@@
 \end{equation}
-  \item The space \(\mathcal H(K)\) can also be identified with the image of the square root of the covariance operator, i.e.:
+  \item[From the covariance operator] The space \(\mathcal H(K)\) can also be identified with the image of the square root of the covariance operator, i.e.:
   \begin{equation}\label{eq:rkhs-square-root}
@@
 
-  \item Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \incomment{cita?}, we can also express
+  \item[From a \(\bf L^2\)-like smoothed norm] Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \citep{mercer1909functions}, we can also express
   \begin{equation}\label{rkhs-sum-lambda}
@@
   \]
-Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita?}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
-\end{enumerate}
+Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
+\end{description}
 
 \begin{comment}
-  Puede que las definiciones 2 y 3 arriba sean innecesarias.
+  Puede que las definiciones 2ª y 3ª arriba sean innecesarias.
 \end{comment}
 
-Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since in general the trajectories do not belong to the RKHS with probability one (see \incomment{cita}).
+Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see][Cor.~7.1]{lukic2001stochastic}.
 As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined when \(x\) is a realization of \(X\).
-However, following Parzen's approach \incomment{cita}\ we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
+However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
 This notation often proves to be useful and convenient.
@@
 
-The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j \in \R\) and \(t_j \in [0, 1]\) as free parameters.
-The general idea will be to impose a prior distribution on these parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data.
-In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_i\}\) separately.
+The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
+Thus, we will search for our functional parameter in the set
+\[
+\mathcal H_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_j \in \R, \ t_1,\dots,t_j \in [0, 1]\right\}.
+\]
+
+The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data.
+In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry.
-Hence, taking into account that \(\Psi_x^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
+Hence, taking into account that \(\Psi_X^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
 \[
@@
 On a separate note, it is worth mentioning that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
-This is why our simplifying assumption on \(\alpha\) is not actually very restricting, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
+This is why our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
 The following result proves that any such extension is indeed unique.
@@
 
-In the case of functional linear regression, the RKHS model considered (with \(\alpha\) as in \eqref{eq:alpha-parameter-rkhs}) is
+In the case of functional linear regression, the RKHS model considered (with \(\alpha\in \mathcal H_{0,p}(K)\)) is
 \begin{equation}\label{eq:rkhs-model-linear}
@@
 On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
+
+\begin{comment}
+¿Es necesario/conveniente escribir aquí la expresión teórica de la distribución a posteriori? O quizás en el apéndice/material suplementario mejor?
+\[
+\pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right) \right\}.
+\]
+
+\end{comment}
 
@@
 
+\begin{comment}
+  En \citet[Sec.~2.5]{grollemund2019bayesian}, los autores proponen un estimador Bayesiano para sus funciones ``escalón'' \(\beta_\theta(t)\) utilizando una pérdida cuadrática, de la forma:
+  \[
+    \hat \beta \in \arg\min_{d\in L^2[0, 1]} \int\int (\beta_\theta(t) - d(t))^2 \pi(\theta\mid \mathcal D_n)\, dt\, d\theta.
+  \]
+  ¿Podríamos hacer nosotros algo parecido con nuestro parámetro funcional \(\alpha\)?
+\end{comment}
+
 \subsection{Functional logistic regression}\label{sec:rkhs-logistic-model}
 
-In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable, and consider an RKHS approach given by the following functional logistic equation:
-
+In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X\), and consider an RKHS approach given by the following functional logistic equation:
 \begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi^{-1}_X(\alpha)\}}, \quad \alpha \in \mathcal H_0(K).
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi^{-1}_X(\alpha)\}}, \quad \alpha \in \mathcal H_{0,p}(K).
 \end{equation}
-After taking into account our assumptions on \(\alpha\) (namely the choice of a fixed value of \(p\)) and incorporating the sample information, we can rewrite the classification model \eqref{eq:rkhs-model-logistic} as
+Following our previous approach, after incorporating the sample information we can rewrite the classification model~\eqref{eq:rkhs-model-logistic} as
 \begin{equation}\label{eq:rkhs-model-logistic-2}
-  \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}},
+Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
 \end{equation}
-with \(\beta_j\in\R\), \(t_j\in[0, 1]\) and \(\alpha_0\in\R\).
+with
+\begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
+  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
+\end{equation}
+where in turn \(\beta_j\in\R\), \(t_j\in[0, 1]\) and \(\alpha_0\in\R\).
 
 In much the same way as the functional linear regression model explained in the previous section, this RKHS-based logistic regression model partially solves some of the issues with the \(L^2\)-model~\eqref{eq:l2-logistic-model} stated in Section~\ref{sec:intro}.
-Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-perfect separation in which the MLE does not exist.
+Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist.
 However, this problem can be easily circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.}.
-Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.
-1]{buenolarraz2021functional}.
+Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.~1]{buenolarraz2021functional}.
 
-As far as prior distributions go, we propose to use the same ones as we did in \eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner}.
+As far as prior distributions go, we propose to use the same ones as we did in \eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner.
+¿Sabéis de alguna que se use en el caso de regresión logística?}.
 However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.
-\incomment{estimators derived are the same except the distributional case, threshold of 0.5, ...}
+
+\begin{comment}
+  Misma pregunta que antes: ¿ponemos la expresión de la posteriori?
+  \begin{align*}
+\log \pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi^{-1}_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}^{-1}(\alpha)\right\}\right)\right]\\
+\quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0).
+\end{align*}
+
+\end{comment}
+
+Another thing to consider is that in this case we are approximating a probability, so before producing an output we need to transform the predicted values to \(\{0, 1\}\).
+We will follow the standard convention of considering a threshold \(\gamma = 0.5\) and predicting \(\hat Y=1\) if and only if \(\mathbb P(Y=1\mid X) \geq \gamma\).
+With this thresholding strategy in mind, the point estimation approach is analogous to the linear regression case:
+\[
+\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} > \gamma \right), \quad i=1,\dots,n',
+\]
+where \(\mathbb I\) is the indicator function and the hat estimates are obtained through a summary statistic of the corresponding marginal posterior distributions.
+On the other hand, the distributional estimates are somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples:
+\begin{description}
+  \item[Average sampled probability] If we choose to average the sampled probabilities \(p_i^{(m)*} = \mathbb P(Y_i \mid X_i,\theta^{(m)*})\), we get the following estimator:
+  \[
+    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \right), \quad i=1,\dots,n'.
+  \]
+  \item[Sampled majority vote] If we choose to average the sampled binary responses \(Y^*\) instead, we compute the predictions as
+\[
+  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*}\right), \quad i=1,\dots,n',
+\]
+where each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
+\end{description}
+Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) variables from each regressor based on the posterior distribution of \(\tau\mid \mathcal D_n\) and feed the reduced dataset to a multivariate binary classification procedure.
 
@@
 
-\begin{comment}
-  - The estimated variance can be used?
-  - The parameter p could be included; but problems: label switching, difficulty in implementation, etc.
-  - g=5 good choice.
-    - Choice of p: equivalent to Dirac's delta prior; chosen via MLE and BIC, ...
-  - Posterior predictive checks, credible intervals (especially for variable selection), ¿Bayesian p-values?, ...
-\end{comment}
+One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model.
+While in principle we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted implications.
+For instance, it would make the implementation more complex, since the dimensionality of the parameters \(\beta\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next.
+In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
+Another possibility would be to adapt a purely Bayesian model selection technique \citep[see][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
+In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
+BIC, AIC, \ldots) to select its optimal value.
+
+As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments can be made:
+\begin{itemize}
+  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation).
+Our numerical studies showed that the results were much better with this choice than, say, with a random or a null vector.
+  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
+  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
+That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter.
+This standardization technique was used previously in the literature; see for example \citet{grollemund2019bayesian}.
+\end{itemize}
+
+To conclude this section, it is worth mentioning that the Bayesian nature of our model allows us to perform some model validation checks straight away.
+For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
+Moreover, we can perform various \textit{visual checks}, such as a plot of the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}).
+In addition, we can compute the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}_m\) that fall below the real value of the statistic.
+They are expected to be around 0.5 when the model accurately represents the data \citep[see][Ch.~6]{gelman1995bayesian}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\textwidth]{img/ppc_lin}
+  \caption{Posterior predictive graphical checks on a fitted model.
+On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the generated samples, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
+\end{figure}
 
 \section{Experimental results}\label{sec:results}
-
 \begin{comment}
-  - comentar el método MCMC usado y por qué (vs. Metropolis and other gradient-based such as NUTS). Comentar que hay varias cadenas en paralelo y luego se hace la media, para aumentar estabilidad.
-  - Algoritmos de comparación
-  - Not only MSE/Acc but also credible intervals, points-of-impact model --> variable selection included.
-
-  - Comparar con Grollemund et al. ?
+  \begin{itemize}
+    \item Comentar el método MCMC usado (\textit{emcee}) y por qué (vs.
+Metropolis u otros basados en gradiente como NUTS).
+Comentar que se lanzan varias cadenas en paralelo y luego se hace la media, para aumentar estabilidad.
+    \item Comentar que se usa el MLE de los parámetros (perturbado) como punto de partida del algoritmo MCMC.
+Comentar la forma de estimación de MLE (varias iteraciones combinando una búsqueda global con una búsqueda local).
+    \item Parámetros comunes de generación de datasets (granularidad de la malla, ...) y de la estrategia de ejecución (dónde se ejecuta, tiempos de ejecución, cuántas veces se repite para eliminar aleatoriedad, ...)
+    \item Métricas usadas para medir el error.
+    \item Contar los algoritmos de comparación usados.
+    \item Mostrar tablas de resultados y gráficas.
+\end{itemize}
 \end{comment}
@@
 \begin{comment}
-  - Details of the log-posterior distribution formula, the use of the \(\log\sigma\) for \(\sigma^2\), the use of the maximum eigenvalue of the Gram matrix in the Zellner prior, etc..
-  - Some implementation decisions/possibilities: burn-in, thinning when computing the distributional estimates, ....
-  - Graphs of traces, posteriors, etc.
+  \begin{itemize}
+  \item Detalles de la distribución log-posterior en cada caso?
+  \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado.
+  \item Comentar algunas decisiones de implementación: burn-in, thinning a la hora de hacer ppcs, ...
+  \item Gráficos de las trazas, de algunas posterioris concretas, etc.
+\end{itemize}
 \end{comment}

### 2022-03-05 paper/paper.tex
Some corrections

@@
 
-The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}.
-It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
+The most common linear regression model is the classical \(L^2\)-model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}.
+It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
 \begin{equation}\label{eq:l2-linear-model}
@@
 For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case \citep[see][]{berrendero2019rkhs}.
-Moreover, the non-invertibility of the covariance operator (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
+Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
 
-A similar \(L^2\)-based logistic regression model can be derived for the binary classification problem via the logistic function:
+A similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
 \begin{equation}\label{eq:l2-logistic-model}
@@
 
-It turns out that a natural alternative to the \(L^2\) model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
+It turns out that in both situations a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-This approach has already been explored in \citet{berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
+These RKHS-based models have already been explored in \citet{berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
+The main contribution of this work is a parametric Bayesian approach to parameter estimation within these recently-proposed models, in which a prior distribution is imposed on the unknown function to  obtain a posterior distribution after seeing the data.
 
@@
 As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-Given the nature of our suggested Bayesian model, we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model.
-In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
+As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model.
+In this way, in addition to making predictions about the input data, we could evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
 These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
@@
 \]
-This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear} which asserts that a centered process \(X\) can be decomposed as
+This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear}, which asserts that a centered process \(X\) can be decomposed as
 \begin{equation}\label{eq:karhunen-loeve}
@@
 where \(\phi_i\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_i = \int X\phi_i\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_i\), the eigenvalue associated with \(\phi_i\).
-This expansion is the main justification of many FDA techniques, such as functional principal component analysis (FPCA)~\cite[see for example][Ch.~8]{ramsay2005functional}.
+This expansion is the main justification for many FDA techniques, such as functional principal component analysis (FPCA)~\cite[see for example][Ch.~8]{ramsay2005functional}.
 
@@
 
-Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see][Cor.~7.1]{lukic2001stochastic}.
-As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined when \(x\) is a realization of \(X\).
+Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
+As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
 However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
@@
 In the sequel, we will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\).
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\) models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
 However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
@@
 The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data.
-In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
+In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry.
@@
 \end{equation}
-In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid \theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
+In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
 
 A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant.
-Nevertheless, we can sample from the posterior distribution\footnote{Provided that the marginal likelihood \(\int \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of MCMC sampling methods.
+Nevertheless, we can sample from the posterior distribution\footnote{Provided that the marginal likelihood \(\int \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of Markov Chain Monte Carlo (MCMC) sampling methods.
 These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known.
-Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the function given.
-In short, they build stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution.
-The Metropolis-Hastings algorithm (e.g.~\cite{chib1995understanding}) is arguably the most well-known method in this regard.
+Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the given function.
+In short, they operate by building stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution.
+The Metropolis-Hastings algorithm \citep[e.g.][]{chib1995understanding} is arguably the most well-known method in this regard.
 
@@
 
-In the case of functional linear regression, the RKHS model considered (with \(\alpha\in \mathcal H_{0,p}(K)\)) is
+In the case of functional linear regression, the RKHS model considered (with a functional parameter \(\alpha\in \mathcal H_{0,p}(K)\)) is
 \begin{equation}\label{eq:rkhs-model-linear}
@@
 When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multivariate linear model with the \(X(t_j)\) as scalar covariates.
-Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information on \(\mathcal D_n\).
+Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
 For a generic element \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (\beta, \tau, \alpha_0, \sigma^2) \in \Theta_p\), the reinterpreted model assumes the form
@@
 It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
-Furthermore, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model \eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, as far as generalizations go the model is easily extensible, in the sense that several functional covariates could be added.
+Furthermore, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
+In addition, as far as generalizations go the model is easily extensible, seeing as how several functional covariates belonging to the RKHS could be added.
 In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
@@
 \end{align}
-where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \eta \in \R^+\) and \(g\in \R\) are hyperparameters of the model.
+where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
 On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
@@
 \begin{description}
-  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\}_m)\) and predict the responses in the usual way by
+  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way by
   \[
@@
   \item[Variable selection estimates] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\).
-Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\}_m)\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\).
+Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\).
 After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known multivariate algorithms suited for this situation.
@@
 
-As far as prior distributions go, we propose to use the same ones as we did in \eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner.
+As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner.
 ¿Sabéis de alguna que se use en el caso de regresión logística?}.
@@
 \[
-\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} > \gamma \right), \quad i=1,\dots,n',
+\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
 \]
@@
 \begin{description}
-  \item[Average sampled probability] If we choose to average the sampled probabilities \(p_i^{(m)*} = \mathbb P(Y_i \mid X_i,\theta^{(m)*})\), we get the following estimator:
+  \item[Average sampled probability] If we choose to average the sampled probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 \mid X_i,\theta^{(m)*})\), we get the following estimator:
   \[
-    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \right), \quad i=1,\dots,n'.
+    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
   \]
@@
 \[
-  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*}\right), \quad i=1,\dots,n',
+  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n',
 \]
@@
 \end{description}
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) variables from each regressor based on the posterior distribution of \(\tau\mid \mathcal D_n\) and feed the reduced dataset to a multivariate binary classification procedure.
+Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a multivariate binary classification procedure.
 
@@
 One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model.
-While in principle we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted implications.
+While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications.
 For instance, it would make the implementation more complex, since the dimensionality of the parameters \(\beta\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next.
@@
 In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, AIC, \ldots) to select its optimal value.
+BIC, DIC, \ldots) to select its optimal value \incomment{deberíamos elegir uno de los criterios?}.
 
@@
   \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation).
-Our numerical studies showed that the results were much better with this choice than, say, with a random or a null vector.
+Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
   \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
@@
 
-To conclude this section, it is worth mentioning that the Bayesian nature of our model allows us to perform some model validation checks straight away.
+To conclude this section, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot of the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}).
-In addition, we can compute the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}_m\) that fall below the real value of the statistic.
+Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}).
+In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}\) that fall below the real value of the statistic.
 They are expected to be around 0.5 when the model accurately represents the data \citep[see][Ch.~6]{gelman1995bayesian}.

### 2022-03-07 paper/paper.tex
Some corrections

@@
 
-The most common linear regression model is the classical \(L^2\)-model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}.
+The most common functional linear regression model is the classical \(L^2\)-model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}.
 It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
@@
 These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
+
+\begin{comment}
+  Quizás falta comentar algo de los resultados experimentales...
+\end{comment}
 
@@
 \[
-\mathcal H_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_j \in \R, \ t_1,\dots,t_j \in [0, 1]\right\}.
+\mathcal H_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_p \in \R, \ t_1,\dots,t_p \in [0, 1]\right\}.
 \]
@@
 At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution \(Y^*\) of the ``sampled responses''
+  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
   \[
-  \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n', \ m=1,\dots,M\right\}
+  Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
   \]
-  following model~\eqref{eq:rkhs-model-linear-2}. Then, we can take the mean of all such samples as a proxy for the response:
+  Then, we can take the mean of all such samples as a proxy for each response variable:
   \[
@@
 
-In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X\), and consider an RKHS approach given by the following functional logistic equation:
+In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1\mid X=x)\).
+Then, an RKHS approach given by the following functional logistic equation:
 \begin{equation}\label{eq:rkhs-model-logistic}
@@
 \end{equation}
-Following our previous approach, after incorporating the sample information we can rewrite the classification model~\eqref{eq:rkhs-model-logistic} as
+Following our previous strategy, after incorporating the sample information we can rewrite the classification model~\eqref{eq:rkhs-model-logistic} as
 \begin{equation}\label{eq:rkhs-model-logistic-2}
@@
 Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.~1]{buenolarraz2021functional}.
+
+\begin{comment}
+  The proposed model holds whenever \(X_0\), \(X_1\) are homoscedastic Gaussian processes (Bueno-Larraz et al.).
+\end{comment}
 
@@
   \]
-  \item[Sampled majority vote] If we choose to average the sampled binary responses \(Y^*\) instead, we compute the predictions as
+  \item[Sampled majority vote] If we choose to average the sampled binary responses \(Y^{(m)*}\) instead, we compute the predictions as
 \[
@@
 That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter.
-This standardization technique was used previously in the literature; see for example \citet{grollemund2019bayesian}.
+This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
 \end{itemize}

### 2022-03-07 paper/paper.tex
Before section 3

@@
     \item Métricas usadas para medir el error.
-    \item Contar los algoritmos de comparación usados.
+    \item Contar los algoritmos de comparación usados (elegir algunos; ver Torrecilla et al 2021, incluir regla de clasificación ``óptima'').
     \item Mostrar tablas de resultados y gráficas.
+En las gráficas, unificar el eje Y para los mismos modelos, y a ser posible, pintar media y desviación típica de varias ejecuciones (haciendo CV con distintos splits train/test).
 \end{itemize}

### 2022-03-15 paper/paper.tex
Add some comments, pending revision

@@
 \begin{comment}
-  Quizás falta comentar algo de los resultados experimentales...
+  Quizás falta comentar algo de los resultados experimentales... también habría que hacer más hincapié en cuál es nuestra propuesta; que se vea bien.
 \end{comment}
@@
 
-There are at least three alternative, equivalent ways of defining the RKHS associated with \(K\), and each of them gives a different insight into the relationship between the process itself and the underlying RKHS.
+There are at least three alternative, equivalent ways of defining the RKHS associated with \(K\), and each of them gives a different insight into the relationship between the process itself and the underlying RKHS:
 
@@
 When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get a sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples:
+Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples:
 
@@
 \]
-where \(\mathbb I\) is the indicator function and the hat estimates are obtained through a summary statistic of the corresponding marginal posterior distributions.
+where \(\mathbb I\) is the indicator function. The hat estimates are again obtained through a summary statistic of the corresponding marginal posterior distributions.
+
 On the other hand, the distributional estimates are somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples:
@@
 \end{description}
+
 Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a multivariate binary classification procedure.
@@
 \begin{itemize}
-  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation).
+  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation) \incomment{La función de verosimilitud es complicada...
+pero nos basta con una aproximación como punto de partida}.
 Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
@@
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}).
+Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}) \incomment{conseguir mejor calidad de la figura}.
 In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}\) that fall below the real value of the statistic.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{supplement}

### 2022-03-22 paper/paper.tex
Some corrections

@@
 \begin{abstract}
-In this work we propose a Bayesian approach for functional linear or logistic  regression models, based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
-These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional alternative to the classical functional regression paradigm.
+We propose a Bayesian approach for functional linear or logistic  regression models, based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
+These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
 The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subset of the RKHS generated by the underlying covariance function.
@@
 
+\begin{comment}
+  \begin{itemize}
+    \item Utilizar \(\alpha\) como parámetro en los modelos \(L^2\)?
+    \item Qué se demuestra exactamente en la proposición 2.1?
+    \item Suponer que los datos \(X\) están centrados desde el principio.
+    \item Definir la isometría de Loève al revés
+  \end{itemize}
+\end{comment}
+
 \section{Introduction}\label{sec:intro}
 
-The problem of inferring a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
-While in principle we could simply treat the functional data as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
-As a consequence, numerous proposals have arisen on how to suitably treat functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on continuous data.
-A recent survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} and \citet{horváth2012inference}.
+The problem of predicting a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
+While in principle we could simply regard the functional data as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
+As a consequence, numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on continuous data.
+A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} and \citet{horváth2012inference}.
 
-In this work we are concerned with (supervised) functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
-Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges that emerge as a result of working in infinite dimensions.
-To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some hidden dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
+Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some more subtle drawbacks that emerge as a result of working in infinite dimensions.
+To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
+In addition, without loss of generality we will suppose that \(X\) is centered, i.e., that its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
+Furthermore, we will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
-The most common functional linear regression model is the classical \(L^2\)-model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}.
+\subsubsection{\(L^2\)-models, shortcomings and alternatives}
+
+ The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the book by~\citet{ramsay2005functional}.
 It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
 \begin{equation}\label{eq:l2-linear-model}
-Y = \beta_0 + \dotprod{X}{\beta} + \epsilon = \beta_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
+Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
 \end{equation}
-\incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\beta_0\in \R\) and \(\epsilon\) is an independent error term with certain simplifying assumptions.
-In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinite-dimensional space \(L^2[0, 1]\).
-Aside from the fact that functional (i.e.
-infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle drawbacks that arise in this \(L^2\)-framework.
-For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case \citep[see][]{berrendero2019rkhs}.
+where \(\alpha_0\in \R\), \(\epsilon\) is an error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is supposed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
+In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
+In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on a linear combination of the marginals, such as \(Y=\beta_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\) \citep[see][]{berrendero2019rkhs, berrendero2020general}.
 Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
@@
 \begin{equation}\label{eq:l2-logistic-model}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{X}{\beta}\}},
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
 \end{equation}
-where \(\beta_0 \in \R\) and \(\beta \in L^2[0, 1]\).
-In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{buenolarraz2021functional}.
+where \(\alpha_0 \in \R\) and \(\beta \in L^2[0, 1]\).
+In this situation, the most common way of estimating the slope function \(\beta(\cdot)\) is via its Maximum Likelihood Estimator (MLE).
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2021functional}.
 
@@
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-These RKHS-based models have already been explored in \citet{berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.
-The main contribution of this work is a parametric Bayesian approach to parameter estimation within these recently-proposed models, in which a prior distribution is imposed on the unknown function to  obtain a posterior distribution after seeing the data.
+In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is particularly appealing to practitioners confronted with functional data problems due to its simplicity.
+These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2021functional} for the case of logistic regression.
 
-Another set of techniques widely studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion.
+A major aim of this work is to motivate these recently-proposed models inside the functional framework, while also providing efficient techniques to apply them in practice.
+Our main contribution is the proposal of a Bayesian approach to parameter estimation within the aforementioned RKHS models, in which a prior distribution is imposed on the unknown functional parameter to obtain a posterior distribution after seeing the data.
+In this way, one can incorporate pre-existing beliefs about the parameters, thus increasing the flexibility of the estimation procedure.
+Even though setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
+\incomment{Comentar resultados experimentales}
+
+Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion.
 As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model.
+As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
 In this way, in addition to making predictions about the input data, we could evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
 These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-
-\begin{comment}
-  Quizás falta comentar algo de los resultados experimentales... también habría que hacer más hincapié en cuál es nuestra propuesta; que se vea bien.
-\end{comment}
+\incomment{Comentar resultados experimentales}
 
@@
 
-\begin{comment}
-¿Mejor poner esto después de la sección 1.1?
-\end{comment}
-
-In Section~\ref{sec:rkhs} we carry out a quick review of the theory of RKHS's from a probabilistic point of view.
-Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models proposed.
+In Section~\ref{sec:rkhs} we review the basics of the theory of RKHS's from a probabilistic point of view.
+Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models we propose.
 The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of computational details.
@@
 
-\subsection{A brief review of the theory of RKHS's}\label{sec:rkhs}
-
-\begin{comment}
-  Es posible que esta sección sea demasiado larga...?
-\end{comment}
+\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
 
 The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and its covariance function \(K(t, s)= \mathbb E[(X(t) - m(t))(X(s) - m(s))]\) are both continuous.
+In what follows, suppose \(X=X(t)\) is a centered \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its covariance function \(K(t, s)= \mathbb E[(X(t) - m(t))(X(s) - m(s))]\) is continuous.
 To construct the RKHS associated with the covariance function, we start by defining the functional space \(\mathcal H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
-\[
+\begin{equation}\label{eq:h0}
 \mathcal H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
-\]
+\end{equation}
 This space is endowed with the inner product
@@
 Then, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.
-As it turns out, functions in this space verify the \textit{reproducing property}, thus giving it its name:
+\incomment{Contar definiciones alternativas (con referencias), y definir el operador lineal \(\mathcal K\) inline, diciendo que al ser K continua es compacto y autoadjunto.} As it turns out, functions in this space verify the \textit{reproducing property}, thus giving it its name:
 \begin{equation}\label{eq:reproducing-property}
@@
 
-As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS \incomment{¿Es necesario destacar alguna suposición/propiedad extra sobre \(K\)?
-Por ejemplo \(\int\int K^2 <\infty\) o que es simétrica y (semi-) definida positiva}.
-A linear operator \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}\ closely related to this covariance function is the so-called covariance operator, namely
+\begin{comment}
+  Un enfoque particularmente útil en estadística es la isometría de Loeve....
+
+  SUPONER QUE X ESTÁ CENTRADA DESDE EL PRINCIPIO.
+\end{comment}
+
+As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
+A compact self-adjoint operator \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}\ closely related to this covariance function is the so-called covariance operator, namely
 \[
@@
 \]
-This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear}, which asserts that a centered process \(X\) can be decomposed as
+\begin{comment}
+  QUITAR K-L Y HABLAR DE LOS AUTOVALORES EN LA DEFINICIÓN ALTERNATIVA.
+\end{comment}
+
+This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear}, which asserts that \(X\) can be decomposed as
 \begin{equation}\label{eq:karhunen-loeve}
@@
   \[
-  \sum_{i=1}^p a_i (X(t_i) - m(t_i)) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
+  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
 \]
-and can be formally defined as
+and can be formally defined as \incomment{Defnir al revés y usar \(\Psi\) directamente en lo sucesivo.}
 \begin{equation}\label{eq:loeves-isometry}
-  \Psi_X(U)(t) = \E[U(X(t) - m(t))], \quad U \in \mathcal L(X).
+  \Psi_X(U)(t) = \E[U X(t)], \quad U \in \mathcal L(X).
 \end{equation}
@@
 \begin{comment}
-  Puede que las definiciones 2ª y 3ª arriba sean innecesarias.
+  Enlazar con la definición de isometría de loeve
 \end{comment}
@@
 In this section we present the precise models and Bayesian methodologies explored in this work.
-In the sequel, we will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\).
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for Loève's isometry \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\) defined in \eqref{eq:h0}, i.e.,
 \begin{equation}\label{eq:alpha-parameter-rkhs}
@@
 
-The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data.
-In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
+The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the information of the available sample data.
+\incomment{Una de las cosas difíciles en estadística bayesiana es poner un prior en un espacio de funciones.
+Una de las ventajas del modelopropuesto (y por el que advocamos por este modelo) es que facilita esta tarea, de manera natural y manejable.
+Poner una subsubsection con: definición de una priori adecuada en nuestro RKHS framework}In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry.
@@
 
-On a separate note, it is worth mentioning that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
+To begin with, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
 This is why our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
@@
 \begin{prop} Let \(H\) be a separable Hilbert space and consider a dense subset \(H_0\subseteq H\).
-Then, any probability distribution on \((H, \mathcal{B}(H))\) is uniquely determined by its restriction to \((H_0, \mathcal B(H_0))\).
+\incomment{Decirlo al revés: la distribución se extiende de forma única a partir del denso.} Then, any probability distribution on \((H, \mathcal{B}(H))\) is uniquely determined by its restriction to \((H_0, \mathcal B(H_0))\).
 \end{prop}
@@
 \end{proof}
+
+\incomment{Mencionar que en realidad trabajamos en \(H_{0,p}\).}
 
@@
 However, this problem can be easily circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.}.
-Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.~1]{buenolarraz2021functional}.
+Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.~1]{berrendero2021functional}.
 

### 2022-03-28 paper/paper.tex
Some corrections bis

@@
   \begin{itemize}
-    \item Utilizar \(\alpha\) como parámetro en los modelos \(L^2\)?
-    \item Qué se demuestra exactamente en la proposición 2.1?
-    \item Suponer que los datos \(X\) están centrados desde el principio.
-    \item Definir la isometría de Loève al revés
+    \item Poner los vectores en negrita.
+    \item Quitar notas a pie de página
   \end{itemize}
@@
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-In addition, without loss of generality we will suppose that \(X\) is centered, i.e., that its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-Furthermore, we will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+We will further suppose without loss of generality that \(X\) is centered, that is, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
+In addition, we will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
@@
 The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-In what follows, suppose \(X=X(t)\) is a centered \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its covariance function \(K(t, s)= \mathbb E[(X(t) - m(t))(X(s) - m(s))]\) is continuous.
+In what follows, suppose \(X=X(t)\) is a centered \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its covariance function \(K(t, s)= \mathbb E[X(t)X(s)]\) is continuous.
 To construct the RKHS associated with the covariance function, we start by defining the functional space \(\mathcal H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
@@
 As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-A compact self-adjoint operator \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}\ closely related to this covariance function is the so-called covariance operator, namely
+An integral operator closely related to this covariance function is the so-called covariance operator, namely
 \[
-\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1].
+\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
 \]
+which is compact and self-adjoint when \(K\) is continuous \incomment{cita}.
+
 \begin{comment}
@@
 \]
-and can be formally defined as \incomment{Defnir al revés y usar \(\Psi\) directamente en lo sucesivo.}
+and can be formally defined as
 \begin{equation}\label{eq:loeves-isometry}
-  \Psi_X(U)(t) = \E[U X(t)], \quad U \in \mathcal L(X).
+  \Psi^{-1}_X(U)(t) = \E[U X(t)], \quad U \in \mathcal L(X).
 \end{equation}
@@
 As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
-However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
+However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
 This notation often proves to be useful and convenient.
@@
 The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for Loève's isometry \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\) defined in \eqref{eq:h0}, i.e.,
-\begin{equation}\label{eq:alpha-parameter-rkhs}
-  \alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot) \in \mathcal H_0(K).
-\end{equation}
-
+However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\mathcal H_0(K)\) defined in~\eqref{eq:h0}.
 The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
@@
 
-The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the information of the available sample data.
+The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
 \incomment{Una de las cosas difíciles en estadística bayesiana es poner un prior en un espacio de funciones.
@@
 Poner una subsubsection con: definición de una priori adecuada en nuestro RKHS framework}In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
+Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry.
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
 \[
@@
 This is why our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
-The following result proves that any such extension is indeed unique.
-
-\begin{prop} Let \(H\) be a separable Hilbert space and consider a dense subset \(H_0\subseteq H\).
-\incomment{Decirlo al revés: la distribución se extiende de forma única a partir del denso.} Then, any probability distribution on \((H, \mathcal{B}(H))\) is uniquely determined by its restriction to \((H_0, \mathcal B(H_0))\).
-\end{prop}
-
-\begin{proof}
-
-First, note that Theorem 7.1.2 in \citet[p.~177]{hsing2015theoretical} establishes that the distribution of any random variable \(X: (\Omega, \mathcal A, \mathbb{P})\to (H, \mathcal B(H))\) is uniquely determined by the distribution of its random projections \(\dotprod{X}{h}_H\) for all \(h \in H\).
-Consider two random variables \(\alpha\) and \(\beta\), taking values in \(H\) and equally distributed on \(H_0\), and let \(h \in H\).
-Since \(H_0\) is dense in \(H\), there is a sequence \(\{h_n\}\subseteq H_0\) such that \(\|h_n - h\|_H \to 0\), and necessarily \(\dotprod{\alpha}{h_n}_H \equiv \dotprod{\beta}{h_n}_H\) for all \(n\), where the equality is in distribution.
-
-Observe now that \(\dotprod{\alpha}{h_n}_H \overset{P}{\to} \dotprod{\alpha}{h}_H\), since
-\[
-\forall \epsilon > 0 \quad \mathbb{P}\{|\dotprod{\alpha}{h_n - h}_H| > \epsilon\} \leq \mathbb{P}\{\|\alpha\|_H \|h_n - h\|_H > \epsilon\} \to 0,
-\]
-and equivalently \(\dotprod{\beta}{h_n}_H \overset{P}{\to} \dotprod{\beta}{h}_H\), so that \(\dotprod{\alpha}{h}_H \equiv \dotprod{\beta}{h}_H\).
-As this is valid for every \(h \in H\), we can conclude that \(\alpha \equiv \beta\) on \(H\).
-\end{proof}
 
@@
 
+\incomment{Disminuir mucho el párrafo de MCMC, simplemente comentarlo de pasada y en todo caso alguna referencia a una survey.}
+
 A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant.
-Nevertheless, we can sample from the posterior distribution\footnote{Provided that the marginal likelihood \(\int \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of Markov Chain Monte Carlo (MCMC) sampling methods.
+Nevertheless, we can sample from the posterior distribution using an algorithm of the family of Markov Chain Monte Carlo (MCMC) sampling methods.
 These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known.
@@
 \begin{equation}\label{eq:rkhs-model-linear}
-  Y = \alpha_0 + \Psi^{-1}_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
+  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
 \end{equation}
 where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term verifying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multivariate linear model with the \(X(t_j)\) as scalar covariates.
+When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
 Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
@@
 On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
+\incomment{Comentar la distribución a priori en su forma completa, decir que son independientes.}
 
 \begin{comment}
-¿Es necesario/conveniente escribir aquí la expresión teórica de la distribución a posteriori? O quizás en el apéndice/material suplementario mejor?
+Pequeña proposición con la cuenta de la posteriori
 \[
@@
 
-When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get a sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples:
+When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
+Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples: \incomment{Recordemos que el algoritmo nos da una observación del vector de parámetros,y hay varias estrategias para usarlas.}
+
+\begin{comment}
+  Comentar que una es promediar y luego predecir, y la otra es al revés.
+Separar un poco lo de selección de variables, comentando que nos concentramos solo en los tiempos y descartamos el resto (se consideran nuisance) (promediar los subvectores de tiempos y luego predecir como en la estrategia A, pero usando modelo multivariante).
+Cambiar los nombres.
+\end{comment}
 
 \begin{description}
-  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way by
+  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way \incomment{Following model...} by
   \[
@@
 
+\begin{comment}
+Comentar después en la sección de resultados cómo se comparan, cuál es más estable.....
+\end{comment}
+
 Note that these estimators can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
@@
 \begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi^{-1}_X(\alpha)\}}, \quad \alpha \in \mathcal H_{0,p}(K).
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha \in \mathcal H_{0,p}(K).
 \end{equation}
@@
   \begin{align*}
-\log \pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi^{-1}_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}^{-1}(\alpha)\right\}\right)\right]\\
+\log \pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}(\alpha)\right\}\right)\right]\\
 \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0).
@@
 \subsection{Model choice and validation}\label{sec:model-choice}
+
+\begin{comment}
+  Distancia máxima a la que puede estar una función de \(H_0,p\) de otra de \(H\)? O de \(H_0\)?
+\end{comment}
 
@@
 En las gráficas, unificar el eje Y para los mismos modelos, y a ser posible, pintar media y desviación típica de varias ejecuciones (haciendo CV con distintos splits train/test).
+    \item Hacer un caso de estudio en el dataset RKHS, mostrando las aproximaciones de los tiempos a los tiempos reales.
 \end{itemize}

### 2022-03-31 paper/paper.tex
Rewriting of section 1.1

@@
 \newcommand{\E}{\mathbb{E}}
+\newcommand{\Hcal}{\ensuremath\mathcal{H}}
 \newcommand{\Var}{\ensuremath\operatorname{Var}}
@@
 
+\begin{comment}
+Comentar los enfoques principales en la literatura: reducción a problema multivariante, desarrollo en bases, splines, etc. Active area of research.... (ver papers para comparar).
+\end{comment}
+
 In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
@@
 We will further suppose without loss of generality that \(X\) is centered, that is, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+In addition, we will tacitly assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
@@
 \end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is an error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is supposed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
+where \(\alpha_0\in \R\), \(\epsilon\) is an error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
 In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on a linear combination of the marginals, such as \(Y=\beta_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\) \citep[see][]{berrendero2019rkhs, berrendero2020general}.
+In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\) (see \citet{berrendero2020general} for additional details on this).
 Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
@@
 The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-In what follows, suppose \(X=X(t)\) is a centered \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its covariance function \(K(t, s)= \mathbb E[X(t)X(s)]\) is continuous.
-To construct the RKHS associated with the covariance function, we start by defining the functional space \(\mathcal H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
+In what follows, suppose that the covariance function of \(X\), \(K(t, s)= \mathbb E[X(t)X(s)]\), is continuous.
+To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
 \begin{equation}\label{eq:h0}
-\mathcal H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
+\Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
 \end{equation}
-This space is endowed with the inner product
-\[
-\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
-\]
-given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.
-\incomment{Contar definiciones alternativas (con referencias), y definir el operador lineal \(\mathcal K\) inline, diciendo que al ser K continua es compacto y autoadjunto.} As it turns out, functions in this space verify the \textit{reproducing property}, thus giving it its name:
-\begin{equation}\label{eq:reproducing-property}
-  \dotprod{K(t, \cdot)}{f}_K = f(t), \quad \text{for all } f \in \mathcal H(K), \ t \in [0, 1].
-\end{equation}
-An important detail to note is that \(\mathcal H(K)\) is a space of genuine functions, and not of equivalence classes of functions.
-Hence, the values of these functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
+This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
+Then, we define \(\Hcal(K)\) to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product defined above.
+As it turns out, functions in this space verify the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\), thus giving it its name.
+An important detail to note is that, since \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, the values of these functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
 
-\begin{comment}
-  Un enfoque particularmente útil en estadística es la isometría de Loeve....
+As we can see, the covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
+An integral operator closely related to this kernel is the so-called covariance operator, namely \(\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds\) for \(f \in L^2[0, 1]\), which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}.
+It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since it is easy to show that we can also express \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with the corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\), where in turn \(\lambda_j\) and \(\phi_j\) are the eigenvalues and eigenfunctions of \(\mathcal K\), respectively.
+Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\subset \R^+_0\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
 
-  SUPONER QUE X ESTÁ CENTRADA DESDE EL PRINCIPIO.
-\end{comment}
-
-As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-An integral operator closely related to this covariance function is the so-called covariance operator, namely
-\[
-\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
-\]
-which is compact and self-adjoint when \(K\) is continuous \incomment{cita}.
-
-\begin{comment}
-  QUITAR K-L Y HABLAR DE LOS AUTOVALORES EN LA DEFINICIÓN ALTERNATIVA.
-\end{comment}
-
-This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear}, which asserts that \(X\) can be decomposed as
-\begin{equation}\label{eq:karhunen-loeve}
-X(t) = \sum_{i=1}^\infty \xi_i \phi_i(t),
-\end{equation}
-where \(\phi_i\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_i = \int X\phi_i\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_i\), the eigenvalue associated with \(\phi_i\).
-This expansion is the main justification for many FDA techniques, such as functional principal component analysis (FPCA)~\cite[see for example][Ch.~8]{ramsay2005functional}.
-
-There are at least three alternative, equivalent ways of defining the RKHS associated with \(K\), and each of them gives a different insight into the relationship between the process itself and the underlying RKHS:
-
-\begin{description}
-  \item[From an isometry] Via \textit{Loève's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see Lemma 1.1 in \citet{lukic2001stochastic}).
+Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
+Specifically, via \textit{Loève's isometry} one can establish a congruence between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see Lemma 1.1 in \citet{lukic2001stochastic}).
 This isometry is essentially the completion of the correspondence
@@
 \]
-and can be formally defined as
-\begin{equation}\label{eq:loeves-isometry}
-  \Psi^{-1}_X(U)(t) = \E[U X(t)], \quad U \in \mathcal L(X).
-\end{equation}
-  \item[From the covariance operator] The space \(\mathcal H(K)\) can also be identified with the image of the square root of the covariance operator, i.e.:
-  \begin{equation}\label{eq:rkhs-square-root}
-  \mathcal H(K) = \mathcal K^{1/2}(L^2[0, 1]).
-\end{equation}
-In this case, the inner product can be expressed as
-\[
-\dotprod{f}{g}_K = \dotprod{\mathcal K^{1/2}(f)}{\mathcal K^{1/2}(g)}.
-\]
-
-  \item[From a \(\bf L^2\)-like smoothed norm] Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \citep{mercer1909functions}, we can also express
-  \begin{equation}\label{rkhs-sum-lambda}
-    \mathcal H(K) = \left\{f \in L^2[0, 1]: \sum_{j=1}^\infty \frac{\dotprod{f}{\phi_j}^2}{\lambda_j} < \infty \right\},
-  \end{equation}
-  with the corresponding inner product
-  \[
-  \dotprod{f}{g}_K = \sum_{j=1}^\infty \frac{\dotprod{f}
-  {\phi_j}\dotprod{g}{\phi_j}}{\lambda_j}.
-  \]
-Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
-\end{description}
-
-\begin{comment}
-  Enlazar con la definición de isometría de loeve
-\end{comment}
-
-Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
+and can be formally defined as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
+Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
 As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
-However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\).
+However, following Parzen's approach in his seminal work on time series analysis \citep[Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\).
 This notation often proves to be useful and convenient.
@@
 In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for Loève's isometry \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\mathcal H_0(K)\) defined in~\eqref{eq:h0}.
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for Loève's isometry \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
 The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
@@
 \[
-\mathcal H_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_p \in \R, \ t_1,\dots,t_p \in [0, 1]\right\}.
+\Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_p \in \R, \ t_1,\dots,t_p \in [0, 1]\right\}.
 \]
@@
 Una de las ventajas del modelopropuesto (y por el que advocamos por este modelo) es que facilita esta tarea, de manera natural y manejable.
-Poner una subsubsection con: definición de una priori adecuada en nuestro RKHS framework}In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
+Poner una subsubsection con: definición de una priori adecuada en nuestro RKHS framework}In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we have
 \[
@@
 
-To begin with, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B \in \mathcal B(\mathcal H(K))\).
-This is why our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\mathcal H_0(K)\) can be directly extended to a prior distribution on \(\mathcal H(K)\).
+To begin with, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B \in \mathcal B(\Hcal(K))\).
+This is why our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
 
@@
 
-In the case of functional linear regression, the RKHS model considered (with a functional parameter \(\alpha\in \mathcal H_{0,p}(K)\)) is
+In the case of functional linear regression, the RKHS model considered (with a functional parameter \(\alpha\in \Hcal_{0,p}(K)\)) is
 \begin{equation}\label{eq:rkhs-model-linear}
@@
 \begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha \in \mathcal H_{0,p}(K).
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha \in \Hcal_{0,p}(K).
 \end{equation}

### 2022-04-03 paper/paper.tex
Start rewriting of section 2 and corrections

@@
 \usepackage{amssymb}
+\usepackage{xcolor}
+\usepackage{bm}
 \usepackage{natbib}
@@
 
-%% Environment for colored comments
-\newenvironment{comment}
+%% Colors
+\definecolor{comment-red}{RGB}{240,30,70}
+
+%% Environment for colored omments
+\newenvironment{comment}[1][comment-red]
 {
-\noindent \em \color{red}
+\noindent \color{#1}
 }
@@
 %% Inline comments
-\newcommand\incomment[1]{\color{red}[\textit{#1}]\color{black}}
+\newcommand\incomment[2][comment-red]{\color{#1}[\textit{#2}]\color{black}}
 \endlocaldefs
@@
 \title{Bayesian RKHS-based methods in functional regression\support{Support information of the article.}}
-\runtitle{Bayesian RKHS-based methods in functional regression }
+\runtitle{Bayesian RKHS-based methods in functional regression}
 
@@
 \begin{abstract}
-We propose a Bayesian approach for functional linear or logistic  regression models, based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
+We propose a novel Bayesian approach for functional linear or logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
 These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
 The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subset of the RKHS generated by the underlying covariance function.
-By imposing a suitable prior distribution on such RKHS, we can perform data-driven inference via standard Bayes methodology.
-The posterior distribution can be approximated from Markov Chain Monte Carlo (MCMC) methods.
-Several estimators derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection method.
+By imposing a suitable prior distribution on such a RKHS, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo (MCMC) methods.
+Several prediction methods derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
 \end{abstract}
@@
 \begin{comment}
+Posibles temas de desarrollo teórico:
   \begin{itemize}
-    \item Poner los vectores en negrita.
-    \item Quitar notas a pie de página
+    \item Enfoque de procesos Gaussianos como priors en modelos RKHS, eligiendo un kernel \(K'\) que tenga buenas propiedades con respecto a \(K\). Distribuciones conjugadas.
+    \item Temas de consistencia de la posteriori: concentración hacia el ``verdadero'' valor el parámetro o bien convergencia hacia la ``verdadera'' posteriori.
+    \item Distancia máxima entre las funciones de \(\Hcal_{0,p}(K)\) y las de \(\Hcal_0(K)\) (ó \(\Hcal(K)\)).
+    \item Robustez de las distribuciones ante pequeñas perturbaciones.
+    \item  En \citet[Sec.~2.5]{grollemund2019bayesian}, los autores proponen un estimador Bayesiano para sus funciones ``escalón'' \(\beta_\theta(\cdot)\) utilizando una pérdida cuadrática, de la forma
+  \[
+    \hat \beta \in \arg\min_{d\in L^2[0, 1]} \int\int (\beta_\theta(t) - d(t))^2 \pi(\theta\mid \mathcal D_n)\, dt\, d\theta.
+  \]
+  ¿Podríamos hacer algo parecido con nuestro parámetro funcional \(\alpha \in \Hcal_{0,p}(K)\)?
   \end{itemize}
@@
 \begin{comment}
-Comentar los enfoques principales en la literatura: reducción a problema multivariante, desarrollo en bases, splines, etc. Active area of research.... (ver papers para comparar).
+Comentar los enfoques principales en la literatura: reducción a problema multivariante, desarrollo en bases, splines, etc.
+Algo parecido: gaussian process regression (multivariante, no funcional, aunque no paramétrico), citar dos ejemplos de la bibliografía (multivariante y funcional).
+Active area of research....
+(ver papers para comparar).
 \end{comment}
@@
 In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\) (see \citet{berrendero2020general} for additional details on this).
+In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
 Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
@@
 
-It turns out that in both situations a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
+It turns out that in both circumstances a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
@@
 Even though setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-\incomment{Comentar resultados experimentales}
+\incomment{Citar grollemund como ejemplo} \incomment{Comentar resultados experimentales}
 
-Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion.
+Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion.
 As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
+As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
 In this way, in addition to making predictions about the input data, we could evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
@@
 The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-In what follows, suppose that the covariance function of \(X\), \(K(t, s)= \mathbb E[X(t)X(s)]\), is continuous.
+Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of \(X\), and in what follows suppose that it is continuous.
 To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
@@
 This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, we define \(\Hcal(K)\) to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product defined above.
-As it turns out, functions in this space verify the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\), thus giving it its name.
+Then, we define \(\Hcal(K)\) to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
+As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\), thus giving it its name.
 An important detail to note is that, since \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, the values of these functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
@@
 It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since it is easy to show that we can also express \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with the corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\), where in turn \(\lambda_j\) and \(\phi_j\) are the eigenvalues and eigenfunctions of \(\mathcal K\), respectively.
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since it is easy to show that we can also express \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and eigenfunctions of \(\mathcal K\), respectively.
+In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
 Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\subset \R^+_0\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
@@
 In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for Loève's isometry \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
 However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
-The value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
+For practical and computational reasons, the value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
 Thus, we will search for our functional parameter in the set
-\[
+\begin{equation}\label{eq:h0p}
 \Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_p \in \R, \ t_1,\dots,t_p \in [0, 1]\right\}.
-\]
+\end{equation}
 
 The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-\incomment{Una de las cosas difíciles en estadística bayesiana es poner un prior en un espacio de funciones.
-Una de las ventajas del modelopropuesto (y por el que advocamos por este modelo) es que facilita esta tarea, de manera natural y manejable.
-Poner una subsubsection con: definición de una priori adecuada en nuestro RKHS framework}In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately.
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry.
@@
 
-To begin with, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B \in \mathcal B(\Hcal(K))\).
-This is why our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-
-\incomment{Mencionar que en realidad trabajamos en \(H_{0,p}\).}
+In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately.
+Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to Gaussian process regression methods).
+Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
+Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
+Even though we actually work on \(\Hcal_{0,p}\), the parameter \(p\) is fundamentally...
 
@@
 \end{equation}
-where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term verifying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
+where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and a generic element will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
 Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
-For a generic element \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (\beta, \tau, \alpha_0, \sigma^2) \in \Theta_p\), the reinterpreted model assumes the form
+For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
 \begin{equation}\label{eq:rkhs-model-linear-2}
@@
 It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
-Furthermore, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
+Also, note that given the time instants \(t_j\) the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
+As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
 In addition, as far as generalizations go the model is easily extensible, seeing as how several functional covariates belonging to the RKHS could be added.
@@
 \begin{comment}
-  En \citet[Sec.~2.5]{grollemund2019bayesian}, los autores proponen un estimador Bayesiano para sus funciones ``escalón'' \(\beta_\theta(t)\) utilizando una pérdida cuadrática, de la forma:
-  \[
-    \hat \beta \in \arg\min_{d\in L^2[0, 1]} \int\int (\beta_\theta(t) - d(t))^2 \pi(\theta\mid \mathcal D_n)\, dt\, d\theta.
-  \]
-  ¿Podríamos hacer nosotros algo parecido con nuestro parámetro funcional \(\alpha\)?
+
 \end{comment}
@@
 In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1\mid X=x)\).
-Then, an RKHS approach given by the following functional logistic equation:
+Then, a RKHS approach given by the following functional logistic equation:
 \begin{equation}\label{eq:rkhs-model-logistic}
@@
 \subsection{Model choice and validation}\label{sec:model-choice}
-
-\begin{comment}
-  Distancia máxima a la que puede estar una función de \(H_0,p\) de otra de \(H\)? O de \(H_0\)?
-\end{comment}
 

### 2022-04-06 paper/paper.tex
More corrections

@@
 \theoremstyle{plain}
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}{Proposition}[section]
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{proposition}{Proposition}[section]
 
@@
 
-%% Environment for colored omments
+%% Environment for colored comments
 \newenvironment{comment}[1][comment-red]
@@
   ¿Podríamos hacer algo parecido con nuestro parámetro funcional \(\alpha \in \Hcal_{0,p}(K)\)?
+  \item Considerar oras distribuciones a priori (especialmente en regresión logística).
   \end{itemize}
@@
 Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} one can establish a congruence between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see Lemma 1.1 in \citet{lukic2001stochastic}).
+Specifically, via \textit{Loève's isometry} one can establish a congruence between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
 This isometry is essentially the completion of the correspondence
@@
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we have
-\[
-\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j).
-\]
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\).
 
 In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately.
-Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to Gaussian process regression methods).
+Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
 Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
 Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}\), the parameter \(p\) is fundamentally...
+Even though we actually work on \(\Hcal_{0,p}(K)\), the parameter \(p\) is fundamentally...\incomment{completar}
 
-In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the parameter vector \(\theta\) on the chosen finite-dimensional parameter space, we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
+In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
 samples amounts to
@@
 In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
-
-\incomment{Disminuir mucho el párrafo de MCMC, simplemente comentarlo de pasada y en todo caso alguna referencia a una survey.}
-
-A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant.
-Nevertheless, we can sample from the posterior distribution using an algorithm of the family of Markov Chain Monte Carlo (MCMC) sampling methods.
-These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known.
-Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the given function.
-In short, they operate by building stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution.
-The Metropolis-Hastings algorithm \citep[e.g.][]{chib1995understanding} is arguably the most well-known method in this regard.
+Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
 
@@
 
-In the case of functional linear regression, the RKHS model considered (with a functional parameter \(\alpha\in \Hcal_{0,p}(K)\)) is
+In the case of functional linear regression, the RKHS model considered (with parameter \(\alpha\in \Hcal_{0,p}(K)\)) is given by
 \begin{equation}\label{eq:rkhs-model-linear}
@@
 \end{equation}
-where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and a generic element will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
+where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term satisfying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
 Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
@@
 As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, as far as generalizations go the model is easily extensible, seeing as how several functional covariates belonging to the RKHS could be added.
+In addition, as far as generalizations go the proposed model is easily extensible, seeing as how several functional covariates could be added to obtain an extended model such as \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
 In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
 
-The prior distribution \(\pi(\theta)\) suggested for the parameter \(\theta \in \Theta_p\) is given by
+The prior distribution suggested for the parameter vector \(\theta \in \Theta_p\) is given by
 \begin{align}\label{eq:prior-linear}
@@
   \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  \beta\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
+  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
 \end{split}
@@
 On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
-In any event, the estimation of \(\alpha_0\) is straightforward, since \(\alpha_0 = \E[Y]\) by~\eqref{eq:rkhs-model-linear-2}, so we could have left this parameter out of the model altogether.
-On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-\incomment{Comentar la distribución a priori en su forma completa, decir que son independientes.}
+In any event, the estimation of \(\alpha_0=\E[Y]\) is straightforward, so it could have been left out of the model altogether.
+On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
+All in all, the proposed prior distribution becomes \(\pi(\theta) = \pi(\alpha_0, \sigma^2)\pi(\tau)\pi(b\mid \tau, \sigma^2)\).
 
-\begin{comment}
-Pequeña proposición con la cuenta de la posteriori
+As for the posterior distribution, a standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
+
+\begin{proposition}
+Under model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the posterior distribution
 \[
-\pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right) \right\}.
+\pi(\theta \mid \mathcal D_n) \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right) \right\},
 \]
+where \(Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
+\end{proposition}
 
-\end{comment}
-
-When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples: \incomment{Recordemos que el algoritmo nos da una observación del vector de parámetros,y hay varias estrategias para usarlas.}
-
-\begin{comment}
-  Comentar que una es promediar y luego predecir, y la otra es al revés.
-Separar un poco lo de selección de variables, comentando que nos concentramos solo en los tiempos y descartamos el resto (se consideran nuisance) (promediar los subvectores de tiempos y luego predecir como en la estrategia A, pero usando modelo multivariante).
-Cambiar los nombres.
-\end{comment}
+Let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
+Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
 
 \begin{description}
-  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way \incomment{Following model...} by
+  \item[Summarize-then-predict] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way \incomment{Following model...} by
   \[
@@
   \]
-  The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+  \incomment{Pasar a la sección 2.3} The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters.
 In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function.
@@
 At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
+  \item[Predict-then-summarize] We can also look at the approximate posterior distribution as a whole and compute the distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
   \[
@@
   \]
-  \item[Variable selection estimates] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\).
-Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\).
-After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known multivariate algorithms suited for this situation.
-\end{description}
 
   \begin{comment}
-Comentar después en la sección de resultados cómo se comparan, cuál es más estable.....
+   Separar un poco lo de selección de variables, comentando que nos concentramos solo en los tiempos y descartamos el resto (se consideran nuisance) (promediar los subvectores de tiempos y luego predecir como en la estrategia A, pero usando modelo de regresión multiple).
   \end{comment}
+  \item[Variable selection] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\).
+Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\).
+After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known prediction algorithms suited for this situation.
+\end{description}
 
@@
 \begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha \in \Hcal_{0,p}(K).
+  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
 \end{equation}
@@
 
-As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner.
-¿Sabéis de alguna que se use en el caso de regresión logística?}.
+As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression approach.
 However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.
@@
   \begin{align*}
-\log \pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}(\alpha)\right\}\right)\right]\\
-\quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0).
+\log \pi(\theta \mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}(\alpha)\right\}\right)\right]\\
+\quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
 \end{align*}
@@
 
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a multivariate binary classification procedure.
+Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a finite-dimensional binary classification procedure.
 
@@
 While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(\beta\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next.
+For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next.
 In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
@@
 \begin{itemize}
-  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation) \incomment{La función de verosimilitud es complicada...
+  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(b\) (see Section~\ref{sec:results} for details on the implementation) \incomment{La función de verosimilitud es complicada...
 pero nos basta con una aproximación como punto de partida}.
@@
     \item Métricas usadas para medir el error.
-    \item Contar los algoritmos de comparación usados (elegir algunos; ver Torrecilla et al 2021, incluir regla de clasificación ``óptima'').
+    \item Comentar cómo se comparan las distintas estrategias de predicción, cuál parece más estable, etc.
+    \item Contar los algoritmos de comparación usados (elegir algunos; ver Torrecilla et al 2021, incluir regla de clasificación ``óptima''?).
     \item Mostrar tablas de resultados y gráficas.
 En las gráficas, unificar el eje Y para los mismos modelos, y a ser posible, pintar media y desviación típica de varias ejecuciones (haciendo CV con distintos splits train/test).
-    \item Hacer un caso de estudio en el dataset RKHS, mostrando las aproximaciones de los tiempos a los tiempos reales.
+    \item Hacer un caso de estudio en el dataset RKHS, midiendo la bondad de las aproximaciones de los instantes de tiempo \(\hat t_j\) a los instantes reales.
+También medir si se recupera el número de componentes subyacente.
 \end{itemize}

### 2022-04-09 paper/paper.tex
Final round of corrections

@@
 
-\begin{comment}
-Comentar los enfoques principales en la literatura: reducción a problema multivariante, desarrollo en bases, splines, etc.
-Algo parecido: gaussian process regression (multivariante, no funcional, aunque no paramétrico), citar dos ejemplos de la bibliografía (multivariante y funcional).
-Active area of research....
-(ver papers para comparar).
-\end{comment}
+FDA is undoubtedly an active area of research; accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
+Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods.
+At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
+Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
+Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
+These ideas extend the theory of Gaussian processes in classical finite-dimensional regression settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative non-parametric Bayesian approach to inference and prediction problems.
 
@@
 These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2021functional} for the case of logistic regression.
+Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
 
@@
 In this way, one can incorporate pre-existing beliefs about the parameters, thus increasing the flexibility of the estimation procedure.
-Even though setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-\incomment{Citar grollemund como ejemplo} \incomment{Comentar resultados experimentales}
+Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
+A similar Bayesian approach has been recently explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
+\incomment{Comentar resultados experimentales}
 
@@
 Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
-Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
+Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) merely by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
 Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the parameter \(p\) is fundamentally...\incomment{completar}
+Even though we actually work on \(\Hcal_{0,p}(K)\), the parameter \(p\) can be selected in several meaningful ways that still make use of the available data, and the range of possibilities is not very wide in practice.
 
@@
 \end{equation}
-In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
+In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y | X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
 Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
@@
 On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-All in all, the proposed prior distribution becomes \(\pi(\theta) = \pi(\alpha_0, \sigma^2)\pi(\tau)\pi(b\mid \tau, \sigma^2)\).
+All in all, the proposed prior distribution becomes \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
 
-As for the posterior distribution, a standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
+As for the posterior distribution, we compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
+A standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
 
 \begin{proposition}
-Under model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the posterior distribution
-\[
-\pi(\theta \mid \mathcal D_n) \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right) \right\},
-\]
-where \(Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
+Under the linear model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
+\begin{align*}
+\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\bm Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
+& + (p+n)\log\sigma - \frac{1}{2}\log |G_\tau|,
+\end{align*}
+where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
 \end{proposition}
 
-Let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\).
+Lastly, in order to generate predictions let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
 Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
@@
 \begin{description}
-  \item[Summarize-then-predict] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way \incomment{Following model...} by
+  \item[Summarize-then-predict] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
   \[
@@
   \]
-  \incomment{Pasar a la sección 2.3} The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability.
-At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-  \item[Predict-then-summarize] We can also look at the approximate posterior distribution as a whole and compute the distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
-  \[
-  Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
-  \]
-  Then, we can take the mean of all such samples as a proxy for each response variable:
+  Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter.
+Although it contributes to measure the uncertainty in the approximations, its estimates are discarded in the final prediction.
+
+  \item[Predict-then-summarize] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
+  \begin{equation}\label{eq:sampled-response-vector}
+  \bm Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
+  \end{equation}
+  Then, we can take the mean of all such samples as a proxy for each response variable, that is,
   \[
@@
   \]
+  This method differs from the previous one in that it takes into account the full approximate posterior distribution instead of summarizing it directly.
 
-  \begin{comment}
-   Separar un poco lo de selección de variables, comentando que nos concentramos solo en los tiempos y descartamos el resto (se consideran nuisance) (promediar los subvectores de tiempos y luego predecir como en la estrategia A, pero usando modelo de regresión multiple).
-  \end{comment}
-  \item[Variable selection] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\).
-Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\).
-After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known prediction algorithms suited for this situation.
+  \item[Variable selection] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
+Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
+After this variable selection has been carried out, we can treat the problem as a finite-dimensional linear regression task and apply any of the well-known prediction algorithms suited for this situation.
 \end{description}
 
-Note that these estimators can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
-As a consequence, what we have in practice is a single algorithm that can produce multiple estimators at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
-Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available estimators is performed to produce a final result.
-
-\begin{comment}
-
-\end{comment}
+Note that these predictors are obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
+As a consequence, what we have in practice is a single algorithm that can produce multiple predictors at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
+Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
 
@@
 
-In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1\mid X=x)\).
-Then, a RKHS approach given by the following functional logistic equation:
+In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
+Then, a RKHS approach is given by the following functional logistic equation:
 \begin{equation}\label{eq:rkhs-model-logistic}
@@
 \end{equation}
-where in turn \(\beta_j\in\R\), \(t_j\in[0, 1]\) and \(\alpha_0\in\R\).
+where in turn \(\beta_j\in\R\) and \(t_j\in[0, 1]\).
 
-In much the same way as the functional linear regression model explained in the previous section, this RKHS-based logistic regression model partially solves some of the issues with the \(L^2\)-model~\eqref{eq:l2-logistic-model} stated in Section~\ref{sec:intro}.
+In much the same way as the linear regression model described before, this RKHS-based logistic regression model partially solves some of the issues with the \(L^2\)-model~\eqref{eq:l2-logistic-model} stated in Section~\ref{sec:intro}.
 Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist.
 However, this problem can be easily circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.}.
-Additionally, under specific conditions this model also entails the logistic \(L^2\)-model in the Gaussian homoscedastic case \citep[see][Th.~1]{berrendero2021functional}.
-
-\begin{comment}
-  The proposed model holds whenever \(X_0\), \(X_1\) are homoscedastic Gaussian processes (Bueno-Larraz et al.).
-\end{comment}
+Additionally, under specific conditions this model also entails the logistic \(L^2\)-model, and in fact it does hold under mild assumptions whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes \citep[see Theorem 1 in][]{berrendero2021functional}.
 
 As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression approach.
-However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.
+However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
+The posterior distribution is in turn derived after a routine calculation.
 
-\begin{comment}
-  Misma pregunta que antes: ¿ponemos la expresión de la posteriori?
+\begin{proposition}
+Under the logistic model~\eqref{eq:rkhs-model-logistic-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
 \begin{align*}
-\log \pi(\theta \mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}(\alpha)\right\}\right)\right]\\
+  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
   \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
 \end{align*}
+Remember that \(\Psi_{X_i}(\alpha) = \sum_j \beta_j X_i(t_j)\).
+\end{proposition}
 
-\end{comment}
-
-Another thing to consider is that in this case we are approximating a probability, so before producing an output we need to transform the predicted values to \(\{0, 1\}\).
-We will follow the standard convention of considering a threshold \(\gamma = 0.5\) and predicting \(\hat Y=1\) if and only if \(\mathbb P(Y=1\mid X) \geq \gamma\).
-With this thresholding strategy in mind, the point estimation approach is analogous to the linear regression case:
+Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to \(\{0, 1\}\).
+We will follow the standard convention of considering a threshold \(\gamma\) (usually \(\gamma = 1/2\)) and predicting \(\hat Y=1\) if and only if \(\mathbb P(Y=1|X) \geq \gamma\).
+With this thresholding strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
 \[
@@
 \]
-where \(\mathbb I\) is the indicator function. The hat estimates are again obtained through a summary statistic of the corresponding marginal posterior distributions.
+where \(\mathbb I\) is the indicator function (\(\mathbb I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
+The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
 
-On the other hand, the distributional estimates are somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples:
+On the other hand, the prediction method that takes into account the entire posterior approximation (i.e.
+the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
+There are primarily two possible outcomes:
 \begin{description}
-  \item[Average sampled probability] If we choose to average the sampled probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 \mid X_i,\theta^{(m)*})\), we get the following estimator:
+  \item[Average sampled probability] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
   \[
@@
   \]
-  \item[Sampled majority vote] If we choose to average the sampled binary responses \(Y^{(m)*}\) instead, we compute the predictions as
+  \item[Average sampled response] Deciding to average the approximate binary responses \(\bm Y^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
 \[
-  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n',
+  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
 \]
-where each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
+In this case, each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
+Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).
 \end{description}
 
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a finite-dimensional binary classification procedure.
+Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced dataset to a finite-dimensional binary classification procedure.
 
@@
 
-One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model.
+One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
 While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next.
+For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
 In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
-Another possibility would be to adapt a purely Bayesian model selection technique \citep[see][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
+Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
 In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, DIC, \ldots) to select its optimal value \incomment{deberíamos elegir uno de los criterios?}.
-
-As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments can be made:
+BIC, DIC, cross-validation, \ldots) to select its optimal value.
+As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
 \begin{itemize}
-  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(b\) (see Section~\ref{sec:results} for details on the implementation) \incomment{La función de verosimilitud es complicada...
-pero nos basta con una aproximación como punto de partida}.
+  \item For the expected value \(b_0\) we propose to use the MLE of \(b\).
+Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation).
 Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
@@
 
+ note that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former optimizes an \(L^2\) loss.
+On the other hand, the mode is also a good candidate because it represents the point of highest probability.
+At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
+
 To conclude this section, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}) \incomment{conseguir mejor calidad de la figura}.
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}\) that fall below the real value of the statistic.
-They are expected to be around 0.5 when the model accurately represents the data \citep[see][Ch.~6]{gelman1995bayesian}.
+Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}).
+In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of approximate estimates \(\{T\{Y^{(m)*}\}: m=1,\dots,M\}\) that fall below the real value of the statistic.
+They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman1995bayesian} for details.
 
@@
   \caption{Posterior predictive graphical checks on a fitted model.
-On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the generated samples, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
+On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the approximate sampled responses, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
 \end{figure}
@@
   \begin{itemize}
-  \item Detalles de la distribución log-posterior en cada caso?
   \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado.

### 2022-04-12 paper/paper.tex
Restructure and more corrections

@@
 \begin{abstract}
-We propose a novel Bayesian approach for functional linear or logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
+We propose a novel Bayesian parametric approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
 These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subset of the RKHS generated by the underlying covariance function.
-By imposing a suitable prior distribution on such a RKHS, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo (MCMC) methods.
-Several prediction methods derived from this posterior distribution turn out to be competitive against other usual  alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
+The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
+By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo (MCMC) methods.
+Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
 \end{abstract}
@@
   ¿Podríamos hacer algo parecido con nuestro parámetro funcional \(\alpha \in \Hcal_{0,p}(K)\)?
-  \item Considerar oras distribuciones a priori (especialmente en regresión logística).
+  \item Considerar otras distribuciones a priori (especialmente en regresión logística).
   \end{itemize}
@@
 The problem of predicting a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
-While in principle we could simply regard the functional data as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
-As a consequence, numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on continuous data.
-A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} and \citet{horváth2012inference}.
+While in principle the functional data could be simply regarded as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
+As a consequence, numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum.
+A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}.
 
-FDA is undoubtedly an active area of research; accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods.
+FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see][]{ullah2013applications}.
+Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
+Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep{aguilera2013comparative}.
 At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
@@
 Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
-These ideas extend the theory of Gaussian processes in classical finite-dimensional regression settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative non-parametric Bayesian approach to inference and prediction problems.
+These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative non-parametric Bayesian approach to inference and prediction problems.
+Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
 
@@
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will further suppose without loss of generality that \(X\) is centered, that is, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
+We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
 In addition, we will tacitly assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
@@
 
- The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the book by~\citet{ramsay2005functional}.
+ The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}.
 It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
@@
 \end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is an error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
+where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
+In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
+In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
 Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
@@
 where \(\alpha_0 \in \R\) and \(\beta \in L^2[0, 1]\).
-In this situation, the most common way of estimating the slope function \(\beta(\cdot)\) is via its Maximum Likelihood Estimator (MLE).
+In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
 However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2021functional}.
@@
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is particularly appealing to practitioners confronted with functional data problems due to its simplicity.
+In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
 These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2021functional} for the case of logistic regression.
-Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
+Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined in the works of \citet{berrendero2018use} and \citet{torrecilla2020optimal}.
 
@@
 Our main contribution is the proposal of a Bayesian approach to parameter estimation within the aforementioned RKHS models, in which a prior distribution is imposed on the unknown functional parameter to obtain a posterior distribution after seeing the data.
-In this way, one can incorporate pre-existing beliefs about the parameters, thus increasing the flexibility of the estimation procedure.
 Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
 A similar Bayesian approach has been recently explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
-\incomment{Comentar resultados experimentales}
+\incomment{Comentar nuestros resultados experimentales}
 
@@
 These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-\incomment{Comentar resultados experimentales}
+\incomment{Comentar nuestros resultados experimentales}
 
@@
 It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since it is easy to show that we can also express \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and eigenfunctions of \(\mathcal K\), respectively.
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since it is easy to show that we can also express \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
 In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
-Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\subset \R^+_0\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
+Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
 
@@
 In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
 However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
@@
 \begin{equation}\label{eq:h0p}
-\Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_1,\dots,\beta_p \in \R, \ t_1,\dots,t_p \in [0, 1]\right\}.
+\Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_j \in \R, \ t_j \in [0, 1]\right\}.
 \end{equation}
@@
 The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry.
+Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi\) is Loève's isometry.
 Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\).
@@
 Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the parameter \(p\) can be selected in several meaningful ways that still make use of the available data, and the range of possibilities is not very wide in practice.
+Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of possible values is not very large in practice.
 
@@
 It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
-Also, note that given the time instants \(t_j\) the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
+Also, note that given the time instants \(t_j\), the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
 As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
@@
 
-As for the posterior distribution, we compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
+As for the posterior distribution, we only compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
 A standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
@@
 
-Lastly, in order to generate predictions let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
+In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
 Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
@@
 In much the same way as the linear regression model described before, this RKHS-based logistic regression model partially solves some of the issues with the \(L^2\)-model~\eqref{eq:l2-logistic-model} stated in Section~\ref{sec:intro}.
-Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist.
+Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}.
 However, this problem can be easily circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.}.
@@
   \]
-  \item[Average sampled response] Deciding to average the approximate binary responses \(\bm Y^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
+  \item[Average sampled response] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
 \[
@@
 
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced dataset to a finite-dimensional binary classification procedure.
+Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
 
@@
   \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter.
+That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
 This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
@@
 
- note that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former optimizes an \(L^2\) loss.
+On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
 On the other hand, the mode is also a good candidate because it represents the point of highest probability.
@@
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of approximate estimates \(\{T\{Y^{(m)*}\}: m=1,\dots,M\}\) that fall below the real value of the statistic.
+Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution \(\bm Y^* \equiv \{\bm Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
+In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\bm Y^{(m)*}\}\) that fall below the real value of the statistic.
 They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman1995bayesian} for details.
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
@@
     \item Comentar cómo se comparan las distintas estrategias de predicción, cuál parece más estable, etc.
-    \item Contar los algoritmos de comparación usados (elegir algunos; ver Torrecilla et al 2021, incluir regla de clasificación ``óptima''?).
+    \item Contar los algoritmos de comparación usados (elegir algunos).
     \item Mostrar tablas de resultados y gráficas.
@@
     \item Hacer un caso de estudio en el dataset RKHS, midiendo la bondad de las aproximaciones de los instantes de tiempo \(\hat t_j\) a los instantes reales.
-También medir si se recupera el número de componentes subyacente.
+También estudiar si se recupera el número de componentes subyacente.
 \end{itemize}
@@
 
-\bibliographystyle{ba}
-\bibliography{bibliography}
+\bibliographystyle{bib/ba}
+\bibliography{bib/bibliography}
 

### 2022-06-06 paper/paper.tex
slides

@@
 
+%% Comentarios
+\usepackage{todonotes}
+\usepackage{ulem}
+
 \startlocaldefs
@@
 \newcommand{\Var}{\ensuremath\operatorname{Var}}
+\definecolor{violet}{rgb}{0.7,0,0.6}
 
@@
 \begin{abstract}
-We propose a novel Bayesian parametric approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
+We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
 These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
 The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo (MCMC) methods.
+By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo methods.
 Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
@@
 \begin{keyword}[class=MSC]
-\kwd[Primary ]{62M20}
+\kwd[primary ]{62M20}
 \kwd[; secondary ]{62F15}
@@
 \end{frontmatter}
-
-\begin{comment}
-Posibles temas de desarrollo teórico:
-  \begin{itemize}
-    \item Enfoque de procesos Gaussianos como priors en modelos RKHS, eligiendo un kernel \(K'\) que tenga buenas propiedades con respecto a \(K\). Distribuciones conjugadas.
-    \item Temas de consistencia de la posteriori: concentración hacia el ``verdadero'' valor el parámetro o bien convergencia hacia la ``verdadera'' posteriori.
-    \item Distancia máxima entre las funciones de \(\Hcal_{0,p}(K)\) y las de \(\Hcal_0(K)\) (ó \(\Hcal(K)\)).
-    \item Robustez de las distribuciones ante pequeñas perturbaciones.
-    \item  En \citet[Sec.~2.5]{grollemund2019bayesian}, los autores proponen un estimador Bayesiano para sus funciones ``escalón'' \(\beta_\theta(\cdot)\) utilizando una pérdida cuadrática, de la forma
-  \[
-    \hat \beta \in \arg\min_{d\in L^2[0, 1]} \int\int (\beta_\theta(t) - d(t))^2 \pi(\theta\mid \mathcal D_n)\, dt\, d\theta.
-  \]
-  ¿Podríamos hacer algo parecido con nuestro parámetro funcional \(\alpha \in \Hcal_{0,p}(K)\)?
-  \item Considerar otras distribuciones a priori (especialmente en regresión logística).
-  \end{itemize}
-\end{comment}
 
@@
 In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
-Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some more subtle drawbacks that emerge as a result of working in infinite dimensions.
+Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions.
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
@@
 
-It turns out that in both circumstances a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
+It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
 In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2021functional} for the case of logistic regression.
-Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined in the works of \citet{berrendero2018use} and \citet{torrecilla2020optimal}.
+These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2021functional} for the case of functional logistic regression.
+Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
 
@@
 Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-A similar Bayesian approach has been recently explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
-\incomment{Comentar nuestros resultados experimentales}
+A similar Bayesian scheme has been recently explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
+\incomment{Comentar nuestros resultados experimentales} % Overall good results
 
@@
 
-\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
-
-The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of \(X\), and in what follows suppose that it is continuous.
+\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
+Let us denote by \todo{Creo que olvidaste decir que la media es 0}\(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of \(X\), and in what follows suppose that it is continuous.
 To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
@@
 Then, we define \(\Hcal(K)\) to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
-As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\), thus giving it its name.
+As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\), \sout{thus giving it its name}.
 An important detail to note is that, since \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, the values of these functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
 
-As we can see, the covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
+Note that he covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
 An integral operator closely related to this kernel is the so-called covariance operator, namely \(\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds\) for \(f \in L^2[0, 1]\), which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}.
 It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since it is easy to show that we can also express \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
 In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
@@
 Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} one can establish a congruence between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
+Specifically, via \textit{Loève's isometry} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
 This isometry is essentially the completion of the correspondence
-  \[
+  \begin{equation}\label{eq:loeves-isometry}
   \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
-\]
-and can be formally defined as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
+  \end{equation}
+and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
 Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
@@
 The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\).
+Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\).
 
@@
 \begin{equation}\label{eq:bayes-theorem}
-  \pi(\theta \mid \mathcal D_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right]\pi(\theta).
+  \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
 \end{equation}
@@
 Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
+Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.
 
@@
 \end{equation}
-where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term satisfying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), \todo{$\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j,\cdot)$, } and \(\epsilon\) is a normally-distributed error term independent from \(X\), satisfying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
 When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
@@
 As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, as far as generalizations go the proposed model is easily extensible, seeing as how several functional covariates could be added to obtain an extended model such as \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
+In addition, \sout {as far as generalizations go the proposed model is easily extensible, seeing as how several functional covariates could be added to obtain an extended model such as} \color{violet} our model could be easily extended to the case of several covariables via an expression of type \color{black} \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
 In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
+
+\subsubsection{Prior selection and posterior derivation}
 
@@
 On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-All in all, the proposed prior distribution becomes \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
+All in all, the proposed prior distribution becomes\color{violet}{, with a slight abuse of notation, } \color{black} \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
 
@@
 
+\subsubsection{Making predictions}
+
 In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
@@
 \begin{description}
-  \item[Summarize-then-predict] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
+  \item[Summarize-then-predict.] \todo{\tiny He añadido un punto}If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
   \[
@@
 
-  \item[Predict-then-summarize] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
+  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the ``sampled responses'' \todo{¿Mejor "simulated responses''?}at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
   \begin{equation}\label{eq:sampled-response-vector}
@@
   \end{equation}
-  Then, we can take the mean of all such samples as a proxy for each response variable, that is,
+  Then, we can take the mean of all such samples \todo{¿No sería mejor "simulated responses", en lugar de "samples" (recuerda que sample es el conjunto de los datos)} as a proxy for each response variable, that is,
   \[
@@
 
-  \item[Variable selection] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
+  \item[Variable selection.] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
 Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
-After this variable selection has been carried out, we can treat the problem as a finite-dimensional linear regression task and apply any of the well-known prediction algorithms suited for this situation.
+After this variable selection has been carried out, we can treat the problem \color{violet} using a \color{black}  finite-dimensional linear regression \color{violet} model\color{black} and apply any of the well-known prediction algorithms suited for this situation.
 \end{description}
 
-Note that these predictors are obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
+Note that these predictors \color{violet} can be \color{black}  obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
 As a consequence, what we have in practice is a single algorithm that can produce multiple predictors at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
 Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
+
+\todo[inline]{Me ha gustado esta sección}
 
@@
 In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
-Then, a RKHS approach is given by the following functional logistic equation:
+\color{violet} Then, following the approach  suggested by \citet{berrendero2021functional}, an RKHS model might be given  by the  equation \color{black}
 \begin{equation}\label{eq:rkhs-model-logistic}
@@
 \end{equation}
-Following our previous strategy, after incorporating the sample information we can rewrite the classification model~\eqref{eq:rkhs-model-logistic} as
+\color{violet} Indeed, note that this can be seen as a finite-dimensional approximation (but, still, with a functional interpretation) to the general RKHS functional logistic model proposed by these authors, which can be obtained by replacing $\Hcal_{0,p}(K)$ with the whole RKHS space $\Hcal(K)$.
+Now, if we aim at a classification problem, our strategy will be similar to that followed the general linear model: after incorporating the sample information we can rewrite the model ~\eqref{eq:rkhs-model-logistic} as\color{black}
 \begin{equation}\label{eq:rkhs-model-logistic-2}
@@
 
-In much the same way as the linear regression model described before, this RKHS-based logistic regression model partially solves some of the issues with the \(L^2\)-model~\eqref{eq:l2-logistic-model} stated in Section~\ref{sec:intro}.
-Moreover, it negates the problem of non-existence of the MLE in the functional case, since our proposed model is finite-dimensional\footnote{Even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}.
-However, this problem can be easily circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.}.
-Additionally, under specific conditions this model also entails the logistic \(L^2\)-model, and in fact it does hold under mild assumptions whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes \citep[see Theorem 1 in][]{berrendero2021functional}.
+Much the same way as the linear regression model described above, \color{violet} this RKHS-based logistic regression model offers some advantages.
+First and foremost, it can be handled allows for a workable Bayesian approach, as we will show below.
+Second, as pointed out in \cite[see Theorem 1 in][]{berrendero2021functional}, under miild conditions the RKHS logistic functional model holds  whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes; this provides a sound theoretical motivation for such model.
+Third, a maximum likelihood approach for parameter estimation (not considered here) is possible as well.
+Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case; however, let us recall that, even in finite-dimensional settings, there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence} though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
+\color{black}
 
-As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression approach.
+\subsubsection{\color{violet} The Bayesian approach: prior selection and posterior derivation}
+
+As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression \color{violet} model\color{black}.
 However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
@@
 
+\subsubsection{Making predictions}
+
 Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to \(\{0, 1\}\).
-We will follow the standard convention of considering a threshold \(\gamma\) (usually \(\gamma = 1/2\)) and predicting \(\hat Y=1\) if and only if \(\mathbb P(Y=1|X) \geq \gamma\).
-With this thresholding strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
+\color{violet} According to the usual criterion of minimizing the missclassification probability, the optimal rule is predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
+Other criteria would lead to general rules of type predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold $\gamma$.
+\color{black}
+\todo[inline]{\footnotesize La frase original era un poco chocante.
+Si no me equivoco no se trata de una "convención" el valor genérico $\gamma$ puede tenr sentido la regla Bayes (con el criterio usual de minimizar la probsbilidad de clasificación errónea) se obtiene siempre com $\gamma=1/2$} With this thresholding strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
 \[
@@
 the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
-There are primarily two possible outcomes:
+\color{violet} Hence, there are primarily two possible outcomes:\color{black}
 \begin{description}
@@
 
-One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
+In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
+
+\paragraph{The choice of \(\bm p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
 While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications.
@@
 In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
-Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
+Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
 In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
 BIC, DIC, cross-validation, \ldots) to select its optimal value.
-As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
+Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}).
+
+\paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
 \begin{itemize}
@@
 Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation).
-Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
+Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
   \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
@@
 
-On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+\paragraph{Summarizing the posterior.} On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
 In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability.
-At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
+On the other hand, the mode is also a good candidate because it represents the point of highest \color{violet}probability density.
+\color{black} At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
 
-To conclude this section, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
+\paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
@@
 In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\bm Y^{(m)*}\}\) that fall below the real value of the statistic.
-They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman1995bayesian} for details.
+They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.
 
@@
 También estudiar si se recupera el número de componentes subyacente.
+    \item En una simulación concreta, mostrar la dependencia del error frente a distintos parámetros del sampler MCMC (por ejemplo, el número de cadenas usadas o el número de iteraciones).
 \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Supplementary Material, if any, should   %%
-%% be provided in {supplement} environment  %%
-%% with title and short description.        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{supplement}
@@
   \begin{itemize}
-  \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado.
+  \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado. Comentar también la posibilidad de transformar \(\tau\) a un espacio no acotado.
   \item Comentar algunas decisiones de implementación: burn-in, thinning a la hora de hacer ppcs, ...
-  \item Gráficos de las trazas, de algunas posterioris concretas, etc.
+  \item Gráficos de las trazas, de algunas posterioris concretas, etc. También de intervalos de confianza hdi3 y hdi97.
 \end{itemize}
@@
 \begin{acks}[Acknowledgments]
-This is an acknowledgements section.
+The authors wish to acknowledge the computational resources provided by the Centro de Computación Científica-Universidad Autónoma de Madrid (CCC-UAM), which were instrumental in performing the simulations in this paper.
 \end{acks}

### 2022-06-06 slides/seio.tex
slides

%
% Copyright (c) 2021 Antonio Coín Castro
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%

\RequirePackage{fix-cm}
\documentclass[10pt, spanish, professionalfonts]{beamer}

\usepackage{pifont}  % xmark, cmark
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

% OPCIONES DE BEAMER

\definecolor{Maroon}{cmyk}{0, 0.87, 0.88, 0.1}
\definecolor{teal}{rgb}{0.0, 0.45, 0.45}

\usetheme[block=fill, subsectionpage=progressbar, titleformat section=smallcaps]{metropolis}
\setbeamertemplate{frametitle continuation}[roman]
\setbeamertemplate{section in toc}[sections numbered]
%\setbeamertemplate{subsection in toc}[subsections unnumbered]
%\setsansfont[BviejoFont={Fira Sans SemiBold}]{Fira Sans Book}  % Increase font weigth
\widowpenalties 1 10000
\raggedbottom

% COLORES
\setbeamercolor{palette primary}{bg=teal}
\setbeamercolor{progress bar}{use=Maroon, fg=Maroon}

% PAQUETES

\usepackage[utf8]{inputenc}
\usepackage[absolute,overlay]{textpos}
\usepackage[spanish, es-nodecimaldot]{babel}
\usepackage{microtype}
\usepackage{epigraph}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{amssymb, amsmath, amsthm, amsfonts, amscd}
\usepackage{listings}

% FONTS
%\usefonttheme{professionalfonts}
%\usepackage{mathpazo}
%\usepackage{eulervm}

\definecolor{backg}{HTML}{F2F2F2} % Fondo
\definecolor{comments}{HTML}{a8a8a8} % Comentarios
\definecolor{keywords}{HTML}{08388c} % Palabras clave
\definecolor{strings}{HTML}{0489B1}  % Strings

\lstset{
language=scala,
basicstyle=\footnotesize\ttfamily,
breaklines=true,
keywordstyle=\color{keywords},
commentstyle=\color{comments},
stringstyle=\color{strings},
xleftmargin=.5cm,
tabsize=2,
% Acentos, ñ, ¿, ¡ (tex.stackexchange.com/questions/24528)
extendedchars=true
}

% COMANDOS PERSONALIZADOS

\renewcommand{\baselinestretch}{1}
\definecolor{mLightBrown}{HTML}{f97e0b}
\newcommand\maroon[1]{\color{mLightBrown}#1\color{black}}
\let\lmin\wedge
\let\lmax\vee
\newtheorem{prop}{Proposición}
\newtheorem{teorema}{Teorema}
\newtheorem{defi}{Definición}
\newcommand\ddfrac[2]{\frac{\displaystyle #1}{\displaystyle #2}}  % Fracción grande

\renewcommand{\epsilon}{\varepsilon}

\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Hcal}{\ensuremath\mathcal{H}}

%% Scalar product
\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}

% PLOTS

%\usepackage{pgfplots}
%\DeclareUnicodeCharacter{2212}{−}
%\usepgfplotslibrary{groupplots,dateplot}
%\usetikzlibrary{patterns,shapes.arrows}
%\pgfplotsset{compat=newest}

% TÍTULO

\title{Métodos Bayesianos en modelos RKHS para regresión funcional}
\providecommand{\subtitle}[1]{}
\subtitle{Marco teórico y experimentación en Python}
\date{Congreso SEIO Granada\\ 8 de mayo de 2022\\}
\author{José R. Berrendero \\ Antonio Coín  \\ Antonio Cuevas\\}
\institute{Universidad Autónoma de Madrid \\ \textit{Departamento de Matemáticas}}

\titlegraphic{
  \begin{textblock*}{2cm}(9.8cm, 6.7cm)
    \includegraphics[width=2cm]{img/logo-uam}
  \end{textblock*}
}
% DOCUMENTO

\begin{document}
\maketitle

\begin{frame}{Índice de contenidos}
  \tableofcontents
\end{frame}

\section{TODO}

\begin{frame}{Problemas a resolver}
  \(\bullet\) Regresión \maroon{lineal} con datos funcionales

\begin{figure}
    \includegraphics[width=.65\textwidth]{img/data_lin}
  \end{figure}

\(\bullet\) Regresión \maroon{logística} con datos funcionales

\begin{figure}
    \includegraphics[width=.65\textwidth]{img/data_log}
  \end{figure}
\end{frame}

\begin{frame}{Modelos usuales}
  \(\bullet\) \(Y\) variable real (continua o binaria).

  \(\bullet\) \(X(t)\) proceso estocástico de segundo orden con trayectorias en \(L^2[0, 1]\).

  \vspace{1em}

\begin{block}{Modelo lineal \(\bm{L^2}\)}
  \[
    Y = \alpha_0 + \maroon{\langle X, \beta \rangle_2} + \varepsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \varepsilon.
  \]
\end{block}
\begin{block}{Modelo logístico \(\bm{L^2}\)}
    \[
    \mathbb P (Y=1\mid X) = \frac{1}{\exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{\exp\{-\alpha_0 - \int_0^1 X(t)\beta(t)\, dt\}}.
  \]
\end{block}
  \(\alpha_0\in\mathbb R\), \(\beta \in L^2[0, 1]\), \(\varepsilon \sim \mathcal N(0, \sigma^2), \mathbb E[\varepsilon]=0, X \bot \varepsilon\).
\end{frame}

\begin{frame}{Algunas desventajas de los modelos \(\bm{L^2}\)}
  \begin{itemize}
    \item \(L^2[0, 1]\) es un espacio extenso, que contiene además funciones muy irregulares.
    \item La teoría usual de mínimos cuadrados no es aplicable directamente. Se requieren técnicas de regularización o reducción de dimensionalidad.
    \item No incluyen como casos particulares modelos sencillos basados en combinaciones lineales de las marginales del proceso, e.g.:
    \[
      Y = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \varepsilon.
    \]
    \item En el caso logístico, no existe el estimador de máxima verosimilitud con probabilidad 1 bajo condiciones muy generales.
  \end{itemize}
\end{frame}

\begin{frame}{Modelo alternativo: RKHS}

Haremos uso de la teoría de los \textit{espacios de Hilbert con núcleo reproductor} (RKHS's, por sus siglas en inglés).

\begin{alertblock}{Propuesta}
  Cambiar el hábitat del parámetro functional \(\beta(\cdot)\).
\end{alertblock}

  En lugar de suponer \(\beta \in L^2[0, 1]\), consideramos \maroon{\(\beta \in \mathcal H(K)\)}, el RKHS asociado a la función de covarianza \(K(t, s)=\mathbb E[X(t)X(s)]\) del proceso centrado \(X\). Además, sustituiremos la evaluación de \(\dotprod{X}{\beta}_2\) por \maroon{\(\dotprod{X}{\beta}_K\)}.
\end{frame}

\begin{frame}{Un repaso muy breve de los RKHS's}
  \begin{definition}
    Suponemos que la función de convarianzas \(K\) (el \textit{kernel}) es continua, y definimos el espacio
    \[
    \Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot),  p \in \N,  a_i \in \R,  t_i \in [0, 1] \right\},
    \]
    al que dotamos del producto escalar \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), dadas \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) y \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). Entonces, \(\Hcal(K)\) se obtiene al completar \(\Hcal_0(K)\) bajo este producto escalar.
  \end{definition}

  \(\Hcal(K)\subset L^2[0, 1]\), pero las funciones en \(\Hcal(K)\) son más simples y manejables en general. Además, es un espacio de \maroon{funciones}, y no de clases de equivalencia.
\end{frame}

\begin{frame}{Un repaso muy breve de los RKHS's II}
Se pueden dar otras definiciones equivalentes de \(\Hcal(K)\), basadas en el \textit{operador de covarianzas}
\[
\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
\]

\begin{definition}
\begin{enumerate}
  \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), con \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
  \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), con \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\). En este caso, \(\lambda_j\) y \(\phi_j\) son los autovalores y autofunciones de \(\mathcal K\).
\end{enumerate}
\end{definition}

La segunda de estas definiciones enfatiza el hecho de que las funciones de \(\Hcal(K)\) son "simples", en el sentido de que sus componentes en una base ortonormal tienden a \(0\) rápidamente.

\end{frame}

\begin{frame}{Una pequeña sutileza}

  En general, las realizaciones \(x=X(\omega)\) \maroon{no pertenecen a \(H(K)\)} \ con probabilidad 1, por lo que la expresión \(\dotprod{x}{\beta}_K\) carece de sentido. Sin embargo, aprovechamos la relación entre \(\Hcal(K)\) y \(\mathcal L(X)\), el espacio de Hilbert generado por el proceso \(X\) en \(L^2(\Omega)\).
  \begin{block}{Isometría de Loève}
    Es la completación de la correspondencia
    \[
\sum_{j=1}^p a_j X(t_j) \longleftrightarrow \sum_{j=1}^p a_j K(t_j, \cdot),
    \]
    Formalmente, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) para \(U \in \mathcal L(X)\).
  \end{block}
  Utilizaremos la notación \maroon{\(\dotprod{x}{\beta}_k \equiv \Psi_x(\beta)\)}.
\end{frame}

\begin{frame}{Un nuevo parámetro funcional...}
  \begin{itemize}
    \item La idea inicial era suponer \(\beta \in \Hcal(K)\).
    \item Para simplificar el modelo, haremos la suposición extra de que \(\beta \in \Hcal_0(K)\), que es denso en \(\Hcal(K)\) por construcción. Es decir,
    \[
      \beta(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot).
    \]
    \item A la hora de evaluar \(\dotprod{x}{\beta}_K\) para \(x=X(\omega)\), utilizaremos la isometría de Loève:
    \[
      \dotprod{x}{\beta}_K = \Psi_x(\beta) = \sum_{j=1}^p \beta_j X(t_j).
    \]
  \end{itemize}

\end{frame}

\begin{frame}{...y dos nuevos modelos}
  Dado un conjunto de datos \(\{(X_i, Y_i): i=1,\dots, n\}\) con muestras i.i.d. de \((X, Y)\), consideramos:

  \vspace{1em}

  \begin{block}{Modelo lineal RKHS}
  \[
    Y_i = \alpha_0 + \maroon{\langle X_i, \beta \rangle_K} + \varepsilon = \alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j) + \varepsilon.
  \]
\end{block}
\begin{block}{Modelo logístico RKHS}
    \[
    \mathbb P (Y_i=1\mid X_i) = \frac{1}{\exp\{-\alpha_0 - \maroon{\langle X_i, \beta\rangle_K}\}} = \frac{1}{\exp\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\}}.
  \]
\end{block}
\end{frame}

\begin{frame}{Estimación de parámetros: un enfoque bayesiano}
  teorema de bayes + aproximación posteriori
  se extiende por densidad
  vector de parámetros theta
\end{frame}

\begin{frame}{Distribuciones a priori}
  Se propone utilizar las mismas distribuciones a priori tanto para el caso de regresión lineal como para el de regresión logística:
  \begin{align*}
    p &\sim \text{Categórica}(p_{\text{max}}),\\
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  \beta\mid \tau, \sigma^2 & \sim \mathcal N\big(b_0, g\sigma^2{\underbrace{\left[\mathcal X_\tau' \mathcal X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
\end{align*}
donde \(I\) es la matriz identidad, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), y \(b_0\in \R^p, \ g \in \R\) y \(\eta \in \R^+\) son hiperparámetros.

\metroset{block=transparent}
\begin{alertblock}{Selección de hiperparámetros}
  \vspace{0.1em}
  Para \(b_0\) se usa una aproximación computacional del MLE; el resto de hiperparámetros se fija por \textit{cross-validation}.
\end{alertblock}
\metroset{block=fill}
\end{frame}

\begin{frame}{Distribuciones a posteriori}
  \begin{align*}
  \log \pi(\theta\mid \boldsymbol{Y}) &\propto \frac{1}{2}\log |G_\tau| - (p+n)\log \sigma\\
  &-\frac{1}{2\sigma^2} \left(\|\boldsymbol{Y}-\alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right).
\end{align*}
\end{frame}

\begin{frame}{Predicciones}
  \textbf{Estimación puntual}: Se resume la distribución a posteriori de los parámetros \(\theta\mid Y\) mediante un estimador puntual (media, mediana, moda), y se utilizan para predecir según el modelo:
  \[
  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j x_i(\hat \tau_j), \quad i=1,\dots, n.
  \]

  \textbf{Estimación distribucional}: Se utiliza la media de \textit{todas}  las muestras generadas de \(Y^*\) como predicción:
  \[
    \hat Y = \frac{1}{M}\sum_{m=1}^M Y^{(j)*}.
  \]

  En ambos casos se puede considerar solo una de cada \(k\) muestras.
\end{frame}

\begin{frame}{Predicciones II}
    Podemos hacer también una selección de \(p\) variables usando los estimadores puntuales de \(\tau \mid Y\), y después aplicar cualquier algoritmo de regresión lineal.

    \begin{figure}
      \includegraphics[width=\textwidth]{img/tau_posterior_lin}
      \caption{Distribución a posteriori estimada para los puntos de impacto.}
    \end{figure}

    En cualquier caso, medimos el error usando el MSE como métrica.
\end{frame}

\begin{frame}{Model checking}
  \begin{itemize}
    \item Análisis de la traza de las cadenas y de la distribución a posteriori de los parámetros. Se obtienen \textbf{intervalos creíbles} para los parámetros.
    \item En cada paso obtenemos una estimación \(\tilde \theta_m\), y podemos generar \(Y^{(m)*} \mid \theta_m, X\) siguiendo el modelo asumido.
    \item \textit{Bayesian p-values}: \(p=P(T(Y^*)\leq T(Y)\mid Y)\) para ciertas elecciones de \(T\): mínimo, máximo, mediana, media. Se calcula contando la proporción de muestras generadas que cumplen la desigualdad, y se espera que esté en torno a \(0.5\).
  \end{itemize}

\end{frame}

\begin{frame}{Model checking II}
  \begin{figure}
    \includegraphics[width=\textwidth]{img/ppc_lin}
    \caption{\textit{Posterior predictive checks} bajo un modelo RKHS.}
  \end{figure}
\end{frame}

\subsection{Experimentos}

\begin{frame}{Datos sintéticos}
  Se consideran 150 ejemplos de \(X \sim GP(0, K(s, t))\) y tres posibles variantes de \(K\): movimiento browniano fraccional (\(H=0.8\)), Ornstein-Uhlenbeck y kernel RBF.

  Se genera la respuesta \(Y\) acorde a dos modelos, RKHS y \(L^2\). Concretamente, se elige
  \[
    Y_i \sim \mathcal N\big(5 -5X_i(0.1) + 10X_i(0.8), \ 0.5\big)
  \]
  ó
  \[
    Y_i \sim \mathcal N\left(5 + \int_0^1 \log(1+4t)X_i(t)\, dt, \ 0.5\right).
  \]

Consideramos una malla regular de \(N=100\) puntos en \([0, 1]\), y un reparto de 100 ejemplos para entrenamiento y 50 para evaluación.
\end{frame}

\begin{frame}{Datos reales}
  Se consideran dos conjuntos de datos reales.
  \begin{itemize}
    \item \textbf{Tecator:} Contiene 215 ejemplos de mediciones de \textit{absorbancia} en muestras de carne para intentar predecir su contenido en grasa.
    \item \textbf{Aemet}: Contiene 73 ejemplos de curvas de temperatura, a partir de las cuales se intenta predecir la precipitación total.
  \end{itemize}

  En ambos casos hacemos una división \(80\%-20\%\) para entrenamiento y \textit{test}.
\end{frame}

\begin{frame}{Preprocesado de los datos}
  \begin{itemize}
    \item Se centran los regresores para que tengan media \(0\). Opcionalmente, se pueden estandarizar en cada punto de la malla.
    \item Se permite \textbf{sustituir los datos \(\boldsymbol{X_i}\) por su desarrollo en una base} de Fourier, con un número determinado de coeficientes. De esta forma realizamos un suavizado de las curvas.
  \end{itemize}
\end{frame}

\begin{frame}{Algoritmos de comparación}
  \textbf{Regresión lineal multivariante:} Regresión Lineal estándar, Lasso (\(L^1\)), Ridge (\(L^2\)).

  \textbf{Regresión no lineal multivariante}: SVM con kernel RBF.

  \textbf{Regresión lineal funcional}: Modelo \(L^2\), KNN Funcional.

  \textbf{Reducción de dimensión:} PCA Funcional (proyección sobre coeficientes).

  \textbf{Selección de variables:} Aleatoria.

  \vspace{1em}

  Se entrenan todos ellos sobre el conjunto de entrenamiento, haciendo \textit{5-fold cross validation} para escoger los mejores hiperparámetros en cada caso (valor de regularización, número de componentes, ...).
\end{frame}

\begin{frame}{Metodología experimental}
  \begin{itemize}
  \item Para cada conjunto de datos consideramos tres posibles valores de \(p\) y cuatro posibles valores de \(\eta\):
  \begin{itemize}
    \item[--] \(p\in \{2,3,4\}\) para los datos sintéticos y \(p\in \{3, 4, 5\}\) para los datos reales.
      \item[--] \(\eta \in \{0.01, 0.1, 1.0, 10.0\}\).
  \end{itemize}

  \item Para aumentar la estabilidad hacemos 5 repeticiones de cada modelo. Es decir, entrenamos un total de 60 modelos sobre el conjunto de entrenamiento

  \item Con el \textbf{mejor modelo obtenido} se realiza también una selección de variables basada en los valores de \(\tau\) estimados mediante los distintos estimadores puntuales.
\end{itemize}

Además, repetimos este proceso con los datos suavizados en una base de Fourier con 11 elementos (también con los algoritmos de comparación).
\end{frame}

\begin{frame}{Metodología experimental II}
  Fijamos algunos hiperparámetros:
  \begin{itemize}
    \item Número de cadenas: 64.
    \item Número de pasos: 100 + 1000.
    \item Movimientos: elección aleatoria (ponderada) entre dos de los recomendados, \textit{stretch} y \textit{walk}.
    \item Inicialización: Entorno aleatorio del MLE + muestras de las distribuciones a priori.
    \item Prior de \(\beta\): \(b_0\) se elige como el MLE de \(\beta\), y fijamos \(g=5\) (recomendado en \textit{Grollemund et al. (2019)}).
  \end{itemize}
  \vspace{1em}

  \textit{Nota:} El valor de \(g\) no parece afectar al resultado, por lo que podríamos eliminar este parámetro.
\end{frame}

\begin{frame}{Resultados RKHS}
\textbf{Sin base}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \cmark & \cmark\\
      O-U & \cmark & \cmark\\
      RBF & \cmark & \cmark
    \end{tabular}
  \end{table}

  \textbf{Base Fourier(11)}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \cmark &  \cmark\\
      O-U & \cmark & \xmark\\
      RBF & \cmark & \cmark
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}{Resultados RKHS II}
  \begin{figure}
    \includegraphics[width=\textwidth]{img/res/reg_emcee_rkhs_fixed}
    \caption{caption.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados RKHS III}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/lin_rkhs_fbm_base11}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/lin_rkhs_ou_base11}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/lin_rkhs_rbf_base11}
    \caption{MSE de los 5 mejores modelos en cada caso con el modelo subyacente RKHS y con base de Fourier.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados \(\boldsymbol{L^2}\)}
\textbf{Sin base}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \xmark & \xmark\\
      O-U & \xmark & \xmark\\
      RBF & \xmark & \xmark
    \end{tabular}
  \end{table}
\textbf{Base Fourier(11)}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \cmark &  \cmark\\
      O-U & \cmark & \cmark\\
      RBF & \xmark & \xmark
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}{Resultados \(\boldsymbol{L^2}\) II}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_fbm_nobase}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_ou_nobase}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_rbf_nobase}
    \caption{MSE de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y sin usar suavizado con bases.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados \(\boldsymbol{L^2}\) III}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_fbm_base11}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_ou_base11}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_rbf_base11}
    \caption{MSE de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y con base de Fourier.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados datos reales}
\textbf{Sin base}
  \begin{table}
    \begin{tabular}{c|cc}
      Dataset & Victoria & Victoria sel. variables \\ \hline
      Tecator & \xmark & \xmark\\
      Aemet & \xmark & \cmark\\
    \end{tabular}
  \end{table}

  \textbf{Base Fourier(11)}
  \begin{table}
    \begin{tabular}{c|cc}
      Dataset & Victoria & Victoria sel. variables \\ \hline
      Tecator & \xmark &  \xmark\\
      Aemet & \xmark & \cmark\\
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}{Resultados datos reales II}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/lin_tecator_nobase}\hspace{5em}
    \includegraphics[width=0.3\textwidth]{img/results/lin_aemet_nobase}
    \caption{MSE de los 5 mejores modelos en cada caso con datos reales y sin usar suavizado con bases.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados datos reales III}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/lin_tecator_base11}\hspace{5em}
    \includegraphics[width=0.3\textwidth]{img/results/lin_aemet_base11}
    \caption{MSE de los 5 mejores modelos en cada caso con datos reales y con base de Fourier.}
  \end{figure}
\end{frame}

\begin{frame}{Algunas observaciones}
\begin{itemize}
  \item Por cada modelo entrenado obtenemos 4 estimadores y 3 estrategias de selección de variables.
  \item En general, se obtienen mejores resultados con el uso de la base.
  \item El tiempo medio de entrenamiento de cada modelo es de unos 20-30 segundos.
  \item En ocasiones, nuestro algoritmo supera con holgura a todos los algoritmos de comparación. Cuando pierde, suele ser frente a uno o dos de ellos (y por poco), pero sigue superando al resto.
  \item Se puede considerar como algoritmo de comparación el estimador de máxima verosimilitud de los parámetros. En casi todas las pruebas experimentales, el desempeño de nuestro algoritmo supera al del MLE.
\end{itemize}

\end{frame}

\begin{frame}{Dificultades}
  \begin{itemize}

  \item Hay multitud de grados de libertad: estandarización de regresores y/o variable respuesta, elección o no de una base (¿cuál?,) algoritmo de estimación del MLE, número, longitud y movimientos de las cadenas MCMC, elección de distribuciones a priori, elección de \(g, b_0\) y \(\eta\), etc.
  \item El algoritmo MCMC es costoso, y en conjuntos de datos grandes las estrategias de \textit{cross-validation} para seleccionar parámetros pueden no ser viables.
  \item Es necesario un procedimiento para seleccionar el valor de \(p\) (BIC, WAIC, WBIC, \textit{cross-validation}, ...).
  \item Debido a la aleatoriedad del algoritmo, los resultados pueden variar sustancialmente de una ejecución a otra.
  \item La distribución a priori para \(\beta\) depende en gran medida del valor de \(b_0\) escogido. Si su estimación inicial no es buena, el algoritmo no funciona demasiado bien.
\end{itemize}
\end{frame}

\begin{frame}{Dificultades II}
  \begin{itemize}
    \item Hay un problema de \textit{identificabilidad} de los coeficientes (son intercambiables), que se agudiza especialmente al usar varias cadenas.
    \item En el caso en el que el modelo subyacente es RKHS, es posible que no se recuperen los verdaderos valores de los parámetros debido a interacciones entre los coeficientes (si \(p\) es mayor que el número real de componentes).
    \item El uso de distribuciones a priori impropias implica comprobar que \(\int_{\Theta} \pi(Y\mid \theta)\pi(\theta)\, d\theta < \infty\).
  \end{itemize}
\end{frame}

\begin{frame}{Alternativas}
  \begin{itemize}
    \item Explorar el concepto de \textit{online learning} aplicado a esta situación, por ejemplo usando como priori de nuevos datos la posteriori aprendida. Estudiar la \textbf{consistencia} de la distribución a posteriori.
    \item Utilizar otras herramientas de suavizado en lugar de bases de Fourier.
    \item Sustitutir la distribución a priori de \(\beta\) por otra que requiera menos hiperparámetros.
    \item Sustituir las distribuciones impropias por otras propias.
  \end{itemize}

\end{frame}

\section{Regresión Logística Funcional Bayesiana}

\subsection{Marco teórico}

\begin{frame}{Planteamiento Bayesiano}
\textbf{Modelo}: Cada \(Y_i\) se puede ver como una variable aleatoria de Bernoulli \(\mathcal B(p(x_i))\), con
\[
p_i \equiv p(x_i)=\mathbb P(Y_i=1\mid X_i=x_i) = \frac{1}{1 + \exp\left\{-\alpha_0-\displaystyle\sum_{j=1}^p \beta_jx_i(\tau_j)\right\}}.
\]

\textbf{Distribuciones a priori}: Igual que en regresión lineal.

\textbf{Log-posterior}:
\begin{align*}
  \log \ &\pi(\beta, \tau, \alpha_0, \log\sigma\mid Y) \propto\\
   &\sum_{i=1}^n \left[ \left(\alpha_0 + \Psi^{-1}_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}^{-1}(\alpha)\right\}\right)\right]\\
   &+
\frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0).
\end{align*}
\end{frame}

\begin{frame}{Model checking}
  \begin{figure}
    \includegraphics[width=\textwidth]{img/ppc_log}
    \caption{\textit{Posterior predictive checks} bajo un modelo RKHS.}
  \end{figure}
\end{frame}

\begin{frame}{Predicciones}

\begin{itemize}
  \item Fijamos un umbral en \(0.5\) para establecer la pertenencia a las clases.
   \item Los estimadores puntuales y la selección de variables son análogos al caso de regresión lineal.
   \item Tenemos ahora dos estimadores basados en la distribución a posteriori (similares a los \textit{ensembles} de clasificadores):
   \begin{itemize}
     \item[--] Basado en el voto mayoritario de las muestras \(Y^*\) generadas.
     \item[--] Basado en la media de las probabilidades \(p_i^*\) generadas (con aplicación posterior del umbral).
   \end{itemize}
  \item Se usa el \textit{accuracy} (precisión) para evaluar los modelos.
\end{itemize}
\end{frame}

\subsection{Experimentos}

\begin{frame}{Datos sintéticos}
Seguimos la misma estrategia de generación de datos antes. Se consideran las mismas tres funciones de covarianza y se genera la respuesta tanto con un modelo RKHS como L2, i.e.:
  \[
    Y_i \sim \mathcal B\left(\frac{1}{1 + \exp\left\{0.5 + 5X_i(0.1) - 10X_i(0.8)\right\}}\right)
  \]
  ó
  \[
    Y_i \sim \mathcal B\left(\frac{1}{1 + \exp\left\{0.5 -\int_0^1 \log(1+4t)X_i(t)\, dt\right\}}\right).
  \]

  \vspace{1em}

  Se introduce un pequeño ruido aleatorio en las etiquetas, y además se intenta que ambas clases estén balanceadas.

\end{frame}

\begin{frame}{Datos reales}
  Se consideran dos conjuntos de datos reales.
  \begin{itemize}
    \item \textbf{Medflies:} Contiene 534 ejemplos de mediciones del número de huevos diario puesto por una serie de moscas, para intentar predecir si viven mucho o poco.
    \item \textbf{Growth}: Contiene 93 ejemplos de curvas de altura en niños y niñas.
  \end{itemize}

  En ambos casos hacemos una división \(80\%-20\%\) para entrenamiento y \textit{test}.
\end{frame}

\begin{frame}{Algoritmos de comparación}
  \textbf{Clasificación lineal multivariante:} Regresión Logística, SVM lineal.

  \textbf{Clasificación no lineal multivariante}: SVM con kernel RBF.

  \textbf{Clasificación funcional}: \textit{Maximum Depth}, \textit{Nearest Centroid} Funcional, KNN Funcional.

  \textbf{Reducción de dimensión:} PCA Funcional (proyección sobre coeficientes).

  \textbf{Selección de variables:} Aleatoria, \textit{Recursive Maxima Hunting}, \textit{RKVS} (Mahalanobis).

  \vspace{1em}

  Mismo preprocesado y metodología experimental que en el caso de regresión lineal, utilizando esta vez 9 coeficientes de Fourier.
\end{frame}

\begin{frame}{Resultados RKHS}
\textbf{Sin base}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \xmark & \xmark\\
      O-U & \xmark & \cmark\\
      RBF & \cmark & \xmark
    \end{tabular}
  \end{table}

  \textbf{Base Fourier(9)}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \cmark &  \cmark\\
      O-U & \cmark & \cmark\\
      RBF & \cmark & \cmark
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}{Resultados RKHS II}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_fbm_nobase}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_ou_nobase}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_rbf_nobase}
    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente RKHS y sin usar suavizado con bases. Distinguimos los algoritmos de comparación, nuestro algoritmo Bayesiano (emcee), y nuestra selección de variables Bayesiana.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados RKHS III}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_fbm_base9}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_ou_base9}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_rbf_base9}
    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente RKHS y con base de Fourier.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados \(\boldsymbol{L^2}\)}
\textbf{Sin base}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \xmark & \xmark\\
      O-U & \cmark & \xmark\\
      RBF & \cmark & \cmark
    \end{tabular}
  \end{table}

  \textbf{Base Fourier(9)}
  \begin{table}
    \begin{tabular}{c|cc}
      Kernel & Victoria & Victoria sel. variables \\ \hline
      fBM & \xmark &  \cmark\\
      O-U & \xmark & \cmark\\
      RBF & \cmark & \cmark
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}{Resultados \(L^2\) II}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/log_l2_fbm_nobase}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_l2_ou_nobase}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_l2_rbf_nobase}
    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y sin usar suavizado con bases.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados \(L^2\) III}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/log_l2_fbm_base9}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_l2_ou_base9}\hfill
    \includegraphics[width=0.3\textwidth]{img/results/log_l2_rbf_base9}
    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y con base de Fourier.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados datos reales}
\textbf{Sin base}
  \begin{table}
    \begin{tabular}{c|cc}
      Dataset & Victoria & Victoria sel. variables \\ \hline
      Medflies & \xmark & \xmark\\
      Growth & \cmark & \cmark\\
    \end{tabular}
  \end{table}

  \textbf{Base Fourier(9)}
  \begin{table}
    \begin{tabular}{c|cc}
      Dataset & Victoria & Victoria sel. variables \\ \hline
      Medflies & \xmark &  \xmark\\
      Growth & \cmark & \cmark\\
    \end{tabular}
  \end{table}

  Tanto en las victorias como en las derrotas, la precisión es muy similar. Podemos considerar un empate en ambos conjuntos.
\end{frame}

\begin{frame}{Resultados datos reales II}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/log_medflies_nobase}\hspace{5em}
    \includegraphics[width=0.3\textwidth]{img/results/log_growth_nobase}
    \caption{Precisión de los 5 mejores modelos en cada caso con datos reales y sin usar suavizado con bases.}
  \end{figure}
\end{frame}

\begin{frame}{Resultados datos reales III}
  \begin{figure}
    \includegraphics[width=0.3\textwidth]{img/results/log_medflies_base9}\hspace{5em}
    \includegraphics[width=0.3\textwidth]{img/results/log_growth_base9}
    \caption{Precisión de los 5 mejores modelos en cada caso con datos reales y con base de Fourier.}
  \end{figure}
\end{frame}

\end{document}

### 2022-06-07 slides/seio.tex
Advance slides SEIO 22

@@
 \RequirePackage{fix-cm}
-\documentclass[10pt, spanish, professionalfonts]{beamer}
+\documentclass[10pt, english, professionalfonts]{beamer}
 
@@
 \definecolor{mLightBrown}{HTML}{f97e0b}
-\newcommand\maroon[1]{\color{mLightBrown}#1\color{black}}
+\newcommand\maroon[1]{\color{mLightBrown}#1\color{mDarkTeal}}
 \let\lmin\wedge
@@
 
-\title{Métodos Bayesianos en modelos RKHS para regresión funcional}
+\title{Bayesian methods in RKHS models for functional regression}
 \providecommand{\subtitle}[1]{}
-\subtitle{Marco teórico y experimentación en Python}
-\date{Congreso SEIO Granada\\ 8 de mayo de 2022\\}
+\subtitle{Theoretical framework and experiments}
+\date{SEIO Congress Granada\\ 8th May 2022\\}
 \author{José R. Berrendero \\ Antonio Coín  \\ Antonio Cuevas\\}
@@
 
-\begin{frame}{Índice de contenidos}
+\begin{frame}{Table of contents}
   \tableofcontents
@@
 
-\section{TODO}
+\section{Introduction}
 
-\begin{frame}{Problemas a resolver}
-  \(\bullet\) Regresión \maroon{lineal} con datos funcionales
+\begin{frame}{FDA: linear regression}
+  \textbf{First problem}: \maroon{Linear } regression with functional data.
+  \vspace{1em}
 
 \begin{figure}
-    \includegraphics[width=.65\textwidth]{img/data_lin}
+    \includegraphics[width=\textwidth]{img/data_lin}
+    \caption{Linear regression data set simulated from a fractional brownian motion.}
   \end{figure}
-
-\(\bullet\) Regresión \maroon{logística} con datos funcionales
+  \vspace{-1em}
+\end{frame}
+\begin{frame}{FDA: logistic regression}
+  \textbf{Second problem}: \maroon{Logistic } regression with functional data.
+\vspace{1em}
 
 \begin{figure}
-    \includegraphics[width=.65\textwidth]{img/data_log}
+    \includegraphics[width=\textwidth]{img/data_log}
+    \caption{Logistic regression data set simulated from a fractional brownian motion.}
   \end{figure}
+  \vspace{-1em}
 \end{frame}
 
-\begin{frame}{Modelos usuales}
-  \(\bullet\) \(Y\) variable real (continua o binaria).
+\begin{frame}{Usual models}
+  \(\bullet\) \(Y\) real variable (continuous or dichotomous).
 
-  \(\bullet\) \(X(t)\) proceso estocástico de segundo orden con trayectorias en \(L^2[0, 1]\).
+  \(\bullet\) \(X(t)\) second order stochastic process with trajectories in \(L^2[0, 1]\).
+
+  \(\bullet\) \(\beta \in L^2[0,1]\) \maroon{functional parameter}.
 
@@
 
-\begin{block}{Modelo lineal \(\bm{L^2}\)}
+\begin{block}{Linear \(\bm{L^2}\)-model}
   \[
@@
 \end{block}
-\begin{block}{Modelo logístico \(\bm{L^2}\)}
+\begin{block}{Logistic \(\bm{L^2}\)-model}
     \[
@@
 \end{block}
-  \(\alpha_0\in\mathbb R\), \(\beta \in L^2[0, 1]\), \(\varepsilon \sim \mathcal N(0, \sigma^2), \mathbb E[\varepsilon]=0, X \bot \varepsilon\).
+  \(\alpha_0\in\mathbb R\), \ \(\varepsilon \sim \mathcal N(0, \sigma^2), \ \mathbb E[\varepsilon]=0, \ X \bot \varepsilon\).
 \end{frame}
 
-\begin{frame}{Algunas desventajas de los modelos \(\bm{L^2}\)}
+\begin{frame}{Some shortcomings of the \(\bm{L^2}\)-models}
   \begin{itemize}
-    \item \(L^2[0, 1]\) es un espacio extenso, que contiene además funciones muy irregulares.
-    \item La teoría usual de mínimos cuadrados no es aplicable directamente. Se requieren técnicas de regularización o reducción de dimensionalidad.
-    \item No incluyen como casos particulares modelos sencillos basados en combinaciones lineales de las marginales del proceso, e.g.:
+    \item \(L^2[0, 1]\) is a extremely wide space, which also contains very ill-behaved functions.
+    \item The usual least-squares theory cannot be applied directly; some regularization or dimensionality reduction techniques are required.
+    \item They do not include simple models based on linear combinations of the marginal of the process as particular cases, such as
     \[
@@
     \]
-    \item En el caso logístico, no existe el estimador de máxima verosimilitud con probabilidad 1 bajo condiciones muy generales.
+    \item In the logistic case, the maximum likelihood estimator does not exist with probability one under fairly general conditions.
   \end{itemize}
@@
 
-\begin{frame}{Modelo alternativo: RKHS}
+\section{A RKHS model for functional regression}
 
-Haremos uso de la teoría de los \textit{espacios de Hilbert con núcleo reproductor} (RKHS's, por sus siglas en inglés).
+\begin{frame}{Alternative model: RKHS}
 
-\begin{alertblock}{Propuesta}
-  Cambiar el hábitat del parámetro functional \(\beta(\cdot)\).
+We will exploit the theory of \textit{reproducing kernel Hilbert spaces} (RKHS's).
+
+\vspace{1em}
+\begin{alertblock}{Proposal}
+  Change the habitat of the functional parameter \(\beta(\cdot)\).
 \end{alertblock}
 
-  En lugar de suponer \(\beta \in L^2[0, 1]\), consideramos \maroon{\(\beta \in \mathcal H(K)\)}, el RKHS asociado a la función de covarianza \(K(t, s)=\mathbb E[X(t)X(s)]\) del proceso centrado \(X\).
-Además, sustituiremos la evaluación de \(\dotprod{X}{\beta}_2\) por \maroon{\(\dotprod{X}{\beta}_K\)}.
+\vspace{1em}
+  Instead of assuming \(\beta \in L^2[0, 1]\), we consider \maroon{\(\beta \in \mathcal H(K)\)}, the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\).
+Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\beta}_K\)}.
 \end{frame}
 
-\begin{frame}{Un repaso muy breve de los RKHS's}
+\begin{frame}{RKHS essentials}
   \begin{definition}
-    Suponemos que la función de convarianzas \(K\) (el \textit{kernel}) es continua, y definimos el espacio
+    Let \(K\) be continuous, and define the preliminary space
     \[
@@
     \]
-    al que dotamos del producto escalar \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), dadas \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) y \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Entonces, \(\Hcal(K)\) se obtiene al completar \(\Hcal_0(K)\) bajo este producto escalar.
+    which is endowed with the scalar product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
+Then, \(\Hcal(K)\) is the completion of \(\Hcal_0(K)\) under this scalar product.
   \end{definition}
 
-  \(\Hcal(K)\subset L^2[0, 1]\), pero las funciones en \(\Hcal(K)\) son más simples y manejables en general. Además, es un espacio de \maroon{funciones}, y no de clases de equivalencia.
+  \vspace{1em}
+
+  Note that \(\Hcal(K)\subset L^2[0, 1]\), but the elements of \(\Hcal(K)\) are simpler and more manageable in general.
+Plus, the latter is a space of \maroon{functions}, and not of equivalence classes.
 \end{frame}
 
-\begin{frame}{Un repaso muy breve de los RKHS's II}
+\begin{frame}{Un repaso muy breve de los RKHS's (cont.)}
 Se pueden dar otras definiciones equivalentes de \(\Hcal(K)\), basadas en el \textit{operador de covarianzas}
@@
 
-La segunda de estas definiciones enfatiza el hecho de que las funciones de \(\Hcal(K)\) son "simples", en el sentido de que sus componentes en una base ortonormal tienden a \(0\) rápidamente.
+La segunda de estas definiciones enfatiza el hecho de que las funciones de \(\Hcal(K)\) son ``simples'', en el sentido de que sus componentes en una base ortonormal tienden a \(0\) rápidamente.
 
@@
 
-\begin{frame}{Una pequeña sutileza}
+\begin{frame}{A small subtlety}
 
-  En general, las realizaciones \(x=X(\omega)\) \maroon{no pertenecen a \(H(K)\)} \ con probabilidad 1, por lo que la expresión \(\dotprod{x}{\beta}_K\) carece de sentido.
-Sin embargo, aprovechamos la relación entre \(\Hcal(K)\) y \(\mathcal L(X)\), el espacio de Hilbert generado por el proceso \(X\) en \(L^2(\Omega)\).
-  \begin{block}{Isometría de Loève}
-    Es la completación de la correspondencia
+  Generally speaking, the concrete realizations \(x=X(\omega)\) \maroon{do not belong to \(\Hcal(K)\) with probability \(1\)}, so the expression \(\dotprod{x}{\beta}_K\) lacks meaning.
+However, we can make use of the relationship between \(\Hcal(K)\) and \(\mathcal L(X)\), the linear span of the process \(X\) within \(L^2(\Omega)\).
+
+  \vspace{1em}
+  \begin{block}{Loève's isometry}
+    It is the completion of the correspondence
     \[
-\sum_{j=1}^p a_j X(t_j) \longleftrightarrow \sum_{j=1}^p a_j K(t_j, \cdot),
+\sum_{j=1}^p a_j X(t_j) \longleftrightarrow \sum_{j=1}^p a_j K(t_j, \cdot).
     \]
-    Formalmente, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) para \(U \in \mathcal L(X)\).
+    Formally, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) for \(U \in \mathcal L(X)\).
   \end{block}
-  Utilizaremos la notación \maroon{\(\dotprod{x}{\beta}_k \equiv \Psi_x(\beta)\)}.
+  With a slight abuse of notation, we will write \maroon{\(\dotprod{x}{\beta}_k \equiv \Psi_x(\beta)\)}. This correspondence will allow us to still call our models ``linear''.
 \end{frame}
 
-\begin{frame}{Un nuevo parámetro funcional...}
+\begin{frame}{A new functional parameter...}
   \begin{itemize}
-    \item La idea inicial era suponer \(\beta \in \Hcal(K)\).
-    \item Para simplificar el modelo, haremos la suposición extra de que \(\beta \in \Hcal_0(K)\), que es denso en \(\Hcal(K)\) por construcción. Es decir,
+    \item The initial idea was to suppose \(\beta \in \Hcal(K)\).
+    \item To further simplify the model, we will assume \(\beta \in \Hcal_0(K)\), which is dense in \(\Hcal(K)\) by construction. That is,
     \[
@@
     \]
-    \item A la hora de evaluar \(\dotprod{x}{\beta}_K\) para \(x=X(\omega)\), utilizaremos la isometría de Loève:
+    \item When evaluating \(\dotprod{x}{\beta}_K\) for a trajectory \(x=X(\omega)\), we resort to Loeve's isometry:
     \[
@@
 
-\begin{frame}{...y dos nuevos modelos}
-  Dado un conjunto de datos \(\{(X_i, Y_i): i=1,\dots, n\}\) con muestras i.i.d. de \((X, Y)\), consideramos:
+\begin{frame}{...and two new models}
+  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), we consider:
 
@@
 
-  \begin{block}{Modelo lineal RKHS}
+  \begin{block}{Linear RKHS model}
   \[
@@
 \end{block}
-\begin{block}{Modelo logístico RKHS}
+\begin{block}{Logistic RKHS model}
     \[
@@
 
-\begin{frame}{Estimación de parámetros: un enfoque bayesiano}
-  teorema de bayes + aproximación posteriori
-  se extiende por densidad
-  vector de parámetros theta
+\begin{frame}{Benefits of the RKHS models}
+  \begin{itemize}
+    \item MLE existe..
+    \item El modelo es verdadero en algunos casos...
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Distribuciones a priori}
-  Se propone utilizar las mismas distribuciones a priori tanto para el caso de regresión lineal como para el de regresión logística:
+\section{A bayesian approach}
+
+\begin{frame}{Parameter estimation}
+
+  We have a parameter vector \(\theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)\), on which we impose a prior distribution \(\pi(\theta)\).
+In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.
+
+  \vspace{1em}
+
+  Afterwards we apply \textit{Bayes'}  theorem in the usual way to obtain the posterior distribution of the parameters given the data.
+
+  \begin{block}{Bayes's theorem}
+      \[
+      \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
+      \]
+  \end{block}
+  \vspace{1em}
+
+  Finally, we obtain samples of the approximate posterior distribution using an algorithm of the family of \textit{Markov chain Monte Carlo} (MCMC) methods.
+
+\end{frame}
+
+\begin{frame}{Prior distributions}
+  Same prior scenario for both linear and logistic models.
+  \begin{block}{Prior de los parámetros}
+    \vspace{-1em}
   \begin{align*}
-    p &\sim \text{Categórica}(p_{\text{max}}),\\
+    p &\sim \text{Categorical}(p_{\text{max}}),\\
   \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
@@
 \end{align*}
-donde \(I\) es la matriz identidad, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), y \(b_0\in \R^p, \ g \in \R\) y \(\eta \in \R^+\) son hiperparámetros.
+where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters.
+\end{block}
 
 \metroset{block=transparent}
-\begin{alertblock}{Selección de hiperparámetros}
+
+\begin{alertblock}{Hyperparameter selection}
   \vspace{0.1em}
-  Para \(b_0\) se usa una aproximación computacional del MLE; el resto de hiperparámetros se fija por \textit{cross-validation}.
+  For \(b_0\) we use a computational approximation of the MLE; the rest of hyperparameters are set via \textit{cross-validation}.
 \end{alertblock}
@@
 
-\begin{frame}{Distribuciones a posteriori}
+\begin{frame}{Posterior distributions}
+  \begin{block}{Linear log-posterior}
+    \vspace{-1em}
     \begin{align*}
-  \log \pi(\theta\mid \boldsymbol{Y}) &\propto \frac{1}{2}\log |G_\tau| - (p+n)\log \sigma\\
-  &-\frac{1}{2\sigma^2} \left(\|\boldsymbol{Y}-\alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right).
+  \log \pi(\theta\mid \mathcal D_n) &\propto \frac{1}{2}\log |G_\tau| - (p+n)\log \sigma\\
+  &-\frac{1}{2\sigma^2} \left(\|\boldsymbol{Y}-\alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right),
 \end{align*}
+where \(\bm Y=(Y_1,\dots,Y_n)'\) y \(\bm{1}\) is an \(n\)-dimensional vector of ones.
+  \end{block}
+
+\vspace{1em}
+  \begin{block}{Logistic log-posterior}
+      \vspace{-1em}
+    \begin{align*}
+  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
+  \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
+\end{align*}
+Remember that \(\Psi_{X_i}(\alpha) = \sum_j \beta_j X_i(t_j)\).
+  \end{block}
 \end{frame}
 
-\begin{frame}{Predicciones}
-  \textbf{Estimación puntual}: Se resume la distribución a posteriori de los parámetros \(\theta\mid Y\) mediante un estimador puntual (media, mediana, moda), y se utilizan para predecir según el modelo:
+\begin{frame}{Posterior approximation}
+  We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
+
+  \vspace{1em}
+
+  \begin{figure}
+    \includegraphics[width=.7\textwidth]{img/posterior}
+    \caption{Estimated posterior distribution \(\theta|\mathcal D_n\).}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Predictions}
+  \textbf{Summarize-then-predict}.
+A summary of the posterior distribution \(\theta| \mathcal D_n\) is obtained via a point-estimate statistic (mean, median, mode), and then predictions are made according to the assumed model, e.g.:
   \[
-  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j x_i(\hat \tau_j), \quad i=1,\dots, n.
+  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j X_i(\hat \tau_j), \quad i=1,\dots, n.
   \]
 
-  \textbf{Estimación distribucional}: Se utiliza la media de \textit{todas}  las muestras generadas de \(Y^*\) como predicción:
+  \textbf{Predict-then-summarize}. For each MCMC iteration we get a sample \(\theta^{(m)*}\) of the approximate posterior. We first construct predictions on each step following our model, i.e.:
   \[
-    \hat Y = \frac{1}{M}\sum_{m=1}^M Y^{(j)*}.
+    Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}, \quad m=1,\dots,M, \quad i=1,\dots, n.
   \]
 
-  En ambos casos se puede considerar solo una de cada \(k\) muestras.
+  Then, the mean of all such intermediate predictions (\textit{posterior predictive distribution}) is used as a proxy for the response variable:
+  \[
+    \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots, n.
+  \]
+
 \end{frame}
 
-\begin{frame}{Predicciones II}
-    Podemos hacer también una selección de \(p\) variables usando los estimadores puntuales de \(\tau \mid Y\), y después aplicar cualquier algoritmo de regresión lineal.
+\begin{frame}{Predictions (cont.)}
+    \textbf{Variable selection}.
+We select \(p\) variables using only a point estimator of \(\tau=(t_1,\dots, t_p)|\mathcal D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
+    \vspace{1em}
 
     \begin{figure}
-      \includegraphics[width=\textwidth]{img/tau_posterior_lin}
-      \caption{Distribución a posteriori estimada para los puntos de impacto.}
+      \includegraphics[width=\textwidth]{img/tau_posterior}
+      \caption{Approximate posterior distribution for the impact points.}
     \end{figure}
 
-    En cualquier caso, medimos el error usando el MSE como métrica.
+    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
 \end{frame}
+
+\subsection{Validación de modelos}
 
@@
   \begin{itemize}
-    \item Análisis de la traza de las cadenas y de la distribución a posteriori de los parámetros. Se obtienen \textbf{intervalos creíbles} para los parámetros.
-    \item En cada paso obtenemos una estimación \(\tilde \theta_m\), y podemos generar \(Y^{(m)*} \mid \theta_m, X\) siguiendo el modelo asumido.
-    \item \textit{Bayesian p-values}: \(p=P(T(Y^*)\leq T(Y)\mid Y)\) para ciertas elecciones de \(T\): mínimo, máximo, mediana, media.
-Se calcula contando la proporción de muestras generadas que cumplen la desigualdad, y se espera que esté en torno a \(0.5\).
+    \item Análisis de la traza de las cadenas y de la distribución a posteriori de los parámetros. Se obtienen directamente \maroon{intervalos creíbles } para los parámetros.
+    \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) para ciertas elecciones de \(T\): mínimo, máximo, mediana, media, etc.
+Se calcula contando la proporción de muestras aproximadas que cumplen la desigualdad, y se espera que esté en torno a \(0.5\).
+    \item Análisis descriptivo y visual de la distribución a posteriori predictiva, tanto con los datos de entrenamiento como con nuevos datos de validación.
   \end{itemize}
@@
 
-\begin{frame}{Model checking II}
+\begin{frame}{Model checking (cont.)}
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/ppc_lin}
-    \caption{\textit{Posterior predictive checks} bajo un modelo RKHS.}
+    \includegraphics[width=\textwidth]{img/ppc_linear}
+    \caption{\textit{Posterior predictive checks} para regresión lineal.}
   \end{figure}
@@
 
-\subsection{Experimentos}
+\begin{frame}{Model checking (cont.)}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{img/ppc_logistic1}
+    \includegraphics[width=\textwidth]{img/ppc_logistic2}
+    \caption{\textit{Posterior predictive checks} para regresión logística.}
+  \end{figure}
+\end{frame}
+
+\section{Experimentos}
+
+\begin{frame}{Metodología}
+  \begin{itemize}
+    \item Por motivos prácticos, consideramos la posibilidad de fijar el valor de \(p\) de antemano, considerándolo como un hiperparámetro más del modelo (equivalente a una distribución a priori degenerada).
+  \end{itemize}
+\end{frame}
 
@@
 
-\begin{frame}{Metodología experimental II}
+\begin{frame}{Metodología experimental (cont.)}
   Fijamos algunos hiperparámetros:
@@
 
-\begin{frame}{Resultados RKHS}
-\textbf{Sin base}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \cmark & \cmark\\
-      O-U & \cmark & \cmark\\
-      RBF & \cmark & \cmark
-    \end{tabular}
-  \end{table}
-
-  \textbf{Base Fourier(11)}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \cmark &  \cmark\\
-      O-U & \cmark & \xmark\\
-      RBF & \cmark & \cmark
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Resultados RKHS II}
+\begin{frame}{Resultados L2}
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/res/reg_emcee_rkhs_fixed}
+    \includegraphics[width=\textwidth]{img/reg_emcee_l2}
     \caption{caption.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados RKHS III}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_rkhs_fbm_base11}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/lin_rkhs_ou_base11}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/lin_rkhs_rbf_base11}
-    \caption{MSE de los 5 mejores modelos en cada caso con el modelo subyacente RKHS y con base de Fourier.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados \(\boldsymbol{L^2}\)}
-\textbf{Sin base}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \xmark & \xmark\\
-      O-U & \xmark & \xmark\\
-      RBF & \xmark & \xmark
-    \end{tabular}
-  \end{table}
-\textbf{Base Fourier(11)}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \cmark &  \cmark\\
-      O-U & \cmark & \cmark\\
-      RBF & \xmark & \xmark
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Resultados \(\boldsymbol{L^2}\) II}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_fbm_nobase}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_ou_nobase}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_rbf_nobase}
-    \caption{MSE de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y sin usar suavizado con bases.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados \(\boldsymbol{L^2}\) III}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_fbm_base11}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_ou_base11}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/lin_l2_rbf_base11}
-    \caption{MSE de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y con base de Fourier.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados datos reales}
-\textbf{Sin base}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Dataset & Victoria & Victoria sel. variables \\ \hline
-      Tecator & \xmark & \xmark\\
-      Aemet & \xmark & \cmark\\
-    \end{tabular}
-  \end{table}
-
-  \textbf{Base Fourier(11)}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Dataset & Victoria & Victoria sel. variables \\ \hline
-      Tecator & \xmark &  \xmark\\
-      Aemet & \xmark & \cmark\\
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Resultados datos reales II}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_tecator_nobase}\hspace{5em}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_aemet_nobase}
-    \caption{MSE de los 5 mejores modelos en cada caso con datos reales y sin usar suavizado con bases.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados datos reales III}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_tecator_base11}\hspace{5em}
-    \includegraphics[width=0.3\textwidth]{img/results/lin_aemet_base11}
-    \caption{MSE de los 5 mejores modelos en cada caso con datos reales y con base de Fourier.}
   \end{figure}
@@
 
-\begin{frame}{Dificultades II}
+\begin{frame}{Dificultades (cont.)}
   \begin{itemize}
@@
 
-\textbf{Distribuciones a priori}: Igual que en regresión lineal.
-
-\textbf{Log-posterior}:
-\begin{align*}
-  \log \ &\pi(\beta, \tau, \alpha_0, \log\sigma\mid Y) \propto\\
-   &\sum_{i=1}^n \left[ \left(\alpha_0 + \Psi^{-1}_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}^{-1}(\alpha)\right\}\right)\right]\\
-   &+
-\frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0).
-\end{align*}
-\end{frame}
-
-\begin{frame}{Model checking}
-  \begin{figure}
-    \includegraphics[width=\textwidth]{img/ppc_log}
-    \caption{\textit{Posterior predictive checks} bajo un modelo RKHS.}
-  \end{figure}
 \end{frame}
@@
 
-\begin{frame}{Resultados RKHS}
-\textbf{Sin base}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \xmark & \xmark\\
-      O-U & \xmark & \cmark\\
-      RBF & \cmark & \xmark
-    \end{tabular}
-  \end{table}
-
-  \textbf{Base Fourier(9)}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \cmark &  \cmark\\
-      O-U & \cmark & \cmark\\
-      RBF & \cmark & \cmark
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Resultados RKHS II}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_fbm_nobase}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_ou_nobase}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_rbf_nobase}
-    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente RKHS y sin usar suavizado con bases.
-Distinguimos los algoritmos de comparación, nuestro algoritmo Bayesiano (emcee), y nuestra selección de variables Bayesiana.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados RKHS III}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_fbm_base9}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_ou_base9}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_rkhs_rbf_base9}
-    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente RKHS y con base de Fourier.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados \(\boldsymbol{L^2}\)}
-\textbf{Sin base}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \xmark & \xmark\\
-      O-U & \cmark & \xmark\\
-      RBF & \cmark & \cmark
-    \end{tabular}
-  \end{table}
-
-  \textbf{Base Fourier(9)}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Kernel & Victoria & Victoria sel. variables \\ \hline
-      fBM & \xmark &  \cmark\\
-      O-U & \xmark & \cmark\\
-      RBF & \cmark & \cmark
-    \end{tabular}
-  \end{table}
-\end{frame}
-
-\begin{frame}{Resultados \(L^2\) II}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/log_l2_fbm_nobase}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_l2_ou_nobase}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_l2_rbf_nobase}
-    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y sin usar suavizado con bases.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados \(L^2\) III}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/log_l2_fbm_base9}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_l2_ou_base9}\hfill
-    \includegraphics[width=0.3\textwidth]{img/results/log_l2_rbf_base9}
-    \caption{Precisión de los 5 mejores modelos en cada caso con el modelo subyacente \(L^2\) y con base de Fourier.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados datos reales}
-\textbf{Sin base}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Dataset & Victoria & Victoria sel. variables \\ \hline
-      Medflies & \xmark & \xmark\\
-      Growth & \cmark & \cmark\\
-    \end{tabular}
-  \end{table}
-
-  \textbf{Base Fourier(9)}
-  \begin{table}
-    \begin{tabular}{c|cc}
-      Dataset & Victoria & Victoria sel. variables \\ \hline
-      Medflies & \xmark &  \xmark\\
-      Growth & \cmark & \cmark\\
-    \end{tabular}
-  \end{table}
-
-  Tanto en las victorias como en las derrotas, la precisión es muy similar. Podemos considerar un empate en ambos conjuntos.
-\end{frame}
-
-\begin{frame}{Resultados datos reales II}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/log_medflies_nobase}\hspace{5em}
-    \includegraphics[width=0.3\textwidth]{img/results/log_growth_nobase}
-    \caption{Precisión de los 5 mejores modelos en cada caso con datos reales y sin usar suavizado con bases.}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Resultados datos reales III}
-  \begin{figure}
-    \includegraphics[width=0.3\textwidth]{img/results/log_medflies_base9}\hspace{5em}
-    \includegraphics[width=0.3\textwidth]{img/results/log_growth_base9}
-    \caption{Precisión de los 5 mejores modelos en cada caso con datos reales y con base de Fourier.}
-  \end{figure}
-\end{frame}
-
 \end{document}

### 2022-06-08 slides/seio.tex
Slides finished

@@
 \usepackage[absolute,overlay]{textpos}
-\usepackage[spanish, es-nodecimaldot]{babel}
+\usepackage[english]{babel}
 \usepackage{microtype}
@@
 \usepackage{listings}
+
+\usepackage[backend=biber,style=numeric-comp,sorting=none]{biblatex}
+\addbibresource{bibliography.bib}
+\setbeamertemplate{bibliography item}[article]
 
@@
 
-\title{Bayesian methods in RKHS models for functional regression}
+\title{Bayesian RKHS-based methods in functional regression}
 \providecommand{\subtitle}[1]{}
@@
 
-\begin{frame}{Usual models}
+\begin{frame}{The usual \(\bm{L^2}\)-models}
   \(\bullet\) \(Y\) real variable (continuous or dichotomous).
@@
     \[
-    \mathbb P (Y=1\mid X) = \frac{1}{\exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{\exp\{-\alpha_0 - \int_0^1 X(t)\beta(t)\, dt\}}.
+    \mathbb P (Y=1\mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \int_0^1 X\beta\}}.
   \]
@@
 
-\begin{frame}{Some shortcomings of the \(\bm{L^2}\)-models}
+\begin{frame}{Shortcomings of the \(\bm{L^2}\)-models}
   \begin{itemize}
@@
 
-\begin{frame}{RKHS essentials}
+\begin{frame}{RKHS reminder}
   \begin{definition}
@@
 
-\begin{frame}{Un repaso muy breve de los RKHS's (cont.)}
-Se pueden dar otras definiciones equivalentes de \(\Hcal(K)\), basadas en el \textit{operador de covarianzas}
-\[
-\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
-\]
-
-\begin{definition}
-\begin{enumerate}
-  \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), con \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
-  \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), con \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\).
+% \begin{frame}{Un repaso muy breve de los RKHS's (cont.)}
+% Se pueden dar otras definiciones equivalentes de \(\Hcal(K)\), basadas en el \textit{operador de covarianzas}
+% \[
+% \mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
+% \]
+%
+% \begin{definition}
+% \begin{enumerate}
+%   \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), con \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
+%   \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), con \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\).
 En este caso, \(\lambda_j\) y \(\phi_j\) son los autovalores y autofunciones de \(\mathcal K\).
-\end{enumerate}
-\end{definition}
-
-La segunda de estas definiciones enfatiza el hecho de que las funciones de \(\Hcal(K)\) son ``simples'', en el sentido de que sus componentes en una base ortonormal tienden a \(0\) rápidamente.
-
-\end{frame}
+% \end{enumerate}
+% \end{definition}
+%
+% La segunda de estas definiciones enfatiza el hecho de que las funciones de \(\Hcal(K)\) son ``simples'', en el sentido de que sus componentes en una base ortonormal tienden a \(0\) rápidamente.
+%
+% \end{frame}
 
@@
 
-  \begin{block}{Linear RKHS model}
+  \begin{block}{Linear RKHS model (Berrendero et al., 2019)}
   \[
@@
 \end{block}
-\begin{block}{Logistic RKHS model}
+\vspace{1em}
+\begin{block}{Logistic RKHS model (Berrendero et al., 2018)}
     \[
@@
 \end{block}
+
 \end{frame}
 
-\begin{frame}{Benefits of the RKHS models}
-  \begin{itemize}
-    \item MLE existe..
-    \item El modelo es verdadero en algunos casos...
-  \end{itemize}
-\end{frame}
+\section{A bayesian approach to parameter estimation}
 
-\section{A bayesian approach}
-
-\begin{frame}{Parameter estimation}
+\begin{frame}{Parameter estimation overview}
 
@@
   Same prior scenario for both linear and logistic models.
-  \begin{block}{Prior de los parámetros}
+  \begin{block}{Prior for the parameters involved}
     \vspace{-1em}
@@
 \end{align*}
-where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters.
+where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R, \ p_{\text{max}}\in \N\) and \(\eta \in \R^+\) are hyperparameters.
 \end{block}
@@
     \includegraphics[width=.7\textwidth]{img/posterior}
-    \caption{Estimated posterior distribution \(\theta|\mathcal D_n\).}
+    \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
   \end{figure}
@@
 
-\subsection{Validación de modelos}
-
 \begin{frame}{Model checking}
   \begin{itemize}
-    \item Análisis de la traza de las cadenas y de la distribución a posteriori de los parámetros. Se obtienen directamente \maroon{intervalos creíbles } para los parámetros.
-    \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) para ciertas elecciones de \(T\): mínimo, máximo, mediana, media, etc.
-Se calcula contando la proporción de muestras aproximadas que cumplen la desigualdad, y se espera que esté en torno a \(0.5\).
-    \item Análisis descriptivo y visual de la distribución a posteriori predictiva, tanto con los datos de entrenamiento como con nuevos datos de validación.
+    \item Analysis of the trace and posterior distribution of the parameters, for example through credible intervals.
+    \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) for any choice of \(T\): min, max, mean, median, mode, etc.
+They are computed by simply measuring the proportion of the approximate posterior samples that fall below the real value of the statistic, and are expected to be around \(0.5\) when the fit is good.
+    \item Descriptive and visual analysis of the posterior predictive distribution, both with training and validation data.
   \end{itemize}
@@
     \includegraphics[width=\textwidth]{img/ppc_linear}
-    \caption{\textit{Posterior predictive checks} para regresión lineal.}
+    \caption{\textit{Posterior predictive checks} for linear regression.}
   \end{figure}
@@
     \includegraphics[width=\textwidth]{img/ppc_logistic2}
-    \caption{\textit{Posterior predictive checks} para regresión logística.}
+    \caption{\textit{Posterior predictive checks} for logistic regression.}
   \end{figure}
@@
 
-\section{Experimentos}
+\section{Experiments}
 
-\begin{frame}{Metodología}
+\begin{frame}{Methodology}
   \begin{itemize}
-    \item Por motivos prácticos, consideramos la posibilidad de fijar el valor de \(p\) de antemano, considerándolo como un hiperparámetro más del modelo (equivalente a una distribución a priori degenerada).
+    \item Experiments on simulated data, with \(150\) samples of different Gaussian processes on a regular grid of \(100\) points. Also on real data sets.
+    \item We perform 10 random train/test partitions to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
+    \item For practical and computational reasons, \textbf{we allow the value of \(\bm{p}\) to be fixed beforehand}, treating it as an hyperparameter (equivalent to a degenerate prior distribution).
+{\color{red} problema de indentificabilidad de coeficientes!!}
+    \item Several functional reference algorithms for comparison, along with variable selection techniques, functional or otherwise.
+
   \end{itemize}
@@
 
-\begin{frame}{Datos sintéticos}
-  Se consideran 150 ejemplos de \(X \sim GP(0, K(s, t))\) y tres posibles variantes de \(K\): movimiento browniano fraccional (\(H=0.8\)), Ornstein-Uhlenbeck y kernel RBF.
-
-  Se genera la respuesta \(Y\) acorde a dos modelos, RKHS y \(L^2\). Concretamente, se elige
-  \[
-    Y_i \sim \mathcal N\big(5 -5X_i(0.1) + 10X_i(0.8), \ 0.5\big)
-  \]
-  ó
-  \[
-    Y_i \sim \mathcal N\left(5 + \int_0^1 \log(1+4t)X_i(t)\, dt, \ 0.5\right).
-  \]
-
-Consideramos una malla regular de \(N=100\) puntos en \([0, 1]\), y un reparto de 100 ejemplos para entrenamiento y 50 para evaluación.
-\end{frame}
-
-\begin{frame}{Datos reales}
-  Se consideran dos conjuntos de datos reales.
-  \begin{itemize}
-    \item \textbf{Tecator:} Contiene 215 ejemplos de mediciones de \textit{absorbancia} en muestras de carne para intentar predecir su contenido en grasa.
-    \item \textbf{Aemet}: Contiene 73 ejemplos de curvas de temperatura, a partir de las cuales se intenta predecir la precipitación total.
-  \end{itemize}
-
-  En ambos casos hacemos una división \(80\%-20\%\) para entrenamiento y \textit{test}.
-\end{frame}
-
-\begin{frame}{Preprocesado de los datos}
-  \begin{itemize}
-    \item Se centran los regresores para que tengan media \(0\). Opcionalmente, se pueden estandarizar en cada punto de la malla.
-    \item Se permite \textbf{sustituir los datos \(\boldsymbol{X_i}\) por su desarrollo en una base} de Fourier, con un número determinado de coeficientes.
-De esta forma realizamos un suavizado de las curvas.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Algoritmos de comparación}
-  \textbf{Regresión lineal multivariante:} Regresión Lineal estándar, Lasso (\(L^1\)), Ridge (\(L^2\)).
-
-  \textbf{Regresión no lineal multivariante}: SVM con kernel RBF.
-
-  \textbf{Regresión lineal funcional}: Modelo \(L^2\), KNN Funcional.
-
-  \textbf{Reducción de dimensión:} PCA Funcional (proyección sobre coeficientes).
-
-  \textbf{Selección de variables:} Aleatoria.
-
+\begin{frame}{Linear regression: simulated data}
   \vspace{1em}
-
-  Se entrenan todos ellos sobre el conjunto de entrenamiento, haciendo \textit{5-fold cross validation} para escoger los mejores hiperparámetros en cada caso (valor de regularización, número de componentes, ...).
-\end{frame}
-
-\begin{frame}{Metodología experimental}
-  \begin{itemize}
-  \item Para cada conjunto de datos consideramos tres posibles valores de \(p\) y cuatro posibles valores de \(\eta\):
-  \begin{itemize}
-    \item[--] \(p\in \{2,3,4\}\) para los datos sintéticos y \(p\in \{3, 4, 5\}\) para los datos reales.
-      \item[--] \(\eta \in \{0.01, 0.1, 1.0, 10.0\}\).
-  \end{itemize}
-
-  \item Para aumentar la estabilidad hacemos 5 repeticiones de cada modelo. Es decir, entrenamos un total de 60 modelos sobre el conjunto de entrenamiento
-
-  \item Con el \textbf{mejor modelo obtenido} se realiza también una selección de variables basada en los valores de \(\tau\) estimados mediante los distintos estimadores puntuales.
-\end{itemize}
-
-Además, repetimos este proceso con los datos suavizados en una base de Fourier con 11 elementos (también con los algoritmos de comparación).
-\end{frame}
-
-\begin{frame}{Metodología experimental (cont.)}
-  Fijamos algunos hiperparámetros:
-  \begin{itemize}
-    \item Número de cadenas: 64.
-    \item Número de pasos: 100 + 1000.
-    \item Movimientos: elección aleatoria (ponderada) entre dos de los recomendados, \textit{stretch} y \textit{walk}.
-    \item Inicialización: Entorno aleatorio del MLE + muestras de las distribuciones a priori.
-    \item Prior de \(\beta\): \(b_0\) se elige como el MLE de \(\beta\), y fijamos \(g=5\) (recomendado en \textit{Grollemund et al. (2019)}).
-  \end{itemize}
-  \vspace{1em}
-
-  \textit{Nota:} El valor de \(g\) no parece afectar al resultado, por lo que podríamos eliminar este parámetro.
-\end{frame}
-
-\begin{frame}{Resultados L2}
   \begin{figure}
     \includegraphics[width=\textwidth]{img/reg_emcee_l2}
-    \caption{caption.}
+    \caption{Mean MSE of predictors for simulated data that has an underlying \(L^2\)-model.}
   \end{figure}
@@
 
-\begin{frame}{Algunas observaciones}
+\begin{frame}{Linear regression: simulated data}
+    \vspace{1em}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{img/reg_emcee_l2}
+    \caption{Mean MSE of predictors for real data sets.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Logistic regression: simulated data}
+    \vspace{1em}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{img/clf_emcee_l2}
+    \caption{Mean accuracy of classifiers for simulated data that has an underlying logistic \(L^2\)-model.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Logistic regression: real data}
+    \vspace{1em}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{img/clf_emcee_real}
+    \caption{Mean accuracy of classifiers for real data sets.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Conclusions and future work}
+
+  \metroset{block=transparent}
+  \begin{alertblock}{Our proposal}
     \begin{itemize}
-  \item Por cada modelo entrenado obtenemos 4 estimadores y 3 estrategias de selección de variables.
-  \item En general, se obtienen mejores resultados con el uso de la base.
-  \item El tiempo medio de entrenamiento de cada modelo es de unos 20-30 segundos.
-  \item En ocasiones, nuestro algoritmo supera con holgura a todos los algoritmos de comparación. Cuando pierde, suele ser frente a uno o dos de ellos (y por poco), pero sigue superando al resto.
-  \item Se puede considerar como algoritmo de comparación el estimador de máxima verosimilitud de los parámetros.
-En casi todas las pruebas experimentales, el desempeño de nuestro algoritmo supera al del MLE.
+      \item Simple model for functional regression, with a solid mathematical background thanks to the theory of RKHS's.
+      \item Computationally feasible Bayesian approach to parameter estimation within the model.
+      \item Competitive with usual alternatives in both simulated and real data sets.
     \end{itemize}
+  \end{alertblock}
+
+  \begin{exampleblock}{The road ahead}
+    \begin{itemize}
+      \item Choose other prior distributions.
+      \item Package the algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}).
+      \item Reduce the variance in predictions, possibly by tackling the \textit{label switching} problem.
+    \end{itemize}
+  \end{exampleblock}
 
@@
 
-\begin{frame}{Dificultades}
-  \begin{itemize}
-
-  \item Hay multitud de grados de libertad: estandarización de regresores y/o variable respuesta, elección o no de una base (¿cuál?,) algoritmo de estimación del MLE, número, longitud y movimientos de las cadenas MCMC, elección de distribuciones a priori, elección de \(g, b_0\) y \(\eta\), etc.
-  \item El algoritmo MCMC es costoso, y en conjuntos de datos grandes las estrategias de \textit{cross-validation} para seleccionar parámetros pueden no ser viables.
-  \item Es necesario un procedimiento para seleccionar el valor de \(p\) (BIC, WAIC, WBIC, \textit{cross-validation}, ...).
-  \item Debido a la aleatoriedad del algoritmo, los resultados pueden variar sustancialmente de una ejecución a otra.
-  \item La distribución a priori para \(\beta\) depende en gran medida del valor de \(b_0\) escogido. Si su estimación inicial no es buena, el algoritmo no funciona demasiado bien.
-\end{itemize}
+\begin{frame}[allowframebreaks]
+    \frametitle{References}
+    \nocite{*}
+    \printbibliography[heading=none]
 \end{frame}
 
-\begin{frame}{Dificultades (cont.)}
-  \begin{itemize}
-    \item Hay un problema de \textit{identificabilidad} de los coeficientes (son intercambiables), que se agudiza especialmente al usar varias cadenas.
-    \item En el caso en el que el modelo subyacente es RKHS, es posible que no se recuperen los verdaderos valores de los parámetros debido a interacciones entre los coeficientes (si \(p\) es mayor que el número real de componentes).
-    \item El uso de distribuciones a priori impropias implica comprobar que \(\int_{\Theta} \pi(Y\mid \theta)\pi(\theta)\, d\theta < \infty\).
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Alternativas}
-  \begin{itemize}
-    \item Explorar el concepto de \textit{online learning} aplicado a esta situación, por ejemplo usando como priori de nuevos datos la posteriori aprendida.
-Estudiar la \textbf{consistencia} de la distribución a posteriori.
-    \item Utilizar otras herramientas de suavizado en lugar de bases de Fourier.
-    \item Sustitutir la distribución a priori de \(\beta\) por otra que requiera menos hiperparámetros.
-    \item Sustituir las distribuciones impropias por otras propias.
-  \end{itemize}
-
-\end{frame}
-
-\section{Regresión Logística Funcional Bayesiana}
-
-\subsection{Marco teórico}
-
-\begin{frame}{Planteamiento Bayesiano}
-\textbf{Modelo}: Cada \(Y_i\) se puede ver como una variable aleatoria de Bernoulli \(\mathcal B(p(x_i))\), con
-\[
-p_i \equiv p(x_i)=\mathbb P(Y_i=1\mid X_i=x_i) = \frac{1}{1 + \exp\left\{-\alpha_0-\displaystyle\sum_{j=1}^p \beta_jx_i(\tau_j)\right\}}.
-\]
-
-\end{frame}
-
-\begin{frame}{Predicciones}
-
-\begin{itemize}
-  \item Fijamos un umbral en \(0.5\) para establecer la pertenencia a las clases.
-   \item Los estimadores puntuales y la selección de variables son análogos al caso de regresión lineal.
-   \item Tenemos ahora dos estimadores basados en la distribución a posteriori (similares a los \textit{ensembles} de clasificadores):
-   \begin{itemize}
-     \item[--] Basado en el voto mayoritario de las muestras \(Y^*\) generadas.
-     \item[--] Basado en la media de las probabilidades \(p_i^*\) generadas (con aplicación posterior del umbral).
-   \end{itemize}
-  \item Se usa el \textit{accuracy} (precisión) para evaluar los modelos.
-\end{itemize}
-\end{frame}
-
-\subsection{Experimentos}
-
-\begin{frame}{Datos sintéticos}
-Seguimos la misma estrategia de generación de datos antes. Se consideran las mismas tres funciones de covarianza y se genera la respuesta tanto con un modelo RKHS como L2, i.e.:
-  \[
-    Y_i \sim \mathcal B\left(\frac{1}{1 + \exp\left\{0.5 + 5X_i(0.1) - 10X_i(0.8)\right\}}\right)
-  \]
-  ó
-  \[
-    Y_i \sim \mathcal B\left(\frac{1}{1 + \exp\left\{0.5 -\int_0^1 \log(1+4t)X_i(t)\, dt\right\}}\right).
-  \]
-
-  \vspace{1em}
-
-  Se introduce un pequeño ruido aleatorio en las etiquetas, y además se intenta que ambas clases estén balanceadas.
-
-\end{frame}
-
-\begin{frame}{Datos reales}
-  Se consideran dos conjuntos de datos reales.
-  \begin{itemize}
-    \item \textbf{Medflies:} Contiene 534 ejemplos de mediciones del número de huevos diario puesto por una serie de moscas, para intentar predecir si viven mucho o poco.
-    \item \textbf{Growth}: Contiene 93 ejemplos de curvas de altura en niños y niñas.
-  \end{itemize}
-
-  En ambos casos hacemos una división \(80\%-20\%\) para entrenamiento y \textit{test}.
-\end{frame}
-
-\begin{frame}{Algoritmos de comparación}
-  \textbf{Clasificación lineal multivariante:} Regresión Logística, SVM lineal.
-
-  \textbf{Clasificación no lineal multivariante}: SVM con kernel RBF.
-
-  \textbf{Clasificación funcional}: \textit{Maximum Depth}, \textit{Nearest Centroid} Funcional, KNN Funcional.
-
-  \textbf{Reducción de dimensión:} PCA Funcional (proyección sobre coeficientes).
-
-  \textbf{Selección de variables:} Aleatoria, \textit{Recursive Maxima Hunting}, \textit{RKVS} (Mahalanobis).
-
-  \vspace{1em}
-
-  Mismo preprocesado y metodología experimental que en el caso de regresión lineal, utilizando esta vez 9 coeficientes de Fourier.
+\begin{frame}[standout]
+  Thank you!
 \end{frame}

### 2022-06-08 slides/seio.tex
Some changes on slides

@@
 \subtitle{Theoretical framework and experiments}
-\date{SEIO Congress Granada\\ 8th May 2022\\}
+\date{SEIO Conference, Granada\\ 8th May 2022\\}
 \author{José R. Berrendero \\ Antonio Coín  \\ Antonio Cuevas\\}
@@
     \includegraphics[width=\textwidth]{img/data_lin}
-    \caption{Linear regression data set simulated from a fractional brownian motion.}
+    \caption{Linear regression data set simulated from a fractional Brownian motion.}
   \end{figure}
@@
     \includegraphics[width=\textwidth]{img/data_log}
-    \caption{Logistic regression data set simulated from a fractional brownian motion.}
+    \caption{Logistic regression data set simulated from a fractional Brownian motion.}
   \end{figure}
@@
     \]
-    \item When evaluating \(\dotprod{x}{\beta}_K\) for a trajectory \(x=X(\omega)\), we resort to Loeve's isometry:
+    \item When evaluating \(\dotprod{x}{\beta}_K\) for a trajectory \(x=X(\omega)\), we resort to Loève's isometry:
     \[
@@
     \[
-    \mathbb P (Y_i=1\mid X_i) = \frac{1}{\exp\{-\alpha_0 - \maroon{\langle X_i, \beta\rangle_K}\}} = \frac{1}{\exp\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\}}.
+    \mathbb P (Y_i=1| X_i) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X_i, \beta\rangle_K}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \sum_j \beta_j X_i(t_j)\}}.
   \]
@@
 \begin{frame}{Posterior approximation}
-  We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
+  % We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
 
@@
   \begin{figure}
-    \includegraphics[width=.7\textwidth]{img/posterior}
+    \includegraphics[width=.9\textwidth]{img/posterior}
     \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
@@
   \[
-  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j X_i(\hat \tau_j), \quad i=1,\dots, n.
+  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j X_i(\hat t_j), \quad i=1,\dots, n.
   \]
 
-  \textbf{Predict-then-summarize}. For each MCMC iteration we get a sample \(\theta^{(m)*}\) of the approximate posterior. We first construct predictions on each step following our model, i.e.:
+  \textbf{Predict-then-summarize}. We first construct predictions on each step of the MCMC algorithm following our model, i.e.:
   \[
@@
   \]
-
   Then, the mean of all such intermediate predictions (\textit{posterior predictive distribution}) is used as a proxy for the response variable:
@@
     \end{figure}
-
-    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
 \end{frame}
 
-\begin{frame}{Model checking}
-  \begin{itemize}
-    \item Analysis of the trace and posterior distribution of the parameters, for example through credible intervals.
-    \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) for any choice of \(T\): min, max, mean, median, mode, etc.
+% \begin{frame}{Model checking}
+%   \begin{itemize}
+%     \item Analysis of the trace and posterior distribution of the parameters, for example through credible intervals.
+%     \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) for any choice of \(T\): min, max, mean, median, mode, etc.
 They are computed by simply measuring the proportion of the approximate posterior samples that fall below the real value of the statistic, and are expected to be around \(0.5\) when the fit is good.
-    \item Descriptive and visual analysis of the posterior predictive distribution, both with training and validation data.
-  \end{itemize}
+%     \item Descriptive and visual analysis of the posterior predictive distribution, both with training and validation data.
+%   \end{itemize}
+%
+% \end{frame}
 
-\end{frame}
-
-\begin{frame}{Model checking (cont.)}
+\begin{frame}{Predictions (cont.)}
+    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/ppc_linear}
-    \caption{\textit{Posterior predictive checks} for linear regression.}
+    \includegraphics[width=.75\textwidth]{img/ppc_lin_standalone}
+    \caption{Example of \textit{posterior predictive distribution} in linear regression.}
   \end{figure}
@@
 
-\begin{frame}{Model checking (cont.)}
-  \begin{figure}
-    \includegraphics[width=\textwidth]{img/ppc_logistic1}
-    \includegraphics[width=\textwidth]{img/ppc_logistic2}
-    \caption{\textit{Posterior predictive checks} for logistic regression.}
-  \end{figure}
-\end{frame}
+% \begin{frame}{Model checking (cont.)}
+%   \begin{figure}
+%     \includegraphics[width=\textwidth]{img/ppc_logistic1}
+%     \includegraphics[width=\textwidth]{img/ppc_logistic2}
+%     \caption{\textit{Posterior predictive checks} for logistic regression.}
+%   \end{figure}
+% \end{frame}
 
@@
 
-\begin{frame}{Methodology}
-  \begin{itemize}
-    \item Experiments on simulated data, with \(150\) samples of different Gaussian processes on a regular grid of \(100\) points. Also on real data sets.
-    \item We perform 10 random train/test partitions to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
-    \item For practical and computational reasons, \textbf{we allow the value of \(\bm{p}\) to be fixed beforehand}, treating it as an hyperparameter (equivalent to a degenerate prior distribution).
+% \begin{frame}{Methodology}
+%   \begin{itemize}
+%     \item Experiments on simulated data, with \(150\) samples of different Gaussian processes on a regular grid of \(100\) points. Also on real data sets.
+%     \item We perform 10 random train/test partitions to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
+%     \item For practical and computational reasons, \textbf{we allow the value of \(\bm{p}\) to be fixed beforehand}, treating it as an hyperparameter (equivalent to a degenerate prior distribution).
 {\color{red} problema de indentificabilidad de coeficientes!!}
-    \item Several functional reference algorithms for comparison, along with variable selection techniques, functional or otherwise.
-
-  \end{itemize}
-\end{frame}
+%     \item Several functional reference algorithms for comparison, along with variable selection techniques, functional or otherwise.
+%
+%   \end{itemize}
+% \end{frame}
 
@@
     \includegraphics[width=\textwidth]{img/reg_emcee_l2}
-    \caption{Mean MSE of predictors for simulated data that has an underlying \(L^2\)-model.}
+    \caption{Mean MSE of predictors for simulated GP data that obeys an underlying \(L^2\)-model.}
   \end{figure}
@@
 
-\begin{frame}{Linear regression: simulated data}
+\begin{frame}{Linear regression: real data}
     \vspace{1em}
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/reg_emcee_l2}
+    \includegraphics[width=\textwidth]{img/reg_emcee_real}
     \caption{Mean MSE of predictors for real data sets.}
@@
     \includegraphics[width=\textwidth]{img/clf_emcee_l2}
-    \caption{Mean accuracy of classifiers for simulated data that has an underlying logistic \(L^2\)-model.}
+    \caption{Mean accuracy of classifiers for simulated GP data that obeys an underlying logistic \(L^2\)-model.}
   \end{figure}
@@
 \end{frame}
+
+\section{Conclusions}
 
@@
     \begin{itemize}
-      \item Choose other prior distributions.
-      \item Package the algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}).
-      \item Reduce the variance in predictions, possibly by tackling the \textit{label switching} problem.
+      \item Reduce the variance in predictions.
+      \item Experiment with other prior distributions.
+      \item Package the Python algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}, \textit{scikit-fda}).
     \end{itemize}

### 2022-06-08 slides/seio.tex
Changes in slides

@@
   \begin{definition}
-    Let \(K\) be continuous, and define the preliminary space
+    Let the covariance function \(K\) be continuous, and define the preliminary space
     \[
@@
 
-  Note that \(\Hcal(K)\subset L^2[0, 1]\), but the elements of \(\Hcal(K)\) are simpler and more manageable in general.
+  Note that \(\Hcal(K)\subset L^2[0, 1]\) as subsets, but the elements of \(\Hcal(K)\) are simpler and more manageable in general.
 Plus, the latter is a space of \maroon{functions}, and not of equivalence classes.
@@
     \[
-      \beta(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot).
+      \beta(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot), \quad \beta_j \in \R, \ t_j \in [0, 1], \ p \in \N.
     \]
@@
 \begin{frame}{...and two new models}
-  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), we consider:
+  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), for \(\beta \in \Hcal_0(K)\) we consider:
 
@@
 
-\begin{frame}{Posterior approximation}
+% \begin{frame}{Posterior approximation}
 %  We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
-
-  \vspace{1em}
-
-  \begin{figure}
-    \includegraphics[width=.9\textwidth]{img/posterior}
-    \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
-  \end{figure}
-\end{frame}
+%
+%   \vspace{1em}
+%
+%   \begin{figure}
+%     \includegraphics[width=.9\textwidth]{img/posterior}
+%     \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
+%   \end{figure}
+% \end{frame}
 
@@
 \begin{frame}{Predictions (cont.)}
+    \vspace{1em}
+  \begin{figure}
+    \includegraphics[width=.75\textwidth]{img/ppc_lin_standalone}
+    \caption{Example of \textit{posterior predictive distribution} in linear regression.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Predictions (cont.)}
     \textbf{Variable selection}.
@@
     \end{figure}
+
+    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
 \end{frame}
@@
 % \end{frame}
-
-\begin{frame}{Predictions (cont.)}
-    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
-  \begin{figure}
-    \includegraphics[width=.75\textwidth]{img/ppc_lin_standalone}
-    \caption{Example of \textit{posterior predictive distribution} in linear regression.}
-  \end{figure}
-\end{frame}
 

### 2022-06-08 slides/seio.tex
Changes in slides bis

@@
 
-\begin{frame}{Table of contents}
-  \tableofcontents
-\end{frame}
+% \begin{frame}{Table of contents}
+%   \tableofcontents
+% \end{frame}
 
@@
 \vspace{1em}
-  Instead of assuming \(\beta \in L^2[0, 1]\), we consider \maroon{\(\beta \in \mathcal H(K)\)}, the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\).
+  Instead of assuming \(\beta \in L^2[0, 1]\), we consider \maroon{\(\beta \in \mathcal H(K)\)}, the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\) (i.e.
+\(\E [X(t)]=0\) for all \(t\)).
 Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\beta}_K\)}.
@@
 
-\begin{frame}{Parameter estimation overview}
-
-  We have a parameter vector \(\theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)\), on which we impose a prior distribution \(\pi(\theta)\).
+% \begin{frame}{Parameter estimation overview}
+%
+%   We have a parameter vector \(\theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)\), on which we impose a prior distribution \(\pi(\theta)\).
 In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.
-
-  \vspace{1em}
-
-  Afterwards we apply \textit{Bayes'} theorem in the usual way to obtain the posterior distribution of the parameters given the data.
-
-  \begin{block}{Bayes's theorem}
-      \[
-      \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
-      \]
-  \end{block}
-  \vspace{1em}
-
-  Finally, we obtain samples of the approximate posterior distribution using an algorithm of the family of \textit{Markov chain Monte Carlo} (MCMC) methods.
-
-\end{frame}
+%
+%   \vspace{1em}
+%
+%   Afterwards we apply \textit{Bayes'} theorem in the usual way to obtain the posterior distribution of the parameters given the data.
+%
+%   \begin{block}{Bayes's theorem}
+%       \[
+%       \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
+%       \]
+%   \end{block}
+%   \vspace{1em}
+%
+%   Finally, we obtain samples of the approximate posterior distribution using an algorithm of the family of \textit{Markov chain Monte Carlo} (MCMC) methods.
+%
+% \end{frame}
 
@@
 
+\vspace{1em}
 \begin{alertblock}{Hyperparameter selection}
@@
 
-\vspace{1em}
   \begin{block}{Logistic log-posterior}
@@
 \end{align*}
-Remember that \(\Psi_{X_i}(\alpha) = \sum_j \beta_j X_i(t_j)\).
   \end{block}
+  They are approximated through \textit{Markov chain Monte Carlo} (MCMC).
 \end{frame}
@@
       \item Reduce the variance in predictions.
-      \item Experiment with other prior distributions.
+      \item Derive theoretical properties of the predictors.
       \item Package the Python algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}, \textit{scikit-fda}).

### 2022-06-24 paper/paper.tex
Reorganize files and make some corrections in paper

@@
 In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2021functional}.
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2018functional}.
 
@@
 In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2021functional} for the case of functional logistic regression.
+These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2018functional} for the case of functional logistic regression.
 Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
@@
 Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-A similar Bayesian scheme has been recently explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
-\incomment{Comentar nuestros resultados experimentales} % Overall good results
+A similar Bayesian scheme has recently been explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
 
@@
 As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
-In this way, in addition to making predictions about the input data, we could evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
+As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
+In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
 These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-\incomment{Comentar nuestros resultados experimentales}
 
@@
 
-\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-Let us denote by \todo{Creo que olvidaste decir que la media es 0}\(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of \(X\), and in what follows suppose that it is continuous.
-To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
+\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
+
+The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
+Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of the centered process \(X\), and in what follows suppose that it is continuous.
+To construct the RKHS \(\Hcal(K)\) associated with the covariance function \(K\), we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
 \begin{equation}\label{eq:h0}
@@
 This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, we define \(\Hcal(K)\) to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
-As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\), \sout{thus giving it its name}.
-An important detail to note is that, since \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, the values of these functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
+Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
+As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\).
+An important consequence is that \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
 
@@
 
+\begin{comment}
+  Explicar mejor de dónde viene la notación de Parzen, y por qué se podría pensar en usarla para expresar un producto escalar (``congruence inner product'', e.g.
+4.17 en Parzen 1961b.
+¿Expresión en Torrecilla et al.
+y Kailath?).
+\end{comment}
+
 \section{Bayesian methodology for RKHS-based functional regression models}\label{sec:methodology}
@@
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\).
+Hence, taking into account that \(\alpha \in \Hcal_{0,p}(K)\) and that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\) when \(\alpha(\cdot)=\sum_j\beta_j K(t_j, \cdot)\).
 
@@
 
-In the case of functional linear regression, the RKHS model considered (with parameter \(\alpha\in \Hcal_{0,p}(K)\)) is given by
+In the case of functional linear regression, the simplified RKHS model considered in this work is given by
 \begin{equation}\label{eq:rkhs-model-linear}
@@
 \end{equation}
-where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), \todo{$\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j,\cdot)$, } and \(\epsilon\) is a normally-distributed error term independent from \(X\), satisfying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+where \(\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j, \cdot) \in \Hcal_{0,p}(K), \beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is a normally-distributed error term independent from \(X\), with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+This model is essentially a finite-dimensional approximation from a functional perspective to the more general RKHS model that assumes \(\alpha \in \Hcal(K)\), proposed in~\citet{berrendero2019rkhs}.
+
 When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
@@
 As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, \sout {as far as generalizations go the proposed model is easily extensible, seeing as how several functional covariates could be added to obtain an extended model such as} \color{violet} our model could be easily extended to the case of several covariables via an expression of type \color{black} \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
+In addition, this model could be easily extended to the case of several covariates via an expression of type \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
 In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
 
-\subsubsection{Prior selection and posterior derivation}
+\subsubsection{The bayesian approach: prior selection and posterior derivation}
 
@@
   \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
+  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2{\underbrace{\left(\mathcal X_\tau' \mathcal X_\tau + \eta I\right)}_{G_\tau}}^{-1}),
 \end{split}
 \end{align}
-where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
+where \(I\) is the identity matrix, \(\mathcal X_\tau\) is the data matrix \((X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
 On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
@@
 On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-All in all, the proposed prior distribution becomes\color{violet}{, with a slight abuse of notation, } \color{black} \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
+All in all, with a slight abuse of notation the proposed prior distribution becomes \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
 
@@
 \begin{description}
-  \item[Summarize-then-predict.] \todo{\tiny He añadido un punto}If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
+  \item[Summarize-then-predict.] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
   \[
@@
 
-  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the ``sampled responses'' \todo{¿Mejor "simulated responses''?}at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
+  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
   \begin{equation}\label{eq:sampled-response-vector}
@@
   \end{equation}
-  Then, we can take the mean of all such samples \todo{¿No sería mejor "simulated responses", en lugar de "samples" (recuerda que sample es el conjunto de los datos)} as a proxy for each response variable, that is,
+  Then, we can take the mean of all such simulated responses as a proxy for each response variable, that is,
   \[
@@
 Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
-After this variable selection has been carried out, we can treat the problem \color{violet} using a \color{black}  finite-dimensional linear regression \color{violet} model\color{black} and apply any of the well-known prediction algorithms suited for this situation.
+After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
 \end{description}
 
-Note that these predictors \color{violet} can be \color{black}  obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
+Note that these predictors can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
 As a consequence, what we have in practice is a single algorithm that can produce multiple predictors at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
 Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
-
-\todo[inline]{Me ha gustado esta sección}
 
@@
 In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
-\color{violet} Then, following the approach  suggested by \citet{berrendero2021functional}, an RKHS model might be given  by the  equation \color{black}
+Then, following the approach  suggested by \citet{berrendero2018functional}, an RKHS model might be given  by the  equation
 \begin{equation}\label{eq:rkhs-model-logistic}
@@
 \end{equation}
-\color{violet} Indeed, note that this can be seen as a finite-dimensional approximation (but, still, with a functional interpretation) to the general RKHS functional logistic model proposed by these authors, which can be obtained by replacing $\Hcal_{0,p}(K)$ with the whole RKHS space $\Hcal(K)$.
-Now, if we aim at a classification problem, our strategy will be similar to that followed the general linear model: after incorporating the sample information we can rewrite the model ~\eqref{eq:rkhs-model-logistic} as\color{black}
+
+Indeed, note that this can be seen as a finite-dimensional approximation (but, still, with a functional interpretation) to the general RKHS functional logistic model proposed by these authors, which can be obtained by replacing \(\Hcal_{0,p}(K)\) with the whole RKHS space \(\Hcal(K)\).
+Now, if we aim at a classification problem, our strategy will be similar to that followed in the functional linear model: after incorporating the sample information, we can rewrite~\eqref{eq:rkhs-model-logistic} as
 \begin{equation}\label{eq:rkhs-model-logistic-2}
@@
 
-Much the same way as the linear regression model described above, \color{violet} this RKHS-based logistic regression model offers some advantages.
-First and foremost, it can be handled allows for a workable Bayesian approach, as we will show below.
-Second, as pointed out in \cite[see Theorem 1 in][]{berrendero2021functional}, under miild conditions the RKHS logistic functional model holds  whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes; this provides a sound theoretical motivation for such model.
-Third, a maximum likelihood approach for parameter estimation (not considered here) is possible as well.
-Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case; however, let us recall that, even in finite-dimensional settings, there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence} though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
+In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model.
+First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
+Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes~\citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for such model.
+Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
+Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case.
+However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
 \color{black}
 
-\subsubsection{\color{violet} The Bayesian approach: prior selection and posterior derivation}
+\subsubsection{The Bayesian approach: prior selection and posterior derivation}
 
-As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression \color{violet} model\color{black}.
+As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model.
 However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
-The posterior distribution is in turn derived after a routine calculation.
+The posterior distribution is then derived after a routine calculation.
 
@@
 \end{align*}
-Remember that \(\Psi_{X_i}(\alpha) = \sum_j \beta_j X_i(t_j)\).
+Remember that \(\Psi_{X_i}(\alpha) = \sum_{j=1}^p \beta_j X_i(t_j)\).
 \end{proposition}
@@
 
-Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to \(\{0, 1\}\).
-\color{violet} According to the usual criterion of minimizing the missclassification probability, the optimal rule is predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
-Other criteria would lead to general rules of type predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold $\gamma$.
-\color{black}
-\todo[inline]{\footnotesize La frase original era un poco chocante.
-Si no me equivoco no se trata de una "convención" el valor genérico $\gamma$ puede tenr sentido la regla Bayes (con el criterio usual de minimizar la probsbilidad de clasificación errónea) se obtiene siempre com $\gamma=1/2$} With this thresholding strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
+Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to a binary output in \(\{0, 1\}\).
+According to the usual criterion of minimizing the misclassification probability, the Bayes optimal rule is recovered by predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
+Nevertheless, for a more general cost function one could consider other criteria that would lead to evaluating whether \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold \(\gamma\in[0, 1]\).
+
+With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
 \[
@@
 the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
-\color{violet} Hence, there are primarily two possible outcomes:\color{black}
+Hence, there are primarily two possible outcomes:
 \begin{description}
-  \item[Average sampled probability] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
+  \item[Average sampled probability.] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
   \[
@@
   \]
-  \item[Average sampled response] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
+  \item[Average sampled response.] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
 \[
@@
 \paragraph{The choice of \(\bm p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications.
+While theoretically we could impose a prior distribution on \(p\) as well (e.g.
+a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
 For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
@@
 BIC, DIC, cross-validation, \ldots) to select its optimal value.
+This is equivalent to imposing a degenerate prior distribution on \(p\).
 Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}).
+\incomment{Reversible MCMC}
 
@@
   \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
+That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
 This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
@@
 In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest \color{violet}probability density.
-\color{black} At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
+On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
+At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
 
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{img/ppc_lin}
+  \includegraphics[width=.75\textwidth]{img/ppc_linear}
   \caption{Posterior predictive graphical checks on a fitted model.
@@
 \section{Experimental results}\label{sec:results}
+
+We show below the results for 10 random train/test partitions and 5-fold cross-validation to select the parameters \(p\) and \(\eta\).
+
+\subsection{Functional Linear Regression}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{img/reg_emcee_rkhs}
+  \caption{Mean MSE for linear regression with simulated data following an underlying RKHS model.}\label{fig:reg-emcee-rkhs}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{img/reg_emcee_l2}
+  \caption{Mean MSE for linear regression with simulated data following an underlying \(L^2\)-model.}\label{fig:reg-emcee-l2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{img/reg_emcee_real}
+  \caption{Mean MSE for linear regression with real data sets.}\label{fig:reg-emcee-real}
+\end{figure}
+
+\subsection{Functional Logistic Regression}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{img/clf_emcee_rkhs}
+  \caption{Mean accuracy for logistic regression with simulated data following an underlying RKHS model.}\label{fig:clf-emcee-rkhs}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{img/clf_emcee_l2}
+  \caption{Mean accuracy for logistic regression with simulated data following an underlying \(L^2\)-model.}\label{fig:clf-emcee-l2}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{img/clf_emcee_real}
+  \caption{Mean accuracy for logistic regression with real data sets.}\label{fig:clf-emcee-real}
+\end{figure}
+
+\newpage
+
 \begin{comment}

### 2022-06-24 slides/seio.tex
Reorganize files and make some corrections in paper

--- 2022-06-08 slides/seio.tex
@@
 \subtitle{Theoretical framework and experiments}
-\date{SEIO Conference, Granada\\ 8th May 2022\\}
+%\date{SEIO Conference, Granada\\ \today\\}
+\date{\today\\ \\}
 \author{José R. Berrendero \\ Antonio Coín  \\ Antonio Cuevas\\}
@@
 \begin{frame}{FDA: linear regression}
-  \textbf{First problem}: \maroon{Linear } regression with functional data.
+  \textbf{First problem}: \maroon{linear } regression with functional data.
   \vspace{1em}
@@
 \begin{frame}{FDA: logistic regression}
-  \textbf{Second problem}: \maroon{Logistic } regression with functional data.
+  \textbf{Second problem}: \maroon{logistic } regression with functional data.
 \vspace{1em}
@@
 \end{block}
+\setbeamerfont{block body}{size=\small}
 \begin{block}{Logistic \(\bm{L^2}\)-model}
     \[
-    \mathbb P (Y=1\mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \int_0^1 X\beta\}}.
+    \mathbb P (Y=1\mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \int_0^1 X(t)\beta(t)\, dt\}}.
   \]
 \end{block}
-  \(\alpha_0\in\mathbb R\), \ \(\varepsilon \sim \mathcal N(0, \sigma^2), \ \mathbb E[\varepsilon]=0, \ X \bot \varepsilon\).
+  \(\alpha_0\in\mathbb R\), \ \(\varepsilon \sim \mathcal N(0, \sigma^2), \ X \bot \varepsilon\).
+  \setbeamerfont{block body}{size=\normalsize}
 \end{frame}
@@
 
-% \begin{frame}{Un repaso muy breve de los RKHS's (cont.)}
-% Se pueden dar otras definiciones equivalentes de \(\Hcal(K)\), basadas en el \textit{operador de covarianzas}
-% \[
-% \mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
-% \]
-%
-% \begin{definition}
-% \begin{enumerate}
-%   \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), con \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
-%   \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), con \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\).
-En este caso, \(\lambda_j\) y \(\phi_j\) son los autovalores y autofunciones de \(\mathcal K\).
-% \end{enumerate}
-% \end{definition}
-%
-% La segunda de estas definiciones enfatiza el hecho de que las funciones de \(\Hcal(K)\) son ``simples'', en el sentido de que sus componentes en una base ortonormal tienden a \(0\) rápidamente.
-%
-% \end{frame}
+\begin{frame}{RKHS reminder (cont.)}
+  There are other equivalent definitions of \(\Hcal(K)\), mainly based on the \textit{covariance operator}
+\[
+\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
+\]
+
+\begin{definition}
+\begin{enumerate}
+  \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
+  \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\).
+Here, \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\).
+\end{enumerate}
+\end{definition}
+
+The second definition emphasizes the fact that functions in \(\Hcal(K)\) are ``simple'', in the sense that their components in an orthonormal basis need to vanish quickly.
+
+\end{frame}
 
@@
   \end{block}
-  With a slight abuse of notation, we will write \maroon{\(\dotprod{x}{\beta}_k \equiv \Psi_x(\beta)\)}. This correspondence will allow us to still call our models ``linear''.
+  With a slight abuse of notation, we will write \maroon{\(\dotprod{x}{\beta}_K \equiv [\Psi_X(\beta)](\omega) := \Psi_x(\beta)\)}\ for \(x=X(\omega)\).
+This correspondence will allow us to still call our models ``linear''.
 \end{frame}
@@
 \vspace{1em}
+
+\setbeamerfont{block body}{size=\small}
 \begin{block}{Logistic RKHS model (Berrendero et al., 2018)}
     \[
-    \mathbb P (Y_i=1| X_i) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X_i, \beta\rangle_K}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \sum_j \beta_j X_i(t_j)\}}.
+    \mathbb P (Y_i=1| X_i) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X_i, \beta\rangle_K}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\}}.
   \]
 \end{block}
+\setbeamerfont{block body}{size=\normalsize}
 
+\end{frame}
+
+\begin{frame}{Advantages of RKHS models}
+  \begin{itemize}
+    \item They are finite-dimensional approximations with a functional interpretaion, which simplifies their understanding and handling.
+    \item Under mild conditions, the general RKHS models include the \(L^2\)-models as particular cases.
+    \item The problem of non-existence of the MLE is mitigated in the logistic model.
+    \item They are particularly suited as a basis for variable selection methods.
+  \end{itemize}
 \end{frame}
@@
 
-  We have a parameter vector \(\theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)\), on which we impose a prior distribution \(\pi(\theta)\).
-In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.
+  We have a parameter vector
+  \[
+  \theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)' \equiv (p, b, \tau, \alpha_0, \sigma^2)',
+  \]
+  on which we impose a prior distribution \(\pi(\theta)\). In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.
 
@@
   \end{block}
-  \vspace{1em}
-
-  Finally, we obtain samples of the approximate posterior distribution using an algorithm of the family of \textit{Markov chain Monte Carlo} (MCMC) methods.
 
@@
   Same prior scenario for both linear and logistic models.
+  \vspace{.5em}
   \begin{block}{Prior for the parameters involved}
@@
   \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  \beta\mid \tau, \sigma^2 & \sim \mathcal N\big(b_0, g\sigma^2{\underbrace{\left[\mathcal X_\tau' \mathcal X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
+  b\mid \tau, \sigma^2 & \sim \mathcal N_p\big(b_0, g\sigma^2{\underbrace{\left[\mathcal X_\tau' \mathcal X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
 \end{align*}
-where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R, \ p_{\text{max}}\in \N\) and \(\eta \in \R^+\) are hyperparameters.
+
+\vspace{-1em}
+where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\) is the data matrix, and \(b_0\in \R^p, \ g \in \R, \ p_{\text{max}}\in \N\) and \(\eta \in \R^+\) are hyperparameters.
 \end{block}
@@
 \begin{frame}{Posterior distributions}
+  \setbeamerfont{block body}{size=\small}
   \begin{block}{Linear log-posterior}
@@
 \end{align*}
-where \(\bm Y=(Y_1,\dots,Y_n)'\) y \(\bm{1}\) is an \(n\)-dimensional vector of ones.
+where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
   \end{block}
-
   \vspace{1em}
@@
   \end{block}
+  \setbeamerfont{block body}{size=\normalsize}
+  They are approximated through \textit{Markov chain Monte Carlo} (MCMC).
 \end{frame}
 
-% \begin{frame}{Posterior approximation}
-%  We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
-%
-%   \vspace{1em}
-%
-%   \begin{figure}
-%     \includegraphics[width=.9\textwidth]{img/posterior}
-%     \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
-%   \end{figure}
-% \end{frame}
+\begin{frame}{Posterior approximation}
+ We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
+
+  \vspace{1em}
+
+  \begin{figure}
+    \includegraphics[width=.7\textwidth]{img/posterior}
+    \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
+  \end{figure}
+\end{frame}
 
@@
     \textbf{Variable selection}.
-We select \(p\) variables using only a point estimator of \(\tau=(t_1,\dots, t_p)|\mathcal D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
+We select \(p\) variables using only a point estimator of \(\tau|\mathcal D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
     \vspace{1em}
@@
 
-% \begin{frame}{Model checking}
-%   \begin{itemize}
-%     \item Analysis of the trace and posterior distribution of the parameters, for example through credible intervals.
-%     \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) for any choice of \(T\): min, max, mean, median, mode, etc.
+\begin{frame}{Bayesian model checking}
+  \begin{itemize}
+    \item It is possible to incorporate existing information through the prior distributions.
+    \item Analysis of the trace and posterior distribution of the parameters, for example through credible intervals.
+    \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) for any choice of \(T\) (min, max, mean, median, mode, etc.).
 They are computed by simply measuring the proportion of the approximate posterior samples that fall below the real value of the statistic, and are expected to be around \(0.5\) when the fit is good.
-%     \item Descriptive and visual analysis of the posterior predictive distribution, both with training and validation data.
-%   \end{itemize}
-%
-% \end{frame}
+    \item Descriptive and visual analysis of the posterior predictive distribution, both with training and validation data.
+  \end{itemize}
 
-% \begin{frame}{Model checking (cont.)}
-%   \begin{figure}
-%     \includegraphics[width=\textwidth]{img/ppc_logistic1}
-%     \includegraphics[width=\textwidth]{img/ppc_logistic2}
-%     \caption{\textit{Posterior predictive checks} for logistic regression.}
-%   \end{figure}
-% \end{frame}
+\end{frame}
 
@@
 
-% \begin{frame}{Methodology}
-%   \begin{itemize}
-%     \item Experiments on simulated data, with \(150\) samples of different Gaussian processes on a regular grid of \(100\) points. Also on real data sets.
-%     \item We perform 10 random train/test partitions to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
-%     \item For practical and computational reasons, \textbf{we allow the value of \(\bm{p}\) to be fixed beforehand}, treating it as an hyperparameter (equivalent to a degenerate prior distribution).
-{\color{red} problema de indentificabilidad de coeficientes!!}
-%     \item Several functional reference algorithms for comparison, along with variable selection techniques, functional or otherwise.
-%
-%   \end{itemize}
-% \end{frame}
+\begin{frame}{Methodology}
+  \begin{itemize}
+    \item Experiments on simulated data, with \(300\) samples of different Gaussian processes on a regular grid of \(100\) points. Also experiments on real data sets.
+    \item We perform 10 random train/test partitions (67\%/33\%) to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
+    \item For practical and computational reasons, \textbf{we allow the value of \(\bm{p}\) to be fixed beforehand}, treating it as an hyperparameter (equivalent to a degenerate prior distribution).
+    \item Several reference methods are considered for comparison, along with variable selection techniques, functional or otherwise.
 
-\begin{frame}{Linear regression: simulated data}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Linear regression: simulated RKHS data}
   \vspace{1em}
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/reg_emcee_l2}
-    \caption{Mean MSE of predictors for simulated GP data that obeys an underlying \(L^2\)-model.}
+    \includegraphics[width=.75\textwidth]{img/reg_emcee_rkhs}
+    \caption{Mean MSE of predictors for simulated GP data that obeys an underlying RKHS model (left is better).}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Linear regression: simulated \(\bm{L^2}\) data}
+  \vspace{1em}
+  \begin{figure}
+    \includegraphics[width=.75\textwidth]{img/reg_emcee_l2}
+    \caption{Mean MSE of predictors for simulated GP data that obeys an underlying \(L^2\)-model (left is better).}
   \end{figure}
@@
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/reg_emcee_real}
-    \caption{Mean MSE of predictors for real data sets.}
+    \includegraphics[width=.75\textwidth]{img/reg_emcee_real}
+    \caption{Mean MSE of predictors for real data sets (left is better).}
   \end{figure}
@@
 
-\begin{frame}{Logistic regression: simulated data}
+\begin{frame}{Logistic regression: simulated RKHS data}
     \vspace{1em}
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/clf_emcee_l2}
-    \caption{Mean accuracy of classifiers for simulated GP data that obeys an underlying logistic \(L^2\)-model.}
+    \includegraphics[width=.75\textwidth]{img/clf_emcee_rkhs}
+    \caption{Mean accuracy of classifiers for simulated GP data that obeys an underlying logistic RKHS model (right is better).}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Logistic regression: simulated \(\bm{L^2}\) data}
+    \vspace{1em}
+  \begin{figure}
+    \includegraphics[width=.75\textwidth]{img/clf_emcee_l2}
+    \caption{Mean accuracy of classifiers for simulated GP data that obeys an underlying logistic \(L^2\)-model (right is better).}
   \end{figure}
@@
   \begin{figure}
-    \includegraphics[width=\textwidth]{img/clf_emcee_real}
-    \caption{Mean accuracy of classifiers for real data sets.}
+    \includegraphics[width=.75\textwidth]{img/clf_emcee_real}
+    \caption{Mean accuracy of classifiers for real data sets (right is better).}
   \end{figure}
@@
     \begin{itemize}
-      \item Simple model for functional regression, with a solid mathematical background thanks to the theory of RKHS's.
+      \item Simple and very interpretable functional regression model, with a solid mathematical background thanks to the theory of RKHS's.
       \item Computationally feasible Bayesian approach to parameter estimation within the model.
@@
     \begin{itemize}
-      \item Reduce the variance in predictions.
-      \item Experiment with other prior distributions.
+      \item Experiment with different prior distributions.
+      \item Derive theoretical properties of the predictors.
       \item Package the Python algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}, \textit{scikit-fda}).

### 2022-08-06 notes/mcmc.tex
Reorganize; edit paper; add mcmc notes

%%%
% Class 'kaobook' from https://github.com/fmarotta/kaobook
%
% Copyright (c) 2022 Antonio Coín Castro
%%%

\documentclass[
  a4paper,
	fontsize=11pt, % Base font size
	twoside=false, % Use different layouts for even and odd pages (in particular, if twoside=true, the margin column will be always on the outside)
	%open=any, % If twoside=true, uncomment this to force new chapters to start on any page, not only on right (odd) pages
	%chapterprefix=true, % Uncomment to use the word "Chapter" before chapter numbers everywhere they appear
	%chapterentrydots=true, % Uncomment to output dots from the chapter name to the page number in the table of contents
  secnumdepth=2,
	numbers=noenddot, % Comment to output dots after chapter numbers; the most common values for this option are: enddot, noenddot and auto (see the KOMAScript documentation for an in-depth explanation)
	%draft=true, % If uncommented, rulers will be added in the header and footer
	%overfullrule=true, % If uncommented, overly long lines will be marked by a black box; useful for correcting spacing problems
]{kaohandt}

% Choose the language
\ifxetexorluatex
	\usepackage{polyglossia}
	\setmainlanguage{spanish}
\else
	\usepackage[spanish]{babel} % Load characters and hyphenation
\fi

% Change default fonts
\setmathfont[range={\mathcal,\mathbfcal, \mathbb}]{NewCMMath-Regular.otf}

% Change page style
\pagestyle{scrheadings}

% Load the bibliography package
\usepackage[
  style=authoryear-comp,
  maxbibnames=99
]{kaobiblio}

% Add comma before year in citations
\renewcommand*{\nameyeardelim}{\addcomma\space}

% Bibliography side cite style
\renewcommand{\formatmargincitation}[1]{%
  \parencite{#1}: \citetitle{#1}.%
}

% Bibliography file
\addbibresource{bibliography.bib}

% Load the package for hyperreferences
\usepackage{kaorefs}

\DefineBibliographyStrings{spanish}{andothers={et~al\adddot}}

% Set reference colors
\hypersetup{
  pdfauthor = {Antonio Coín},
  pdftitle = {MCMC y la librería emcee},
  citecolor=teal,
  urlcolor=blue!50!black,
  %anchorcolor=OliveGreen,
  linkcolor=teal
}

% Load mathematical packages for theorems and related environments
\usepackage{kaotheorems}

% Set the paths where to look for images
\usepackage{subcaption}
\graphicspath{{img/}}

% Only show section titles in margintoc
\setcounter{margintocdepth}{\sectiontocdepth}

% Load additional packages
\usepackage{dsfont}  %\mathds

% Some custom commands
\newcommand{\R} {\ensuremath{\mathds{R}}}
\newcommand{\E} {\ensuremath{\mathds{E}}}
\newcommand{\eps}{\varepsilon}
\newcommand{\bm}{\symbf}

%% Scalar product
\newcommand\dotprod[2]{\left\langle #1,\, #2 \right\rangle}

\DeclareMathOperator{\var} {Var}
\DeclareMathOperator{\cov} {Cov}

% Do not re-initialize sidenote counter after each section
\counterwithout{sidenote}{section}

%%%%%%%%%%%%%%%
% DOCUMENT
%%%%%%%%%%%%%%%

\begin{document}

\title{Markov chain Monte Carlo y la librería emcee}
\author[FM]{Antonio Coín Castro}
\date{\normalsize{Julio 2022}}
\maketitle

\section{Problema a resolver}

Queremos obtener muestras de una distribución que no conocemos en forma cerrada\sidenote[][*-1]{También es posible que conozcamos \(p(x)\), pero que obtener muestras de ella directamente sea difícil computacionalmente.}, pero sí conocemos una función proporcional a su densidad. Formalmente, queremos muestrear de una densidad desconocida \(p(x)\) a través de una función conocida \(f(x)\), dado que \(p(x) \propto f(x)\). Esta situación es común, por ejemplo, en inferencia bayesiana, donde la distribución a posteriori es muchas veces intratable debido a la constante de normalización, pero sí sabemos que \(posterior \propto prior \times likelihood\).

Cuando \(x \in \R\), podemos emplear algoritmos Monte Carlo de muestreo por rechazo o \textit{rejection sampling}, que se basan en muestrear uniformemente del área que encierra \(f(x)\) a través de otra función auxiliar que recubre este área. Sin embargo, en dimensión superior estos algoritmos sufren rápidamente la llamada \textit{maldición de la dimensionalidad}. Es por esto que para muestrear de distribuciones multidimensionales surgen nuevos métodos: los algoritmos de Markov chain Monte Carlo (MCMC).

\begin{marginfigure}[*-9]
  \centering
  \includegraphics[width=.6\textwidth]{img/cube_sphere}
  \caption*{Intuitivamente, para recubrir el volumen de \(f(x)\) en dimensión alta necesitamos un volumen que escala exponencialmente con la dimensión.}
\end{marginfigure}

\section{MCMC}

Estos métodos se basan en la construcción iterativa de una cadena de Markov cuya distribución estacionaria es la distribución objetivo \(p(x)\). De esta forma, en una cadena suficientemente avanzada podemos tomar los distintos estados por los que se ha pasado como muestras aproximadas de \(p(x)\). A la hora de construir las cadenas, las probabilidades de transición se asignan de forma que se favorezcan regiones donde se acumula más densidad de la distribución objetivo (evaluando la función \(f(x)\) conocida). \marginnote[*-4]{Se consigue paliar un poco el problema de la maldición de la dimensionalidad, e incluso se puede reducir más la influencia de la dimensión controlando el tamaño de los pasos en la cadena.}

A diferencia de otros algoritmos de Monte Carlo, aquí obtenemos muestras correladas. Si se desean muestras i.i.d., se puede hacer \textit{thinning} de la traza y considerar solo una de cada \(n\) muestras.

\subsection{Algoritmo Metropolis-Hastings}

Se trata en realidad de un algoritmo marco (\sidecite[*-1]{metropolis1953equation}) a partir del cual han surgido muchos otros. Se basa en el uso de una distribución auxiliar simétrica\sidenote[][*2]{La función auxiliar se conoce como densidad propuesta o distribución de salto (\textit{proposal density} o \textit{jumping distribution}).} \(g(x'| x_t)\) que genera una propuesta para el estado siguiente a partir del estado actual (e.g. una distribución normal centrada en \(x_t\)). En cada iteración se genera una propuesta \(x'\), se calcula el \textit{ratio de aceptación} \(\alpha=f(x')/f(x_t)=p(x')/p(x_t)\), y finalmente se acepta el nuevo estado con probabilidad \(\min(1, \alpha)\). De esta forma se recorre aleatoriamente el espacio muestral, tendiendo a quedarse en las regiones de mayor densidad, y visitando ocasionalmente regiones poco probables con respecto a \(p(x)\).

\begin{figure}[h!]
  \includegraphics[width=.7\textwidth]{mh}
  \caption{Esquema del algoritmo Metropolis-Hastings~(extraída de \cite{jaewook2015metamodel}).}
\end{figure}

Una extensión directa es considerar distribuciones no simétricas para \(g\), donde el valor de \(\alpha\) se sustituye por
\[
\alpha=\frac{f(x')g(x_t \mid x')}{f(x_t)g(x'\mid x_t)},
\]
para que se sigan cumpliendo las ecuaciones de equilibrio de la cadena de Markov. Cabe destacar que cuando se rechaza una propuesta, se mantiene el estado anterior.

\subsection{Algoritmo de Gibbs}

Cuando el número de dimensiones es alto, encontrar una distribución de salto adecuada puede ser difícil (por ejemplo, puede ser necesario ajustar el tamaño del salto en cada dimensión individual). Un enfoque alternativo es el algoritmo de Gibbs (\sidecite[*-2]{geman1984stochastic}), que considera muestras separadas en cada dimensión. Concretamente, en cada paso se obtienen \textit{en orden} muestras unidimensionales (por ejemplo, mediante muestreo por rechazo) de cada variable mediante su distribución condicionada al resto de variables:
\[
p(x_k^{(i+1)} \mid x_{1}^{(i+1)}, \dots, x_{k-1}^{(i+1)}, x_{k+1}^{(i)}, \dots, x_L^{(i)}), \quad k=1,\dots,L.
\]
Notamos que usamos los valores más recientes de las variables disponibles en cada momento.

\subsection{Hamiltonian Monte Carlo}

Esta sub-familia de algoritmos introduce información del gradiente de la función objetivo para guiar la elección de propuestas de nuevos puntos. Se basan en la teoría de mecánica hamiltoniana, y en general consiguen aumentar la velocidad de convergencia debido a una elección inteligente de los nuevos puntos del espacio. Dentro de estos métodos destaca el \textit{No U-Turn Sampler} (NUTS, \sidecite[*-4]{hoffman2014no}), que tiene la ventaja añadida de que realiza un ajuste automático de hiperparámetros.

\subsection{Reversible-jump MCMC}

Estos métodos permiten variar la dimensionalidad subyacente del espacio muestral en cada propuesta (\sidecite[*0]{green1995reversible}). Teóricamente son útiles para muestrear de mixturas de gaussianas, procesos de Dirichlet o, en general, de cualquier distribución donde haya una variable latente que represente la dimensión del espacio. Sin embargo, no hay una implementación de referencia establecida.

\section{Affine-invariant ensemble sampler}

Una propiedad deseable de los algoritmos de muestreo es que sean \textit{invariantes por transformaciones afines}. Es decir, que vean dos distribuciones que difieren en una transformación afín, \(p(x)\) y \(p_{A, b}(Ax + b)\), como de igual dificultad. Esto es útil, por ejemplo, cuando uno tiene distribuciones muy asimétricas en las que es difícil encontrar un tamaño de paso que funcione bien en todas las dimensiones, pero que tras un reescalado se simplifican notablemente.

\begin{marginfigure}[*-6]
  \centering
  \includegraphics[width=.9\textwidth]{img/skewed}
  \caption*{Contornos de una distribución muy asimétrica en \(\R^2\).}
\end{marginfigure}

En general, un algoritmo de MCMC se puede describir como \(X(t+1)=R(X(t), \xi(t), p)\), donde \(X(t)\) es el estado de la cadena en el instante \(t\), \(p(x)\) es la distribución objetivo, y \(\xi(t)\) es una secuencia de variables aleatorias i.i.d. (asociadas a los procedimientos aleatorios de la cadena). La propiedad de invarianza afín se puede expresar como
\[
R(Ax+b, \xi(t), p_{A,b}) = AR(x, \xi(t), p) + b,
\]
para todo \(A, b\) y \(x\), y para casi todo \(\xi(t)\). Intuitivamente, esto quiere decir que si fijamos un generador aleatorio y ejecutamos dos veces el algoritmo, uno usando la densidad \(p\) empezando en \(X(0)\) y otro usando la densidad \(p_{A,b}\) con punto inicial \(Y(0)=AX(0)+b\), tendremos que \(Y(t)=AX(t)+b\).

En \sidetextcite[*-0.6]{goodman2010ensemble} los autores consideran un conjunto de cadenas (\textit{ensemble sampler}) con la propiedad de invarianza afín, inspirados en el algoritmo de optimización de Nelder-Mead. Específicamente, se tiene un conjunto \(X=(X_1, \dots, X_L)\) de \textit{walkers}, donde \(X_k(t)\) denota una cadena individual en el tiempo \(t\). En cada iteración, se usa una transformación invariante-afín contruida usando las posiciones \textit{actuales}\sidenote{En clara analogía con el algoritmo de Gibbs.} del resto de cadenas, mediante el llamado \textit{conjunto complementario}
\[
X_{-k}(t) = \{X_1(t+1), \dots, X_{k-1}(t+1), X_{k+1}(t), \dots, X_L(t)\}, \quad k=1,\dots, L.
\]

Para mantener la invarianza afín se hacen avanzar las cadenas de una en una y se utiliza un esquema Metropolis-Hastings de propuesta y aceptación para garantizar que se preserva la distribución conjunta del \textit{ensemble}. Se consideran principalmente dos tipos de movimientos:
\begin{itemize}
  \item \textit{Stretch move}. Para cada cadena \(1\leq k \leq L\) se elige aleatoriamente \(X_j \in X_{-k}(t)\), y la propuesta toma la forma
  \[
    X_k(t) \to Y = X_j + Z(X_k(t) - X_j),
  \]
  donde \(Z \stackrel{i.i.d.}{\sim} g(z)\) verificando\sidenote{Esta condición se impone para mantener la simetría de la propuesta.} \(g(z^{-1})=zg(z)\). Concretamente, se utiliza la densidad auxiliar parametrizada por \(a > 1\) dada por
  \[
  g(z) \propto \begin{cases}
    \frac{1}{\sqrt{z}}, & \text{si } z \in [a^{-1}, a],\\
    0, & \text{en otro caso.}
\end{cases}
  \]
  \begin{marginfigure}[*-1]
  \centering
  \includegraphics[width=.8\textwidth]{img/goodman_weare}
  \caption*{Esquema del \textit{stretch move} (extraído de \cite{goodman2010ensemble}).}
\end{marginfigure}

  La probabilidad de aceptación correspondiente para que se verifiquen las ecuaciones de equilibrio de las cadenas, suponiendo \(\R^N\) como espacio muestral, es
  \[
    \alpha = \min\left(1, \ Z^{N-1}\frac{p(Y)}{p(X_k(t))}\right).
  \]

  \item \textit{Walk move}. Para cada cadena \(1\leq k \leq L\) se elige un subconjunto aleatorio \(S_k \subseteq X_{-k}(t)\) con \(|S_k| \geq 2\), y se propone el movimiento
\[
X_k(t) \to Y = X_k(t) + W,
\]
donde \(W\) es una normal centrada en \(0\) y con la misma covarianza que la covarianza muestral de todas las cadenas de \(S_k\). La probabilidad de aceptación es directamente el ratio de aceptación de Metropolis, i.e., \(\min(1, p(Y)/p(X_k(t)))\).
\end{itemize}

\section{Librería emcee en Python}

El paquete \textit{emcee}\footnote[2]{\url{https://emcee.readthedocs.io/en/stable/}} de Python (\sidecite{foreman2013emcee}) proporciona una implementación paralela del algoritmo de \textcite{goodman2010ensemble}. La idea es dividir el conjunto de cadenas \(X\) en dos subconjuntos \(X^{(0)}\) y \(X^{(1)}\) de igual tamaño, y en cada iteración seguir un enfoque alternante de la siguiente forma:
\begin{enumerate}
  \item Se actualizan a la vez \textit{todas} las cadenas en \(X^{(0)}\) (mediante algunos de los movimientos disponibles, e.g. \textit{stretch move}) usando como conjunto complementario \(X^{(1)}\).
  \item Se utilizan las nuevas posiciones en \(X^{(0)}\) para actualizar \(X^{(1)}\).
\end{enumerate}
De esta forma se mantienen las ecuaciones de equilibrio de las cadenas, y cada uno de estos sub-pasos puede beneficiarse de un número arbitrario de procesadores\sidenote{Acotado superiormente por \(L/2\).} para su cómputo. También en el artículo citado se aconseja inicializar las cadenas en una bola gaussiana alrededor de un punto que se espere que tenga alta probabilidad según \(p(x)\).

Otra de las ventajas que ofrece este método frente a algoritmos clásicos es que solo requiere fijar unos pocos hiperparámetros (independientes de la dimensión subyacente), en lugar de los \(O(N^2)\) correspondientes, por ejemplo, a la matriz de covarianza de la distribución de salto \(N\)-dimensional en Metropolis-Hastings. Además, el paquete ha sido extendido para admitir otro tipo de movimientos paralelos, y el objetivo del proyecto es proporcionar un método de muestreo de propósito general que funcione correctamente en una clase amplia de problemas. Es muy usado por ejemplo en el campo de la astrofísica, y está bien establecido y testeado.

\printbibliography[title=Referencias]

\end{document}

### 2022-08-06 notes/notes.tex
Reorganize; edit paper; add mcmc notes

--- 2022-02-19 notes.tex
@@
 \newcommand{\bm}{\symbf}
+
+%% Scalar product
+\newcommand\dotprod[2]{\left\langle #1,\, #2 \right\rangle}
+
 \DeclareMathOperator{\var} {Var}
@@
 
+  \textbf{Why choose the parameter in the RKHS?}
+
+  - Small RKHS norm implies small variations and hence a higher degree of smoothness.
+Specifically, \(\|f\|_K\) controls how fast the function varies over the input space with respect to the geometry defined by the kernel:
+  \[
+  |f(t) - f(s)| = |\dotprod{f}{K_t - K_s}_K| \leq \|f\|_K \|K_t - K_s\|_K.
+  \]
+
+  - Note that in an RKHS the evaluation functional is continuous, so a small perturbation \(f\mapsto f + \delta f\) results in a small change in \(f(x)\) for all \(x\). Not equivalence classes.
+
+  - The model is \textit{linear} because of the correspondence between the RKHS and the linear span of the process (Loève's isometry).
+
 \input{notes-rkhs.tex}

### 2022-08-06 paper/paper.tex
Reorganize; edit paper; add mcmc notes

@@
 The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo methods.
+By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated through Markov chain Monte Carlo methods.
 Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
@@
 Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep{aguilera2013comparative}.
+Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
 At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
@@
 Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
-These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative non-parametric Bayesian approach to inference and prediction problems.
+These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems.
 Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
@@
 We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
@@
 where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
+In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
 In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
 Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
+Thus, some regularization or dimensionality reduction technique is needed for parameter estimation.
 
@@
 
-It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
+It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
 In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2018functional} for the case of functional logistic regression.
+These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional, berrendero2018use} for the case of functional logistic regression.
 Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
@@
 In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
-These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
+These points-of-impact selection models for functional predictors have also been considered in the  literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
+Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain.
 
@@
 Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of the centered process \(X\), and in what follows suppose that it is continuous.
-To construct the RKHS \(\Hcal(K)\) associated with the covariance function \(K\), we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
+To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
 \begin{equation}\label{eq:h0}
@@
 Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
+Specifically, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
 This isometry is essentially the completion of the correspondence
@@
 As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
-However, following Parzen's approach in his seminal work on time series analysis \citep[Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\).
-This notation often proves to be useful and convenient.
-
-\begin{comment}
-  Explicar mejor de dónde viene la notación de Parzen, y por qué se podría pensar en usarla para expresar un producto escalar (``congruence inner product'', e.g.
-4.17 en Parzen 1961b.
-¿Expresión en Torrecilla et al.
-y Kailath?).
-\end{comment}
+However, following Parzen's approach in his seminal work \citep[e.g.][Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with the image \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\).
+This notation, viewed as a formal extension of the inner product, often proves to be useful and convenient.
 
@@
 In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
 However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
@@
 Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of possible values is not very large in practice.
+Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of feasible values is not very large in practice.
 
@@
 \log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\bm Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
-& + (p+n)\log\sigma - \frac{1}{2}\log |G_\tau|,
+& + (p+n+2)\log\sigma - \frac{1}{2}\log |G_\tau|,
 \end{align*}
@@
   \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
-  \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
+  \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
 \end{align*}
@@
 
+\begin{comment}
+LO QUE NO SIRVA, GUARDAR PARA EL TFM.
+
+Reescribir label switching (nos pasa; lo que pasa que con p free se acentua).
+
+Ocurre porque el likelihood es simetrico: L(theta mid y) = L(thetaperm mid y) para cualquier permutacion (hay p! posibles) --> no-identificabilidad.
+
+"In these situations, parameter estimates and other output are inevitably aggregated
+across numerous replications, and because the labels of the enumerated classes are
+arbitrarily assigned, there is no guarantee that the same class would have the same
+label from replication to replication."
+
+From a Bayesian pointof view, this amounts to truncating the original prior distribution... (to verify the constraints)
+
+As a result, in any MCMC algorithm, labels of the
+components can permute multiple times between iterations of the sampler.
+Paradoxically, as noted by Celeux, Hurn and Robert (2000), Fruhwirth-Schnatter (2001),
+Jasra, Holmes and Stephens (2005) and Papastamoulis and Iliopoulos (2010), among others,
+label switching is a prerequisite for MCMC convergence. If there is no label switching it
+means that the sampler is not exploring all the modes of the posterior distribution of (1)
+
+"When sampling from the unconstrained posterior via MCMC methods, which component of the sampled parameter corresponds to which state is unknown due to potential label switching." Si usamos la marginal para haacer inferencia --> nonsensical answers (Stephens, 2000)
+
+Solución: añadir constraints de identificabilidad artificiales (ej relabeling, mantener ordenados, posprocesado).
+
+Imposing an identifiability constraint on the parameters breaks the sym-
+metry of the posterior distribution of the parameters, and so we might hope that it would allow us to perform sensible inference for the individual components of the mixture.
+Inference conditional on such a constraint may be performed by post-processing the sample (proposition 3.1 in Stephens 1997 phD thesis).
+
+OTRA OPCION: Elegir en cada iteración como parámetro para ordenar aquel que deja mayor separación (tras escalarlos todos a [0,1]): \verb@https://helda.helsinki.fi/bitstream/handle/10138/329485/Approximate_Bayesian_computation_for_finite_mixture_models.pdf?sequence=1@
+
+OTRA SOLUCION: "Adaptive Metropolis with Online Relabeling" --> The key step is the selection, after each proposal of a permutation (relabeling) of the proposed vector by minimizing a quadratic cost function
+$L_{t-1} (x)$ over all permutations of the proposed vector, where L is defined by
+$L_t(x) = (x - \mu)'\Sigma^{-1}(x - \mu)$
+This step forces the posterior sample to look as unimodal as possible.
+
+Citas: Stephens (2000), Papastamoulis \& Iliopoulos, 2010; Rodriguez \& Walker, 2014
+\end{comment}
+
 \paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
@@
 At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
+\incomment{Argue that, because of the label switching and multimodality, summarizing the posterior might be impractical, unhelpful, misleading, etc.
+So the best strategy is the posterior mean.}
 

### 2022-08-10 paper/paper.tex
Reorganize and start writing

@@
 
-FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see][]{ullah2013applications}.
+FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}.
 Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.

### 2022-08-10 thesis/src/abstract.tex
Reorganize and start writing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2022 Antonio Coín
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
  \begin{Large}
  \textsc{Abstract}
\end{Large}
\end{center}
%
\noindent We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's). These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm. The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest. By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated through Markov chain Monte Carlo methods. Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.\\

\noindent
\textsc{keywords:} functional data, linear regression, logistic regression, reproducing kernel Hilbert space, Bayesian inference, Markov chain Monte Carlo.

\incomment{add blank page}

### 2022-08-10 thesis/src/background-related-work.tex
Reorganize and start writing

--- 2022-08-06 paper/paper.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Copyright (c) 2022 Antonio Coín Castro
 %
-\pubyear{2022}
-\volume{TBA}
-\issue{TBA}
-\doi{0000}
-\arxiv{2010.00000}
-\firstpage{1}
-\lastpage{1}
+% This work is licensed under a
+% Creative Commons Attribution-ShareAlike 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{xcolor}
-\usepackage{bm}
-\usepackage{natbib}
-\usepackage[
-  colorlinks,
-  citecolor=blue,
-  urlcolor=blue,
-  filecolor=blue,
-  backref=page
-]{hyperref}
-\usepackage{graphicx}
-
-%% Comentarios
-\usepackage{todonotes}
-\usepackage{ulem}
-
-\startlocaldefs
-\numberwithin{equation}{section}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}{Proposition}[section]
-
-\renewcommand{\epsilon}{\varepsilon}
-
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\Hcal}{\ensuremath\mathcal{H}}
-\newcommand{\Var}{\ensuremath\operatorname{Var}}
-\definecolor{violet}{rgb}{0.7,0,0.6}
-
-%% Scalar product
-\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}
-
-%% Colors
-\definecolor{comment-red}{RGB}{240,30,70}
-
-%% Environment for colored comments
-\newenvironment{comment}[1][comment-red]
-{
-\noindent \color{#1}
-}
-{
-\color{black}
-}
-
-%% Inline comments
-\newcommand\incomment[2][comment-red]{\color{#1}[\textit{#2}]\color{black}}
-\endlocaldefs
-
-\begin{document}
-
-\begin{frontmatter}
-\title{Bayesian RKHS-based methods in functional regression\support{Support information of the article.}}
-\runtitle{Bayesian RKHS-based methods in functional regression}
-
-\begin{aug}
-\author{\fnms{First} \snm{Author}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{first@somewhere.com}},
-\author{\fnms{Second} \snm{Author}\thanksref{addr1,t3,m1,m2}\ead[label=e2]{second@somewhere.com}}
-\and
-\author{\fnms{Third} \snm{Author}\thanksref{addr2,t1,m2}%
-\ead[label=e3]{third@somewhere.com}%
-\ead[label=u1,url]{http://www.foo.com}}
-
-\runauthor{F. Author et al.}
-
-\address[addr1]{Address of the First and Second authors
-     Usually a few lines long
-    \printead{e1} % print email address of "e1"
-    \printead*{e2}
-}
-
-\address[addr2]{Address of the Third author
-    Usually a few lines long
-    Usually a few lines long
-    \printead{e3}
-    \printead{u1}
-}
-
-\thankstext{t1}{Some comment}
-\thankstext{t2}{First supporter of the project}
-\thankstext{t3}{Second supporter of the project}
-
-\end{aug}
-
-\begin{abstract}
-We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
-These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated through Markov chain Monte Carlo methods.
-Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
-\end{abstract}
-
-\begin{keyword}[class=MSC]
-\kwd[primary ]{62M20}
-\kwd[; secondary ]{62F15}
-\end{keyword}
-
-\begin{keyword}
-\kwd{functional data}
-\kwd{linear regression}
-\kwd{logistic regression}
-\kwd{reproducing kernel Hilbert space}
-\kwd{Bayesian inference}
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{sec:intro}
-
-The problem of predicting a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
-While in principle the functional data could be simply regarded as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
-As a consequence, numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum.
-A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}.
-
-FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see][]{ullah2013applications}.
-Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
-At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
-Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
-These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems.
-Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
-
-In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
-Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions.
-To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Background and related work}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{equation}\label{eq:l2-logistic-model}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
+  \Prob(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
 \end{equation}
@@
 
-\subsubsection{Organization of the paper}
+\section{Reproducing kernel Hilbert spaces}\label{sec:rkhs}
 
-In Section~\ref{sec:rkhs} we review the basics of the theory of RKHS's from a probabilistic point of view.
-Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models we propose.
-The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of computational details.
-Lastly, the conclusions drawn from this work are presented in Section~\ref{sec:conclusion}.
+For a more detailed account, see for example~\citet{berlinet2004reproducing}.
 
-\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
-
-The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of the centered process \(X\), and in what follows suppose that it is continuous.
+Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and without loss of generality we can assume \(\E[X(t)]=0\) for all \(t\in[0,1]\).
+Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\), and in what follows suppose that it is continuous.
 To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
 \begin{equation}\label{eq:h0}
-\Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
+\Hcal_0(K) = \left\{ f: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
 \end{equation}
-This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
-As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\).
+Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\).
+However, this space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\) for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
+
+\begin{proposition} \(\dotprod{\cdot}{\cdot}_K\) is an inner product.
+  \begin{proof}
+    Symmetry and linearity are clear from the definition, while non-negativity is a direct consequence of \(K\) being a covariance function, and thus positive semidefinite.
+  \end{proof}
+\end{proposition}
+
+Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\), turning it into a genuine Hilbert space.
+As it turns out, functions in this space satisfy the so-called \textit{reproducing property}, which can be easily checked from the definitions.
+
+\begin{proposition}
+  If \(f\in \Hcal(K)\), then \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(t \in [0, 1]\).
+  \begin{proof}
+
+\end{proof}
+\end{proposition}
+
+This is where the name ``reproducing kernel'' comes from.
 An important consequence is that \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
+
+\begin{outcomment}
+  A kernel defines an RKHS and vice-versa.
+\end{outcomment}
 
@@
 This notation, viewed as a formal extension of the inner product, often proves to be useful and convenient.
-
-\section{Bayesian methodology for RKHS-based functional regression models}\label{sec:methodology}
-
-In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
-For practical and computational reasons, the value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
-Thus, we will search for our functional parameter in the set
-\begin{equation}\label{eq:h0p}
-\Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_j \in \R, \ t_j \in [0, 1]\right\}.
-\end{equation}
-
-The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\alpha \in \Hcal_{0,p}(K)\) and that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\) when \(\alpha(\cdot)=\sum_j\beta_j K(t_j, \cdot)\).
-
-In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately.
-Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
-Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) merely by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
-Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of feasible values is not very large in practice.
-
-In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
-samples amounts to
-\begin{equation}\label{eq:bayes-theorem}
-  \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
-\end{equation}
-In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y | X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
-Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
-Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.
-
-\subsection{Functional linear regression}\label{sec:rkhs-linear-model}
-
-In the case of functional linear regression, the simplified RKHS model considered in this work is given by
-\begin{equation}\label{eq:rkhs-model-linear}
-  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
-\end{equation}
-where \(\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j, \cdot) \in \Hcal_{0,p}(K), \beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is a normally-distributed error term independent from \(X\), with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
-This model is essentially a finite-dimensional approximation from a functional perspective to the more general RKHS model that assumes \(\alpha \in \Hcal(K)\), proposed in~\citet{berrendero2019rkhs}.
-
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
-Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
-For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
-\begin{equation}\label{eq:rkhs-model-linear-2}
-  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
-\end{equation}
-
-It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
-Also, note that given the time instants \(t_j\), the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
-As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, this model could be easily extended to the case of several covariates via an expression of type \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
-In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
-
-\subsubsection{The bayesian approach: prior selection and posterior derivation}
-
-The prior distribution suggested for the parameter vector \(\theta \in \Theta_p\) is given by
-\begin{align}\label{eq:prior-linear}
-  \begin{split}
-  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
-  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2{\underbrace{\left(\mathcal X_\tau' \mathcal X_\tau + \eta I\right)}_{G_\tau}}^{-1}),
-\end{split}
-\end{align}
-where \(I\) is the identity matrix, \(\mathcal X_\tau\) is the data matrix \((X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
-On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
-In any event, the estimation of \(\alpha_0=\E[Y]\) is straightforward, so it could have been left out of the model altogether.
-On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-All in all, with a slight abuse of notation the proposed prior distribution becomes \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
-
-As for the posterior distribution, we only compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
-A standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
-
-\begin{proposition}
-Under the linear model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
-\begin{align*}
-\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\bm Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
-& + (p+n+2)\log\sigma - \frac{1}{2}\log |G_\tau|,
-\end{align*}
-where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
-\end{proposition}
-
-\subsubsection{Making predictions}
-
-In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
-
-\begin{description}
-  \item[Summarize-then-predict.] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
-  \[
-    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
-  \]
-  Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter.
-Although it contributes to measure the uncertainty in the approximations, its estimates are discarded in the final prediction.
-
-  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
-  \begin{equation}\label{eq:sampled-response-vector}
-  \bm Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
-  \end{equation}
-  Then, we can take the mean of all such simulated responses as a proxy for each response variable, that is,
-  \[
-  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
-  \]
-  This method differs from the previous one in that it takes into account the full approximate posterior distribution instead of summarizing it directly.
-
-  \item[Variable selection.] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
-Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
-After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
-\end{description}
-
-Note that these predictors can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
-As a consequence, what we have in practice is a single algorithm that can produce multiple predictors at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
-Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
-
-\subsection{Functional logistic regression}\label{sec:rkhs-logistic-model}
-
-In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, an RKHS model might be given  by the  equation
-\begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
-\end{equation}
-
-Indeed, note that this can be seen as a finite-dimensional approximation (but, still, with a functional interpretation) to the general RKHS functional logistic model proposed by these authors, which can be obtained by replacing \(\Hcal_{0,p}(K)\) with the whole RKHS space \(\Hcal(K)\).
-Now, if we aim at a classification problem, our strategy will be similar to that followed in the functional linear model: after incorporating the sample information, we can rewrite~\eqref{eq:rkhs-model-logistic} as
-\begin{equation}\label{eq:rkhs-model-logistic-2}
-Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
-\end{equation}
-with
-\begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
-  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
-\end{equation}
-where in turn \(\beta_j\in\R\) and \(t_j\in[0, 1]\).
-
-In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model.
-First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes~\citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for such model.
-Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
-Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case.
-However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
-\color{black}
-
-\subsubsection{The Bayesian approach: prior selection and posterior derivation}
-
-As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model.
-However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
-The posterior distribution is then derived after a routine calculation.
-
-\begin{proposition}
-Under the logistic model~\eqref{eq:rkhs-model-logistic-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
-\begin{align*}
-  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
-  \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
-\end{align*}
-Remember that \(\Psi_{X_i}(\alpha) = \sum_{j=1}^p \beta_j X_i(t_j)\).
-\end{proposition}
-
-\subsubsection{Making predictions}
-
-Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to a binary output in \(\{0, 1\}\).
-According to the usual criterion of minimizing the misclassification probability, the Bayes optimal rule is recovered by predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
-Nevertheless, for a more general cost function one could consider other criteria that would lead to evaluating whether \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold \(\gamma\in[0, 1]\).
-
-With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
-\[
-\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
-\]
-where \(\mathbb I\) is the indicator function (\(\mathbb I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
-The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
-
-On the other hand, the prediction method that takes into account the entire posterior approximation (i.e.
-the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
-Hence, there are primarily two possible outcomes:
-\begin{description}
-  \item[Average sampled probability.] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
-  \[
-    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
-  \]
-  \item[Average sampled response.] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
-\[
-  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
-\]
-In this case, each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
-Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).
-\end{description}
-
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
-
-\subsection{Model choice and validation}\label{sec:model-choice}
-
-In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
-
-\paragraph{The choice of \(\bm p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we could impose a prior distribution on \(p\) as well (e.g.
-a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
-In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
-Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
-In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, DIC, cross-validation, \ldots) to select its optimal value.
-This is equivalent to imposing a degenerate prior distribution on \(p\).
-Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}).
-\incomment{Reversible MCMC}
-
-\begin{comment}
-LO QUE NO SIRVA, GUARDAR PARA EL TFM.
-
-Reescribir label switching (nos pasa; lo que pasa que con p free se acentua).
-
-Ocurre porque el likelihood es simetrico: L(theta mid y) = L(thetaperm mid y) para cualquier permutacion (hay p! posibles) --> no-identificabilidad.
-
-"In these situations, parameter estimates and other output are inevitably aggregated
-across numerous replications, and because the labels of the enumerated classes are
-arbitrarily assigned, there is no guarantee that the same class would have the same
-label from replication to replication."
-
-From a Bayesian pointof view, this amounts to truncating the original prior distribution... (to verify the constraints)
-
-As a result, in any MCMC algorithm, labels of the
-components can permute multiple times between iterations of the sampler.
-Paradoxically, as noted by Celeux, Hurn and Robert (2000), Fruhwirth-Schnatter (2001),
-Jasra, Holmes and Stephens (2005) and Papastamoulis and Iliopoulos (2010), among others,
-label switching is a prerequisite for MCMC convergence. If there is no label switching it
-means that the sampler is not exploring all the modes of the posterior distribution of (1)
-
-"When sampling from the unconstrained posterior via MCMC methods, which component of the sampled parameter corresponds to which state is unknown due to potential label switching." Si usamos la marginal para haacer inferencia --> nonsensical answers (Stephens, 2000)
-
-Solución: añadir constraints de identificabilidad artificiales (ej relabeling, mantener ordenados, posprocesado).
-
-Imposing an identifiability constraint on the parameters breaks the sym-
-metry of the posterior distribution of the parameters, and so we might hope that it would allow us to perform sensible inference for the individual components of the mixture.
-Inference conditional on such a constraint may be performed by post-processing the sample (proposition 3.1 in Stephens 1997 phD thesis).
-
-OTRA OPCION: Elegir en cada iteración como parámetro para ordenar aquel que deja mayor separación (tras escalarlos todos a [0,1]): \verb@https://helda.helsinki.fi/bitstream/handle/10138/329485/Approximate_Bayesian_computation_for_finite_mixture_models.pdf?sequence=1@
-
-OTRA SOLUCION: "Adaptive Metropolis with Online Relabeling" --> The key step is the selection, after each proposal of a permutation (relabeling) of the proposed vector by minimizing a quadratic cost function
-$L_{t-1} (x)$ over all permutations of the proposed vector, where L is defined by
-$L_t(x) = (x - \mu)'\Sigma^{-1}(x - \mu)$
-This step forces the posterior sample to look as unimodal as possible.
-
-Citas: Stephens (2000), Papastamoulis \& Iliopoulos, 2010; Rodriguez \& Walker, 2014
-\end{comment}
-
-\paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
-\begin{itemize}
-  \item For the expected value \(b_0\) we propose to use the MLE of \(b\).
-Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation).
-Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
-  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
-  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
-This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
-\end{itemize}
-
-\paragraph{Summarizing the posterior.} On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
-At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-\incomment{Argue that, because of the label switching and multimodality, summarizing the posterior might be impractical, unhelpful, misleading, etc.
-So the best strategy is the posterior mean.}
-
-\paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
-For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution \(\bm Y^* \equiv \{\bm Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\bm Y^{(m)*}\}\) that fall below the real value of the statistic.
-They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=.75\textwidth]{img/ppc_linear}
-  \caption{Posterior predictive graphical checks on a fitted model.
-On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the approximate sampled responses, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
-\end{figure}
-
-\section{Experimental results}\label{sec:results}
-
-We show below the results for 10 random train/test partitions and 5-fold cross-validation to select the parameters \(p\) and \(\eta\).
-
-\subsection{Functional Linear Regression}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_rkhs}
-  \caption{Mean MSE for linear regression with simulated data following an underlying RKHS model.}\label{fig:reg-emcee-rkhs}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_l2}
-  \caption{Mean MSE for linear regression with simulated data following an underlying \(L^2\)-model.}\label{fig:reg-emcee-l2}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_real}
-  \caption{Mean MSE for linear regression with real data sets.}\label{fig:reg-emcee-real}
-\end{figure}
-
-\subsection{Functional Logistic Regression}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_rkhs}
-  \caption{Mean accuracy for logistic regression with simulated data following an underlying RKHS model.}\label{fig:clf-emcee-rkhs}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_l2}
-  \caption{Mean accuracy for logistic regression with simulated data following an underlying \(L^2\)-model.}\label{fig:clf-emcee-l2}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_real}
-  \caption{Mean accuracy for logistic regression with real data sets.}\label{fig:clf-emcee-real}
-\end{figure}
-
-\newpage
-
-\begin{comment}
-  \begin{itemize}
-    \item Comentar el método MCMC usado (\textit{emcee}) y por qué (vs.
-Metropolis u otros basados en gradiente como NUTS).
-Comentar que se lanzan varias cadenas en paralelo y luego se hace la media, para aumentar estabilidad.
-    \item Comentar que se usa el MLE de los parámetros (perturbado) como punto de partida del algoritmo MCMC.
-Comentar la forma de estimación de MLE (varias iteraciones combinando una búsqueda global con una búsqueda local).
-    \item Parámetros comunes de generación de datasets (granularidad de la malla, ...) y de la estrategia de ejecución (dónde se ejecuta, tiempos de ejecución, cuántas veces se repite para eliminar aleatoriedad, ...)
-    \item Métricas usadas para medir el error.
-    \item Comentar cómo se comparan las distintas estrategias de predicción, cuál parece más estable, etc.
-    \item Contar los algoritmos de comparación usados (elegir algunos).
-    \item Mostrar tablas de resultados y gráficas.
-En las gráficas, unificar el eje Y para los mismos modelos, y a ser posible, pintar media y desviación típica de varias ejecuciones (haciendo CV con distintos splits train/test).
-    \item Hacer un caso de estudio en el dataset RKHS, midiendo la bondad de las aproximaciones de los instantes de tiempo \(\hat t_j\) a los instantes reales.
-También estudiar si se recupera el número de componentes subyacente.
-    \item En una simulación concreta, mostrar la dependencia del error frente a distintos parámetros del sampler MCMC (por ejemplo, el número de cadenas usadas o el número de iteraciones).
-\end{itemize}
-\end{comment}
-
-\subsection{Simulation studies}
-
-\subsection{Application to real data}
-
-\section{Conclusion}\label{sec:conclusion}
-
-\begin{supplement}
-\stitle{Title of Supplement A}
-\sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}
-
-\begin{comment}
-  \begin{itemize}
-  \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado. Comentar también la posibilidad de transformar \(\tau\) a un espacio no acotado.
-  \item Comentar algunas decisiones de implementación: burn-in, thinning a la hora de hacer ppcs, ...
-  \item Gráficos de las trazas, de algunas posterioris concretas, etc. También de intervalos de confianza hdi3 y hdi97.
-\end{itemize}
-\end{comment}
-
-\end{supplement}
-
-\bibliographystyle{bib/ba}
-\bibliography{bib/bibliography}
-
-\begin{acks}[Acknowledgments]
-The authors wish to acknowledge the computational resources provided by the Centro de Computación Científica-Universidad Autónoma de Madrid (CCC-UAM), which were instrumental in performing the simulations in this paper.
-\end{acks}
-
-\end{document}

### 2022-08-10 thesis/src/bayesian-functional-regression.tex
Reorganize and start writing

--- 2022-08-06 paper/paper.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Copyright (c) 2022 Antonio Coín Castro
 %
-\pubyear{2022}
-\volume{TBA}
-\issue{TBA}
-\doi{0000}
-\arxiv{2010.00000}
-\firstpage{1}
-\lastpage{1}
-
+% This work is licensed under a
+% Creative Commons Attribution-ShareAlike 4.0 International License.
 %
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{xcolor}
-\usepackage{bm}
-\usepackage{natbib}
-\usepackage[
-  colorlinks,
-  citecolor=blue,
-  urlcolor=blue,
-  filecolor=blue,
-  backref=page
-]{hyperref}
-\usepackage{graphicx}
-
-%% Comentarios
-\usepackage{todonotes}
-\usepackage{ulem}
-
-\startlocaldefs
-\numberwithin{equation}{section}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}{Proposition}[section]
-
-\renewcommand{\epsilon}{\varepsilon}
-
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\Hcal}{\ensuremath\mathcal{H}}
-\newcommand{\Var}{\ensuremath\operatorname{Var}}
-\definecolor{violet}{rgb}{0.7,0,0.6}
-
-%% Scalar product
-\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}
-
-%% Colors
-\definecolor{comment-red}{RGB}{240,30,70}
-
-%% Environment for colored comments
-\newenvironment{comment}[1][comment-red]
-{
-\noindent \color{#1}
-}
-{
-\color{black}
-}
-
-%% Inline comments
-\newcommand\incomment[2][comment-red]{\color{#1}[\textit{#2}]\color{black}}
-\endlocaldefs
-
-\begin{document}
-
-\begin{frontmatter}
-\title{Bayesian RKHS-based methods in functional regression\support{Support information of the article.}}
-\runtitle{Bayesian RKHS-based methods in functional regression}
-
-\begin{aug}
-\author{\fnms{First} \snm{Author}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{first@somewhere.com}},
-\author{\fnms{Second} \snm{Author}\thanksref{addr1,t3,m1,m2}\ead[label=e2]{second@somewhere.com}}
-\and
-\author{\fnms{Third} \snm{Author}\thanksref{addr2,t1,m2}%
-\ead[label=e3]{third@somewhere.com}%
-\ead[label=u1,url]{http://www.foo.com}}
-
-\runauthor{F. Author et al.}
-
-\address[addr1]{Address of the First and Second authors
-     Usually a few lines long
-    \printead{e1} % print email address of "e1"
-    \printead*{e2}
-}
-
-\address[addr2]{Address of the Third author
-    Usually a few lines long
-    Usually a few lines long
-    \printead{e3}
-    \printead{u1}
-}
-
-\thankstext{t1}{Some comment}
-\thankstext{t2}{First supporter of the project}
-\thankstext{t3}{Second supporter of the project}
-
-\end{aug}
-
-\begin{abstract}
-We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
-These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated through Markov chain Monte Carlo methods.
-Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
-\end{abstract}
-
-\begin{keyword}[class=MSC]
-\kwd[primary ]{62M20}
-\kwd[; secondary ]{62F15}
-\end{keyword}
-
-\begin{keyword}
-\kwd{functional data}
-\kwd{linear regression}
-\kwd{logistic regression}
-\kwd{reproducing kernel Hilbert space}
-\kwd{Bayesian inference}
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{sec:intro}
-
-The problem of predicting a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
-While in principle the functional data could be simply regarded as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
-As a consequence, numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum.
-A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}.
-
-FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see][]{ullah2013applications}.
-Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
-At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
-Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
-These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems.
-Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
-
-In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
-Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions.
-To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
-
-\subsubsection{\(L^2\)-models, shortcomings and alternatives}
-
- The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}.
-It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
-\begin{equation}\label{eq:l2-linear-model}
-Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
-\end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
-Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
-Thus, some regularization or dimensionality reduction technique is needed for parameter estimation.
-
-A similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
-\begin{equation}\label{eq:l2-logistic-model}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
-\end{equation}
-where \(\alpha_0 \in \R\) and \(\beta \in L^2[0, 1]\).
-In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2018functional}.
-
-It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
-As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional, berrendero2018use} for the case of functional logistic regression.
-Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
-
-A major aim of this work is to motivate these recently-proposed models inside the functional framework, while also providing efficient techniques to apply them in practice.
-Our main contribution is the proposal of a Bayesian approach to parameter estimation within the aforementioned RKHS models, in which a prior distribution is imposed on the unknown functional parameter to obtain a posterior distribution after seeing the data.
-Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-A similar Bayesian scheme has recently been explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
-
-Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion.
-As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
-In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
-These points-of-impact selection models for functional predictors have also been considered in the  literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain.
-
-\subsubsection{Organization of the paper}
-
-In Section~\ref{sec:rkhs} we review the basics of the theory of RKHS's from a probabilistic point of view.
-Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models we propose.
-The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of computational details.
-Lastly, the conclusions drawn from this work are presented in Section~\ref{sec:conclusion}.
-
-\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
-
-The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of the centered process \(X\), and in what follows suppose that it is continuous.
-To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
-\begin{equation}\label{eq:h0}
-\Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
-\end{equation}
-This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
-As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\).
-An important consequence is that \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
-
-Note that he covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-An integral operator closely related to this kernel is the so-called covariance operator, namely \(\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds\) for \(f \in L^2[0, 1]\), which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}.
-It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
-In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
-Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
-
-Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
-This isometry is essentially the completion of the correspondence
-  \begin{equation}\label{eq:loeves-isometry}
-  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
-  \end{equation}
-and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
-Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
-As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
-However, following Parzen's approach in his seminal work \citep[e.g.][Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with the image \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\).
-This notation, viewed as a formal extension of the inner product, often proves to be useful and convenient.
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{align*}
-\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\bm Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
+\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\symbf Y- \alpha_0\symbf{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
 & + (p+n+2)\log\sigma - \frac{1}{2}\log |G_\tau|,
 \end{align*}
-where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
+where \(\symbf Y=(Y_1,\dots,Y_n)'\) and \(\symbf{1}\) is an \(n\)-dimensional vector of ones.
 \end{proposition}
@@
   \begin{equation}\label{eq:sampled-response-vector}
-  \bm Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
+  \symbf Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
   \end{equation}
@@
 
-\paragraph{The choice of \(\bm p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
+\paragraph{The choice of \(\symbf p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
 While theoretically we could impose a prior distribution on \(p\) as well (e.g.
@@
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution \(\bm Y^* \equiv \{\bm Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\bm Y^{(m)*}\}\) that fall below the real value of the statistic.
+Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\symbf Y^*)\) of the posterior predictive distribution \(\symbf Y^* \equiv \{\symbf Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
+In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the real value of the statistic.
 They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.
@@
   \centering
-  \includegraphics[width=.75\textwidth]{img/ppc_linear}
+  %\includegraphics[width=.75\textwidth]{img/ppc_linear}
   \caption{Posterior predictive graphical checks on a fitted model.
@@
 \end{figure}
-
-\section{Experimental results}\label{sec:results}
-
-We show below the results for 10 random train/test partitions and 5-fold cross-validation to select the parameters \(p\) and \(\eta\).
-
-\subsection{Functional Linear Regression}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_rkhs}
-  \caption{Mean MSE for linear regression with simulated data following an underlying RKHS model.}\label{fig:reg-emcee-rkhs}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_l2}
-  \caption{Mean MSE for linear regression with simulated data following an underlying \(L^2\)-model.}\label{fig:reg-emcee-l2}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_real}
-  \caption{Mean MSE for linear regression with real data sets.}\label{fig:reg-emcee-real}
-\end{figure}
-
-\subsection{Functional Logistic Regression}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_rkhs}
-  \caption{Mean accuracy for logistic regression with simulated data following an underlying RKHS model.}\label{fig:clf-emcee-rkhs}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_l2}
-  \caption{Mean accuracy for logistic regression with simulated data following an underlying \(L^2\)-model.}\label{fig:clf-emcee-l2}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_real}
-  \caption{Mean accuracy for logistic regression with real data sets.}\label{fig:clf-emcee-real}
-\end{figure}
-
-\newpage
-
-\begin{comment}
-  \begin{itemize}
-    \item Comentar el método MCMC usado (\textit{emcee}) y por qué (vs.
-Metropolis u otros basados en gradiente como NUTS).
-Comentar que se lanzan varias cadenas en paralelo y luego se hace la media, para aumentar estabilidad.
-    \item Comentar que se usa el MLE de los parámetros (perturbado) como punto de partida del algoritmo MCMC.
-Comentar la forma de estimación de MLE (varias iteraciones combinando una búsqueda global con una búsqueda local).
-    \item Parámetros comunes de generación de datasets (granularidad de la malla, ...) y de la estrategia de ejecución (dónde se ejecuta, tiempos de ejecución, cuántas veces se repite para eliminar aleatoriedad, ...)
-    \item Métricas usadas para medir el error.
-    \item Comentar cómo se comparan las distintas estrategias de predicción, cuál parece más estable, etc.
-    \item Contar los algoritmos de comparación usados (elegir algunos).
-    \item Mostrar tablas de resultados y gráficas.
-En las gráficas, unificar el eje Y para los mismos modelos, y a ser posible, pintar media y desviación típica de varias ejecuciones (haciendo CV con distintos splits train/test).
-    \item Hacer un caso de estudio en el dataset RKHS, midiendo la bondad de las aproximaciones de los instantes de tiempo \(\hat t_j\) a los instantes reales.
-También estudiar si se recupera el número de componentes subyacente.
-    \item En una simulación concreta, mostrar la dependencia del error frente a distintos parámetros del sampler MCMC (por ejemplo, el número de cadenas usadas o el número de iteraciones).
-\end{itemize}
-\end{comment}
-
-\subsection{Simulation studies}
-
-\subsection{Application to real data}
-
-\section{Conclusion}\label{sec:conclusion}
-
-\begin{supplement}
-\stitle{Title of Supplement A}
-\sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}
-
-\begin{comment}
-  \begin{itemize}
-  \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado. Comentar también la posibilidad de transformar \(\tau\) a un espacio no acotado.
-  \item Comentar algunas decisiones de implementación: burn-in, thinning a la hora de hacer ppcs, ...
-  \item Gráficos de las trazas, de algunas posterioris concretas, etc. También de intervalos de confianza hdi3 y hdi97.
-\end{itemize}
-\end{comment}
-
-\end{supplement}
-
-\bibliographystyle{bib/ba}
-\bibliography{bib/bibliography}
-
-\begin{acks}[Acknowledgments]
-The authors wish to acknowledge the computational resources provided by the Centro de Computación Científica-Universidad Autónoma de Madrid (CCC-UAM), which were instrumental in performing the simulations in this paper.
-\end{acks}
-
-\end{document}

### 2022-08-10 thesis/src/introduction.tex
Reorganize and start writing

--- 2022-06-24 paper/paper.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Copyright (c) 2022 Antonio Coín Castro
 %
-\pubyear{2022}
-\volume{TBA}
-\issue{TBA}
-\doi{0000}
-\arxiv{2010.00000}
-\firstpage{1}
-\lastpage{1}
+% This work is licensed under a
+% Creative Commons Attribution-ShareAlike 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{xcolor}
-\usepackage{bm}
-\usepackage{natbib}
-\usepackage[
-  colorlinks,
-  citecolor=blue,
-  urlcolor=blue,
-  filecolor=blue,
-  backref=page
-]{hyperref}
-\usepackage{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% Comentarios
-\usepackage{todonotes}
-\usepackage{ulem}
-
-\startlocaldefs
-\numberwithin{equation}{section}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}{Proposition}[section]
-
-\renewcommand{\epsilon}{\varepsilon}
-
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\Hcal}{\ensuremath\mathcal{H}}
-\newcommand{\Var}{\ensuremath\operatorname{Var}}
-\definecolor{violet}{rgb}{0.7,0,0.6}
-
-%% Scalar product
-\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}
-
-%% Colors
-\definecolor{comment-red}{RGB}{240,30,70}
-
-%% Environment for colored comments
-\newenvironment{comment}[1][comment-red]
-{
-\noindent \color{#1}
-}
-{
-\color{black}
-}
-
-%% Inline comments
-\newcommand\incomment[2][comment-red]{\color{#1}[\textit{#2}]\color{black}}
-\endlocaldefs
-
-\begin{document}
-
-\begin{frontmatter}
-\title{Bayesian RKHS-based methods in functional regression\support{Support information of the article.}}
-\runtitle{Bayesian RKHS-based methods in functional regression}
-
-\begin{aug}
-\author{\fnms{First} \snm{Author}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{first@somewhere.com}},
-\author{\fnms{Second} \snm{Author}\thanksref{addr1,t3,m1,m2}\ead[label=e2]{second@somewhere.com}}
-\and
-\author{\fnms{Third} \snm{Author}\thanksref{addr2,t1,m2}%
-\ead[label=e3]{third@somewhere.com}%
-\ead[label=u1,url]{http://www.foo.com}}
-
-\runauthor{F. Author et al.}
-
-\address[addr1]{Address of the First and Second authors
-     Usually a few lines long
-    \printead{e1} % print email address of "e1"
-    \printead*{e2}
-}
-
-\address[addr2]{Address of the Third author
-    Usually a few lines long
-    Usually a few lines long
-    \printead{e3}
-    \printead{u1}
-}
-
-\thankstext{t1}{Some comment}
-\thankstext{t2}{First supporter of the project}
-\thankstext{t3}{Second supporter of the project}
-
-\end{aug}
-
-\begin{abstract}
-We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
-These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated from Markov Chain Monte Carlo methods.
-Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
-\end{abstract}
-
-\begin{keyword}[class=MSC]
-\kwd[primary ]{62M20}
-\kwd[; secondary ]{62F15}
-\end{keyword}
-
-\begin{keyword}
-\kwd{functional data}
-\kwd{linear regression}
-\kwd{logistic regression}
-\kwd{reproducing kernel Hilbert space}
-\kwd{Bayesian inference}
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{sec:intro}
+% \begin{center}
+%   \begin{minipage}{0.5\textwidth}
+%     \begin{small}
+%       \emph{In which the reasons for creating this package are laid bare for the
+%       whole world to see and we encounter some usage guidelines.}
+%     \end{small}
+%   \end{minipage}
+%   \vspace{0.5cm}
+% \end{center}
 
@@
 
-FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see][]{ullah2013applications}.
+FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}.
 Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep{aguilera2013comparative}.
+Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
 At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
@@
 We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a ``labeled'' data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
 
-\subsubsection{\(L^2\)-models, shortcomings and alternatives}
+\section{Objectives and scope}
 
- The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}.
-It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
-\begin{equation}\label{eq:l2-linear-model}
-Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
-\end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be very hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
-Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
-
-A similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
-\begin{equation}\label{eq:l2-logistic-model}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
-\end{equation}
-where \(\alpha_0 \in \R\) and \(\beta \in L^2[0, 1]\).
-In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2018functional}.
-
-It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
-As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use, berrendero2018functional} for the case of functional logistic regression.
-Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
-
-A major aim of this work is to motivate these recently-proposed models inside the functional framework, while also providing efficient techniques to apply them in practice.
-Our main contribution is the proposal of a Bayesian approach to parameter estimation within the aforementioned RKHS models, in which a prior distribution is imposed on the unknown functional parameter to obtain a posterior distribution after seeing the data.
-Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-A similar Bayesian scheme has recently been explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
-
-Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion.
-As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
-In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
-These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-
-\subsubsection{Organization of the paper}
+\section{Structure overview}
 
@@
 Lastly, the conclusions drawn from this work are presented in Section~\ref{sec:conclusion}.
-
-\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
-
-The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of the centered process \(X\), and in what follows suppose that it is continuous.
-To construct the RKHS \(\Hcal(K)\) associated with the covariance function \(K\), we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
-\begin{equation}\label{eq:h0}
-\Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
-\end{equation}
-This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
-As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\).
-An important consequence is that \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
-
-Note that he covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-An integral operator closely related to this kernel is the so-called covariance operator, namely \(\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds\) for \(f \in L^2[0, 1]\), which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}.
-It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
-In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
-Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
-
-Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
-This isometry is essentially the completion of the correspondence
-  \begin{equation}\label{eq:loeves-isometry}
-  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
-  \end{equation}
-and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
-Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
-As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
-However, following Parzen's approach in his seminal work on time series analysis \citep[Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\).
-This notation often proves to be useful and convenient.
-
-\begin{comment}
-  Explicar mejor de dónde viene la notación de Parzen, y por qué se podría pensar en usarla para expresar un producto escalar (``congruence inner product'', e.g.
-4.17 en Parzen 1961b.
-¿Expresión en Torrecilla et al.
-y Kailath?).
-\end{comment}
-
-\section{Bayesian methodology for RKHS-based functional regression models}\label{sec:methodology}
-
-In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
-For practical and computational reasons, the value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
-Thus, we will search for our functional parameter in the set
-\begin{equation}\label{eq:h0p}
-\Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_j \in \R, \ t_j \in [0, 1]\right\}.
-\end{equation}
-
-The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\alpha \in \Hcal_{0,p}(K)\) and that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\) when \(\alpha(\cdot)=\sum_j\beta_j K(t_j, \cdot)\).
-
-In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately.
-Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
-Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) merely by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
-Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of possible values is not very large in practice.
-
-In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
-samples amounts to
-\begin{equation}\label{eq:bayes-theorem}
-  \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
-\end{equation}
-In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y | X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
-Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
-Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.
-
-\subsection{Functional linear regression}\label{sec:rkhs-linear-model}
-
-In the case of functional linear regression, the simplified RKHS model considered in this work is given by
-\begin{equation}\label{eq:rkhs-model-linear}
-  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
-\end{equation}
-where \(\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j, \cdot) \in \Hcal_{0,p}(K), \beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is a normally-distributed error term independent from \(X\), with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
-This model is essentially a finite-dimensional approximation from a functional perspective to the more general RKHS model that assumes \(\alpha \in \Hcal(K)\), proposed in~\citet{berrendero2019rkhs}.
-
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
-Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
-For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
-\begin{equation}\label{eq:rkhs-model-linear-2}
-  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
-\end{equation}
-
-It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
-Also, note that given the time instants \(t_j\), the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
-As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, this model could be easily extended to the case of several covariates via an expression of type \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
-In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
-
-\subsubsection{The bayesian approach: prior selection and posterior derivation}
-
-The prior distribution suggested for the parameter vector \(\theta \in \Theta_p\) is given by
-\begin{align}\label{eq:prior-linear}
-  \begin{split}
-  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
-  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2{\underbrace{\left(\mathcal X_\tau' \mathcal X_\tau + \eta I\right)}_{G_\tau}}^{-1}),
-\end{split}
-\end{align}
-where \(I\) is the identity matrix, \(\mathcal X_\tau\) is the data matrix \((X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
-On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
-In any event, the estimation of \(\alpha_0=\E[Y]\) is straightforward, so it could have been left out of the model altogether.
-On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-All in all, with a slight abuse of notation the proposed prior distribution becomes \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
-
-As for the posterior distribution, we only compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
-A standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
-
-\begin{proposition}
-Under the linear model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
-\begin{align*}
-\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\bm Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
-& + (p+n)\log\sigma - \frac{1}{2}\log |G_\tau|,
-\end{align*}
-where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
-\end{proposition}
-
-\subsubsection{Making predictions}
-
-In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
-
-\begin{description}
-  \item[Summarize-then-predict.] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
-  \[
-    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
-  \]
-  Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter.
-Although it contributes to measure the uncertainty in the approximations, its estimates are discarded in the final prediction.
-
-  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
-  \begin{equation}\label{eq:sampled-response-vector}
-  \bm Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
-  \end{equation}
-  Then, we can take the mean of all such simulated responses as a proxy for each response variable, that is,
-  \[
-  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
-  \]
-  This method differs from the previous one in that it takes into account the full approximate posterior distribution instead of summarizing it directly.
-
-  \item[Variable selection.] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
-Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
-After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
-\end{description}
-
-Note that these predictors can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
-As a consequence, what we have in practice is a single algorithm that can produce multiple predictors at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
-Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
-
-\subsection{Functional logistic regression}\label{sec:rkhs-logistic-model}
-
-In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, an RKHS model might be given  by the  equation
-\begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
-\end{equation}
-
-Indeed, note that this can be seen as a finite-dimensional approximation (but, still, with a functional interpretation) to the general RKHS functional logistic model proposed by these authors, which can be obtained by replacing \(\Hcal_{0,p}(K)\) with the whole RKHS space \(\Hcal(K)\).
-Now, if we aim at a classification problem, our strategy will be similar to that followed in the functional linear model: after incorporating the sample information, we can rewrite~\eqref{eq:rkhs-model-logistic} as
-\begin{equation}\label{eq:rkhs-model-logistic-2}
-Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
-\end{equation}
-with
-\begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
-  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
-\end{equation}
-where in turn \(\beta_j\in\R\) and \(t_j\in[0, 1]\).
-
-In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model.
-First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes~\citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for such model.
-Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
-Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case.
-However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
-\color{black}
-
-\subsubsection{The Bayesian approach: prior selection and posterior derivation}
-
-As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model.
-However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
-The posterior distribution is then derived after a routine calculation.
-
-\begin{proposition}
-Under the logistic model~\eqref{eq:rkhs-model-logistic-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
-\begin{align*}
-  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
-  \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
-\end{align*}
-Remember that \(\Psi_{X_i}(\alpha) = \sum_{j=1}^p \beta_j X_i(t_j)\).
-\end{proposition}
-
-\subsubsection{Making predictions}
-
-Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to a binary output in \(\{0, 1\}\).
-According to the usual criterion of minimizing the misclassification probability, the Bayes optimal rule is recovered by predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
-Nevertheless, for a more general cost function one could consider other criteria that would lead to evaluating whether \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold \(\gamma\in[0, 1]\).
-
-With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
-\[
-\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
-\]
-where \(\mathbb I\) is the indicator function (\(\mathbb I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
-The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
-
-On the other hand, the prediction method that takes into account the entire posterior approximation (i.e.
-the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
-Hence, there are primarily two possible outcomes:
-\begin{description}
-  \item[Average sampled probability.] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
-  \[
-    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
-  \]
-  \item[Average sampled response.] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
-\[
-  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
-\]
-In this case, each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
-Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).
-\end{description}
-
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
-
-\subsection{Model choice and validation}\label{sec:model-choice}
-
-In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
-
-\paragraph{The choice of \(\bm p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we could impose a prior distribution on \(p\) as well (e.g.
-a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
-In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
-Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
-In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, DIC, cross-validation, \ldots) to select its optimal value.
-This is equivalent to imposing a degenerate prior distribution on \(p\).
-Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}).
-\incomment{Reversible MCMC}
-
-\paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
-\begin{itemize}
-  \item For the expected value \(b_0\) we propose to use the MLE of \(b\).
-Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation).
-Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
-  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
-  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
-This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
-\end{itemize}
-
-\paragraph{Summarizing the posterior.} On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
-At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-
-\paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
-For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution \(\bm Y^* \equiv \{\bm Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\bm Y^{(m)*}\}\) that fall below the real value of the statistic.
-They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=.75\textwidth]{img/ppc_linear}
-  \caption{Posterior predictive graphical checks on a fitted model.
-On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the approximate sampled responses, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
-\end{figure}
-
-\section{Experimental results}\label{sec:results}
-
-We show below the results for 10 random train/test partitions and 5-fold cross-validation to select the parameters \(p\) and \(\eta\).
-
-\subsection{Functional Linear Regression}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_rkhs}
-  \caption{Mean MSE for linear regression with simulated data following an underlying RKHS model.}\label{fig:reg-emcee-rkhs}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_l2}
-  \caption{Mean MSE for linear regression with simulated data following an underlying \(L^2\)-model.}\label{fig:reg-emcee-l2}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/reg_emcee_real}
-  \caption{Mean MSE for linear regression with real data sets.}\label{fig:reg-emcee-real}
-\end{figure}
-
-\subsection{Functional Logistic Regression}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_rkhs}
-  \caption{Mean accuracy for logistic regression with simulated data following an underlying RKHS model.}\label{fig:clf-emcee-rkhs}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_l2}
-  \caption{Mean accuracy for logistic regression with simulated data following an underlying \(L^2\)-model.}\label{fig:clf-emcee-l2}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{img/clf_emcee_real}
-  \caption{Mean accuracy for logistic regression with real data sets.}\label{fig:clf-emcee-real}
-\end{figure}
-
-\newpage
-
-\begin{comment}
-  \begin{itemize}
-    \item Comentar el método MCMC usado (\textit{emcee}) y por qué (vs.
-Metropolis u otros basados en gradiente como NUTS).
-Comentar que se lanzan varias cadenas en paralelo y luego se hace la media, para aumentar estabilidad.
-    \item Comentar que se usa el MLE de los parámetros (perturbado) como punto de partida del algoritmo MCMC.
-Comentar la forma de estimación de MLE (varias iteraciones combinando una búsqueda global con una búsqueda local).
-    \item Parámetros comunes de generación de datasets (granularidad de la malla, ...) y de la estrategia de ejecución (dónde se ejecuta, tiempos de ejecución, cuántas veces se repite para eliminar aleatoriedad, ...)
-    \item Métricas usadas para medir el error.
-    \item Comentar cómo se comparan las distintas estrategias de predicción, cuál parece más estable, etc.
-    \item Contar los algoritmos de comparación usados (elegir algunos).
-    \item Mostrar tablas de resultados y gráficas.
-En las gráficas, unificar el eje Y para los mismos modelos, y a ser posible, pintar media y desviación típica de varias ejecuciones (haciendo CV con distintos splits train/test).
-    \item Hacer un caso de estudio en el dataset RKHS, midiendo la bondad de las aproximaciones de los instantes de tiempo \(\hat t_j\) a los instantes reales.
-También estudiar si se recupera el número de componentes subyacente.
-    \item En una simulación concreta, mostrar la dependencia del error frente a distintos parámetros del sampler MCMC (por ejemplo, el número de cadenas usadas o el número de iteraciones).
-\end{itemize}
-\end{comment}
-
-\subsection{Simulation studies}
-
-\subsection{Application to real data}
-
-\section{Conclusion}\label{sec:conclusion}
-
-\begin{supplement}
-\stitle{Title of Supplement A}
-\sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}
-
-\begin{comment}
-  \begin{itemize}
-  \item Comentar el uso de \(\log\sigma\) en lugar de \(\sigma^2\) para que el espacio de búsqueda sea no acotado. Comentar también la posibilidad de transformar \(\tau\) a un espacio no acotado.
-  \item Comentar algunas decisiones de implementación: burn-in, thinning a la hora de hacer ppcs, ...
-  \item Gráficos de las trazas, de algunas posterioris concretas, etc. También de intervalos de confianza hdi3 y hdi97.
-\end{itemize}
-\end{comment}
-
-\end{supplement}
-
-\bibliographystyle{bib/ba}
-\bibliography{bib/bibliography}
-
-\begin{acks}[Acknowledgments]
-The authors wish to acknowledge the computational resources provided by the Centro de Computación Científica-Universidad Autónoma de Madrid (CCC-UAM), which were instrumental in performing the simulations in this paper.
-\end{acks}
-
-\end{document}

### 2022-08-23 thesis/src/background-related-work.tex
Finish background section

@@
 
+\let\epsilon\varepsilon
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Background and related work}
+\chapter{Background and related work}\label{ch:background}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsubsection{\(L^2\)-models, shortcomings and alternatives}
+FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}.
+Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
+Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
+At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
+Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
+Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
+These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems.
+Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
 
- The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}.
-It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
+\subsection*{\(L^2\)-models}
+
+ The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}.
+It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
 \begin{equation}\label{eq:l2-linear-model}
@@
 \end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
-Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
-Thus, some regularization or dimensionality reduction technique is needed for parameter estimation.
+where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\) and \(\E[X(t)\epsilon]=0\) almost everwhere; and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
+A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\mathcal K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\mathcal K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
+This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\mathcal K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).
 
-A similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
+Technical details aside, the inference on \(\beta\) is also hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
+Although it can be tempting to simply discretize the observed values on a grid and proceed with standard multiple linear regression, this would result in an under-determined model in which the estimated parameters lose meaning (essentially because the number of available parameters is infinite, but the number of equations is finite).
+Thus, some regularization or dimensionality reduction techniques are needed for parameter estimation; see \citet{reiss2017methods} for a summary of several widespread methods.
+
+A common inference strategy is to expand both \(X\) and \(\beta\) on a certain data-driven orthonormal basis of \(L^2[0,1]\), say \(\{\phi_j\}\), up to a certain value \(p\in\N\):
+\[
+  X(t)=\sum_{j=1}^p x_j \phi_j(t), \quad \beta(t) = \sum_{j=1}^p \beta_j \phi_j(t).
+\]
+In this way, model~\eqref{eq:l2-linear-model} simplifies to
+\[
+Y = \alpha_0 + \sum_{j=1}^p \beta_j x_j,
+\]
+which can then be solved in the usual manner.
+For example, when the chosen basis is composed of eigenfunctions of the covariance operator \(\mathcal K\), this method is known as Functional Principal Component Regression (FPCR).
+Another class of methods focus on selecting an \textit{a priori} basis for \(\beta\) (e.g.
+a spline basis) and solving a penalized least squares problem.
+In particular, for a truncated basis representation of the form \(\beta(t)=\phi(t)'b\), with \(\phi(t)=(\phi_1(t), \dots, \phi_p(t))'\) and \(b\in\R^p\), a typical minimization problem would be
+\[
+\argmin_{(\alpha_0, b)\in\R\times\R^p} \ \left\{ \sum_{i=1}^n \left( y_i - \alpha_0 - \int_0^1 x_i(t)[\phi(t)'b]\, dt\right)^2 + \lambda \Omega(b) \right\},
+\]
+where \(\Omega\) is some penalty function and \(\lambda>0\) is a regularization parameter.
+
+Turning our attention to logistic regression, a similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
 \begin{equation}\label{eq:l2-logistic-model}
@@
 
-It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
+\subsection*{RKHS models}
+
+In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
+It turns out that in both linear and logistic scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
+In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
 These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional, berrendero2018use} for the case of functional logistic regression.
-Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
 
-A major aim of this work is to motivate these recently-proposed models inside the functional framework, while also providing efficient techniques to apply them in practice.
+There are other proposals for models that change the habitat of the functional parameter, and there are even some in which \(\beta\) is supposed to live in a RKHS, most notably \citet{yuan2010reproducing}.
+However, these are arbitrary RKHS's that do not directly exploit the relation with the process \(X\) (though the authors use said connection to derive some asymptotic properties of their estimators), and therefore the approach is somewhat different.
+Incidentally, the RKHS associated with \(X\) also has some interesting properties that contribute to shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
+
+A major aim of this work is to motivate these recently-proposed RKHS models inside the functional framework, while also providing efficient techniques to apply them in practice.
 Our main contribution is the proposal of a Bayesian approach to parameter estimation within the aforementioned RKHS models, in which a prior distribution is imposed on the unknown functional parameter to obtain a posterior distribution after seeing the data.
-Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
+Although setting a prior distribution on a functional space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Chapter~\ref{ch:bayesian} for details).
 A similar Bayesian scheme has recently been explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
@@
 As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
+As will become apparent in the forthcoming chapters, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
 In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
 These points-of-impact selection models for functional predictors have also been considered in the  literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain.
+Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain (a sort of ``region selection'' algorithm).
+
+\subsection*{Bayesian inference}
+
+\begin{outcomment}
+  Rellenar?
+\end{outcomment}
 
@@
 
-For a more detailed account, see for example~\citet{berlinet2004reproducing}.
+In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces (RKHS's).
+Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory.
+For a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}.
 
-Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and without loss of generality we can assume \(\E[X(t)]=0\) for all \(t\in[0,1]\).
-Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\), and in what follows suppose that it is continuous.
-To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
+We start by defining the concept of kernel functions, which are the foundation of these spaces.
+Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity.
+
+\begin{definition}
+  We say that a function of two variables \(K: \mathcal T \times \mathcal T \to \R\) is positive semidefinite\footnote{Sometimes in the literature these functions are known simply as positive definite.} if
+  \[
+    \sum_{i,j=1}^p a_ia_jK(t_i, t_j) \geq 0
+  \]
+  for any \(p\in\N\), any \((a_1,\dots, a_p)\in \R^p\) and any \((t_1,\dots, t_p)\in\mathcal T^p\).
+Note that this is equivalent to saying that the matrix \((K(t_i, t_j))_{i,j}\) is positive semidefinite for any choice of \(p\in\N\) and \((t_1,\dots, t_p)\in\mathcal T^p\).
+\end{definition}
+
+We will be interested mainly in positive semidefinite functions that are symmetric, which are usually referred to as \textit{kernel functions}, and which happen to be the class of covariance functions of second order stochastic processes \citep[][Th.
+27]{berlinet2004reproducing}.
+Let us now show how a RKHS arises from a kernel function.
+
+Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and without loss of generality assume \(\E[X(t)]=0\) for all \(t\in[0,1]\).
+Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\).
+To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional vector space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
 \begin{equation}\label{eq:h0}
@@
 \end{equation}
-Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\).
-However, this space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\) for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
+Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\). However, this new space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely
+\[
+\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
+\]
+for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
+We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}: \(f(t) = \dotprod{K(t, \cdot)}{f}_K\) for all \(t \in [0, 1]\).
+In particular, \(K(t, s)=\dotprod{K(t, \cdot)}{K(s, \cdot)}_K\), which can be understood as saying that the kernel \textit{reproduces} itself, hence the term ``reproducing kernel''.
 
-\begin{proposition} \(\dotprod{\cdot}{\cdot}_K\) is an inner product.
+\begin{proposition} \((\Hcal_0(K), \dotprod{\cdot}{\cdot}_K)\) is an inner product space.
+\end{proposition}
   \begin{proof}
-    Symmetry and linearity are clear from the definition, while non-negativity is a direct consequence of \(K\) being a covariance function, and thus positive semidefinite.
+    Firstly, consider \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\), and observe that
+    \[
+      \dotprod{f}{g}_K = \sum_j b_j f(s_j) = \sum_i a_i g(t_i).
+    \]
+    These equalities show that \(\dotprod{f}{g}_K\) depends only on \(f\) and \(g\) through their values, so it is independent of their representation in \(\Hcal_0(K)\).
+From this expression (and also from the original definition) it is straightforward to check linearity and symmetry, while positive semidefiniteness is a direct consequence of \(K\) being a covariance function, and thus positive semidefinite.
+It remains to prove that \(\dotprod{f}{f}_K=0\) implies \(f=0\).
+Indeed, for all \(t\in[0,1]\) and \(\epsilon \in \R\), we have
+    \[
+    0 \leq \dotprod{f + \epsilon K(t, \cdot)}{f + \epsilon K(t, \cdot)}_K = \dotprod{f}{f}_K + 2\epsilon\dotprod{f}{K(t, \cdot)}_K + \epsilon^2 K(t, t).
+    \]
+    Now, if \(\dotprod{f}{f}_K=0\), by letting \(\epsilon\to 0^{\pm}\) we see that \(f(t)=\dotprod{f}{K(t, \cdot)}_K\) necessarily vanishes for all \(t\), as desired.
   \end{proof}
+
+At this point, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\), which informally amounts to adding all the limits of Cauchy sequences in \(\Hcal_0(K)\), turning it into a genuine Hilbert space with the inner product extended accordingly.
+Then it is immediate to see that the reproducing property is retained.
+An important consequence is that \(\Hcal(K)\) is a space of actual functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
+
+Indeed, another way of defining a RKHS is via the continuity of all the \textit{evaluation operators}, i.e., \(\delta_t(f) := f(t)\) for \(f\in\Hcal(K)\) and \(t\in[0,1]\).
+In this case, the interpretation is that if two functions are close in the RKHS norm, they are also pointwise close.
+Although in this definition there is no explicit mention of the kernel, it can be recovered through the Riesz representation theorem.
+We summarize below some interesting topological properties of \(\Hcal(K)\).
+
+\begin{proposition} The following properties hold for the space \(\Hcal(K)\):
+  \begin{enumerate}
+    \item The evaluation operator \(\delta_t\) is bounded for all \(t\in[0,1]\).
+    \item Norm convergence implies pointwise convergence, and if \(K\) is continuous, it also implies uniform convergence.
+    \item If \(K\) is \(m\)-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
+  \end{enumerate}
 \end{proposition}
+\begin{proof}
+  To prove \((i)\), simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
+  \[
+  |\delta_t(f)| = |f(t)| = |\dotprod{f}{K(t, \cdot)}_K| \leq \|K(t, \cdot)\|_K \|f\|_K = \sqrt{K(t, t)}\|f\|_K,
+  \]
+  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \((ii)\), observe that for any \(t\in[0,1]\) we have
+  \[
+    |f_n(t) - f(t)| = |\delta_t(f_n - f)| \leq \|\delta_t\|\|f_n - f\|_K,
+  \]
+  and we have just shown above that the evaluation operator is bounded.
+Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform.
+Finally, we prove the second statement in \((iii)\), and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof.
+If \(K\) is continuous, then it is clear that every function in \(\Hcal_0(K)\) is continuous.
+Let us now fix an arbitrary \(f\in \Hcal(K)\).
+By definition, there exists a sequence \(\{f_n\}\subset \Hcal_0(K)\) of continuous functions such that \(\|f_n - f\|_K \to 0\).
+But then \(f_n \to f\) uniformly by \((ii)\), and therefore \(f\) is continuous as the uniform limit of continuous functions.
+Note that this statement remains valid even when the underlying domain is not compact.
+\end{proof}
 
-Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\), turning it into a genuine Hilbert space.
-As it turns out, functions in this space satisfy the so-called \textit{reproducing property}, which can be easily checked from the definitions.
+For the sake of completeness, it is worth mentioning that the RKHS associated with a kernel function is unique, and indeed the Moore-Aronszajn theorem \citep[e.g.][Th.
+3]{berlinet2004reproducing} states that there is a one-to-one correspondence between kernel functions and reproducing kernel Hilbert spaces.
+Let us now illustrate the definition in a few simple cases; see \citet[][Ch.~1]{saitoh2016theory} for more involved examples.
 
-\begin{proposition}
-  If \(f\in \Hcal(K)\), then \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(t \in [0, 1]\).
-  \begin{proof}
+\begin{example}\label{ex:bm}
+  If \(X\) is the standard Brownian motion on \([0, 1]\), it is well known that the associated covariance function is \(K_{\text{bm}}(t, s) = \min\{t,s\}\).
+The corresponding RKHS is given by \citep[][Ex.
+8.19]{janson1997gaussian}:
+  \[
+    \Hcal(K_{\text{bm}}) = \{f: f \text{ is absolutely continuous, } f(0)=0 \text{ and } f'\in L^2[0, 1]\},
+  \]
+  with inner product \(\dotprod{f}{g}_{K_{\text{bm}}} = \int f'g'\).
+\end{example}
 
-\end{proof}
-\end{proposition}
+\begin{example}[Finite-dimensional RKHS]
+  Setting aside our stochastic process framework for a moment, we can conceive examples outside of \(\mathcal T =[0,1]\).
+Consider a vector-valued random variable \(X = (X_1, \dots, X_p)\) with non-singular covariance matrix \(\Sigma\).
+In this case the index set would be \(\mathcal T_p = \{1,\dots,p\}\), so identifying functions on \(\mathcal T_p\) with points in the \(p\)-dimensional Euclidean space, we have \(\Hcal(\Sigma)=\R^p\), with inner product \(\dotprod{x}{y}_\Sigma=x\Sigma^{-1}y\).
+\end{example}
 
-This is where the name ``reproducing kernel'' comes from.
-An important consequence is that \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
+\begin{example}[Non-RKHS Hilbert space]
+  As we pointed out before, the space of square integrable functions \(L^2(\R)\) \textit{is not} a RKHS.
+There are many additional ways of seeing this; for example, in this space norm convergence does not imply pointwise convergence.
+Another possibility is to observe that the would-be reproducing kernel must be the Dirac delta function:
+  \[
+    f(s) = \int_{-\infty}^\infty f(t)\delta(t, s)\, dt \quad \forall s \in \R.
+  \]
+  However, \(\delta(t, \cdot) \notin L^2(\R)\).
+\end{example}
 
-\begin{outcomment}
-  A kernel defines an RKHS and vice-versa.
-\end{outcomment}
+\subsection*{The covariance operator}
 
-Note that he covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-An integral operator closely related to this kernel is the so-called covariance operator, namely \(\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds\) for \(f \in L^2[0, 1]\), which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}.
-It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
-In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
-Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
+ Note that, as expected, the covariance function of \(X\) plays a crucial role in characterizing the associated RKHS.
+An integral operator closely related to the kernel is the so-called covariance operator, namely
+\[
+\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
+\]
+which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}, so for the remaining of this work we will indeed suppose that \(K\) is continuous.
+As it turns out, the covariance function admits a spectral decomposition in terms of the eigenvalues and eigenfunctions of this integral operator, a sort of continuous generalization of the eigendecomposition of a symmetric positive semidefinite matrix.
 
-Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
+\begin{theorem}[Mercer's theorem]\label{th:mercer}
+    Let \(K\) be a continuous kernel on \([0,1]^2\).
+Then, there exists an orthonormal basis \(\{\phi_j\}\) of \(L^2[0,1]\) consisting of eigenfunctions of \(\mathcal K\), whose corresponding eigenvalues are all positive and form a decreasing sequence \(\{\lambda_j\}\to 0\), and such that
+    \[
+      K(t, s) = \sum_{j=1}^\infty \lambda_j \phi_j(t)\phi_j(s)
+    \]
+    for all \(t\) and \(s\), where the convergence is absolute and uniform.
+\end{theorem}
+
+Furthermore, a similar orthogonal decomposition holds for our stochastic process \(X\), an expansion which lies at the base of many \(L^2\) functional regression models.
+The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Ch.~3.2]{berlinet2004reproducing}.
+
+\begin{theorem}[Karhunen-Loève expansion]
+  With the assumptions and notations of Theorem~\eqref{th:mercer}, the centered process \(X=X(t)\) with covariance function \(K\) admits the quadratic-mean representation
+  \[
+  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
+  \]
+  where the convergence is in \(L^2(\Omega)\) and uniform in \(t\).
+The \(\zeta_j\) are independent zero-mean random variables with \(\E[\zeta_i\zeta_j]=\delta_{ij}\lambda_j\), given explictly by the formula
+  \[
+    \zeta_j = \int_0^1 X(t)\phi_j(t)\, dt.
+  \]
+\end{theorem}
+
+In addition to these results, it is worth mentioning that the covariance operator \(\mathcal K\) provides several alternative definitions of \(\Hcal(K)\).
+For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
+Note that, since \(\{\lambda_j\}\) tends to zero, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, not only in that they inherit the regularity of the kernel \(K\), but also in the sense that their components in an orthonormal basis, namely \(\dotprod{f}{\phi_j}\), need to vanish quickly.
+
+\subsection*{Loève's isometry}
+
+Until now we have seen that \(\Hcal(K)\) is a space related to \(X\) only through its covariance function.
+Nevertheless, we will now hopefully clarify that this space can indeed be seen as \textit{the} Hilbert space inherently associated with the process \(X\) itself.
+Let us shift the perspective momentarily and consider the linear span of \(X\) in the space of all zero-mean random variables with finite second moment, i.e.:
+\[
+\mathcal L_0(X) = \left\{\sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
+\]
+It is well known that its completion in \(L^2(\Omega)\), denoted \(\mathcal L(X)\), is a Hilbert space, sometimes called the Hilbert space generated by \(X\).
+In the words of \citet{berlinet2004reproducing}, ``\(\mathcal L(X)\) contains the random variables attainable by linear operations, including limits, on the measurements of the process''.
+As it happens, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and \(\mathcal L(X)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
 This isometry is essentially the completion of the correspondence
@@
 and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
-Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
+Thus, \(\Hcal(K)\) can be regarded (and it is often done in statistics) as an isometric copy of \(\mathcal L(X)\).
+
+However, despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process.
+It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
+For instance, the trajectories of the Brownian motion are known to be nowhere differentiable, but as shown in Example \ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere.
+An heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
+\[
+  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
+\]
+and the truncated version \(X_N(t)\) up to the \(N\)-th term.
+On the one hand, for each fixed \(t\) we have \(X_N(t)\to X(t)\) in the quadratic mean sense (by the Karhunen-Loève theorem), but on the other hand observe that
+\[
+  \E\|X_N(\cdot)\|^2_K = \E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
+\]
+
 As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
@@
 This notation, viewed as a formal extension of the inner product, often proves to be useful and convenient.
+Some properties stemming from this interpretation are stated below \citep[see][p.~974]{parzen1961approach}.
+
+\begin{proposition}
+  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations are satisfied:
+\begin{enumerate}
+  \item \(\dotprod{X}{K(t, \cdot)}_K = X(t)\), a particular reproducing property.
+  \item \(\E[\dotprod{X}{f}_K] = 0\).
+  \item \(\E[\dotprod{X}{f}_K\dotprod{X}{g}_K] = \dotprod{f}{g}_K\).
+\end{enumerate}
+\end{proposition}
+Note that the statement \((iii)\) above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\mathcal L(X)\), respectively.
+
+\subsection*{Applications in machine learning}
+
+Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning.
+For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \mathcal H\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS.
+Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
+\[
+K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\mathcal H}.
+\]
+The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with an simple expression but a complex feature space (e.g.
+infinite-dimensional), such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).
+
+Another example of the use of RKHS's is in the context of regularization problems, where the celebrated \textit{representer theorems} provide a way of obtaining closed-form solutions to a penalized optimization problem.
+For instance, learning techniques that rely on empirical risk minimization \citep[e.g.][]{vapnik1991principles} benefit from one such result by \citet{scholkopf2001generalized}, stated below.
+
+\begin{theorem}[Generalized representer theorem]
+  Consider a RKHS \(\Hcal(K)\) of functions defined over a non-empty set \(\mathcal X\), a training set \(\{(x_i, y_i)\}_{i=1}^p \subset (\mathcal X \times \R)^p\), an arbitrary error function \(R: (\mathcal X  \times \R^2)^p \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
+Then, any minimizer \(f^\ast\) of the regularized empirical risk
+  \[
+    R(\{(x_i, y_i, f(x_i))\}_{i=1}^p) + \Omega(\|f\|_K)
+  \]
+  is of the form \(f^\ast(\cdot) = \sum_{i=1}^p a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,p\).
+\end{theorem}
+
+\section{Markov chain Monte Carlo}\label{sec:mcmc}
+
+In the context of Bayesian inference, there are many situations in which we would like to sample from a distribution without using its explicit closed-form density, either because we do not know it, or because it is computationally difficult to do so.
+In these cases, we could work with a function \(f(x)\) proportional to our target density \(\pi(x)\).
+More specifically, in a Bayesian framework the posterior distribution is often intractable due to the normalizing integral constant, but we do know that \(posterior \propto prior \times likelihood\).
+
+When \(x\in\R\), we can use Monte Carlo algorithms such as \textit{rejection sampling}.
+In a nutshell, this method samples uniformly from the area under \(f(x)\) by means of an auxiliary density that encompasses this area.
+Specifically, one needs to find a density function \(g\) with \(\supp f \subset \supp g\) and \(f \leq Mg\) for some \(M>1\).
+Then, sampling from \(g\) and accepting the proposal, say \(x\), with probability \(f(x)/Mg(x)\) yields an approximate set of samples from \(f\).
+Nevertheless, these kinds of algorithms are rapidly affected by the so-called \textit{curse of dimensionality}, so new techniques are needed when sampling from multidimensional distributions.
+This is where Markov chain Monte Carlo (MCMC) methods come into play \citep[see e.g.][]{brooks2011handbook}.
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=.25\textwidth]{cube_sphere}
+  \caption{To cover the volume of \(f(x)\) in higher dimensions we intuitively need a volume that grows exponentially with the dimension itself.}
+\end{figure}
+
+MCMC methods are based on the iterative construction of a Markov chain whose stationary distribution is the objective distribution \(\pi(x)\).
+In this way, we can take the samples of a sufficiently advanced chain as approximate realizations of \(X\sim \pi(x)\).
+When it comes to building the chains, transition probabilities are assigned in such a way that regions with higher density with respect to \(\pi(x)\) are favored (by means of the known proportional function \(f(x)\)).
+The dimensionality issues are somewhat mitigated, and on top of that there are specific tuning techniques to tackle them.
+However, the samples obtained from this procedure are not independent, though they can be made so by \textit{thinning} the chain and considering, say, only one every \(M\) samples.
+
+Another advantage of these methods is that they recover the marginal distributions directly.
+Suppose that after a successful MCMC run we have \(M\) multivariate samples \(\{\theta^{(m)}=(\theta_1^{(m)},\dots, \theta_p^{(m)})\}\) of a joint \(p\)-dimensional distribution.
+The marginal distribution of each variable, which would be theoretically computed as
+\[
+\pi(\theta_i) = \int \pi(\theta)\,d\theta_1\,\cdots\,d\theta_{i-1}\,d\theta_{i+1}\,\cdots \,d\theta_p,
+\]
+can be approximated by just retaining the samples \(\{\theta_i^{(m)}\}\) corresponding to \(\theta_i\).
+
+\subsection*{Metropolis-Hastings}
+
+The Metropolis-Hastings algorithm \citep{metropolis1953equation} is arguably the most well-known algorithm in this context.
+As a matter of fact, it is actually more of a general framework from which many other methods are derived.
+In its original formulation, it relies on a symmetric \textit{proposal} distribution \(g(x'|x_t)\) that generates a proposal for the next state given the current one (e.g.
+a Gaussian distribution centered on \(x_t\)).
+Approximate samples from the intended distribution can then be generated through the following iterative acceptance-based procedure:
+\begin{enumerate}[1.]
+  \item Generate a proposal \(x' \sim g(x'|x_t)\).
+  \item Accept the proposal with probability \(\alpha=\min\{1, f(x')/f(x_t)\}\). If accepted, set \(x_{t+1}=x'\); otherwise set \(x_{t+1}=x_t\).
+\end{enumerate}
+
+ \begin{figure}[ht!]
+   \centering
+   \includegraphics[width=.6\textwidth]{mh}
+   \caption{Representation of the Metropolis-Hastings algorithm (taken from   \citet{jaewook2015metamodel}). The discarded proposals are represented by dashed circles.}
+ \end{figure}
+
+Note that \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\).
+This specific acceptance rate is also chosen so that the detailed balance equations of the chain are satisfied.
+An immediate generalization consists on choosing a general (possibly non-symmetric) jump distribution \(g\), in which case the acceptance probability is rewritten as
+\[
+\alpha = \min\left\{1, \frac{f(x')g(x_t \mid x')}{f(x_t)g(x'\mid x_t)}\right\}.
+\]
+
+Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high.
+If we know the conditional distribution of each variable, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension.
+Specifically, in each step we get unidimensional samples \textit{in turn} (e.g.
+via rejection sampling) of each variable conditional on the rest of them, using always the more up-to-date values available:
+
+\[
+\pi(x_l^{(i+1)} \mid x_{1}^{(i+1)}, \dots, x_{l-1}^{(i+1)}, x_{l+1}^{(i)}, \dots, x_L^{(i)}), \quad l=1,\dots,L.
+\]
+
+Lastly, there are several methods based on Metropolis-Hastings that are the subject of active research, both from a theoretical and computational standpoint.
+A good example are \textit{Hamiltonian Monte Carlo} methods \citep[e.g.][]{neal2011mcmc}, which are a family of algorithms that introduce gradient information to guide the proposals in the sample space.
+They are based on Hamiltonian dynamics, and in general they improve convergence speed due to a more intelligent choice of new points on the chain.
+With automatic differentiation being developed in the last few decades, these algorithms have become computationally feasible and have been widely adopted by the scientific community.
+
+\subsection*{Affine-invariant ensemble sampler}
+
+An interesting and often desirable property of sampling algorithms is that they be \textit{affine-invariant}, which means that they regard two distributions that differ in an affine transformation, say \(\pi(x)\) and \(\pi_{A, b}(Ax + b)\), as equally difficult to sample from.
+This is useful when one is working with very asymmetrical or skewed distributions, for an affine transformation can turn them into ones with simpler shapes.
+
+Generally speaking, a MCMC algorithm can be described at time \(t+1\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi(x)\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d.
+random variables that represent the random behavior of the chain.
+With this notation, the property of affine-invariant can be characterized as \citep{goodman2010ensemble}:
+ \[
+ R(A\lambda+b, \xi(t), \pi_{A,b}) = AR(\lambda, \xi(t), \pi) + b,
+ \]
+ for all \(A,b\) and \(\lambda\), and for allmost all \(\xi(t)\).
+This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Gamma(t)+b\) for all \(t\).
+In \citep{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property.
+Specifically, they work with a set \(\Lambda=(\Lambda_1, \dots, \Lambda_L)\) of \textit{walkers}, where \(\Lambda_l(t)\) represents an individual chain at time \(t\).
+At each iteration, an affine-invariant transformation is used to find the next point, which is constructed using the current values of the rest of the walkers (similar to Gibb's algorithm), namely the \textit{complementary ensemble}
+ \[
+  \Lambda_{-l}(t) = \{\Lambda_1(t+1), \dots, \Lambda_{l-1}(t+1), \Lambda_{l+1}(t), \dots, \Lambda_L(t)\}, \quad l=1,\dots, L.
+  \]
+
+To maintain the affine invariance and the joint distribution of the ensemble, the walkers are advanced one by one following a Metropolis-Hastings acceptance scheme.
+There are mainly two types of moves.
+
+\begin{itemize}
+  \item \textit{Stretch move}. For each walker \(1\leq l \leq L\) another walker \(\Lambda_j \in \Lambda_{-l}(t)\) is chosen at random, and the proposal is constructed as
+  \[
+    \Lambda_l(t) \to \Gamma = \Lambda_j + Z(\Lambda_l(t) - \Lambda_j),
+  \]
+  where \(Z \stackrel{i.i.d.}{\sim} g(z)\) satisfying the symmetry condition \(g(z^{-1})=zg(z)\). In particular, the suggested density is
+  \[
+  g_a(z) \propto \begin{cases}
+    \frac{1}{\sqrt{z}}, & \text{if } z \in [a^{-1}, a],\\
+    0, & \text{otherwise.}
+\end{cases}, \quad a > 1.
+  \]
+
+Supposing \(\R^p\) is the sample space, the corresponding acceptance probability (chosen so that the detailed balance equations are satisfied) is:
+  \[
+    \alpha = \min\left\{1, \ Z^{p-1}\frac{\pi(\Gamma)}{\pi(\Lambda_l(t))}\right\}.
+  \]
+
+  \item \textit{Walk move}. For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is chosen, and the proposed move is
+\[
+\Lambda_l(t) \to \Gamma = \Lambda_l(t) + W,
+\]
+where \(W\) is a normal distribution with mean \(0\) and the same covariance as the sample covariance of all walkers in \(S_l\).
+The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).
+\end{itemize}
+
+From a computational perspective, the Python library \textit{emcee} \footnote{\url{https://emcee.readthedocs.io/en/stable/}} \citep{foreman2013emcee} provides a parallel implementation of this algorithm.
+The idea is to divide the ensemble \(\Lambda\) into two equally-sized subsets \(\Lambda^{(0)}\) and \(\Lambda^{(1)}\), and then proceed on each iteration in the following alternate fashion:
+
+\begin{enumerate}[1.]
+  \item Update \textit{all} walkers in \(\Lambda^{(0)}\) through one of the available moves explained above, using \(\Lambda^{(1)}\) as the complementary ensemble.
+  \item Use the new values in \(\Lambda^{(0)}\) to update \(\Lambda^{(1)}\).
+\end{enumerate}
+
+In this way the detailed balance equations are still satisfied, and each of the steps can benefit from the computing power of an arbitrary number of processors (up to \(L/2\)).

### 2022-08-23 thesis/src/bayesian-functional-regression.tex
Finish background section

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Copyright (c) 2022 Antonio Coín Castro
+% Copyright (c) 2022 Antonio Coín
 %
@@
 
-\section{Bayesian methodology for RKHS-based functional regression models}\label{sec:methodology}
+\let\epsilon\varepsilon
 
-In this section we present the precise models and Bayesian methodologies explored in this work.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Bayesian methodology for RKHS-based functional regression models}\label{ch:bayesian}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{outcomment}
+  Reescribir dejando p libre.
+\end{outcomment}
+
+In this chapter we present the precise models and Bayesian methodologies explored in this work.
 The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
@@
 In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y | X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
-Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
+Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via MCMC methods \incomment{Referir al capítulo de model choice para detalles}.
 Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.
 
-\subsection{Functional linear regression}\label{sec:rkhs-linear-model}
+\section{Functional linear regression}\label{sec:rkhs-linear-model}
 
@@
 
-\begin{description}
-  \item[Summarize-then-predict.] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
+  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
   \[
@@
 
-  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
+  \paragraph{Predict-then-summarize.} Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
   \begin{equation}\label{eq:sampled-response-vector}
@@
 
-  \item[Variable selection.] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
+  \paragraph{Variable selection.} Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
 Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
 After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
-\end{description}
 
@@
 
-\subsection{Functional logistic regression}\label{sec:rkhs-logistic-model}
+\section{Functional logistic regression}\label{sec:rkhs-logistic-model}
 
@@
 Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
-
-\subsection{Model choice and validation}\label{sec:model-choice}
-
-In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
-
-\paragraph{The choice of \(\symbf p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we could impose a prior distribution on \(p\) as well (e.g.
-a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
-In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
-Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
-In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, DIC, cross-validation, \ldots) to select its optimal value.
-This is equivalent to imposing a degenerate prior distribution on \(p\).
-Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}).
-\incomment{Reversible MCMC}
-
-\begin{comment}
-LO QUE NO SIRVA, GUARDAR PARA EL TFM.
-
-Reescribir label switching (nos pasa; lo que pasa que con p free se acentua).
-
-Ocurre porque el likelihood es simetrico: L(theta mid y) = L(thetaperm mid y) para cualquier permutacion (hay p! posibles) --> no-identificabilidad.
-
-"In these situations, parameter estimates and other output are inevitably aggregated
-across numerous replications, and because the labels of the enumerated classes are
-arbitrarily assigned, there is no guarantee that the same class would have the same
-label from replication to replication."
-
-From a Bayesian pointof view, this amounts to truncating the original prior distribution... (to verify the constraints)
-
-As a result, in any MCMC algorithm, labels of the
-components can permute multiple times between iterations of the sampler.
-Paradoxically, as noted by Celeux, Hurn and Robert (2000), Fruhwirth-Schnatter (2001),
-Jasra, Holmes and Stephens (2005) and Papastamoulis and Iliopoulos (2010), among others,
-label switching is a prerequisite for MCMC convergence. If there is no label switching it
-means that the sampler is not exploring all the modes of the posterior distribution of (1)
-
-"When sampling from the unconstrained posterior via MCMC methods, which component of the sampled parameter corresponds to which state is unknown due to potential label switching." Si usamos la marginal para haacer inferencia --> nonsensical answers (Stephens, 2000)
-
-Solución: añadir constraints de identificabilidad artificiales (ej relabeling, mantener ordenados, posprocesado).
-
-Imposing an identifiability constraint on the parameters breaks the sym-
-metry of the posterior distribution of the parameters, and so we might hope that it would allow us to perform sensible inference for the individual components of the mixture.
-Inference conditional on such a constraint may be performed by post-processing the sample (proposition 3.1 in Stephens 1997 phD thesis).
-
-OTRA OPCION: Elegir en cada iteración como parámetro para ordenar aquel que deja mayor separación (tras escalarlos todos a [0,1]): \verb@https://helda.helsinki.fi/bitstream/handle/10138/329485/Approximate_Bayesian_computation_for_finite_mixture_models.pdf?sequence=1@
-
-OTRA SOLUCION: "Adaptive Metropolis with Online Relabeling" --> The key step is the selection, after each proposal of a permutation (relabeling) of the proposed vector by minimizing a quadratic cost function
-$L_{t-1} (x)$ over all permutations of the proposed vector, where L is defined by
-$L_t(x) = (x - \mu)'\Sigma^{-1}(x - \mu)$
-This step forces the posterior sample to look as unimodal as possible.
-
-Citas: Stephens (2000), Papastamoulis \& Iliopoulos, 2010; Rodriguez \& Walker, 2014
-\end{comment}
-
-\paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
-\begin{itemize}
-  \item For the expected value \(b_0\) we propose to use the MLE of \(b\).
-Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation).
-Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
-  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
-  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
-This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
-\end{itemize}
-
-\paragraph{Summarizing the posterior.} On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
-At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-\incomment{Argue that, because of the label switching and multimodality, summarizing the posterior might be impractical, unhelpful, misleading, etc.
-So the best strategy is the posterior mean.}
-
-\paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
-For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\symbf Y^*)\) of the posterior predictive distribution \(\symbf Y^* \equiv \{\symbf Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the real value of the statistic.
-They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.
-
-\begin{figure}[ht]
-  \centering
-  %\includegraphics[width=.75\textwidth]{img/ppc_linear}
-  \caption{Posterior predictive graphical checks on a fitted model.
-On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the approximate sampled responses, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
-\end{figure}

### 2022-08-23 thesis/src/introduction.tex
Finish background section

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Introduction}
+\chapter{Introduction}\label{ch:introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %     \begin{small}
-%       \emph{In which the reasons for creating this package are laid bare for the
-%       whole world to see and we encounter some usage guidelines.}
+%       \emph{Quote}
 %     \end{small}
@@
 % \end{center}
+
+\begin{outcomment}
+  - mirar cosas que puedan ser paragraphs
+  - Listas enumeradas con numeros 1., 2., etc.
+  - Descripción corta de cada capítulo justo debajo?
+\end{outcomment}
 
@@
 
-FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}.
-Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
-At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
-Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
-These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems.
-Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
+\begin{outcomment}
+  ejemplo de Tecator en la introducción (información de las derivadas+smoothing technique)
+  Gráficas
+  Introducción a FDA
+\end{outcomment}
+
+\begin{center}
+\color{teal}\FourStar
+\end{center}
 
@@
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+\incomment{Definir proceso estocástico L2, el espacio L2(Omega), el espacio de probabilidad subyacente (decir que no es relevante) y el espacio L2[0,1]} We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
+In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(x_i, y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
+
+\begin{outcomment}
+  Comentar el caso function-on-function regression.
+\end{outcomment}
 

### 2022-08-23 thesis/src/model-choice.tex
Finish background section

--- 2022-08-06 paper/paper.tex
@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Copyright (c) 2022 Antonio Coín
 %
-\pubyear{2022}
-\volume{TBA}
-\issue{TBA}
-\doi{0000}
-\arxiv{2010.00000}
-\firstpage{1}
-\lastpage{1}
+% This work is licensed under a
+% Creative Commons Attribution-ShareAlike 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{xcolor}
-\usepackage{bm}
-\usepackage{natbib}
-\usepackage[
-  colorlinks,
-  citecolor=blue,
-  urlcolor=blue,
-  filecolor=blue,
-  backref=page
-]{hyperref}
-\usepackage{graphicx}
-
-%% Comentarios
-\usepackage{todonotes}
-\usepackage{ulem}
-
-\startlocaldefs
-\numberwithin{equation}{section}
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}{Proposition}[section]
-
-\renewcommand{\epsilon}{\varepsilon}
-
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\Hcal}{\ensuremath\mathcal{H}}
-\newcommand{\Var}{\ensuremath\operatorname{Var}}
-\definecolor{violet}{rgb}{0.7,0,0.6}
-
-%% Scalar product
-\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}
-
-%% Colors
-\definecolor{comment-red}{RGB}{240,30,70}
-
-%% Environment for colored comments
-\newenvironment{comment}[1][comment-red]
-{
-\noindent \color{#1}
-}
-{
-\color{black}
-}
-
-%% Inline comments
-\newcommand\incomment[2][comment-red]{\color{#1}[\textit{#2}]\color{black}}
-\endlocaldefs
-
-\begin{document}
-
-\begin{frontmatter}
-\title{Bayesian RKHS-based methods in functional regression\support{Support information of the article.}}
-\runtitle{Bayesian RKHS-based methods in functional regression}
-
-\begin{aug}
-\author{\fnms{First} \snm{Author}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{first@somewhere.com}},
-\author{\fnms{Second} \snm{Author}\thanksref{addr1,t3,m1,m2}\ead[label=e2]{second@somewhere.com}}
-\and
-\author{\fnms{Third} \snm{Author}\thanksref{addr2,t1,m2}%
-\ead[label=e3]{third@somewhere.com}%
-\ead[label=u1,url]{http://www.foo.com}}
-
-\runauthor{F. Author et al.}
-
-\address[addr1]{Address of the First and Second authors
-     Usually a few lines long
-    \printead{e1} % print email address of "e1"
-    \printead*{e2}
-}
-
-\address[addr2]{Address of the Third author
-    Usually a few lines long
-    Usually a few lines long
-    \printead{e3}
-    \printead{u1}
-}
-
-\thankstext{t1}{Some comment}
-\thankstext{t2}{First supporter of the project}
-\thankstext{t3}{Second supporter of the project}
-
-\end{aug}
-
-\begin{abstract}
-We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
-These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated through Markov chain Monte Carlo methods.
-Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.
-\end{abstract}
-
-\begin{keyword}[class=MSC]
-\kwd[primary ]{62M20}
-\kwd[; secondary ]{62F15}
-\end{keyword}
-
-\begin{keyword}
-\kwd{functional data}
-\kwd{linear regression}
-\kwd{logistic regression}
-\kwd{reproducing kernel Hilbert space}
-\kwd{Bayesian inference}
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{sec:intro}
-
-The problem of predicting a scalar response from a functional covariate is one that has gained traction over the last few decades, as more and more data is being generated with an ever-increasing level of granularity in the measurements.
-While in principle the functional data could be simply regarded as a discretized vector in a very high dimension, there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
-As a consequence, numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum.
-A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}.
-
-FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see][]{ullah2013applications}.
-Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint.
-Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
-At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
-Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
-These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems.
-Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.
-
-In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations.
-Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions.
-To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\).
-We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\mathcal D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), and our aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).
-
-\subsubsection{\(L^2\)-models, shortcomings and alternatives}
-
- The most common functional linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}.
-It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
-\begin{equation}\label{eq:l2-linear-model}
-Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
-\end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-In this case, the inference on \(\beta\) is hampered by the fact that \(L^2[0,1]\) is an extremely wide space that also contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
-In spite of this, model~\eqref{eq:l2-linear-model} is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
-Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.
-Thus, some regularization or dimensionality reduction technique is needed for parameter estimation.
-
-A similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
-\begin{equation}\label{eq:l2-logistic-model}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
-\end{equation}
-where \(\alpha_0 \in \R\) and \(\beta \in L^2[0, 1]\).
-In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2018functional}.
-
-It turns out that in both scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
-As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
-In fact, it includes a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional, berrendero2018use} for the case of functional logistic regression.
-Incidentally, these models also shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
-
-A major aim of this work is to motivate these recently-proposed models inside the functional framework, while also providing efficient techniques to apply them in practice.
-Our main contribution is the proposal of a Bayesian approach to parameter estimation within the aforementioned RKHS models, in which a prior distribution is imposed on the unknown functional parameter to obtain a posterior distribution after seeing the data.
-Although setting a prior distribution on a function space is generally a hard task, the specific parametric formulation of the RKHS models we propose greatly facilitates this (see Section~\ref{sec:methodology} for details).
-A similar Bayesian scheme has recently been explored in \citet{grollemund2019bayesian}, albeit not within a RKHS framework.
-
-Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion.
-As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
-As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
-In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information.
-These points-of-impact selection models for functional predictors have also been considered in the  literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.
-Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain.
-
-\subsubsection{Organization of the paper}
-
-In Section~\ref{sec:rkhs} we review the basics of the theory of RKHS's from a probabilistic point of view.
-Section~\ref{sec:methodology} is devoted to explaining the Bayesian methodology and the functional regression models we propose.
-The empirical results of the experimentation are contained in Section~\ref{sec:results}, which also includes a short discussion of computational details.
-Lastly, the conclusions drawn from this work are presented in Section~\ref{sec:conclusion}.
-
-\subsection{Some essentials on RKHS's and notation}\label{sec:rkhs}
-
-The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces (for a more detailed account, see for example~\citet{berlinet2004reproducing}).
-Let us denote by \(K(t, s)= \mathbb E[X(t)X(s)]\) the covariance function of the centered process \(X\), and in what follows suppose that it is continuous.
-To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
-\begin{equation}\label{eq:h0}
-\Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
-\end{equation}
-This space is endowed with the inner product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-Then, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\).
-As it turns out, functions in this space satisfy the \textit{reproducing property} \(\dotprod{K(t, \cdot)}{f}_K = f(t)\) for all \(f \in \Hcal(K)\) and \(t \in [0, 1]\).
-An important consequence is that \(\Hcal(K)\) is a space of genuine functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
-
-Note that he covariance function of \(X\) (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS.
-An integral operator closely related to this kernel is the so-called covariance operator, namely \(\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds\) for \(f \in L^2[0, 1]\), which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}.
-It is worth mentioning that this operator provides several alternative definitions of \(\Hcal(K)\); for example, the RKHS can be identified with the image of the square root of the covariance operator, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), where \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\), respectively.
-In this case, the corresponding inner product is \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
-Note that, since the spectral theorem for compact operators tells us that the sequence of eigenvalues \(\{\lambda_j\}\) tends to zero \citep[e.g.][Th.~4.2.4]{hsing2015theoretical}, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly.
-
-Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space.
-Specifically, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
-This isometry is essentially the completion of the correspondence
-  \begin{equation}\label{eq:loeves-isometry}
-  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
-  \end{equation}
-and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
-Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
-As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\).
-However, following Parzen's approach in his seminal work \citep[e.g.][Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with the image \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\).
-This notation, viewed as a formal extension of the inner product, often proves to be useful and convenient.
-
-\section{Bayesian methodology for RKHS-based functional regression models}\label{sec:methodology}
-
-In this section we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
-For practical and computational reasons, the value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
-Thus, we will search for our functional parameter in the set
-\begin{equation}\label{eq:h0p}
-\Hcal_{0,p}(K)=\left\{ \sum_{j=1}^p \beta_j K(t_j, \cdot): \ \beta_j \in \R, \ t_j \in [0, 1]\right\}.
-\end{equation}
-
-The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\alpha \in \Hcal_{0,p}(K)\) and that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\) when \(\alpha(\cdot)=\sum_j\beta_j K(t_j, \cdot)\).
-
-In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately.
-Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
-Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) merely by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
-Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of feasible values is not very large in practice.
-
-In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
-samples amounts to
-\begin{equation}\label{eq:bayes-theorem}
-  \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
-\end{equation}
-In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y | X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
-Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.
-Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.
-
-\subsection{Functional linear regression}\label{sec:rkhs-linear-model}
-
-In the case of functional linear regression, the simplified RKHS model considered in this work is given by
-\begin{equation}\label{eq:rkhs-model-linear}
-  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
-\end{equation}
-where \(\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j, \cdot) \in \Hcal_{0,p}(K), \beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is a normally-distributed error term independent from \(X\), with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
-This model is essentially a finite-dimensional approximation from a functional perspective to the more general RKHS model that assumes \(\alpha \in \Hcal(K)\), proposed in~\citet{berrendero2019rkhs}.
-
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
-Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
-For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
-\begin{equation}\label{eq:rkhs-model-linear-2}
-  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
-\end{equation}
-
-It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
-Also, note that given the time instants \(t_j\), the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
-As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, this model could be easily extended to the case of several covariates via an expression of type \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
-In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
-
-\subsubsection{The bayesian approach: prior selection and posterior derivation}
-
-The prior distribution suggested for the parameter vector \(\theta \in \Theta_p\) is given by
-\begin{align}\label{eq:prior-linear}
-  \begin{split}
-  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
-  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2{\underbrace{\left(\mathcal X_\tau' \mathcal X_\tau + \eta I\right)}_{G_\tau}}^{-1}),
-\end{split}
-\end{align}
-where \(I\) is the identity matrix, \(\mathcal X_\tau\) is the data matrix \((X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
-On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
-In any event, the estimation of \(\alpha_0=\E[Y]\) is straightforward, so it could have been left out of the model altogether.
-On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.
-All in all, with a slight abuse of notation the proposed prior distribution becomes \(\pi(\theta) = \pi(b| \tau, \sigma^2)\pi(\tau)\pi(\alpha_0, \sigma^2)\).
-
-As for the posterior distribution, we only compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
-A standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
-
-\begin{proposition}
-Under the linear model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
-\begin{align*}
-\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\bm Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
-& + (p+n+2)\log\sigma - \frac{1}{2}\log |G_\tau|,
-\end{align*}
-where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
-\end{proposition}
-
-\subsubsection{Making predictions}
-
-In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
-
-\begin{description}
-  \item[Summarize-then-predict.] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
-  \[
-    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
-  \]
-  Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter.
-Although it contributes to measure the uncertainty in the approximations, its estimates are discarded in the final prediction.
-
-  \item[Predict-then-summarize.] Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
-  \begin{equation}\label{eq:sampled-response-vector}
-  \bm Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
-  \end{equation}
-  Then, we can take the mean of all such simulated responses as a proxy for each response variable, that is,
-  \[
-  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
-  \]
-  This method differs from the previous one in that it takes into account the full approximate posterior distribution instead of summarizing it directly.
-
-  \item[Variable selection.] Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
-Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
-After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
-\end{description}
-
-Note that these predictors can be obtained all at once after only one round of training (that is, an individual MCMC run to approximate the posterior distribution).
-As a consequence, what we have in practice is a single algorithm that can produce multiple predictors at the same computational cost, so that any of them can be chosen (or even switched back and forth) depending on the particularities of the problem at hand.
-Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
-
-\subsection{Functional logistic regression}\label{sec:rkhs-logistic-model}
-
-In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, an RKHS model might be given  by the  equation
-\begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
-\end{equation}
-
-Indeed, note that this can be seen as a finite-dimensional approximation (but, still, with a functional interpretation) to the general RKHS functional logistic model proposed by these authors, which can be obtained by replacing \(\Hcal_{0,p}(K)\) with the whole RKHS space \(\Hcal(K)\).
-Now, if we aim at a classification problem, our strategy will be similar to that followed in the functional linear model: after incorporating the sample information, we can rewrite~\eqref{eq:rkhs-model-logistic} as
-\begin{equation}\label{eq:rkhs-model-logistic-2}
-Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
-\end{equation}
-with
-\begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
-  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
-\end{equation}
-where in turn \(\beta_j\in\R\) and \(t_j\in[0, 1]\).
-
-In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model.
-First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes~\citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for such model.
-Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
-Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case.
-However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
-\color{black}
-
-\subsubsection{The Bayesian approach: prior selection and posterior derivation}
-
-As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model.
-However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
-The posterior distribution is then derived after a routine calculation.
-
-\begin{proposition}
-Under the logistic model~\eqref{eq:rkhs-model-logistic-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the log-posterior distribution
-\begin{align*}
-  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
-  \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
-\end{align*}
-Remember that \(\Psi_{X_i}(\alpha) = \sum_{j=1}^p \beta_j X_i(t_j)\).
-\end{proposition}
-
-\subsubsection{Making predictions}
-
-Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to a binary output in \(\{0, 1\}\).
-According to the usual criterion of minimizing the misclassification probability, the Bayes optimal rule is recovered by predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
-Nevertheless, for a more general cost function one could consider other criteria that would lead to evaluating whether \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold \(\gamma\in[0, 1]\).
-
-With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
-\[
-\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
-\]
-where \(\mathbb I\) is the indicator function (\(\mathbb I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
-The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
-
-On the other hand, the prediction method that takes into account the entire posterior approximation (i.e.
-the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
-Hence, there are primarily two possible outcomes:
-\begin{description}
-  \item[Average sampled probability.] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
-  \[
-    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
-  \]
-  \item[Average sampled response.] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
-\[
-  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
-\]
-In this case, each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
-Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).
-\end{description}
-
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
-
-\subsection{Model choice and validation}\label{sec:model-choice}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Model choice, implementation and validation}\label{ch:model-choice}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
 
-\paragraph{The choice of \(\bm p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
+\paragraph{The choice of \(\symbf p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
 While theoretically we could impose a prior distribution on \(p\) as well (e.g.
@@
 
-\begin{comment}
-LO QUE NO SIRVA, GUARDAR PARA EL TFM.
+%Reversible-jump MCMC
+% Estos métodos permiten variar la dimensionalidad subyacente del espacio muestral en cada propuesta (\sidecite[*0]{green1995reversible}).
+Teóricamente son útiles para muestrear de mixturas de gaussianas, procesos de Dirichlet o, en general, de cualquier distribución donde haya una variable latente que represente la dimensión del espacio.
+Sin embargo, no hay una implementación de referencia establecida.
 
+\begin{outcomment}
 Reescribir label switching (nos pasa; lo que pasa que con p free se acentua).
@@
 Citas: Stephens (2000), Papastamoulis \& Iliopoulos, 2010; Rodriguez \& Walker, 2014
-\end{comment}
+\end{outcomment}
+
+Aunque se puede tener en cuenta solo las primeras 'p' posiciones, en el propio algoritmo de MCMC no se puede 'apagar' fácilmente el valor de los betas que no se usan, por lo que crecen indefinidamente y generan problemas (lo mismo ocurre con los 'tau', que incluso se pueden salir de rango).
+Aunque una posible solución es limitar todos los taus y no solo los p primeros, esto solo retrasa el problema (con cadenas más largas vuelve a ocurrir, además de ocasionar el problema "RuntimeWarning:covariance is not positive-semidefinite." con el WalkMove), ya que los betas no están limitados.
+Podríamos intentar limitarlos a un valor razonable.
+
+Por otro lado, en algunas cadenas el sigma2 es más grande de la cuenta, ya que tiende a "compensar" la falta de flexibilidad del modelo (e.j.
+si el valor de p es más bajo).
+Esto hace que los estimadores basados en la media y en el posterior mean sean ligeramente peores (sobre todo el de la media puntual, donde la varianza estimada sí se ve notablemente afectada).
+Quizás para contrarrestar esto habría que tener un modelo con un $\sigma^2_p$ para cada valor de p?
+
+Además, aunque todo esto funcionase, si por ejemplo en una iteración el modelo elige p=2, está optimizando beta1 y beta2 para ese caso; si luego miramos solo la marginal de beta1, se habrán mezclado los casos en los que p=1 y beta1 debe valer una cosa, y en los que p=2 y beta1 vale otra cosa potencialmente distinta (se puede compensar con beta2, hay más grados de libertad).
 
@@
 
+\begin{outcomment}
+  Sección para elección del algoritmo mcmc.
+Contar que también se usa pymc (PPL, probabilistic programming language) y que en particular se probaron metropolis y NUTS: "Dentro de estos métodos destaca el \textit{No U-Turn Sampler} (NUTS, \citet{hoffman2014no}), que tiene la ventaja añadida de que realiza un ajuste automático de hiperparámetros."
+
+  % Otra de las ventajas que ofrece este método frente a algoritmos clásicos es que solo requiere fijar unos pocos hiperparámetros (independientes de la dimensión subyacente), en lugar de los \(O(N^2)\) correspondientes, por ejemplo, a la matriz de covarianza de la distribución de salto \(N\)-dimensional en Metropolis-Hastings.
+Además, el paquete ha sido extendido para admitir otro tipo de movimientos paralelos, y el objetivo del proyecto es proporcionar un método de muestreo de propósito general que funcione correctamente en una clase amplia de problemas.
+Es muy usado por ejemplo en el campo de la astrofísica, y está bien establecido y testeado.
+\end{outcomment}
+
+\begin{outcomment}
+  Sección para inicialización.
+
+  También en el artículo citado se aconseja inicializar las cadenas en una bola gaussiana alrededor de un punto que se espere que tenga alta probabilidad según \(p(x)\).
+\end{outcomment}
+
 \paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
 For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\bm Y^*)\) of the posterior predictive distribution \(\bm Y^* \equiv \{\bm Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\bm Y^*)\leq T(\bm Y)| \bm Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\bm Y^{(m)*}\}\) that fall below the real value of the statistic.
+Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\symbf Y^*)\) of the posterior predictive distribution \(\symbf Y^* \equiv \{\symbf Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
+In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the real value of the statistic.
 They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.

### 2022-08-26 thesis/chapters/background-related-work.tex
Revision up to chapter 4 complete

--- 2022-08-23 thesis/src/background-related-work.tex
@@
 where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\) and \(\E[X(t)\epsilon]=0\) almost everwhere; and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
-A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\mathcal K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\mathcal K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
-This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\mathcal K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).
+A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
+This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).
 
@@
 \[
-  X(t)=\sum_{j=1}^p x_j \phi_j(t), \quad \beta(t) = \sum_{j=1}^p \beta_j \phi_j(t).
+  X(t)=\sum_{j=1}^p Z_j \phi_j(t), \quad \beta(t) = \sum_{j=1}^p \beta_j \phi_j(t).
 \]
@@
 \[
-Y = \alpha_0 + \sum_{j=1}^p \beta_j x_j,
+Y = \alpha_0 + \sum_{j=1}^p \beta_j Z_j,
 \]
 which can then be solved in the usual manner.
-For example, when the chosen basis is composed of eigenfunctions of the covariance operator \(\mathcal K\), this method is known as Functional Principal Component Regression (FPCR).
+For example, when the chosen basis is composed of eigenfunctions of the covariance operator \(\K\), this method is known as Functional Principal Component Regression (FPCR).
 Another class of methods focus on selecting an \textit{a priori} basis for \(\beta\) (e.g.
@@
 \begin{equation}\label{eq:l2-logistic-model}
-  \Prob(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
+  \P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\beta}\}},
 \end{equation}
@@
 
-\begin{outcomment}
-  Rellenar?
-\end{outcomment}
+The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis when new information is available.
+It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}.
+Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model.
+
+One difficulty found in almost all Bayesian methods is that the posterior distribution, the main object of interest, is usually intractable due to the integral that appears as the normalizing constant in Bayes' rule.
+A way to bypass this limitation is to use conjugate distributions, where it can be rigorously proven that a certain combination of likelihood and prior distributions produces a posterior distribution in the same family as the prior.
+However, unless conjugate priors are used, a closed-form expression is generally unattainable and some type of approximation is required.
+Apart from basic numerical integration, some well-performing methods in this regard include variational inference approaches \citep[e.g.][]{blei2017variational} and MCMC methods \citep[e.g.][]{brooks2011handbook}.
+The former are based on approximating the posterior by another distribution restricted to a certain parametric family so that the Kullback-Leibler divergence between them is minimized, whereas the latter are iterative methods that directly provide approximate samples of the posterior distribution.
+
+As a comment, there are some works that tackle Bayesian inference from a functional perspective; a recent example are the functional Bayesian neural networks proposed by~\citet{sun2019functional}.
 
@@
 
-In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces (RKHS's).
+In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces.
 Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory.
@@
 \begin{definition}
-  We say that a function of two variables \(K: \mathcal T \times \mathcal T \to \R\) is positive semidefinite\footnote{Sometimes in the literature these functions are known simply as positive definite.} if
+  We say that a function of two variables \(K: \T \times \T \to \R\) is positive semidefinite\footnote{Sometimes in the literature these functions are known simply as positive definite.} if
   \[
@@
   \]
-  for any \(p\in\N\), any \((a_1,\dots, a_p)\in \R^p\) and any \((t_1,\dots, t_p)\in\mathcal T^p\).
-Note that this is equivalent to saying that the matrix \((K(t_i, t_j))_{i,j}\) is positive semidefinite for any choice of \(p\in\N\) and \((t_1,\dots, t_p)\in\mathcal T^p\).
+  for any \(p\in\N\), any \((a_1,\dots, a_p)\in \R^p\) and any \((t_1,\dots, t_p)\in\T^p\).
+Note that this is equivalent to saying that the matrix \((K(t_i, t_j))_{i,j}\) is positive semidefinite for any choice of \(p\in\N\) and \((t_1,\dots, t_p)\in\T^p\).
 \end{definition}
@@
     \item Norm convergence implies pointwise convergence, and if \(K\) is continuous, it also implies uniform convergence.
-    \item If \(K\) is \(m\)-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
+    \item If \(K\) is m-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
   \end{enumerate}
@@
 But then \(f_n \to f\) uniformly by \((ii)\), and therefore \(f\) is continuous as the uniform limit of continuous functions.
-Note that this statement remains valid even when the underlying domain is not compact.
+Note that this statement remains valid, with a slightly different proof, even when the underlying domain is not compact.
 \end{proof}
@@
 \begin{example}[Finite-dimensional RKHS]
-  Setting aside our stochastic process framework for a moment, we can conceive examples outside of \(\mathcal T =[0,1]\).
+  Setting aside our stochastic process framework for a moment, we can conceive examples outside of \(\T =[0,1]\).
 Consider a vector-valued random variable \(X = (X_1, \dots, X_p)\) with non-singular covariance matrix \(\Sigma\).
-In this case the index set would be \(\mathcal T_p = \{1,\dots,p\}\), so identifying functions on \(\mathcal T_p\) with points in the \(p\)-dimensional Euclidean space, we have \(\Hcal(\Sigma)=\R^p\), with inner product \(\dotprod{x}{y}_\Sigma=x\Sigma^{-1}y\).
+In this case the index set would be \(\T_p = \{1,\dots,p\}\), so identifying functions on \(\T_p\) with points in the \(p\)-dimensional Euclidean space, we have \(\Hcal(\Sigma)=\R^p\), with inner product \(\dotprod{x}{y}_\Sigma=x\Sigma^{-1}y\).
 \end{example}
@@
 \[
-\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
+\K f(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
 \]
@@
     Let \(K\) be a continuous kernel on \([0,1]^2\).
-Then, there exists an orthonormal basis \(\{\phi_j\}\) of \(L^2[0,1]\) consisting of eigenfunctions of \(\mathcal K\), whose corresponding eigenvalues are all positive and form a decreasing sequence \(\{\lambda_j\}\to 0\), and such that
+Then, there exists an orthonormal basis \(\{\phi_j\}\) of \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive and form a decreasing sequence \(\{\lambda_j\}\to 0\), and such that
     \[
@@
   \[
-    \zeta_j = \int_0^1 X(t)\phi_j(t)\, dt.
+    \zeta_j = \int_0^1 X(t)\phi_j(t)\, dt, \quad j\geq 1.
   \]
@@
 
-In addition to these results, it is worth mentioning that the covariance operator \(\mathcal K\) provides several alternative definitions of \(\Hcal(K)\).
-For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}\).
+In addition to these results, it is worth mentioning that the covariance operator \(\K\) provides several alternative definitions of \(\Hcal(K)\).
+For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\K^{-1/2}(f)}{\K^{-1/2}(g)}\).
 Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
@@
 \[
-\mathcal L_0(X) = \left\{\sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
+\Lcal_0(X) = \left\{\sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
 \]
-It is well known that its completion in \(L^2(\Omega)\), denoted \(\mathcal L(X)\), is a Hilbert space, sometimes called the Hilbert space generated by \(X\).
-In the words of \citet{berlinet2004reproducing}, ``\(\mathcal L(X)\) contains the random variables attainable by linear operations, including limits, on the measurements of the process''.
-As it happens, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and \(\mathcal L(X)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
+It is well known that its completion in \(L^2(\Omega)\), denoted \(\Lcal(X)\), is a Hilbert space, sometimes called the Hilbert space generated by \(X\).
+In the words of \citet{berlinet2004reproducing}, ``\(\Lcal(X)\) \textit{contains the random variables attainable by linear operations, including limits, on the measurements of the process}''.
+As it happens, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and \(\Lcal(X)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
 This isometry is essentially the completion of the correspondence
@@
   \end{equation}
-and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
-Thus, \(\Hcal(K)\) can be regarded (and it is often done in statistics) as an isometric copy of \(\mathcal L(X)\).
+and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \Lcal(X)\).
+Thus, \(\Hcal(K)\) can be regarded as an isometric copy of \(\Lcal(X)\), an approach that is often useful in statistics.
 
@@
 It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
-For instance, the trajectories of the Brownian motion are known to be nowhere differentiable, but as shown in Example \ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere.
+For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere.
 An heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
@@
 \end{proposition}
-Note that the statement \((iii)\) above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\mathcal L(X)\), respectively.
+Note that the statement \((iii)\) above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.
 
@@
 Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning.
-For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \mathcal H\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS.
+For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS.
 Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
 \[
-K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\mathcal H}.
+K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
 \]
@@
 \begin{theorem}[Generalized representer theorem]
-  Consider a RKHS \(\Hcal(K)\) of functions defined over a non-empty set \(\mathcal X\), a training set \(\{(x_i, y_i)\}_{i=1}^p \subset (\mathcal X \times \R)^p\), an arbitrary error function \(R: (\mathcal X  \times \R^2)^p \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
+  Consider a RKHS \(\Hcal(K)\) of functions defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
 Then, any minimizer \(f^\ast\) of the regularized empirical risk
   \[
-    R(\{(x_i, y_i, f(x_i))\}_{i=1}^p) + \Omega(\|f\|_K)
+    R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
   \]
-  is of the form \(f^\ast(\cdot) = \sum_{i=1}^p a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,p\).
+  is of the form \(f^\ast(\cdot) = \sum_{i=1}^n a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,n\).
 \end{theorem}
@@
 Then, sampling from \(g\) and accepting the proposal, say \(x\), with probability \(f(x)/Mg(x)\) yields an approximate set of samples from \(f\).
-Nevertheless, these kinds of algorithms are rapidly affected by the so-called \textit{curse of dimensionality}, so new techniques are needed when sampling from multidimensional distributions.
-This is where Markov chain Monte Carlo (MCMC) methods come into play \citep[see e.g.][]{brooks2011handbook}.
+Nevertheless, these kinds of algorithms are rapidly affected by the so-called \textit{curse of dimensionality} (see Figure \ref{fig:curse_dimensionality}), so new techniques are needed when sampling from multidimensional distributions.
+This is where Markov chain Monte Carlo (MCMC) methods come into play.
 
@@
   \includegraphics[width=.25\textwidth]{cube_sphere}
-  \caption{To cover the volume of \(f(x)\) in higher dimensions we intuitively need a volume that grows exponentially with the dimension itself.}
+  \caption{To cover the volume of \(f(x)\) in higher dimensions we intuitively need a volume that grows exponentially with the dimension itself.}\label{fig:curse_dimensionality}
 \end{figure}
@@
 The dimensionality issues are somewhat mitigated, and on top of that there are specific tuning techniques to tackle them.
-However, the samples obtained from this procedure are not independent, though they can be made so by \textit{thinning} the chain and considering, say, only one every \(M\) samples.
+However, the samples obtained from this procedure are not independent, though they can be made so by \textit{thinning} the chain and considering only one every few samples.
 
@@
    \centering
-   \includegraphics[width=.6\textwidth]{mh}
-   \caption{Representation of the Metropolis-Hastings algorithm (taken from   \citet{jaewook2015metamodel}). The discarded proposals are represented by dashed circles.}
+   \includegraphics[width=.55\textwidth]{mh}
+   \caption{Representation of the Metropolis-Hastings algorithm (taken from \citet{jaewook2015metamodel}). The discarded proposals are represented by dashed circles.}\label{fig:mh}
  \end{figure}
 
+Figure~\ref{fig:mh} shows a schematic representation of the process.
 Note that \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\).

### 2022-08-26 thesis/chapters/bayesian-functional-regression.tex
Revision up to chapter 4 complete

--- 2022-08-23 thesis/src/bayesian-functional-regression.tex
@@
 
-\begin{outcomment}
-  Reescribir dejando p libre.
-\end{outcomment}
+In this chapter we present the precise functional models and Bayesian methodologies explored in this work.
+The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by substituting the functional parameter \(\beta\in L^2[0,1]\) for \(\alpha \in \Hcal(K)\), replacing also the scalar product \(\dotprod{X}{\beta}\) for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}, i.e.:
+\begin{equation}\label{eq:alpha_h0}
+\alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot), \text{ for some } p \in \N, \ \beta_j \in \R \text{ and } t_j \in [0,1].
+\end{equation}
+Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j)\) when \(\alpha\) is as in \eqref{eq:alpha_h0}.
 
-In this chapter we present the precise models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by taking a functional parameter \(\alpha \in \Hcal(K)\) and replacing the scalar product for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
-However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}.
-For practical and computational reasons, the value of \(p\), the dimensionality of the model, will be fixed beforehand in a suitable way (see Section~\ref{sec:model-choice} for details), and we will regard \(\beta_j\) and \(t_j\) as free parameters.
-Thus, we will search for our functional parameter in the set
+In this way we get a simpler, finite-dimensional approximation of the functional RKHS model, which we argue reduces the overall complexity of the model while still capturing most of the relevant information.
+When it comes to parameter estimation, a direct optimization of some loss function would probably require a tailored algorithm that took into account the continuous nature of the times \(t_j\).
+Indeed, such an idea is explored in \citet{berrendero2018functional} for the logistic regression case, where the authors propose a ``greedy max-max'' method reminescent of the EM algorithm that alternates between estimating the coefficients and the time instants through a maximum likelihood approach.
+
+At this point we propose to follow a Bayesian approach to estimate the parameters of the model, which we believe is in line with the idea of simplicity we pursue, and also introduces an additional layer of flexibility in the model.
+In this way, we can include problem-specific information in the model through the use of prior distributions, and on top of that, this method works almost unaltered for both linear and logistic regression models.
+The general idea will be to impose a prior distribution on the functional parameter to eventually derive a posterior model after incorporating the available sample information.
+In view of~\eqref{eq:alpha_h0}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_0(K)\)) it suffices to first consider a discrete distribution on \(p\), and then impose \(p\)-dimensional continuous prior distributions on the coefficients \(\beta_j\) and the times \(t_j\) given \(p\).
+Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
+Moreover, note that starting from a probability distribution \(\P_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\P\) on \(\Hcal(K)\) merely by defining \(\P(B) = \P_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
+Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
+
+However, after some intial experimentation we found that, for practical and computational reasons, the value of \(p\), the dimensionality of the model, is best fixed beforehand in a suitable way (see Chapter~\ref{ch:model-choice} for details).
+Thus, we will regard only the \(\beta_j\) and \(t_j\) as free parameters, and search for our functional parameter in the space
 \begin{equation}\label{eq:h0p}
@@
 \end{equation}
+Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of feasible values is not very large in practice.
+Moreover, we could think of this approach as imposing a degenerate prior distribution on \(p\), so it is in a way a particular case of the more general model discussed above.
 
-The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.
-Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\alpha \in \Hcal_{0,p}(K)\) and that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\) when \(\alpha(\cdot)=\sum_j\beta_j K(t_j, \cdot)\).
-
-In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately.
-Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods).
-Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) merely by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
-Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
-Even though we actually work on \(\Hcal_{0,p}(K)\), the discrete parameter \(p\) can still be selected in several meaningful ways that make use of the available data, and the set of feasible values is not very large in practice.
-
-In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
+In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\) (which will be specified shortly), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d.
 samples amounts to
 \begin{equation}\label{eq:bayes-theorem}
-  \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
+  \pi(\theta \mid \D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
 \end{equation}
 In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y | X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.
-Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via MCMC methods \incomment{Referir al capítulo de model choice para detalles}.
-Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.
+Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can still be accomplished via MCMC methods (see Chapter~\ref{ch:model-choice} for implementation details).
 
@@
 \end{equation}
-where \(\alpha(\cdot)=\sum_{j=1}^p\beta_jK(t_j, \cdot) \in \Hcal_{0,p}(K), \beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is a normally-distributed error term independent from \(X\), with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\).
+where \(\alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot) \in \Hcal_{0,p}(K)\), \(\alpha_0\in\R\), and \(\epsilon \sim \mathcal N(0, \sigma^2)\) is an error term independent from \(X\).
 This model is essentially a finite-dimensional approximation from a functional perspective to the more general RKHS model that assumes \(\alpha \in \Hcal(K)\), proposed in~\citet{berrendero2019rkhs}.
 
-When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
-Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\).
+When \(p\) is fixed, the parameter space becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this \((2p+2)\)-dimensional space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
+Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\D_n\).
 For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
@@
 
-It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
+It is worth mentioning that the model remains linear in the sense that it fundamentally involves random variables belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
 Also, note that given the time instants \(t_j\), the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
@@
 
-\subsubsection{The bayesian approach: prior selection and posterior derivation}
+\subsection*{The bayesian approach: prior and posterior}
 
@@
   \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
-  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2{\underbrace{\left(\mathcal X_\tau' \mathcal X_\tau + \eta I\right)}_{G_\tau}}^{-1}),
+  \tau                     & \sim \U([0, 1]^p),                                              \\
+  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2{\underbrace{\left(\X_\tau' \X_\tau + \eta I\right)}_{G_\tau}}^{-1}),
 \end{split}
 \end{align}
-where \(I\) is the identity matrix, \(\mathcal X_\tau\) is the data matrix \((X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
-On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}.
+where \(I\) is the identity matrix, \(\X_\tau\) is the data matrix \((X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model.
+On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known in the standard linear regression setting as Jeffrey's prior \citep{jeffreys1946invariant}.
 In any event, the estimation of \(\alpha_0=\E[Y]\) is straightforward, so it could have been left out of the model altogether.
@@
 
-As for the posterior distribution, we only compute a function proportional to its log-density, since that is all that is needed for the MCMC algorithm to work.
+As for the posterior distribution, we only compute a function proportional to its log-density, since that is all that is needed for a MCMC algorithm to work.
 A standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.
@@
 \begin{align*}
-\log \pi(\theta \mid \mathcal D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\symbf Y- \alpha_0\symbf{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
+\log \pi(\theta \mid \D_n) \propto {} & \frac{1}{2\sigma^2}\left(\|\symbf Y- \alpha_0\symbf{1} - \X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right)\\
 & + (p+n+2)\log\sigma - \frac{1}{2}\log |G_\tau|,
@@
 
-\subsubsection{Making predictions}
+\subsection*{Making predictions}
 
-In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \mathcal D_n)\).
-Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
+In order to generate predictions, let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}=(b^{(m)*}, \tau^{(m)*}, \alpha_0^{(m)*}, (\sigma^2)^{(m)*})\) of the posterior distribution \(\pi(\theta| \D_n)\).
+Assuming now a previously unseen test set \(\D'_{n'}\) in the same conditions as \(\D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.
 
-  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\) and predict the responses in the usual way following model \eqref{eq:rkhs-model-linear}, i.e.:
+  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) that acts as a summary of the marginal posterior distributions, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\), and then predict the responses in the usual way following model~\eqref{eq:rkhs-model-linear}, i.e.:
   \[
@@
 
-  \paragraph{Variable selection.} Lastly, we can focus only on the marginal posterior distribution of \(\tau|\mathcal D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
-Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
+  \paragraph{Variable selection.} Lastly, we can focus only on the marginal posterior distribution of \(\tau|\D_n\) and select \(p\) time instants using a point-estimate statistic \(T\) as in our first strategy, but discarding the rest of the parameters.
+Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix given by \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
 After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
@@
 
-In the case of functional logistic regression, we regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, an RKHS model might be given  by the  equation
+In the case of functional logistic regression, we can regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and as suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
+Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model is given  by the  equation
 \begin{equation}\label{eq:rkhs-model-logistic}
-  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
+  \P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
 \end{equation}
@@
 \begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
-  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
+  p_i = \P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
 \end{equation}
@@
 First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=i\) (\(i=0,1\)) are homoscedastic Gaussian processes~\citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for such model.
+Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model.
 Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
@@
 However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.
-\color{black}
 
-\subsubsection{The Bayesian approach: prior selection and posterior derivation}
+\subsection*{The Bayesian approach: prior and posterior}
 
@@
 \begin{align*}
-  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
+  \log \pi(\theta \mid \D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
   \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
@@
 
-\subsubsection{Making predictions}
+\subsection*{Making predictions}
 
 Bear in mind that in this case we are essentially approximating the probabilities \(p_i\) in~\eqref{eq:rkhs-model-logistic-2-parameter}, so before producing a response we need to transform the predicted values to a binary output in \(\{0, 1\}\).
-According to the usual criterion of minimizing the misclassification probability, the Bayes optimal rule is recovered by predicting \(\hat Y=1\) whenever \(\mathbb P(Y=1|X) \geq 1/2\).
-Nevertheless, for a more general cost function one could consider other criteria that would lead to evaluating whether \(\mathbb P(Y=1|X) \geq \gamma\) for some threshold \(\gamma\in[0, 1]\).
+According to the usual criterion of minimizing the misclassification probability, it is known that the Bayes optimal rule is recovered by predicting \(\hat Y=1\) whenever \(\P(Y=1|X) \geq 1/2\).
+Nevertheless, for a more general cost function one could consider other criteria that would lead to evaluating whether \(\P(Y=1|X) \geq \gamma\) for some threshold \(\gamma\in[0, 1]\).
 
@@
 \[
-\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
+\hat Y_i = \I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
 \]
-where \(\mathbb I\) is the indicator function (\(\mathbb I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
+where \(\I\) is the indicator function (\(\I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
 The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
@@
 the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
-Hence, there are primarily two possible outcomes:
-\begin{description}
-  \item[Average sampled probability.] If we choose to average the approximate probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
+Hence, there are primarily two possible outcomes.
+
+  \paragraph{Average sampled probability.} If we choose to average the approximate probabilities \(p_i^{(m)*} = \P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
   \[
-    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
+    \hat Y_i = \I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
   \]
-  \item[Average sampled response.] Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
+  \paragraph{Average sampled response.} Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
   \[
-  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
+    \hat Y_i = \I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
   \]
@@
 Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).
-\end{description}
 
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \mathcal D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
+Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.

### 2022-08-26 thesis/chapters/experiments.tex
Revision up to chapter 4 complete

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2022 Antonio Coín
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Experiments}\label{ch:experiments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{outcomment}
  Incluir figuras de los datasets, y en general repasar las figuras del notebook.

  Enlazar al github, y comentar que hay demos en jupyter.

  - Note that the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure.

  - Ejemplo en RKHS con p=3 y 2 componentes subyacentes: mostrar todas las gráficas (+ppc).

  - Ejemplo dejando todo constante y aumentando \(p\). Quizás también otro variando el tamaño del grid y/o el número de ejemplos de entrenamiento? Bidimensional en ntrain/ngrid?
\end{outcomment}

### 2022-08-26 thesis/chapters/introduction.tex
Revision up to chapter 4 complete

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2022 Antonio Coín
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}\label{ch:introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{center}
%   \begin{minipage}{0.5\textwidth}
%     \begin{small}
%       \emph{Quote}
%     \end{small}
%   \end{minipage}
%   \vspace{0.5cm}
% \end{center}

Over the last few decades, situations involving data in the form of functions have become commonplace in many statistical scenarios, as more and more information is available worldwide with an ever-increasing level of granularity in the measurements. In particular, functional data problems are far from unheard of in the data science an machine learning community, since they have attracted the attention of researchers and practitioners equally. Medical data, weather indicators or stock exchange indices are examples of elements that benefit from a functional treatment, where the observations are regarded as single entities rather than as a conglomerate of individual points.

Under a functional framework, the objects of interest are \textit{random functions} instead of random points in a finite-dimensional space. While in principle the functional data could be simply regarded as a discretized vector in a very high dimension (an indeed such a discretization is performed in practice), there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question. Thus, the general idea is to assume the existence of an underlying sufficiently smooth function that generates each (possibly noisy) functional observation, even though we only record it on a finite grid of points.

To see what this kind of data looks like, Figure~\ref{fig:tecator_orig} shows an example of a functional data set, which is known in the literature as the Tecator data set, and whose elements represent near-infrared absorbance curves of meat samples. The objective here is to predict the fat content based on this absorbance spectrum, separating the samples into those with ``high'' and ``low'' fat content. At first glance it does not seem that the trajectories contain much relevant information to help classify the samples. However, after a suitable smoothing of the data (e.g. by representing each function in a Fourier basis) we can take the derivatives of the curves. In this case, after differentiating twice a clearer pattern emerges (see Figure~\ref{fig:tecator_derivatives}), one from which inference and prediction will surely be easier.

\begin{figure}[ht!]
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{tecator}
    \caption{Original curves}\label{fig:tecator_orig}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{tecator_derivative}
    \caption{Second derivatives}\label{fig:tecator_derivatives}
  \end{subfigure}
  \caption{Curves in the Tecator data set and their second derivatives (after smoothing).}\label{fig:tecator}
\end{figure}

In recent years numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum. A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}. As the name suggests, FDA techniques are heavily inspired by functional analysis tools and methods: Hilbert spaces, orthonormal systems, linear operators, and so on. In particular, a notion that also intersects with the classical theory of machine learning and pattern recognition is that of reproducing kernel Hilbert spaces (RKHS's). We will demonstrate throughout this work how these spaces of functions possess properties that allow for an efficient treatment of functional data. On the other hand, Bayesian inference methods are ubiquitous in the realm of statistics, and their usual non-parametric approach also makes use of random functions, though in a slightly different manner than in the FDA context. However, the two methodologies can certainly interact and benefit from one another, as we intend to show in this thesis. We will be particularly interested in Markov chain Monte Carlo (MCMC) methods, which allow us to approximate an arbitrary posterior distribution through a function proportional to its density.

\begin{center}
\color{teal}\FourStar
\end{center}

In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations. Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions. Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios amply explored in the literature. To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e. a process with finite second moments and whose realizations are square-integrable functions indexed on \([0,1]\)). The underlying probability space \((\Omega, \mathcal A, \P)\) is not important. We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\). In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(x_i, y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\). Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\). Figure~\ref{fig:linear_data_example} depicts a typical data set used in functional regression, and we already saw in Figure~\ref{fig:tecator} what a functional classification data set may look like.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\textwidth]{linear_data_example}
  \caption{Simulated data set for functional regression. On the left we have \(n=50\) functional observations on an equispaced grid of \(N=100\) points on \([0,1]\). To each observation corresponds a real number; the distribution of these responses is shown on the right.}\label{fig:linear_data_example}
\end{figure}

\section{Objectives and scope}

We list below the main objectives of this work in no particular order.

\begin{enumerate}[1.]
  \item To perform a brief but thorough literature review that contextualizes functional linear and logistic regression in statistics and machine learning, exploring the predominant models and techniques.
  \item To propose a novel functional model for linear and logistic regression that builds on existing work and focuses on simplicity, both in terms of interpretation and implementation.
  \item To describe a general Bayesian approach for parameter estimation within the suggested model.
  \item To introduce the tools needed to specify and implement the proposed model, mainly reproducing kernel Hilbert spaces and Markov chain Monte Carlo methods.
  \item To carry out a comprehensive experimental study to test these new models and compare them to existing methods, both in simulations and in real-world scenarios.
\end{enumerate}

It is beyond the scope of this thesis to provide an extensive review of FDA as a whole, or to delve too deeply into the details and inner workings of most models and techniques mentioned. Nevertheless, references are provided frequently throughout the text to point the interested reader towards more specialized resources.

\section{Structure overview}

In Chapter~\ref{ch:background} we summarize the relevant literature related to our problem, along with a review of the basics of RKHS's and MCMC methods. Chapter~\ref{ch:bayesian} is devoted to explaining the Bayesian methodology and the functional regression models we propose. Then, in Chapter~\ref{ch:model-choice} we present a short discussion of theoretical and computational details that have led to the concrete specification of the model, as well as some validation techniques. The empirical results of the experimentation are contained in Chapter~\ref{ch:experiments}. Lastly, the conclusions drawn from this work and future paths of research are presented in Chapter~\ref{ch:conclusions}.

### 2022-08-26 thesis/chapters/model-choice.tex
Revision up to chapter 4 complete

--- 2022-08-23 thesis/src/model-choice.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
 
-\paragraph{The choice of \(\symbf p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
+\section{Hyperparameter selection}
+
+\subsection*{Label switching and the choice of \(p\)}
+
+One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
 While theoretically we could impose a prior distribution on \(p\) as well (e.g.
@@
 
-\paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
+\subsection*{Other hyperparameters}
+
+As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
 \begin{itemize}
@@
 
+\section{MCMC implementation}
+
 \begin{outcomment}
@@
 
+\section{Validation techniques}
+
 \paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.

### 2022-08-26 thesis/chapters/background-related-work.tex
Start Ch.4

@@
 However, unless conjugate priors are used, a closed-form expression is generally unattainable and some type of approximation is required.
-Apart from basic numerical integration, some well-performing methods in this regard include variational inference approaches \citep[e.g.][]{blei2017variational} and MCMC methods \citep[e.g.][]{brooks2011handbook}.
+Apart from basic numerical integration, some well-performing methods in this regard are variational inference approaches \citep[e.g.][]{blei2017variational} and MCMC methods \citep[e.g.][]{brooks2011handbook}.
 The former are based on approximating the posterior by another distribution restricted to a certain parametric family so that the Kullback-Leibler divergence between them is minimized, whereas the latter are iterative methods that directly provide approximate samples of the posterior distribution.
 
-As a comment, there are some works that tackle Bayesian inference from a functional perspective; a recent example are the functional Bayesian neural networks proposed by~\citet{sun2019functional}.
+As a comment, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by Bayesian neural networks.
+Some examples are the functional Bayesian neural networks proposed by~\citet{sun2019functional}, variational implicit processes \citep{ma2019variational} and their deep variants \citep{ortega2022deep}, or the functional variational inference approach suggested in~\citet{ma2021functional}.
 

### 2022-08-26 thesis/chapters/experiments.tex
Start Ch.4

@@
   - Ejemplo dejando todo constante y aumentando \(p\). Quizás también otro variando el tamaño del grid y/o el número de ejemplos de entrenamiento? Bidimensional en ntrain/ngrid?
+
+  - Sobre experimentos con p-free: Por otro lado, en algunas cadenas el sigma2 es más grande de la cuenta, ya que tiende a "compensar" la falta de flexibilidad del modelo (e.j.
+si el valor de p es más bajo).
+Esto hace que los estimadores basados en la media y en el posterior mean sean ligeramente peores (sobre todo el de la media puntual, donde la varianza estimada sí se ve notablemente afectada).
+Quizás para contrarrestar esto habría que tener un modelo con un $\sigma^2_p$ para cada valor de p?
 \end{outcomment}

### 2022-08-26 thesis/chapters/model-choice.tex
Start Ch.4

@@
 
-In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
+In this chapter we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.\incomment{completar}
 
-\section{Hyperparameter selection}
+\section{Hyperparameter selection \incomment{cambiar}}
 
@@
 
-One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we could impose a prior distribution on \(p\) as well (e.g.
-a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
-In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time.
-Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
-In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, DIC, cross-validation, \ldots) to select its optimal value.
-This is equivalent to imposing a degenerate prior distribution on \(p\).
-Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}).
-\incomment{Reversible MCMC}
+A phenomenon found in  label switching, ...
+leads to non-identifiability of the components.
+This happens because the likelihood is symmetric with respect to the ordering of the parameters \(b\) and \(\tau\).
+For this reason, since the components are arbitrarily ordered, they may be exchanged from one iteration to the next in any MCMC algorithm.
+This can cause nonsensical answers when using the marginal posterior distributions to perform inference, as different labelings might be mixed on each component.
 
-%Reversible-jump MCMC
-% Estos métodos permiten variar la dimensionalidad subyacente del espacio muestral en cada propuesta (\sidecite[*0]{green1995reversible}).
-Teóricamente son útiles para muestrear de mixturas de gaussianas, procesos de Dirichlet o, en general, de cualquier distribución donde haya una variable latente que represente la dimensión del espacio.
-Sin embargo, no hay una implementación de referencia establecida.
+However, this occurrence is a condition for the convergence: no switching indicates that not all modes of the posterior distribution are being explored by the sampler, as pointed out by many authors \citep[see for example][]{celeux2000computational}.
+For this reason, many ad-hoc solutions revolve around post-processing or relabeling the samples to eliminate the switching effect.
 
 \begin{outcomment}
-Reescribir label switching (nos pasa; lo que pasa que con p free se acentua).
-
-Ocurre porque el likelihood es simetrico: L(theta mid y) = L(thetaperm mid y) para cualquier permutacion (hay p! posibles) --> no-identificabilidad.
-
-"In these situations, parameter estimates and other output are inevitably aggregated
-across numerous replications, and because the labels of the enumerated classes are
-arbitrarily assigned, there is no guarantee that the same class would have the same
-label from replication to replication."
-
-From a Bayesian pointof view, this amounts to truncating the original prior distribution... (to verify the constraints)
-
-As a result, in any MCMC algorithm, labels of the
-components can permute multiple times between iterations of the sampler.
-Paradoxically, as noted by Celeux, Hurn and Robert (2000), Fruhwirth-Schnatter (2001),
-Jasra, Holmes and Stephens (2005) and Papastamoulis and Iliopoulos (2010), among others,
-label switching is a prerequisite for MCMC convergence. If there is no label switching it
-means that the sampler is not exploring all the modes of the posterior distribution of (1)
-
-"When sampling from the unconstrained posterior via MCMC methods, which component of the sampled parameter corresponds to which state is unknown due to potential label switching." Si usamos la marginal para haacer inferencia --> nonsensical answers (Stephens, 2000)
-
-Solución: añadir constraints de identificabilidad artificiales (ej relabeling, mantener ordenados, posprocesado).
-
 Imposing an identifiability constraint on the parameters breaks the sym-
@@
 
-Aunque se puede tener en cuenta solo las primeras 'p' posiciones, en el propio algoritmo de MCMC no se puede 'apagar' fácilmente el valor de los betas que no se usan, por lo que crecen indefinidamente y generan problemas (lo mismo ocurre con los 'tau', que incluso se pueden salir de rango).
-Aunque una posible solución es limitar todos los taus y no solo los p primeros, esto solo retrasa el problema (con cadenas más largas vuelve a ocurrir, además de ocasionar el problema "RuntimeWarning:covariance is not positive-semidefinite." con el WalkMove), ya que los betas no están limitados.
-Podríamos intentar limitarlos a un valor razonable.
+\incomment{conector} one of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
+While theoretically we can impose a prior distribution on \(p\) as well (e.g.
+a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
+For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
+In this case we would have no immediate way of tracking down which set of parameters is ``active'' at any given time.
+A simple approach would be to always consider the first \(p\) parameters and ignore the rest, and we did indeed try this technique, but it gave rise to new difficulties and the results obtained were not good.
+In fact, the label switching issue is accentuated when \(p\) is allowed to vary \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, and on top of that, the interpretation of, say, the first coefficient \(\beta_1\) in a model with \(3\) components is different than the interpretation of the same coefficient in a model with only \(2\).
 
-Por otro lado, en algunas cadenas el sigma2 es más grande de la cuenta, ya que tiende a "compensar" la falta de flexibilidad del modelo (e.j.
-si el valor de p es más bajo).
-Esto hace que los estimadores basados en la media y en el posterior mean sean ligeramente peores (sobre todo el de la media puntual, donde la varianza estimada sí se ve notablemente afectada).
-Quizás para contrarrestar esto habría que tener un modelo con un $\sigma^2_p$ para cada valor de p?
+\incomment{Reversible MCMC}
 
-Además, aunque todo esto funcionase, si por ejemplo en una iteración el modelo elige p=2, está optimizando beta1 y beta2 para ese caso; si luego miramos solo la marginal de beta1, se habrán mezclado los casos en los que p=1 y beta1 debe valer una cosa, y en los que p=2 y beta1 vale otra cosa potencialmente distinta (se puede compensar con beta2, hay más grados de libertad).
+% Estos métodos permiten variar la dimensionalidad subyacente del espacio muestral en cada propuesta (\sidecite[*0]{green1995reversible}).
+Teóricamente son útiles para muestrear de mixturas de gaussianas, procesos de Dirichlet o, en general, de cualquier distribución donde haya una variable latente que represente la dimensión del espacio.
+Sin embargo, no hay una implementación de referencia establecida.
+
+Another possibility would be to adapt a purely Bayesian model selection technique to our framework \citep[see][]{piironen2017comparison, gelman2013bayesian}, \incomment{más detalles?} or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
+In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
+BIC, DIC, cross-validation, \ldots) to select its optimal value.
+Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Chapter~\ref{ch:experiments}).
 
@@
 
+  Sección para autocorrelation times.
+
   También en el artículo citado se aconseja inicializar las cadenas en una bola gaussiana alrededor de un punto que se espere que tenga alta probabilidad según \(p(x)\).

### 2022-08-30 chapters/appendix-code.tex
Finish up to Ch.6

--- 2022-08-26 thesis/chapters/model-choice.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Model choice, implementation and validation}\label{ch:model-choice}
+\chapter{On the code developed}\label{ch:code}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-In this chapter we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.\incomment{completar}
+\begin{outcomment}
+  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks.
 
-\section{Hyperparameter selection \incomment{cambiar}}
+  Contar que son clases a la sklearn y compatibles con ska-fda. Enlazar al github.
 
-\subsection*{Label switching and the choice of \(p\)}
+  El proceso es reproducible (semilla 2022)
 
-A phenomenon found in  label switching, ...
-leads to non-identifiability of the components.
-This happens because the likelihood is symmetric with respect to the ordering of the parameters \(b\) and \(\tau\).
-For this reason, since the components are arbitrarily ordered, they may be exchanged from one iteration to the next in any MCMC algorithm.
-This can cause nonsensical answers when using the marginal posterior distributions to perform inference, as different labelings might be mixed on each component.
+  Se usan 4 cores.
 
-However, this occurrence is a condition for the convergence: no switching indicates that not all modes of the posterior distribution are being explored by the sampler, as pointed out by many authors \citep[see for example][]{celeux2000computational}.
-For this reason, many ad-hoc solutions revolve around post-processing or relabeling the samples to eliminate the switching effect.
-
-\begin{outcomment}
-Imposing an identifiability constraint on the parameters breaks the sym-
-metry of the posterior distribution of the parameters, and so we might hope that it would allow us to perform sensible inference for the individual components of the mixture.
-Inference conditional on such a constraint may be performed by post-processing the sample (proposition 3.1 in Stephens 1997 phD thesis).
-
-OTRA OPCION: Elegir en cada iteración como parámetro para ordenar aquel que deja mayor separación (tras escalarlos todos a [0,1]): \verb@https://helda.helsinki.fi/bitstream/handle/10138/329485/Approximate_Bayesian_computation_for_finite_mixture_models.pdf?sequence=1@
-
-OTRA SOLUCION: "Adaptive Metropolis with Online Relabeling" --> The key step is the selection, after each proposal of a permutation (relabeling) of the proposed vector by minimizing a quadratic cost function
-$L_{t-1} (x)$ over all permutations of the proposed vector, where L is defined by
-$L_t(x) = (x - \mu)'\Sigma^{-1}(x - \mu)$
-This step forces the posterior sample to look as unimodal as possible.
-
-Citas: Stephens (2000), Papastamoulis \& Iliopoulos, 2010; Rodriguez \& Walker, 2014
 \end{outcomment}
-
-\incomment{conector} one of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we can impose a prior distribution on \(p\) as well (e.g.
-a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
-For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next.
-In this case we would have no immediate way of tracking down which set of parameters is ``active'' at any given time.
-A simple approach would be to always consider the first \(p\) parameters and ignore the rest, and we did indeed try this technique, but it gave rise to new difficulties and the results obtained were not good.
-In fact, the label switching issue is accentuated when \(p\) is allowed to vary \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, and on top of that, the interpretation of, say, the first coefficient \(\beta_1\) in a model with \(3\) components is different than the interpretation of the same coefficient in a model with only \(2\).
-
-\incomment{Reversible MCMC}
-
-% Estos métodos permiten variar la dimensionalidad subyacente del espacio muestral en cada propuesta (\sidecite[*0]{green1995reversible}).
-Teóricamente son útiles para muestrear de mixturas de gaussianas, procesos de Dirichlet o, en general, de cualquier distribución donde haya una variable latente que represente la dimensión del espacio.
-Sin embargo, no hay una implementación de referencia establecida.
-
-Another possibility would be to adapt a purely Bayesian model selection technique to our framework \citep[see][]{piironen2017comparison, gelman2013bayesian}, \incomment{más detalles?} or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
-In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
-BIC, DIC, cross-validation, \ldots) to select its optimal value.
-Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Chapter~\ref{ch:experiments}).
-
-\subsection*{Other hyperparameters}
-
-As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
-\begin{itemize}
-  \item For the expected value \(b_0\) we propose to use the MLE of \(b\).
-Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation).
-Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
-  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
-  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
-This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
-\end{itemize}
-
-\paragraph{Summarizing the posterior.} On a separate note, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
-At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
-\incomment{Argue that, because of the label switching and multimodality, summarizing the posterior might be impractical, unhelpful, misleading, etc.
-So the best strategy is the posterior mean.}
-
-\section{MCMC implementation}
-
-\begin{outcomment}
-  Sección para elección del algoritmo mcmc.
-Contar que también se usa pymc (PPL, probabilistic programming language) y que en particular se probaron metropolis y NUTS: "Dentro de estos métodos destaca el \textit{No U-Turn Sampler} (NUTS, \citet{hoffman2014no}), que tiene la ventaja añadida de que realiza un ajuste automático de hiperparámetros."
-
-  % Otra de las ventajas que ofrece este método frente a algoritmos clásicos es que solo requiere fijar unos pocos hiperparámetros (independientes de la dimensión subyacente), en lugar de los \(O(N^2)\) correspondientes, por ejemplo, a la matriz de covarianza de la distribución de salto \(N\)-dimensional en Metropolis-Hastings.
-Además, el paquete ha sido extendido para admitir otro tipo de movimientos paralelos, y el objetivo del proyecto es proporcionar un método de muestreo de propósito general que funcione correctamente en una clase amplia de problemas.
-Es muy usado por ejemplo en el campo de la astrofísica, y está bien establecido y testeado.
-\end{outcomment}
-
-\begin{outcomment}
-  Sección para inicialización.
-
-  Sección para autocorrelation times.
-
-  También en el artículo citado se aconseja inicializar las cadenas en una bola gaussiana alrededor de un punto que se espere que tenga alta probabilidad según \(p(x)\).
-\end{outcomment}
-
-\section{Validation techniques}
-
-\paragraph{Model validation and Bayesian checks.} To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away.
-For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.
-Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(\symbf Y^*)\) of the posterior predictive distribution \(\symbf Y^* \equiv \{\symbf Y^{(m)*}\}\) (see Figure~\ref{fig:ppc}).
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) approximate estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the real value of the statistic.
-They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.
-
-\begin{figure}[ht]
-  \centering
-  %\includegraphics[width=.75\textwidth]{img/ppc_linear}
-  \caption{Posterior predictive graphical checks on a fitted model.
-On the left there is a comparison between the observed distribution of the response variable and the posterior predictive distribution of the approximate sampled responses, while the distribution of the average of the posterior predictive responses is depicted on the right.}\label{fig:ppc}
-\end{figure}

### 2022-08-30 chapters/background-related-work.tex
Finish up to Ch.6

--- 2022-08-26 thesis/chapters/background-related-work.tex
@@
 
-As a comment, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by Bayesian neural networks.
+On a separate note, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by Bayesian neural networks.
 Some examples are the functional Bayesian neural networks proposed by~\citet{sun2019functional}, variational implicit processes \citep{ma2019variational} and their deep variants \citep{ortega2022deep}, or the functional variational inference approach suggested in~\citet{ma2021functional}.
+Moreover, there are approaches to functional regression from a Bayesian perspective that impose a Gaussian process prior on the functional parameter \(\beta\) \citep[e.g][]{lian2016posterior}, and it turns out that the RKHS's corresponding to these Gaussian processes, described for example in~\citet{van2008reproducing}, are useful when studying contraction rates of the posterior distribution.
 
@@
 
-\begin{itemize}
-  \item \textit{Stretch move}. For each walker \(1\leq l \leq L\) another walker \(\Lambda_j \in \Lambda_{-l}(t)\) is chosen at random, and the proposal is constructed as
+  \paragraph{Stretch move.} For each walker \(1\leq l \leq L\) another walker \(\Lambda_j \in \Lambda_{-l}(t)\) is chosen at random, and the proposal is constructed as
   \[
@@
   \]
-
 Supposing \(\R^p\) is the sample space, the corresponding acceptance probability (chosen so that the detailed balance equations are satisfied) is:
@@
 
-  \item \textit{Walk move}. For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is chosen, and the proposed move is
+  \paragraph{Walk move.} For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is chosen, and the proposed move is
 \[
@@
 where \(W\) is a normal distribution with mean \(0\) and the same covariance as the sample covariance of all walkers in \(S_l\).
-The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).
-\end{itemize}
+The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).\\
 

### 2022-08-30 chapters/bayesian-functional-regression.tex
Finish up to Ch.6

--- 2022-08-26 thesis/chapters/bayesian-functional-regression.tex
@@
 Specifically, we can consider the times \(\hat t_j = T\{t_j^{(m)*}\}\) and reduce the original data set to just the \(n\times p\) real matrix given by \(\{X_i(\hat t_j): i=1, \dots,n, \ j=1,\dots,p\}\).
-After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.
+After this variable selection has been carried out, we can tackle the problem using a finite-dimensional linear regression model and apply any of the well-known prediction algorithms suited for this situation.\\
 
@@
 Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
+Also, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
+On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
+At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
 
@@
   In this case, each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
-Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).
+Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).\\
 

### 2022-08-30 chapters/conclusions.tex
Finish up to Ch.6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2022 Antonio Coín
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusions}\label{ch:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{outcomment}
  - Destacar también las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc. En la introducción, conclusiones, experimentos, ..
- Destacar interpretabilidad en los experimentos: e.g. un valor de beta=3.4 se puede interpretar fácilmente en un caso concreto. En general, nuestros modelos tienen menor dimensión.
- Logistic un poco mejor que linear?
- Tiempo de ejecución alto (característico de los algoritmos MCMC), a cambio mucha sencillez.
- CV ha sido pequeño. Con más parámetros, probablemente mejores resultados. También flexibilidad para cambiar las prioris.
- We do not advise to use the mean as a summary statistic.
\end{outcomment}

\section{Future work}

\begin{outcomment}
  - Mejorar pymc.
  - consistencia de la posteriori, robustez de la priori
  - Otras prioris

\end{outcomment}

### 2022-08-30 chapters/experiments.tex
Finish up to Ch.6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2022 Antonio Coín
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\let\epsilon\varepsilon

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Experiments}\label{ch:experiments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this chapter we present the results of the experiments carried out to test the performance of our models in different scenarios. The structure of the code and other minor computational details are discussed in Appendix~\ref{ch:code}.

\subsection*{Experimental setting}

First and foremost, we consider \(T=\{\text{mean, median, mode}\}\) for the summarize-then-predict approach to prediction. In this way in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach, which will we call ``\textit{posterior\_mean}'' for short) and 3 variable selection methods (one for each statistic), while in logistic regression there are 5 prediction methods and 3 variable selection procedures. In each case, after variable selection is performed, we use a penalized multiple linear/logistic regression method to generate the corresponding predictions. All in all, we are looking at 7 (8 in the case of logistic regression) prediction methods, and although all of them are derived from a single MCMC run, we will treat them as separate for the purposes of experimentation. In what follows we relax the notation and use the expressions ``model'', ``prediction method'' and ``algorithm'' almost interchangeably when referring to these 7 (or 8) cases.

We consider \(n=150\) training samples and \(n'=100\) testing samples on an equispaced grid of \(N=100\) points on \([0, 1]\) for the simulated data sets, and we do a 66\%/33\% train/test split on the real data sets. We then perform \(5\)-fold cross validation (CV) on the training set to select the best values of \(p\) and \(\eta\) for each model, and after refitting the best model in each case on the whole training set, we evaluate it and measure the predictive performance on the test set. We look for \(p\) in the set \(\{1,2,\dots,10\}\), while the possible values of \(\eta\) are \(\{10^{-4}, 10^{-3}, \dots, 10^2\}\).

Because of execution time constraints, the hyperparameters of the MCMC method are not part of the CV process, and are selected manually based on an initial set of experiments, as well as recommendations from the original article. We use 64 chains and run them for 900 iterations in total, discarding the first 500 iterations as burn-in. Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains at each iteration, selecting the stretch move (the default) with probability \(0.7\), or the walk move with probability \(0.3\).

Lastly, since the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure, we independently repeat the whole process 10 times (each with a different train/test configuration), and average the results across these executions. The metrics used to evaluate the perfomance of the models are the RMSE for linear regression and the accuracy for logistic regression. Recall that
\[
\operatorname{RMSE}(\{\hat y_i\}) = \sqrt{\frac{1}{n'}\sum_{i=1}^{n'} (y_i - \hat y_i)^2}
\]
and
\[
\operatorname{Accuracy}(\{\hat y_i\}) = \frac{1}{n'} \sum_{i=1}^{n'} \I(y_i = \hat y_i).
\]

\subsection*{Data sets}

We consider a set of functional regressors common to linear and logistic regression problems. They are 4 Gaussian processes (GP's), each with a different kernel function:
\begin{enumerate}
  \item A Brownian motion \(K_1(t,s)=\min\{t,s\}\).
  \item A fractional Browmian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
  \item An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
  \item A Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), with \(\nu=0.2\).
\end{enumerate}
Also, when applicable, we fix a variance \(\sigma^2=0.5\) for the error terms \(\epsilon\).

\paragraph{Linear regression data sets.} We consider three different types of data sets, all with a common value of \(\alpha_0=5\).
\begin{itemize}
  \item A RKHS model where the response is generated as \(Y= 0.5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
  \item A \(L^2\)-model with underlying coefficient function \(\beta(t)=\log(1+4t)\), again for the same four GP's.
  \item A model based on non-GP regressors. Specifically, we use a geometric Brownian motion (GBM) defined as \(X(t)=e^{BM(t)}\), where \(BM(t)\) is a standard Brownian motion. In this case, we consider two data sets, one with a RKHS response and one with a \(L^2\) response, with the same parameters as above.
\end{itemize}
As for the real data sets, we use the (twice-differentiated) Tecator data set introduced in Chapter~\ref{ch:introduction} to predict fat content of 193 meat samples, as well as what we call the Moisture \citep{kalivas1997two} and Sugar \citep{bro1999exploratory} data sets. The first consists of near-infrared spectra of 100 wheat samples and the objective is to predict the samples' moisture content, whereas the second contains 268 samples of sugar fluorescence data in order to predict ash content. The three data sets are measured on a grid of 100, 101 and 115 equispaced points in \([0, 1]\), respectively.

\paragraph{Logistic regression data sets.} Again we consider three different types of data sets,  with a common value of \(\alpha_0=-0.5\).
\begin{itemize}
  \item Four logistic RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
  \[
  \P(Y=1\mid X) = \frac{1}{e^{0.5 +5X(0.1) - X(0.4) - 10X(0.8)}}.
  \]
  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case.
  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's and label them according to their origin. Firstly, we consider a homoscedastic case to distinguish between a standard Browmian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\). Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance \(2\), that is, with kernel \(K(t,s)=2\min\{t,s\}\).
\end{itemize}
In this case there are again three real data sets. The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, trying to predict if their longevity is high or low. The second one is the Berkeley Growth Study data set \citep{tuddenham1954physical}, which records the height of 54 girls and 39 boys over 31 different points in their lives. Finally, we have selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.

\subsection*{Comparison algorithms}

We have included a fairly comprehensive suite of comparison algorithms, chosen among the most common methods used in machine learning and FDA. There are finite-dimensional models that work on the discretized data, variable selection/dimension reduction procedures and purley functional methods.

\paragraph{Linear regression comparison algorithms.} We consider the following methods.
\begin{itemize}
  \item Manual. Dummy variable selection method with a pre-specified number of components (equispaced on \([0, 1]\)).
  \item Lasso. Linear least squares with \(l^1\) regularization.
  \item Ridge. Linear least squares with \(l^2\) regularization.
  \item PLS. Partial least squares for dimension reduction.
  \item PCA. Principal component analysis for dimension reduction.
  \item PLS1. Partial least squares regression \citep[e.g.][]{wegelin2000survey}.
  \item APLS. Functional partial least squares regression proposed by \citet{delaigle2012methodology}.
  \item RMH. Recursive maxima hunting variable selection method proposed by \citet{torrecilla2016feature}.
  \item FLin. Functional \(L^2\) linear regression model with fixed basis expansion.
  \item FPLS1. Functional PLS regression through basis expansion, implemented as in \citet{aguilera2010using}.
  \item FPCA. Functional principal component analysis.
\end{itemize}

\paragraph{Logistic regression comparison algorithms.} All the variable selection and dimension reduction algorithms from above are also considered in this case.
\begin{itemize}
  \item Log. Standard multiple logistic regression with \(l^2\) regularization.
  \item LDA. Linear discriminant analysis.
  \item QDA. Quadratic discriminant analysis.
  \item RKVS. RKHS-based variable selection method proposed in \citet{berrendero2018use}.
  \item APLS. Functional PLS used as a dimension reduction method, as proposed in \citet{delaigle2012achieving} in combination with the nearest centroid algorithm.
  \item FLog. Functional RKHS-based logistic regression algorithm proposed in \citet{berrendero2018functional}.
  \item FLDA. Implementation of the so-called functional linear discriminant analysis in \citet{preda2007pls}.
  \item MDC. Maximum depth classifier \citep[e.g.][]{ghosh2005maximum}.
  \item FKNN. Functional K-nearest neighbors classifier with the \(L^2\)-distance.
  \item FNC. Functional nearest centroid classifier with the \(L^2\)-distance.
\end{itemize}

The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair. In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is in \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,11\}\).

Most algorithms have been taken from the libraries \textit{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/}} and \textit{scikit-fda}\footnote{\url{https://fda.readthedocs.io/en/latest/}}, the first oriented to machine learning in general and the second to functional data analysis in particular. However, some methods were not found in these packages and had to be implemented from scratch. This is the case of the FLDA, FPLS and APLS methods, which we coded following the corresponding articles.

\subsection*{Results display}

We have adopted a visual approach to presenting the experimentation results, using colored graphs to help visualize them. Alternatively, tables representing these results can be consulted instead in Appendix~\ref{ch:tables}.

We show the mean and standard deviation of the RMSE obtained across the 10 random runs, depicting our models in orange and the comparison algorithms in blue. We also show the global mean of all the comparison algorithms with a dashed vertical line, and we exclude extreme negative results from this mean to avoid distortion. Moreover, we separate complete prediction algorithms from two-stage methods, the latter being the ones that perform variable selection or dimension reduction prior to a multiple linear/logistic regression method.

We do not expect our models to defeat all the comparison algorithms in all cases (and indeed they do not), but we intend to show that they are competitive methods that on average match or improve other usual alternatives. After all, \textit{there is no such thing as a free lunch}\footnote{The phrase alludes to the No Free Lunch theorem, which loosely speaking asserts that all machine learning methods are equivalent when their performance is averaged across all possible problems. However, the practical implications of this statement are questionable \citep[see e.g.][]{giraud2005toward}.}!

\section{Functional linear regression}

\subsection*{Simulated data sets}

In Figure~\ref{fig:reg_emcee_rkhs} we see the results for the four GP regressors considered in the RKHS case. This is the most favorable case for us, as the underlying model coincides with our assumed model. Indeed, we can see that in most cases our algorithms are the ones with lower RMSE, save for a few exceptions, notably with the Gaussian kernel. A subsequent analysis showed that this particular data set is especially sensitive to the value of the hyperparameter \(\eta\) in our prior distribution, so a more tailored approach would be needed to obtain better results.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth]{reg_emcee_rkhs}
  \caption{RMSE of predictors for simulated GP data that obeys an underlying RKHS model (lower is better).}\label{fig:reg_emcee_rkhs}
\end{figure}

In Figure~\ref{fig:reg_emcee_l2} we see the results for the case with an underlying \(L^2\)-model, which would be our most direct competitor. In this case the outcome is satisfactory, since for the most part our models are on a par with the rest, even beating other methods that were designed with the \(L^2\)-model in mind. Moreover, whenever one of our models has a higher RMSE, the difference is pretty small in general.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth]{reg_emcee_l2}
  \caption{RMSE of predictors for simulated GP data that obeys an underlying \(L^2\) model (lower is better).}\label{fig:reg_emcee_l2}
\end{figure}

Note that our Bayesian models have a higher standard deviation, partly because there is an intrinsic randomness in the methods (apart from the train/test splits), and it can be the cause of the occasional worse performance. In relation to this, we observe that the methods that use the mean as a summary statistic tend to perform much worse. This is because the mean is very sensitive to outliers, so if at some point there is a chain that randomly deviates from the rest, the mean of the posterior distribution will be greatly impacted. In particular, very high values of the parameter \(\sigma^2\) were reported sometimes, producing misleading results when averaged with the rest of the values.

Lastly, the results for the non-GP case can be seen in Figure~\ref{fig:reg_emcee_nongp}, in which the regressors are realizations of a GBM. In this case we still get better results under the RKHS model, while the results under the \(L^2\)-model are slightly worse. However, as before, the difference is small (except for the \textit{emcee\_mean} methods).

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.7\textwidth]{reg_emcee_nongp}
  \caption{RMSE of predictors for simulated data with GBM regressors (lower is better).}\label{fig:reg_emcee_nongp}
\end{figure}

\subsection*{Real data}

Figure~\ref{fig:reg_emcee_real} shows the results for the real data sets. In these data sets there is a substantial difference in perfomance between some of our methods and the reference algorithms. However, the predict-then-summarize approach (represented as \textit{posterior\_mean}) seems to work quite well, always near the mean RMSE of all the comparison algorithms. Moreover, our two-stage methods seem to perform better than our other methods in these cases, scoring again very close to the mean of the reference methods.

We have to bear in mind that real data is more complex and noisy than simulated data, and it is possible that after a suitable pre-preprocessing we could obtain better results with our methods. However, our goal was to perform a general comparison without exploring too much to the specifics of any particular data set.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.85\textwidth]{reg_emcee_real}
  \caption{RMSE of predictors for real data sets (lower is better).}\label{fig:reg_emcee_real}
\end{figure}

\section{Functional logistic regression}

\subsection*{Simulated data sets}

In Figure~\ref{fig:clf_emcee_rkhs} we see the results for the GP regressors in the logistic RKHS case. Our models perform fairly well in this advantageous case, although the two-stage methods fare somewhat poorly in these data sets. However, in most cases the differences observed account for only one or two misclassified samples.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth]{clf_emcee_rkhs}
  \caption{Accuracy of classifiers for simulated GP data that obeys an underlying logistic RKHS model (higher is better).}\label{fig:clf_emcee_rkhs}
\end{figure}

Continuing with Figure~\ref{fig:clf_emcee_l2}, we see that in the \(L^2\) case the results are again promising, since our models score consistently on or above the mean of the reference models, and in many cases surpassing most of them. Moreover, the overall accuracy of all methods is low (below 60\%), so this is indeed a difficult problem in which even small increases in accuracy are relevant.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth]{clf_emcee_l2}
  \caption{Accuracy of classifiers for simulated GP data that obeys an underlying logistic \(L^2\)-model (higher is better).}\label{fig:clf_emcee_l2}
\end{figure}

Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case. Our variable selection methods are close to the global mean accuracy in both cases.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.7\textwidth]{clf_emcee_nongp}
  \caption{Accuracy of classifiers for simulated data coming from two different GP's, labeled according to their origin (higher is better).}\label{fig:clf_emcee_nongp}
\end{figure}

\subsection*{Real data}

As for the real data sets, in Figure~\ref{fig:clf_emcee_real} we see positive results in general, obtaining in most cases accuracies well above the mean of the reference models, and sometimes above most of them. However, our variable selection methods seem to be a bit worse in this case. Again, the models that use \textit{emcee\_mean} are the exception, and in all these data sets they perform steadily worse than the rest of our models.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.85\textwidth]{clf_emcee_real}
  \caption{Accuracy of classifiers for real data sets (higher is better).}\label{fig:clf_emcee_real}
\end{figure}

\section{Additional experiments}

In this section we present some experiments that are not directly related to measuring the prediction performance relative to other methods.

\subsection*{Model misspecification}

One requirement that our model should satisfy is that it be able to recover the true parameter function when the underlying model is a RKHS model. This is generally true when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?

Take for example a data set generated according to the formula \(Y=-5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\). Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(\tau_1, \tau_2, \tau_3)\) for a model with 3 components.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.8\textwidth]{beta_trace_3}
  \caption{Posterior distribution in a linear model with \(p=3\).}\label{fig:beta_trace_3}
\end{figure}

As we can see, one of the coefficients has gone to zero to account for the overspecification of the model, while the other two have stabilized very close to the true parameters. The same goes for the time instants, except that in this case there is no default value to represent that a component is unused, so the time corresponding to the null coefficient oscillates back and forth. The estimated function, based for example in the mode of the posterior distributions, will not perfect, essentially because of the noise in the response. But it should be close to the true parameter function \(\alpha(t)=-5K(t, 0.1) + 10K(t, 0.8)\), providing a good predictive performance.

In contrast, if we consider now a model with \(p=4\) with the same data, we might obtain posterior distributions such as the ones in Figure~\ref{fig:beta_trace_4}. In this situation two components should go to zero, but that is no longer the case. For example, while the green component has a high density around 0, it also has a considerable mass around 10, effectively ``competing'' with the red component. This is a manifestation of the label switching issue, caused in this case by an excessive number of degrees of freedom in the model.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.8\textwidth]{beta_trace_4}
  \caption{Posterior distribution in a linear model with \(p=4\).}\label{fig:beta_trace_4}
\end{figure}

There is still another possible situation, one in which there is no label switching but the four components are non-negligible in the estimated function. This can happen because the components exploit the additional freedom and ``work together'', so to speak. In this way we could obtain an estimate that might not resemble the true coefficient function, but that could have a very low prediction error. However, this could also work to our detriment and cause the estimated function to be worse prediction-wise than simpler alternatives. This phenomenon is expected to strengthen as the difference between the true and assumed value of \(p\) grows larger.

\incomment{Me gustaría añadir aquí un último experimento en el que se muestran gráficas de evolución del RMSE o del Accuracy según va variando alguno de los hiperparámetros (especialmente el valor de p y el número de ejemplos de entrenamiento).}

\subsection*{Non-reported experiments}

Apart from all the experiments described above, we performed several tests with different configurations and alternatives. Although these trials did not produce better results in general, we mention here some of them to show how the model could be easily extended. In fact, many of these modifications, if not all, are still available in the provided code and can be enabled through specific parameters.

The first thing tried were cases in which the parameter \(p\) was part of the Bayesian model, with a categorical prior distribution. As explained in Chapter~\ref{ch:model-choice} this approach led to many complications and generally poor results. However, the idea of letting the model select how many components it needs to have based on the data is an interesting one.

We also tried changing some prior distributions. For instance, we swapped the improper prior distribution on \(\sigma^2\) for a Cauchy distribution on \(\R^+\). The results were similar, but we ultimately discarded this approach because it introduced yet another hyperparameter into the model with no clear benefit. In addition, we tried to impose a \(\beta\) distribution on each of the time instants \(t_j\), selecting the hyperparameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable). This approach does work, but in our estimation the computational overhead introduced by the beta density function (about 1.5x increase in execution time) is not worth the mild improvements, if any.

A lot of minor things were tried, such as standardizing the regressors and/or the responses, smoothing the functional data, or transforming the values of \(\tau\) during sampling so that the domain was unconstrained. It is likely that some of these strategies would lead to better results in specific cases, but overall they did not improve the general performance of the prediction procedure.

### 2022-08-30 chapters/model-choice.tex
Finish up to Ch.6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2022 Antonio Coín
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Model choice, implementation and validation}\label{ch:model-choice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this chapter we gather together several remarks on the main choices made throughout the design of our model, as well as some examples of validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.

\section{Model specification}

First we describe some problems and modeling decisions that we had to deal with throughout the development of the model.

\subsection*{Label switching}

A well-known issue found in mixture-like models like the ones we propose is \textit{label switching}, which in short refers to the non-identifiability of the components of the model caused by their interchangeability. In our case, this happens because the likelihood is symmetric with respect to the ordering of the parameters \(b\) and \(\tau\), i.e., \(\pi(Y|X,\theta)=\pi(Y|X, \nu(\theta))\) for any permutation \(\nu\) that rearranges the indices \(j=1,\dots, p\). Thus, since the components are arbitrarily ordered, they may be inadvertedly exchanged from one iteration to the next in any MCMC algorithm. This can cause nonsensical answers when summarizing the marginal posterior distributions to perform inference, as different labelings might be mixed on each component \citep{stephens2000dealing}.

However, this phenomenon is perhaps surprisingly a condition for the convergence of the MCMC method: as pointed out by many authors \citep[e.g.][]{celeux2000computational}, a lack of switching would indicate that not all modes of the posterior distribution were being explored by the sampler. For this reason, many ad-hoc solutions revolve around post-processing and relabeling the samples to eliminate the switching effect, but they generally do not prevent it from happening in the first place.

The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein. A common approach that seems to work well is to simply impose an ordering in the offending parameters, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\). We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the most separation between any two of them (suitably averaged across all iterations of the chains).

This is an area of ongoing research, and thus there are other, more complex relabeling strategies, both deterministic and probabilistic. A summary of several such methods can be found for example in \citet{rodriguez2014label} and \citet{papastamoulis2015label}. In particular, we tested the pivot method proposed by \citet{marin2005bayesian}, in which all samples are aligned to minimize their distance to a reference element (the ``pivot''), which is chosen as the sample that maximizes the posterior density. However, the process of finding the appropriate permutation in each case was time-consuming, and the results were similar, if not worse, than the ones obtained with the simpler order constraints, so the latter were chosen as the default relabeling method of our algorithm (though the pivot method can still be enabled through a dedicated argument in the sampling procedure).

\subsection*{The choice of \(p\)}

One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model. While theoretically we can impose a prior distribution on \(p\) as well (e.g. a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next. In this case we would have no immediate way of tracking down which set of parameters is ``active'' at any given time. A simple approach would be to always consider the first \(p\) parameters and ignore the rest, and we did indeed try this technique, but it gave rise to new difficulties and the results obtained were not good. In fact, the label switching issue is accentuated when \(p\) is allowed to vary \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, and on top of that, the interpretation of, say, the first coefficient \(\beta_1\) in a model with \(3\) components is different than the interpretation of the same coefficient in a model with only \(2\) components.

This inconsistency in the interpretation of the components when the dimensionality of the model increases or decreases can be mitigated using a particular type of MCMC method known as reversible-jump MCMC \citep{green1995reversible}. Theoretically, these algorithms are specifically designed to approximate the posterior distribution in mixture-like models when the number of components is unknown, allowing the underlying dimensionality to change between iterations. However, since they are not yet widely adopted in practice and a reference implementation is not available, we decided against using them in our applications.

Another possibility would be to adapt a purely Bayesian model selection technique to our framework \citep[see][]{piironen2017comparison, gelman2013bayesian}, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. These methods are usually based in computing a quantity known as the \textit{Bayes factor}, which in turn needs the normalizing integral constant we have been trying to avoid all along. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, cross-validation, \ldots) to select its optimal value. As we will see shortly in Chapter~\ref{ch:experiments}, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios.

\subsection*{Other hyperparameters}

As for the default values of the rest of hyperparameters in the prior distributions in~\eqref{eq:prior-linear}, several comments are in order:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use the MLE of \(b\). Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes. Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation indicates that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}

\section{MCMC implementation}

The MCMC method chosen for approximating the posterior distribution in our models is the affine-invariant ensemble sampler described in Section~\ref{sec:mcmc}. As mentioned there, we use the implementation in the Python library \textit{emcee}, which is both reliable and easy to use; it aims to be a general-purpose package that performs well in a wide class of problems. One advantage of this method, apart from the property of affine-invariance, is that it only requires us to specify a few hyperparameters, irrespective of the underlying dimension. This contrasts to, say, the \(\mathcal O(N^2)\) degrees of freedom corresponding to the covariance matrix of an \(N\)-dimensional jump distribution in Metropolis-Hastings.
Furthermore, following \citet{goodman2010ensemble} we can use the \textit{integrated autocorrelation time} to compute an estimate of the effective sample size (i.e. the number of independent samples) after the sampling is done.

After selecting the number of iterations and the number of chains we want, we need to specify the initial points for each of them. As pointed out in \citet{foreman2013emcee}, a good initial choice is a Gaussian ball around a point in \(\Theta_p\) that is expected to have a high probability with respect to the objective distribution. In our implementation we adopt this method, choosing an approximation of the MLE of \(\theta\) as the central point in each case. To perform this approximation we employ the optimization suite of the Python library \textit{scipy}\footnote{\url{https://docs.scipy.org/doc/scipy/}}, and in particular we use the Basin-hopping algorithm \citep{wales1997global}. This is a two-phase stochastic method that combines global steps with local optimization, in the hope of avoiding getting stuck too quickly in local maxima. To reduce the effects of randomness, we run the algorithm a few times and retain the point with the highest likelihood, and to avoid biasing the sampler too much towards the specific point selected, we let a fraction of the initial points be random (within reasonable bounds).

Other less relevant hyperparameters include the burn-in period for the chains, which is the number of initial samples discarded, or the amount of thinning performed, which is the number of consecutive samples discarded to reduce the correlation among them. Another computational decision we made is working with \(\log \sigma\) instead of \(\sigma^2\), so that the domain of this parameter is an unconstrained space, which apparently helps increase the efficiency of the method.

Finally, it is worth mentioning that we experimented with another MCMC framework called \textit{pymc}\footnote{\url{https://docs.pymc.io}}. This is a well-established project aimed at developing a probabilistic programming language for Bayesian modeling with emphasis on MCMC methods \citep{salvatier2016probabilistic}. This package is more complex to use and has a steeper learning curve, but at the same time offers a richer experience and a wide variety of samplers. Apart from standard Metropolis-Hastings techniques, the main algorithm in this library is the \textit{No U-Turn Sampler} (NUTS), a Hamiltonian Monte Carlo method proposed in \citet{hoffman2014no} that performs an automatic adjustment of hyperparameters. However, while the results obtained were similar to that of \textit{emcee}, the execution time of the NUTS sampler was considerably higher (most likely due to parametrization issues with the model). Nevertheless, we decided to include this package as part of the developed code for future use (the Metropolis-Hastings sampler is fully functional and shows comparable performance to \textit{emcee}), although we do not report the corresponding experimental results.

\section{Validation techniques}

To conclude, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away. For example, we can derive credible intervals for each of the parameters (see Figure~\ref{fig:credible_intervals}), and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions.

\begin{figure}[ht!]
  \centering
    \setlength{\fboxrule}{.5pt}%
  \fbox{\includegraphics[width=.9\textwidth]{credible_intervals}}
  \caption{Example of 94\% credible intervals on the posterior distribution of the impact points}\label{fig:credible_intervals}
\end{figure}

Moreover, we can perform various visual checks, such as a plot comparing both the observed and posterior predictive distribution of the responses (see Figure~\ref{fig:ppc}). The posterior predictive distribution for a new sample \(x\) is formally defined as
\[
\pi(y\mid x, \mathcal D_n) = \int_{\Theta} \pi(y\mid x, \theta)\pi(\theta\mid \mathcal D_n)\, d\theta,
\]
and in our context it can be approximated by the simulated responses \(\symbf Y^* \equiv \{\symbf Y^{(m)*}\}\), following the notation of~\eqref{eq:sampled-response-vector}. This distribution accounts for the uncertainty about \(\theta\), and we do indeed use it for prediction in our predict-then-summarize approach.

\begin{figure}[ht!]
  \centering
  \setlength{\fboxrule}{.5pt}%
  \fbox{\includegraphics[width=.65\textwidth]{ppc_linear}}
  \caption{Example of posterior predictive graphical checks on a fitted linear regression model. We show a comparison between the observed distribution of the response variable and the posterior predictive distribution consisting on the approximate sampled responses.}\label{fig:ppc_linear}
\end{figure}

In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the real value of the statistic. They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.

### 2022-08-30 chapters/abstract.tex
Finish up to Ch.6
Revision

--- 2022-08-10 thesis/src/abstract.tex
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{center}
@@
 \end{center}
-%
-\noindent We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of Reproducing Kernel Hilbert Spaces (RKHS's).
-These new models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest.
-By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology, and the posterior distribution can be estimated through Markov chain Monte Carlo methods.
-Several prediction strategies derived from this posterior distribution turn out to be competitive against other usual alternatives in both simulated examples and real data sets, including a Bayesian-motivated variable selection procedure.\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\noindent This work is framed within the field of Functional Data Analysis, a branch of statistics in which the objects of interest are random functions in a functional space instead of, say, random points in \(\R^p\).
+Due to the infinite-dimensional nature of the data, the most common \(L^2\)-models for functional linear and logistic regression present some complications that require further simplifications, usually in the form of regularization or dimensionality reduction.
+
+We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of reproducing kernel Hilbert spaces (RKHS's).
+These recently-proposed models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
+The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest, which has a tractable parametric form.
+By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology.
+The posterior distribution can be estimated through Markov chain Monte Carlo methods, which do not require a complete specification of the posterior density.
+
+We derive several prediction strategies from the approximate posterior distribution, including a Bayesian-motivated variable selection procedure.
+We show through a comprehensive set of experiments that these methods are competitive against other usual alternatives in terms of predictive performance, both in simulated examples and real data sets.
+Overall, our proposed model is simple with regard to interpretation and implementation, and has an added flexibility thanks to the ambient Bayesian framework.\\
 
@@
 \textsc{keywords:} functional data, linear regression, logistic regression, reproducing kernel Hilbert space, Bayesian inference, Markov chain Monte Carlo.
-
-\incomment{add blank page}

### 2022-08-30 chapters/appendix-code.tex
Revision

@@
 \begin{outcomment}
-  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks.
+  Aquí irá un pequeño resumen de la estructura del código y los archivos, y un mini-tutorial de cómo usarlo para reproducir los experimentos. No debería ocupar más de una página.
 
-  Contar que son clases a la sklearn y compatibles con ska-fda. Enlazar al github.
-
-  El proceso es reproducible (semilla 2022)
-
-  Se usan 4 cores.
-
+  \begin{itemize}
+    \item  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks.
+    \item Contar que son clases a la sklearn y compatibles con ska-fda. Enlazar al github.
+    \item Se usan 4 cores.
+    \item El proceso es reproducible (semilla 2022).
+  \end{itemize}
 \end{outcomment}

### 2022-08-30 chapters/background-related-work.tex
Revision

@@
 \end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\) and \(\E[X(t)\epsilon]=0\) almost everwhere; and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
+where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\) and \(\E[X(t)\epsilon]=0\) almost everywhere; and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
 A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
@@
   where the convergence is in \(L^2(\Omega)\) and uniform in \(t\).
-The \(\zeta_j\) are independent zero-mean random variables with \(\E[\zeta_i\zeta_j]=\delta_{ij}\lambda_j\), given explictly by the formula
+The \(\zeta_j\) are independent zero-mean random variables with \(\E[\zeta_i\zeta_j]=\delta_{ij}\lambda_j\), given explicitly by the formula
   \[
@@
  \]
- for all \(A,b\) and \(\lambda\), and for allmost all \(\xi(t)\).
+ for all \(A,b\) and \(\lambda\), and for almost all \(\xi(t)\).
 This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Gamma(t)+b\) for all \(t\).
@@
 
-From a computational perspective, the Python library \textit{emcee} \footnote{\url{https://emcee.readthedocs.io/en/stable/}} \citep{foreman2013emcee} provides a parallel implementation of this algorithm.
+From a computational perspective, the Python library \textit{emcee}\footnote{\url{https://emcee.readthedocs.io/en/stable/}} \citep{foreman2013emcee} provides a parallel implementation of this algorithm.
 The idea is to divide the ensemble \(\Lambda\) into two equally-sized subsets \(\Lambda^{(0)}\) and \(\Lambda^{(1)}\), and then proceed on each iteration in the following alternate fashion:

### 2022-08-30 chapters/bayesian-functional-regression.tex
Revision

@@
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j)\) when \(\alpha\) is as in \eqref{eq:alpha_h0}.
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j)\) when \(\alpha\) is as in~\eqref{eq:alpha_h0}.
 
@@
 When it comes to parameter estimation, a direct optimization of some loss function would probably require a tailored algorithm that took into account the continuous nature of the times \(t_j\).
-Indeed, such an idea is explored in \citet{berrendero2018functional} for the logistic regression case, where the authors propose a ``greedy max-max'' method reminescent of the EM algorithm that alternates between estimating the coefficients and the time instants through a maximum likelihood approach.
+Indeed, such an idea is explored in \citet{berrendero2018functional} for the logistic regression case, where the authors propose a ``greedy max-max'' method reminiscent of the EM algorithm that alternates between estimating the coefficients and the time instants through a maximum likelihood approach.
 
@@
 
-\subsection*{The bayesian approach: prior and posterior}
+\subsection*{The Bayesian approach: prior and posterior}
 

### 2022-08-30 chapters/conclusions.tex
Revision

@@
 \begin{outcomment}
-  - Destacar también las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc. En la introducción, conclusiones, experimentos, ..
-- Destacar interpretabilidad en los experimentos: e.g. un valor de beta=3.4 se puede interpretar fácilmente en un caso concreto. En general, nuestros modelos tienen menor dimensión.
-- Logistic un poco mejor que linear?
-- Tiempo de ejecución alto (característico de los algoritmos MCMC), a cambio mucha sencillez.
-- CV ha sido pequeño. Con más parámetros, probablemente mejores resultados. También flexibilidad para cambiar las prioris.
-- We do not advise to use the mean as a summary statistic.
+  Algo que no he comentado en ninguna parte es que el tiempo de ejecución de nuestros modelos es bastante alto (del orden de 30-40 segundos por ejecución) frente al de los algoritmos de comparación (unos pocos segundos en general).
+Como es algo negativo, no sé muy bien cómo ponerlo, si comentarlo de pasada aquí en las conclusiones, o dedicarle un poco más de espacio en el capítulo de experimentos.\\
+\end{outcomment}
+
+\begin{outcomment}
+  Algunos comentarios generales:
+  \begin{itemize}
+    \item Destacar las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc.
+    \item Destacar la interpretabilidad y sencillez de nuestro modelo. En general, conseguimos resultados competitivos usando modelos con baja dimensión.
+    \item Los resultados en regresión logística parecen en media un poco mejores que en regresión lineal.
+    \item No se recomienda usar la media como estadístico de resumen de la distribución a posteriori.
+    \item La búsqueda de hiperparámetros en CV no ha sido demasiado grande. Probablemente con más recursos computacionales se podrían conseguir mejores resultados.
+  \end{itemize}
 \end{outcomment}
@@
 \begin{outcomment}
-  - Mejorar pymc.
-  - consistencia de la posteriori, robustez de la priori
-  - Otras prioris
-
+  \begin{itemize}
+    \item Ahondar en la relación entre RKHS y regresión funcional.
+    \item Experimentar con otros algoritmos de MCMC (en particular, reversible-jump MCMC), y con otros paquetes (pymc).
+    \item Considerar otras distribuciones a priori.
+    \item Derivar algunos resultados de consistencia y/o robustez de la posteriori.
+    \item Extender el modelo lineal a un modelo GLM.
+  \end{itemize}
 \end{outcomment}

### 2022-08-30 chapters/experiments.tex
Revision

@@
 Lastly, since the use of MCMC algorithms introduces a source of stochasticity in the prediction procedure, we independently repeat the whole process 10 times (each with a different train/test configuration), and average the results across these executions.
-The metrics used to evaluate the perfomance of the models are the RMSE for linear regression and the accuracy for logistic regression.
+The metrics used to evaluate the performance of the models are the RMSE for linear regression and the accuracy for logistic regression.
 Recall that
@@
   \item A Brownian motion \(K_1(t,s)=\min\{t,s\}\).
-  \item A fractional Browmian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
+  \item A fractional Brownian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
   \item An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
@@
   \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's and label them according to their origin.
-Firstly, we consider a homoscedastic case to distinguish between a standard Browmian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\).
+Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\).
 Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance \(2\), that is, with kernel \(K(t,s)=2\min\{t,s\}\).
@@
 We have included a fairly comprehensive suite of comparison algorithms, chosen among the most common methods used in machine learning and FDA.
-There are finite-dimensional models that work on the discretized data, variable selection/dimension reduction procedures and purley functional methods.
+There are finite-dimensional models that work on the discretized data, variable selection/dimension reduction procedures and purely functional methods.
 
@@
 \begin{itemize}
-  \item Manual. Dummy variable selection method with a pre-specified number of components (equispaced on \([0, 1]\)).
-  \item Lasso. Linear least squares with \(l^1\) regularization.
-  \item Ridge. Linear least squares with \(l^2\) regularization.
-  \item PLS. Partial least squares for dimension reduction.
-  \item PCA. Principal component analysis for dimension reduction.
-  \item PLS1. Partial least squares regression \citep[e.g.][]{wegelin2000survey}.
-  \item APLS. Functional partial least squares regression proposed by \citet{delaigle2012methodology}.
-  \item RMH. Recursive maxima hunting variable selection method proposed by \citet{torrecilla2016feature}.
-  \item FLin. Functional \(L^2\) linear regression model with fixed basis expansion.
-  \item FPLS1. Functional PLS regression through basis expansion, implemented as in \citet{aguilera2010using}.
-  \item FPCA. Functional principal component analysis.
+  \item Manual.\hspace{.3em} Dummy variable selection method with a pre-specified number of components (equispaced on \([0, 1]\)).
+  \item Lasso.\hspace{.3em} Linear least squares with \(l^1\) regularization.
+  \item Ridge.\hspace{.3em} Linear least squares with \(l^2\) regularization.
+  \item PLS.\hspace{.3em} Partial least squares for dimension reduction.
+  \item PCA.\hspace{.3em} Principal component analysis for dimension reduction.
+  \item PLS1.\hspace{.3em} Partial least squares regression \citep[e.g.][]{wegelin2000survey}.
+  \item APLS.\hspace{.3em} Functional partial least squares regression proposed by \citet{delaigle2012methodology}.
+  \item RMH.\hspace{.3em} Recursive maxima hunting variable selection method proposed by \citet{torrecilla2016feature}.
+  \item FLin.\hspace{.3em} Functional \(L^2\) linear regression model with fixed basis expansion.
+  \item FPLS1.\hspace{.3em} Functional PLS regression through basis expansion, implemented as in \citet{aguilera2010using}.
+  \item FPCA.\hspace{.3em} Functional principal component analysis.
 \end{itemize}
@@
 \begin{itemize}
-  \item Log. Standard multiple logistic regression with \(l^2\) regularization.
-  \item LDA. Linear discriminant analysis.
-  \item QDA. Quadratic discriminant analysis.
-  \item RKVS. RKHS-based variable selection method proposed in \citet{berrendero2018use}.
-  \item APLS. Functional PLS used as a dimension reduction method, as proposed in \citet{delaigle2012achieving} in combination with the nearest centroid algorithm.
-  \item FLog. Functional RKHS-based logistic regression algorithm proposed in \citet{berrendero2018functional}.
-  \item FLDA. Implementation of the so-called functional linear discriminant analysis in \citet{preda2007pls}.
-  \item MDC. Maximum depth classifier \citep[e.g.][]{ghosh2005maximum}.
-  \item FKNN. Functional K-nearest neighbors classifier with the \(L^2\)-distance.
-  \item FNC. Functional nearest centroid classifier with the \(L^2\)-distance.
+  \item Log.\hspace{.3em} Standard multiple logistic regression with \(l^2\) regularization.
+  \item LDA.\hspace{.3em} Linear discriminant analysis.
+  \item QDA.\hspace{.3em} Quadratic discriminant analysis.
+  \item RKVS.\hspace{.3em} RKHS-based variable selection method proposed in \citet{berrendero2018use}.
+  \item APLS.\hspace{.3em} Functional PLS used as a dimension reduction method, as proposed in \citet{delaigle2012achieving} in combination with the nearest centroid algorithm.
+  \item FLog.\hspace{.3em} Functional RKHS-based logistic regression algorithm proposed in \citet{berrendero2018functional}.
+  \item FLDA.\hspace{.3em} Implementation of the so-called functional linear discriminant analysis in \citet{preda2007pls}.
+  \item MDC.\hspace{.3em} Maximum depth classifier \citep[e.g.][]{ghosh2005maximum}.
+  \item FKNN.\hspace{.3em} Functional K-nearest neighbors classifier with the \(L^2\)-distance.
+  \item FNC.\hspace{.3em} Functional nearest centroid classifier with the \(L^2\)-distance.
 \end{itemize}
@@
 In this case the outcome is satisfactory, since for the most part our models are on a par with the rest, even beating other methods that were designed with the \(L^2\)-model in mind.
-Moreover, whenever one of our models has a higher RMSE, the difference is pretty small in general.
+Moreover, whenever one of our models has a higher RMSE, the difference is pretty small in comparison.
 \begin{figure}[ht!]
@@
 Figure~\ref{fig:reg_emcee_real} shows the results for the real data sets.
-In these data sets there is a substantial difference in perfomance between some of our methods and the reference algorithms.
-However, the predict-then-summarize approach (represented as \textit{posterior\_mean}) seems to work quite well, always near the mean RMSE of all the comparison algorithms.
+In these data sets there is a substantial difference in performance between some of our methods and the reference algorithms.
+However, the predict-then-summarize approach (represented as \textit{posterior\_mean}) seems to work quite well, always scoring near the mean RMSE of all the comparison algorithms.
 Moreover, our two-stage methods seem to perform better than our other methods in these cases, scoring again very close to the mean of the reference methods.
@@
 The same goes for the time instants, except that in this case there is no default value to represent that a component is unused, so the time corresponding to the null coefficient oscillates back and forth.
-The estimated function, based for example in the mode of the posterior distributions, will not perfect, essentially because of the noise in the response.
+The estimated function (based for example in the mode of the posterior distributions) will not be perfect, essentially because of the noise in the response.
 But it should be close to the true parameter function \(\alpha(t)=-5K(t, 0.1) + 10K(t, 0.8)\), providing a good predictive performance.
@@
 In contrast, if we consider now a model with \(p=4\) with the same data, we might obtain posterior distributions such as the ones in Figure~\ref{fig:beta_trace_4}.
-In this situation two components should go to zero, but that is no longer the case.
+In this situation two coefficients should go to zero, but that is no longer the case.
 For example, while the green component has a high density around 0, it also has a considerable mass around 10, effectively ``competing'' with the red component.
@@
 Apart from all the experiments described above, we performed several tests with different configurations and alternatives.
-Although these trials did not produce better results in general, we mention here some of them to show how the model could be easily extended.
+Although these trials did not produce better results in general, we mention some of them here to show how the model could be easily extended.
 In fact, many of these modifications, if not all, are still available in the provided code and can be enabled through specific parameters.
 
-The first thing tried were cases in which the parameter \(p\) was part of the Bayesian model, with a categorical prior distribution.
+The first thing we tried was to make the parameter \(p\) part of the Bayesian model, imposing a categorical prior distribution on it.
 As explained in Chapter~\ref{ch:model-choice} this approach led to many complications and generally poor results.
-However, the idea of letting the model select how many components it needs to have based on the data is an interesting one.
+However, the idea of letting the model select how many components it should have based on the data is an interesting one, and in some situations this could greatly improve the results of variable selection procedures based on other model selection criteria.
 

### 2022-08-30 chapters/introduction.tex
Revision

--- 2022-08-26 thesis/chapters/introduction.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{center}
-%   \begin{minipage}{0.5\textwidth}
-%     \begin{small}
-%       \emph{Quote}
-%     \end{small}
-%   \end{minipage}
-%   \vspace{0.5cm}
-% \end{center}
 
@@
   \item To perform a brief but thorough literature review that contextualizes functional linear and logistic regression in statistics and machine learning, exploring the predominant models and techniques.
-  \item To propose a novel functional model for linear and logistic regression that builds on existing work and focuses on simplicity, both in terms of interpretation and implementation.
-  \item To describe a general Bayesian approach for parameter estimation within the suggested model.
+  \item To propose a novel RKHS-based functional model for linear and logistic regression that builds on existing work and focuses on simplicity, both in terms of interpretation and implementation.
+  \item To describe and implement a general Bayesian approach for parameter estimation within the suggested model.
   \item To introduce the tools needed to specify and implement the proposed model, mainly reproducing kernel Hilbert spaces and Markov chain Monte Carlo methods.

### 2022-09-01 chapters/abstract.tex
Añade contenido en Ch.5; alguna tabla

@@
 
-\noindent This work is framed within the field of Functional Data Analysis, a branch of statistics in which the objects of interest are random functions in a functional space instead of, say, random points in \(\R^p\).
+\noindent This work is framed within the field of Functional Data Analysis, a branch of statistics in which the objects of interest are random functions in a functional space, instead of, say, random points in \(\R^p\).
 Due to the infinite-dimensional nature of the data, the most common \(L^2\)-models for functional linear and logistic regression present some complications that require further simplifications, usually in the form of regularization or dimensionality reduction.
 
-We propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of reproducing kernel Hilbert spaces (RKHS's).
+In this thesis we propose a novel Bayesian approach for functional linear and logistic regression models, based on the theory of reproducing kernel Hilbert spaces (RKHS's).
 These recently-proposed models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
@@
 We show through a comprehensive set of experiments that these methods are competitive against other usual alternatives in terms of predictive performance, both in simulated examples and real data sets.
-Overall, our proposed model is simple with regard to interpretation and implementation, and has an added flexibility thanks to the ambient Bayesian framework.\\
+Overall, our proposed model is simple with regard to interpretation and feasible with regard to implementation, while enjoying the added flexibility of an ambient Bayesian framework.\\
 

### 2022-09-01 chapters/appendix-code.tex
Añade contenido en Ch.5; alguna tabla

@@
 \begin{outcomment}
-  Aquí irá un pequeño resumen de la estructura del código y los archivos, y un mini-tutorial de cómo usarlo para reproducir los experimentos. No debería ocupar más de una página.
+  Aquí irá un pequeño resumen de la estructura del código y los archivos, y un mini-tutorial de cómo usarlo para reproducir los experimentos.
 
   \begin{itemize}
-    \item  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks.
+    \item Características de la máquina donde se ejecuta. Se usan 4 cores.
+    \item  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks (decir que hay más gráficas y análisis y cosas así).
     \item Contar que son clases a la sklearn y compatibles con ska-fda. Enlazar al github.
-    \item Se usan 4 cores.
     \item El proceso es reproducible (semilla 2022).

### 2022-09-01 chapters/background-related-work.tex
Añade contenido en Ch.5; alguna tabla

@@
 Then, sampling from \(g\) and accepting the proposal, say \(x\), with probability \(f(x)/Mg(x)\) yields an approximate set of samples from \(f\).
-Nevertheless, these kinds of algorithms are rapidly affected by the so-called \textit{curse of dimensionality} (see Figure \ref{fig:curse_dimensionality}), so new techniques are needed when sampling from multidimensional distributions.
+Nevertheless, these kinds of algorithms are rapidly affected by the so-called \textit{curse of dimensionality} (see the schematic in Figure~\ref{fig:curse_dimensionality}), so new techniques are needed when sampling from multidimensional distributions.
 This is where Markov chain Monte Carlo (MCMC) methods come into play.

### 2022-09-01 chapters/conclusions.tex
Añade contenido en Ch.5; alguna tabla

@@
 \begin{outcomment}
-  Algo que no he comentado en ninguna parte es que el tiempo de ejecución de nuestros modelos es bastante alto (del orden de 30-40 segundos por ejecución) frente al de los algoritmos de comparación (unos pocos segundos en general).
-Como es algo negativo, no sé muy bien cómo ponerlo, si comentarlo de pasada aquí en las conclusiones, o dedicarle un poco más de espacio en el capítulo de experimentos.\\
-\end{outcomment}
-
-\begin{outcomment}
   Algunos comentarios generales:
   \begin{itemize}
-    \item Destacar las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc.
-    \item Destacar la interpretabilidad y sencillez de nuestro modelo. En general, conseguimos resultados competitivos usando modelos con baja dimensión.
+    \item Destacar las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc. Computationally feasible model.
+    \item Destacar la interpretabilidad y sencillez de nuestro modelo.
+En general, conseguimos resultados competitivos usando modelos con baja dimensión.
+Aunque el modelo final es simple, la teoría subyacente no lo es.
+Modelo simple útil para practitioners.
+También se presta a post-interpretación (ej.
+si salen dos modas)
+    \item Algo del tiempo de ejecución. Se podría aproximar mejor el error medio con más repeticiones aleatorias (en lugar de 10), pero sería muy costoso computacionalmente.
     \item Los resultados en regresión logística parecen en media un poco mejores que en regresión lineal.
-    \item No se recomienda usar la media como estadístico de resumen de la distribución a posteriori.
+    \item No se recomienda usar la media como estadístico de resumen de la distribución a posteriori. En general, los que usan la posteriori entera son más estables.
     \item La búsqueda de hiperparámetros en CV no ha sido demasiado grande. Probablemente con más recursos computacionales se podrían conseguir mejores resultados.

### 2022-09-01 chapters/experiments.tex
Añade contenido en Ch.5; alguna tabla

@@
 First and foremost, we consider \(T=\{\text{mean, median, mode}\}\) for the summarize-then-predict approach to prediction.
-In this way in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach, which will we call ``\textit{posterior\_mean}'' for short) and 3 variable selection methods (one for each statistic), while in logistic regression there are 5 prediction methods and 3 variable selection procedures.
+In this way in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach, which will we call ``\textit{posterior\_mean}'' for short) and 3 variable selection methods (one for each statistic), while in logistic regression there are similarly 5 prediction methods and 3 variable selection procedures.
 In each case, after variable selection is performed, we use a penalized multiple linear/logistic regression method to generate the corresponding predictions.
-All in all, we are looking at 7 (8 in the case of logistic regression) prediction methods, and although all of them are derived from a single MCMC run, we will treat them as separate for the purposes of experimentation.
+All in all, we are looking at 7 (8 in the case of logistic regression) prediction methods, and although all of them are derived from a single MCMC run, we will treat them as separate for the purposes of the experimentation.
 In what follows we relax the notation and use the expressions ``model'', ``prediction method'' and ``algorithm'' almost interchangeably when referring to these 7 (or 8) cases.
@@
 \begin{enumerate}
-  \item A Brownian motion \(K_1(t,s)=\min\{t,s\}\).
-  \item A fractional Brownian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
-  \item An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
-  \item A Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), with \(\nu=0.2\).
+  \item BM.\hspace{0.3em} A Brownian motion \(K_1(t,s)=\min\{t,s\}\).
+  \item fBM.\hspace{0.3em} A fractional Brownian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
+  \item O-U.\hspace{0.3em} An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
+  \item Gaussian.\hspace{0.3em} A Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), with \(\nu=0.2\).
 \end{enumerate}
@@
 \paragraph{Logistic regression data sets.} Again we consider three different types of data sets,  with a common value of \(\alpha_0=-0.5\).
+In this case we randomly permute 10\% of the labels to introduce some noise in the simulations.
 \begin{itemize}
@@
 We have adopted a visual approach to presenting the experimentation results, using colored graphs to help visualize them.
-Alternatively, tables representing these results can be consulted instead in Appendix~\ref{ch:tables}.
+We felt that this was a better way of summarizing a large empirical study such as the one we have carried out.
+Alternatively, tables representing these results can be consulted in Appendix~\ref{ch:tables}.
 
-We show the mean and standard deviation of the RMSE obtained across the 10 random runs, depicting our models in orange and the comparison algorithms in blue.
+We show the mean and standard deviation of the score obtained in each case across the 10 random runs, depicting our models in orange and the comparison algorithms in blue.
 We also show the global mean of all the comparison algorithms with a dashed vertical line, and we exclude extreme negative results from this mean to avoid distortion.
@@
 
-We do not expect our models to defeat all the comparison algorithms in all cases (and indeed they do not), but we intend to show that they are competitive methods that on average match or improve other usual alternatives.
+We do not expect our models to defeat all the comparison algorithms in all cases (and indeed they do not), but we intend to show that they are competitive methods that on average match or improve the performance of other usual alternatives.
 After all, \textit{there is no such thing as a free lunch}\footnote{The phrase alludes to the No Free Lunch theorem, which loosely speaking asserts that all machine learning methods are equivalent when their performance is averaged across all possible problems.
@@
 
-\section{Functional linear regression}
+\section{Functional linear regression}\label{sec:experiments-linear}
 
@@
 
-Note that our Bayesian models have a higher standard deviation, partly because there is an intrinsic randomness in the methods (apart from the train/test splits), and it can be the cause of the occasional worse performance.
+Note that some of our Bayesian models have a higher standard deviation, partly because there is an intrinsic randomness in the methods (apart from the train/test splits), and it can be the cause of the occasional worse performance.
 In relation to this, we observe that the methods that use the mean as a summary statistic tend to perform much worse.
@@
 
-\section{Functional logistic regression}
+\section{Functional logistic regression}\label{sec:experiments-logistic}
 
@@
 Continuing with Figure~\ref{fig:clf_emcee_l2}, we see that in the \(L^2\) case the results are again promising, since our models score consistently on or above the mean of the reference models, and in many cases surpassing most of them.
+The predict-then-summarize approaches (\textit{emcee\_posterior\_mean} and \textit{emcee\_posterior\_vote}) are particularly good in this case, and in general have low standard errors.
 Moreover, the overall accuracy of all methods is low (below 60\%), so this is indeed a difficult problem in which even small increases in accuracy are relevant.
@@
 As for the real data sets, in Figure~\ref{fig:clf_emcee_real} we see positive results in general, obtaining in most cases accuracies well above the mean of the reference models, and sometimes above most of them.
+In particular, the predict-then-summarize tend to have a good performance and achieve a lower standard error across executions, which is a trend that we also saw in the simulated data sets.
 However, our variable selection methods seem to be a bit worse in this case.
-Again, the models that use \textit{emcee\_mean} are the exception, and in all these data sets they perform steadily worse than the rest of our models.
+Again, the models that use \textit{emcee\_mean} are the exception, and in all these data sets they perform steadily worse than the rest of our Bayesian models.
 
@@
 
-In this section we present some experiments that are not directly related to measuring the prediction performance relative to other methods.
+In this section we present some experiments that are not directly related to measuring the prediction performance of our models relative to other methods.
 
@@
 One requirement that our model should satisfy is that it be able to recover the true parameter function when the underlying model is a RKHS model.
-This is generally true when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?
+This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?
 
@@
 Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(\tau_1, \tau_2, \tau_3)\) for a model with 3 components.
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=.8\textwidth]{beta_trace_3}
-  \caption{Posterior distribution in a linear model with \(p=3\).}\label{fig:beta_trace_3}
-\end{figure}
-
 As we can see, one of the coefficients has gone to zero to account for the overspecification of the model, while the other two have stabilized very close to the true parameters.
@@
 
-In contrast, if we consider now a model with \(p=4\) with the same data, we might obtain posterior distributions such as the ones in Figure~\ref{fig:beta_trace_4}.
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=.8\textwidth]{beta_trace_3}
+  \caption{Estimated posterior distribution for \(\beta\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
+\end{figure}
+
+In contrast, if we consider now a model with \(p=4\) with the same data, we might obtain posterior distributions like the ones in Figure~\ref{fig:beta_trace_4}.
 In this situation two coefficients should go to zero, but that is no longer the case.
@@
   \includegraphics[width=.8\textwidth]{beta_trace_4}
-  \caption{Posterior distribution in a linear model with \(p=4\).}\label{fig:beta_trace_4}
+  \caption{Estimated posterior distribution for \(\beta\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
 \end{figure}
 
-There is still another possible situation, one in which there is no label switching but the four components are non-negligible in the estimated function.
-This can happen because the components exploit the additional freedom and ``work together'', so to speak.
-In this way we could obtain an estimate that might not resemble the true coefficient function, but that could have a very low prediction error.
+There is still another possible situation, one in which there is no label switching but the estimated function has four non-negligible components.
+This can happen because the different components exploit the additional freedom and ``work together'', so to speak.
+In this way we might obtain an estimate that does not resemble the true coefficient function, but that has a very low prediction error.
 However, this could also work to our detriment and cause the estimated function to be worse prediction-wise than simpler alternatives.
@@
 
-\incomment{Me gustaría añadir aquí un último experimento en el que se muestran gráficas de evolución del RMSE o del Accuracy según va variando alguno de los hiperparámetros (especialmente el valor de p y el número de ejemplos de entrenamiento).}
+\subsection*{Dependence on \(p\)}
+
+Another thing we wanted to look at was the dependence of the final prediction result on the chosen value of \(p\), especially when there is no concept of ``components'' in the underlying model.
+We can take for example the homoscedastic mixture data set in the logistic regression problem, and fix the parameter \(\eta=0.01\).
+Then the corresponding mean accuracies (in 10 repetitions) for RKHS models with \(p=1,2,\dots 10\) components are shown in Figure~\ref{fig:dependence_acc_p}, along with their standard errors (which are arguably not very informative).
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.9\textwidth]{mixture_dependence_acc_p}
+  \caption{Mean accuracy in 10 repetitions for our logistic RKHS methods as a function of \(p\), using \(\eta=0.01\) and the homoscedastic mixture data set.
+The corresponding standard errors are shown in faded colors.}\label{fig:dependence_acc_p}
+\end{figure}
+
+It would appear that in the methods that use the whole posterior distribution are more stable and somewhat independent of the value of \(p>1\).
+The other methods show a slight downward trend as \(p\) increases (although not so much in the variable selection methods), and in general their best results are obtained at \(p=2\) and \(p=8\).
+We expect that this effect or some small variation of it will remain valid in other situations; however, a profound study of this would be the subject of a different experiment altogether.
+
+\subsection*{Execution times}
+
+We summarize below the mean execution time for each independent run of our CV experiment, grouping the simulated data sets by the underlying data generation strategy (since the results within these groups were very similar).
+
+\begin{table}[ht!]
+  \centering
+  \begin{tabular}{lcccccc}
+\toprule
+  &            RKHS &           \(L^2\) &           GBM &        Tecator & Moisture & Sugar \\
+\midrule
+Linear regression & 260 & 245 & 264 & 230 & 264 & 285\\
+\bottomrule
+
+  &            RKHS &           \(L^2\) &           Mixture &        Medflies & Growth & Phoneme \\
+\toprule
+Logistic regression & 277 & 263 & 246 & 250 & 218 & 277\\
+\bottomrule
+\end{tabular}
+  \caption{Mean execution times (in minutes) for the independent runs of the CV experiments in Sections~\ref{sec:experiments-linear} and~\ref{sec:experiments-logistic}, grouped when possible by the underlying data generation strategy.}
+\end{table}
+
+The reference algorithms used were all fast methods that can complete their whole CV loop in just a few seconds, so any comparison with them in this regard is uninteresting.
+However, this outcome was to be expected, since in general the MCMC methods are known to have a high computational cost in exchange for the simplicity in modeling they provide.
+An individual MCMC run is not that expensive time-wise, but even for a modest 5-fold cross-validation loop the effects add up quickly to a considerable amount of time.
+In Figure~\ref{fig:dependence_time_n_train} we show the evolution of the execution time versus the number of training samples in a generic case, which not surprisingly has a more or less linear behavior.
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.6\textwidth]{l2_dependence_time_n_train}
+  \caption{Mean training times of a logistic RKHS model, averaged across 10 independent runs, as a function of the number of training samples.
+The values are surrounded by the corresponding standard deviations.
+}\label{fig:dependence_time_n_train}
+\end{figure}
 
@@
 The first thing we tried was to make the parameter \(p\) part of the Bayesian model, imposing a categorical prior distribution on it.
-As explained in Chapter~\ref{ch:model-choice} this approach led to many complications and generally poor results.
+As explained in Chapter~\ref{ch:model-choice}, this approach led to many complications and generally poor results.
 However, the idea of letting the model select how many components it should have based on the data is an interesting one, and in some situations this could greatly improve the results of variable selection procedures based on other model selection criteria.
@@
 In addition, we tried to impose a \(\beta\) distribution on each of the time instants \(t_j\), selecting the hyperparameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable).
-This approach does work, but in our estimation the computational overhead introduced by the beta density function (about 1.5x increase in execution time) is not worth the mild improvements, if any.
+This approach does work, but in our estimation the computational overhead introduced by the beta density function (sometimes resulting in a 1.5x increase in execution time) is not worth the mild improvements, if any.
 

### 2022-09-01 chapters/introduction.tex
Añade contenido en Ch.5; alguna tabla

@@
 \begin{enumerate}[1.]
-  \item To perform a brief but thorough literature review that contextualizes functional linear and logistic regression in statistics and machine learning, exploring the predominant models and techniques.
+  \item To perform a brief but thorough literature review that contextualizes functional linear and logistic regression within statistics and machine learning, exploring the predominant models and techniques.
   \item To propose a novel RKHS-based functional model for linear and logistic regression that builds on existing work and focuses on simplicity, both in terms of interpretation and implementation.
   \item To describe and implement a general Bayesian approach for parameter estimation within the suggested model.
-  \item To introduce the tools needed to specify and implement the proposed model, mainly reproducing kernel Hilbert spaces and Markov chain Monte Carlo methods.
-  \item To carry out a comprehensive experimental study to test these new models and compare them to existing methods, both in simulations and in real-world scenarios.
+  \item To introduce the tools needed to specify the proposed model and put it into practice, mainly reproducing kernel Hilbert spaces and Markov chain Monte Carlo methods.
+  \item To carry out an extensive experimental study to test these new models and compare them to existing methods, both in simulations and in real-world scenarios.
 \end{enumerate}
 
-It is beyond the scope of this thesis to provide an extensive review of FDA as a whole, or to delve too deeply into the details and inner workings of most models and techniques mentioned.
+It is beyond the scope of this thesis to provide a complete review of FDA as a whole, or to delve too deeply into the details and inner workings of most models and techniques mentioned.
 Nevertheless, references are provided frequently throughout the text to point the interested reader towards more specialized resources.
@@
 The empirical results of the experimentation are contained in Chapter~\ref{ch:experiments}.
-Lastly, the conclusions drawn from this work and future paths of research are presented in Chapter~\ref{ch:conclusions}.
+Lastly, the conclusions drawn from this work and future paths of research are reviewed in Chapter~\ref{ch:conclusions}.

### 2022-09-02 chapters/appendix-code.tex
Finish document

--- 2022-08-30 chapters/appendix-code.tex
@@
 
-\begin{outcomment}
-  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks.
+The Python source code developed for this work is available at the GitHub repository \url{https://github.com/antcc/rk-bfr}.
+The code is adequately documented and is structured in several directories as follows:
 
-  Contar que son clases a la sklearn y compatibles con ska-fda. Enlazar al github.
+\begin{itemize}
+  \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our models, separated according to the functionality they provide.
+  \item The \texttt{reference\_methods} folder contains the implementation of the functional comparison algorithms that were not available through a standard Python library.
+  \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g.
+number of iterations, type of data set, \ldots), along with some utility files.
+In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our models without a cross-validation loop.
+\end{itemize}
 
-  El proceso es reproducible (semilla 2022)
-
-  Se usan 4 cores.
-
-\end{outcomment}
+When possible, the code was implemented in a generic way that would allow for easy extensions or derivations, and also with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries.
+Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them.
+The code for the experiments was executed in a machine with 4 cores, and a random seed with value 2022 was set for reproducibility.
+We provide a script file \texttt{launch.sh} that illustrates a typical execution.
+Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way.
+Inside these notebooks there is a step-by-step guide on how one might execute our algorithms, accompanied by many graphical representations, and offering the possibility of changing multiple parameters to experiment with the code.
+In addition, there is also a notebook that can be used to generate all the tables and figures of this document pertaining to the experimental results.

### 2022-09-02 chapters/background-related-work.tex
Finish document

@@
 \end{equation}
-where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\) and \(\E[X(t)\epsilon]=0\) almost everywhere; and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
+where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\).
 A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
@@
 In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional, berrendero2018use} for the case of functional logistic regression.
+These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional} for the case of functional logistic regression.
 
 There are other proposals for models that change the habitat of the functional parameter, and there are even some in which \(\beta\) is supposed to live in a RKHS, most notably \citet{yuan2010reproducing}.
-However, these are arbitrary RKHS's that do not directly exploit the relation with the process \(X\) (though the authors use said connection to derive some asymptotic properties of their estimators), and therefore the approach is somewhat different.
+However, theirs are arbitrary RKHS's that do not directly exploit the relation with the process \(X\) (though the authors do use this connection to derive some asymptotic properties of their estimators), and therefore the approach is somewhat different.
 Incidentally, the RKHS associated with \(X\) also has some interesting properties that contribute to shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.
@@
   \]
-  and we have just shown above that the evaluation operator is bounded.
+  as we have just shown in \((i)\) that the evaluation operator is bounded.
 Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform.
@@
     Let \(K\) be a continuous kernel on \([0,1]^2\).
-Then, there exists an orthonormal basis \(\{\phi_j\}\) of \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive and form a decreasing sequence \(\{\lambda_j\}\to 0\), and such that
+Then, there exists an orthonormal system \(\{\phi_j\}\) in \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive and form a non-increasing sequence \(\{\lambda_j\}\to 0\), and such that
     \[
@@
 
+In connection with our RKHS theory, it can be shown that the set \(\{\sqrt{\lambda_j}\phi_j\}\) constitutes an orthonormal basis of \(\Hcal(K)\) \citep[see e.g.][Sec.~4.4]{cucker2007learning}.
 Furthermore, a similar orthogonal decomposition holds for our stochastic process \(X\), an expansion which lies at the base of many \(L^2\) functional regression models.
-The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Ch.~3.2]{berlinet2004reproducing}.
+The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Sec.~3.2]{berlinet2004reproducing}.
 
 \begin{theorem}[Karhunen-Loève expansion]
-  With the assumptions and notations of Theorem~\eqref{th:mercer}, the centered process \(X=X(t)\) with covariance function \(K\) admits the quadratic-mean representation
+  With the assumptions and notations of Theorem~\ref{th:mercer}, the centered process \(X=X(t)\) with covariance function \(K\) admits the quadratic-mean representation
   \[
@@
 For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\K^{-1/2}(f)}{\K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can also be seen as the set \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
 Note that, since \(\{\lambda_j\}\) tends to zero, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, not only in that they inherit the regularity of the kernel \(K\), but also in the sense that their components in an orthonormal basis, namely \(\dotprod{f}{\phi_j}\), need to vanish quickly.
@@
 \[
-  \E\|X_N(\cdot)\|^2_K = \E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
+  \E\left[\|X_N(\cdot)\|^2_K\right] = \E \left[\sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j}\right] = N \to \infty \quad (N\to\infty).
 \]
@@
  for all \(A,b\) and \(\lambda\), and for almost all \(\xi(t)\).
-This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Gamma(t)+b\) for all \(t\).
-In \citep{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property.
+This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Lambda(t)+b\) for all \(t\).
+In \citet{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property.
 Specifically, they work with a set \(\Lambda=(\Lambda_1, \dots, \Lambda_L)\) of \textit{walkers}, where \(\Lambda_l(t)\) represents an individual chain at time \(t\).

### 2022-09-02 chapters/bayesian-functional-regression.tex
Finish document

@@
 As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model.
-However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model.
+However, in this case the nuisance parameter \(\sigma^2\) only appears as part of the hierarchical prior distribution, and not in the final model.
 The posterior distribution is then derived after a routine calculation.

### 2022-09-02 chapters/conclusions.tex
Finish document

--- 2022-08-30 chapters/conclusions.tex
@@
 
-\begin{outcomment}
-  - Destacar también las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc. En la introducción, conclusiones, experimentos, ..
-- Destacar interpretabilidad en los experimentos: e.g. un valor de beta=3.4 se puede interpretar fácilmente en un caso concreto. En general, nuestros modelos tienen menor dimensión.
-- Logistic un poco mejor que linear?
-- Tiempo de ejecución alto (característico de los algoritmos MCMC), a cambio mucha sencillez.
-- CV ha sido pequeño. Con más parámetros, probablemente mejores resultados. También flexibilidad para cambiar las prioris.
-- We do not advise to use the mean as a summary statistic.
-\end{outcomment}
+In this work we have addressed some inference and prediction issues involved in the task of learning from functional data.
+In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios.
+Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective that are conceptually simpler and facilitate the interpretation and subsequent analysis, an attribute that practitioners surely find especially compelling.
+The underlying theoretical aspects are nevertheless far from trivial, so we have also explored some of the related mathematical background in this text.
+
+With the aid of Bayesian methods, we have introduced a computationally feasible approach to parameter estimation within the proposed RKHS models.
+The flexibility is improved through the use of prior distributions that can introduce pre-existing information into the model.
+In addition, the fact that we have a complete posterior distribution at our disposal enables us to derive different prediction and variable selection methods, or even perform other types of inference such as the construction of credible intervals.
+Moreover, the use of MCMC methods to approximate the posterior distribution results in an implementation that is simpler than most existing alternatives.
+Although this simplicity comes at the expense of increased computational requirements, this computational cost is still reasonable and the model can be used in practice.
+
+We implemented the whole inference and prediction pipeline in the Python programming language, and then used it to carry out a large empirical study to assess the performance of our models against other usual alternatives, functional and otherwise.
+The results were generally satisfactory, proving that we have competitive methods that achieve good results in different scenarios, often with low dimensionality and thus increased tractability and interpretability.
+Lastly, we believe that with more computational resources we could consider a larger pool of hyperparameters in the model selection phase, which would most likely contribute to better results.
 
@@
 
-\begin{outcomment}
-  - Mejorar pymc.
-  - consistencia de la posteriori, robustez de la priori
-  - Otras prioris
+Some lines of further research suggested by our work are the following.
 
-\end{outcomment}
+\begin{itemize}
+  \item To delve further into the relationship between RKHS's and functional data problems, a connection that has proven to be fruitful in many scenarios.
+A first idea would be to extend the RKHS-based logistic regression model to a generalized functional linear model with an arbitrary link function.
+  \item To try to derive some theoretical properties of our Bayesian predictors.
+For example, consistency and/or robustness results regarding the posterior distribution would be an excellent complement to the practical side of this work.
+  \item To find other prior distributions for our parameters that perform better in general, or to eliminate the need of the tuning hyperparameter \(\eta\).
+  \item To experiment with other MCMC algorithms for posterior approximation.
+Specifically, it would be very interesting to implement an efficient and reliable reversible-jump MCMC method in Python, which as we have already mentioned is a better fit for our particular Bayesian model.
+\end{itemize}

### 2022-09-02 chapters/experiments.tex
Finish document

@@
   \[
-  \P(Y=1\mid X) = \frac{1}{e^{0.5 +5X(0.1) - X(0.4) - 10X(0.8)}}.
+  \P(Y=1\mid X) = \frac{1}{1 + e^{0.5 +5X(0.1) - X(0.4) - 10X(0.8)}}.
   \]
   \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case.
-  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's and label them according to their origin.
+  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's (with equal probability \(p=1/2\)) and label them according to their origin.
 Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\).
@@
 The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair.
-In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is in \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,11\}\).
+In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is one of \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,9,11\}\).
 
@@
 Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case.
-Our variable selection methods are close to the global mean accuracy in both cases.
+Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e.
+with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}.
+Our results are in line with the empirical studies of said article, where the authors conclude that even though the asymptotic theoretical error is zero, most methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.
 
@@
 
-One requirement that our model should satisfy is that it be able to recover the true parameter function when the underlying model is a RKHS model.
+One requirement that our model should satisfy is that it should be able to recover the true parameter function when the underlying model is a RKHS model.
 This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?

### 2022-09-02 chapters/introduction.tex
Finish document

@@
 Over the last few decades, situations involving data in the form of functions have become commonplace in many statistical scenarios, as more and more information is available worldwide with an ever-increasing level of granularity in the measurements.
-In particular, functional data problems are far from unheard of in the data science an machine learning community, since they have attracted the attention of researchers and practitioners equally.
+In particular, functional data problems are far from unheard of in the data science and machine learning community, since they have attracted the attention of researchers and practitioners equally.
 Medical data, weather indicators or stock exchange indices are examples of elements that benefit from a functional treatment, where the observations are regarded as single entities rather than as a conglomerate of individual points.
@@
 In recent years numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum.
-A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}.
+A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{ramsay2005functional}, \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}.
 As the name suggests, FDA techniques are heavily inspired by functional analysis tools and methods: Hilbert spaces, orthonormal systems, linear operators, and so on.
@@
 We will demonstrate throughout this work how these spaces of functions possess properties that allow for an efficient treatment of functional data.
+
 On the other hand, Bayesian inference methods are ubiquitous in the realm of statistics, and their usual non-parametric approach also makes use of random functions, though in a slightly different manner than in the FDA context.
@@
 Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios amply explored in the literature.
+
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e.
@@
 It is beyond the scope of this thesis to provide a complete review of FDA as a whole, or to delve too deeply into the details and inner workings of most models and techniques mentioned.
-Nevertheless, references are provided frequently throughout the text to point the interested reader towards more specialized resources.
+Nevertheless, references are often provided throughout the text to point the interested reader towards more specialized resources.
 

### 2022-09-02 chapters/model-choice.tex
Revision
Finish document

@@
 In our case, this happens because the likelihood is symmetric with respect to the ordering of the parameters \(b\) and \(\tau\), i.e., \(\pi(Y|X,\theta)=\pi(Y|X, \nu(\theta))\) for any permutation \(\nu\) that rearranges the indices \(j=1,\dots, p\).
-Thus, since the components are arbitrarily ordered, they may be inadvertedly exchanged from one iteration to the next in any MCMC algorithm.
+Thus, since the components are arbitrarily ordered, they may be inadvertently exchanged from one iteration to the next in any MCMC algorithm.
 This can cause nonsensical answers when summarizing the marginal posterior distributions to perform inference, as different labelings might be mixed on each component \citep{stephens2000dealing}.
@@
 A common approach that seems to work well is to simply impose an ordering in the offending parameters, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\).
-We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the most separation between any two of them (suitably averaged across all iterations of the chains).
+We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).
 
@@
   \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter.
+That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter (which can be selected for example by cross-validation strategies).
 This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
@@
   \caption{Example of posterior predictive graphical checks on a fitted linear regression model.
-We show a comparison between the observed distribution of the response variable and the posterior predictive distribution consisting on the approximate sampled responses.}\label{fig:ppc_linear}
+We show a comparison between the observed distribution of the response variable and the posterior predictive distribution consisting on the approximate sampled responses.}\label{fig:ppc}
 \end{figure}
 
-In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the real value of the statistic.
+In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(\symbf Y^*)\leq T(\symbf Y)| \symbf Y)\), and are computed by simply measuring the proportion of the \(M\) estimates \(T\{\symbf Y^{(m)*}\}\) that fall below the observed value of the statistic, \(T(\symbf{Y})\).
 They are expected to be around 0.5 when the model accurately represents the data, and a deviation either way can be indicative of modeling issues; see Chapter 6 of \citet{gelman2013bayesian} for details.

### 2022-09-04 chapters/appendix-code.tex
Finish first complete revision

@@
 
-When possible, the code was implemented in a generic way that would allow for easy extensions or derivations, and also with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries.
+When possible, the code was implemented in a generic way that would allow for easy extensions or derivations.
+It was also developed with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries.
 Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them.

### 2022-09-04 chapters/background-related-work.tex
Finish first complete revision

@@
 Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}.
-At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}.
+At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework.
+Examples of this endeavor include the definition of centraility measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others.
+Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.
+
 Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}.
@@
 A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
-This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).
+This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study \citep[see][]{cardot2011functional}.
 
 Technical details aside, the inference on \(\beta\) is also hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
-Although it can be tempting to simply discretize the observed values on a grid and proceed with standard multiple linear regression, this would result in an under-determined model in which the estimated parameters lose meaning (essentially because the number of available parameters is infinite, but the number of equations is finite).
+Although it can be tempting to simply discretize the observed values on a grid and proceed with standard multiple linear regression, this would result in an under-determined model in which the estimated parameters lose meaning.
+This happens essentially because the number of available parameters is infinite, but the number of equations is finite \citep[see][Sec.~15.2]{ramsay2005functional}.
 Thus, some regularization or dimensionality reduction techniques are needed for parameter estimation; see \citet{reiss2017methods} for a summary of several widespread methods.
@@
 A common inference strategy is to expand both \(X\) and \(\beta\) on a certain data-driven orthonormal basis of \(L^2[0,1]\), say \(\{\phi_j\}\), up to a certain value \(p\in\N\):
+
 \[
@@
 
-The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis when new information is available.
+The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis (or a parameter) when new information is available.
 It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}.
@@
 For a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}.
-
+Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity.
 We start by defining the concept of kernel functions, which are the foundation of these spaces.
-Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity.
 
@@
 
-Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and without loss of generality assume \(\E[X(t)]=0\) for all \(t\in[0,1]\).
+Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and for simplicity assume \(\E[X(t)]=0\) for all \(t\in[0,1]\).
 Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\).
@@
   As we pointed out before, the space of square integrable functions \(L^2(\R)\) \textit{is not} a RKHS.
-There are many additional ways of seeing this; for example, in this space norm convergence does not imply pointwise convergence.
+There are many ways of seeing this; for example, in this space norm convergence does not imply pointwise convergence.
 Another possibility is to observe that the would-be reproducing kernel must be the Dirac delta function:
@@
 \]
-The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with an simple expression but a complex feature space (e.g.
+The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex feature space (e.g.
 infinite-dimensional), such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).
@@
 \begin{theorem}[Generalized representer theorem]
-  Consider a RKHS \(\Hcal(K)\) of functions defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
+  Consider a RKHS of functions \(\Hcal(K)\) defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
 Then, any minimizer \(f^\ast\) of the regularized empirical risk
@@
 
-The Metropolis-Hastings algorithm \citep{metropolis1953equation} is arguably the most well-known algorithm in this context.
+The Metropolis-Hastings algorithm \citep{metropolis1953equation} is arguably the best known algorithm in this context.
 As a matter of fact, it is actually more of a general framework from which many other methods are derived.
@@
 Note that \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\).
-This specific acceptance rate is also chosen so that the detailed balance equations of the chain are satisfied.
+This specific acceptance rate is also chosen so that the detailed balance equations of the chain are satisfied, i.e., \(\pi(x)P(x'|x)=\pi(x')P(x|x')\), where \(P(x|y)\) represents the transition probability from state \(x\) to state \(y\).
+These conditions essentially guarantee that \(\pi\) is the stationary distribution of the chain; see \citet{robert1999monte} for additional details.
 An immediate generalization consists on choosing a general (possibly non-symmetric) jump distribution \(g\), in which case the acceptance probability is rewritten as
@@
 Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high.
-If we know the conditional distribution of each variable, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension.
+If we know the conditional distribution of each variable on the rest of them, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension.
 Specifically, in each step we get unidimensional samples \textit{in turn} (e.g.

### 2022-09-04 chapters/bayesian-functional-regression.tex
Finish first complete revision

@@
   \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) that acts as a summary of the marginal posterior distributions, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\), and then predict the responses in the usual way following model~\eqref{eq:rkhs-model-linear}, i.e.:
-  \[
+  \begin{equation}\label{eq:summarize-predict-linear}
     \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
-  \]
+  \end{equation}
   Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter.
@@
 With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
-\[
+\begin{equation}\label{eq:summarize-predict-logistic}
 \hat Y_i = \I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
-\]
+\end{equation}
 where \(\I\) is the indicator function (\(\I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise).
@@
 
-On the other hand, the prediction method that takes into account the entire posterior approximation (i.e.
-the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
+On the other hand, the prediction method that takes into account the entire posterior approximation (that is, the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
 Hence, there are primarily two possible outcomes.

### 2022-09-04 chapters/conclusions.tex
Finish first complete revision

@@
 In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios.
-Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective that are conceptually simpler and facilitate the interpretation and subsequent analysis, an attribute that practitioners surely find especially compelling.
+Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective.
+This leads to a conceptually simpler approach that facilitates the interpretation and subsequent analysis, which we believe is a compelling feature for most practitioners.
 The underlying theoretical aspects are nevertheless far from trivial, so we have also explored some of the related mathematical background in this text.
@@
 Although this simplicity comes at the expense of increased computational requirements, this computational cost is still reasonable and the model can be used in practice.
+The execution time only starts to be of some relevance when performing extensive simulations and comparisons, but it is not a serious burden for the use of our methods in practical cases.
 

### 2022-09-04 chapters/experiments.tex
Finish first complete revision

@@
 
-First and foremost, we consider \(T=\{\text{mean, median, mode}\}\) for the summarize-then-predict approach to prediction.
-In this way in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach, which will we call ``\textit{posterior\_mean}'' for short) and 3 variable selection methods (one for each statistic), while in logistic regression there are similarly 5 prediction methods and 3 variable selection procedures.
-In each case, after variable selection is performed, we use a penalized multiple linear/logistic regression method to generate the corresponding predictions.
+First and foremost, we consider \(T=\{\text{mean, median, mode}\}\) for our summarize-then-predict approach to prediction (see~\eqref{eq:summarize-predict-linear} and~\eqref{eq:summarize-predict-logistic}).
+As a result, in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach) and 3 variable selection methods (one for each statistic), while in logistic regression there are similarly 5 prediction methods and 3 variable selection procedures.
+In each case, after variable selection is performed, we use a \(l^2\)-penalized multiple linear/logistic regression method to generate the corresponding predictions.
 All in all, we are looking at 7 (8 in the case of logistic regression) prediction methods, and although all of them are derived from a single MCMC run, we will treat them as separate for the purposes of the experimentation.
@@
 
-We consider a set of functional regressors common to linear and logistic regression problems. They are 4 Gaussian processes (GP's), each with a different kernel function:
+We consider a set of functional regressors common to linear and logistic regression problems. They are 4 Gaussian processes (GP's), each with a different covariance function:
 \begin{enumerate}
@@
   \item O-U.\hspace{0.3em} An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
-  \item Gaussian.\hspace{0.3em} A Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), with \(\nu=0.2\).
+  \item Gaussian.\hspace{0.3em} A Gaussian process with Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), where \(\nu=0.2\).
 \end{enumerate}
@@
 In Figure~\ref{fig:dependence_time_n_train} we show the evolution of the execution time versus the number of training samples in a generic case, which not surprisingly has a more or less linear behavior.
+This, however, is not an impediment for using our methods in practice.
 

### 2022-09-04 chapters/introduction.tex
Finish first complete revision

@@
 As the name suggests, FDA techniques are heavily inspired by functional analysis tools and methods: Hilbert spaces, orthonormal systems, linear operators, and so on.
-In particular, a notion that also intersects with the classical theory of machine learning and pattern recognition is that of reproducing kernel Hilbert spaces (RKHS's).
+In particular, a notion that also intersects with the classical theory of machine learning and pattern recognition, and that has gained traction in recent years, is that of reproducing kernel Hilbert spaces (RKHS's).
 We will demonstrate throughout this work how these spaces of functions possess properties that allow for an efficient treatment of functional data.
@@
 Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions.
-Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios amply explored in the literature.
-
+Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios widely explored in the literature.
 To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e.
@@
 The underlying probability space \((\Omega, \mathcal A, \P)\) is not important.
-We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
+We will further suppose that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
 In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(x_i, y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\).
@@
 
-\begin{figure}[ht]
+\enlargethispage{1\baselineskip}
+
+\begin{figure}[htbp!]
   \centering

### 2022-09-04 chapters/model-choice.tex
Finish first complete revision

@@
 The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein.
-A common approach that seems to work well is to simply impose an ordering in the offending parameters, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\).
+A common approach that seems to work well is to simply impose an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\).
 We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).

### 2022-09-05 chapters/appendix-code.tex
Finish document

@@
   \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our models, separated according to the functionality they provide.
-  \item The \texttt{reference\_methods} folder contains the implementation of the functional comparison algorithms that were not available through a standard Python library.
+  \item The \texttt{reference\_methods} folder contains our implementation of the functional comparison algorithms that were not available through a standard Python library.
   \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g.

### 2022-09-05 chapters/background-related-work.tex
Finish document

@@
 At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include the definition of centraility measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others.
+Examples of this endeavor include the definition of centrality measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others.
 Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.
@@
 A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}).
-This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study \citep[see][]{cardot2011functional}.
+This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and uniqueness of \(\beta\) need further study \citep[see][]{cardot2011functional}.
 
@@
 Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion.
-As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures.
+As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs, bueno2019variable}, but in general they have their own dedicated algorithms and procedures.
 As will become apparent in the forthcoming chapters, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model.
@@
 A way to bypass this limitation is to use conjugate distributions, where it can be rigorously proven that a certain combination of likelihood and prior distributions produces a posterior distribution in the same family as the prior.
-However, unless conjugate priors are used, a closed-form expression is generally unattainable and some type of approximation is required.
+However, unless conjugate priors are used, a closed-form expression of the posterior is generally unattainable and some type of approximation is required.
 Apart from basic numerical integration, some well-performing methods in this regard are variational inference approaches \citep[e.g.][]{blei2017variational} and MCMC methods \citep[e.g.][]{brooks2011handbook}.
@@
 
-On a separate note, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by Bayesian neural networks.
-Some examples are the functional Bayesian neural networks proposed by~\citet{sun2019functional}, variational implicit processes \citep{ma2019variational} and their deep variants \citep{ortega2022deep}, or the functional variational inference approach suggested in~\citet{ma2021functional}.
+On a separate note, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by a Bayesian neural network.
+Some examples are the functional Bayesian neural networks proposed by~\citet{sun2019functional}, variational implicit processes \citep{ma2019variational} and their deep variants \citep{ortega2022deep}, or the functional variational inference techniques suggested in~\citet{ma2021functional}.
 Moreover, there are approaches to functional regression from a Bayesian perspective that impose a Gaussian process prior on the functional parameter \(\beta\) \citep[e.g][]{lian2016posterior}, and it turns out that the RKHS's corresponding to these Gaussian processes, described for example in~\citet{van2008reproducing}, are useful when studying contraction rates of the posterior distribution.
@@
 Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory.
-For a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}.
-Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity.
+Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity; for a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}.
 We start by defining the concept of kernel functions, which are the foundation of these spaces.
@@
 for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).
-We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}: \(f(t) = \dotprod{K(t, \cdot)}{f}_K\) for all \(t \in [0, 1]\).
+We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}:
+\[
+f(t) = \dotprod{K(t, \cdot)}{f}_K, \quad \text{for all } t \in [0, 1].
+\]
 In particular, \(K(t, s)=\dotprod{K(t, \cdot)}{K(s, \cdot)}_K\), which can be understood as saying that the kernel \textit{reproduces} itself, hence the term ``reproducing kernel''.
@@
     \]
-    Now, if \(\dotprod{f}{f}_K=0\), by letting \(\epsilon\to 0^{\pm}\) we see that \(f(t)=\dotprod{f}{K(t, \cdot)}_K\) necessarily vanishes for all \(t\), as desired.
+    Now, if \(\dotprod{f}{f}_K=0\), by letting \(\epsilon\to 0^{\pm}\) it follows that \(f(t)=\dotprod{f}{K(t, \cdot)}_K\) necessarily vanishes for all \(t\), as desired.
   \end{proof}
@@
 \begin{proof}
-  To prove \((i)\), simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
+  To prove \textit{(i)}, simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
   \[
@@
   \]
-  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \((ii)\), observe that for any \(t\in[0,1]\) we have
+  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \textit{(ii)}, observe that for any \(t\in[0,1]\) we have
   \[
@@
   \]
-  as we have just shown in \((i)\) that the evaluation operator is bounded.
+  as we have just shown in \textit{(i)} that the evaluation operator is bounded.
 Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform.
-Finally, we prove the second statement in \((iii)\), and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof.
+Finally, we prove the second statement in \textit{(iii)}, and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof.
 If \(K\) is continuous, then it is clear that every function in \(\Hcal_0(K)\) is continuous.
@@
 By definition, there exists a sequence \(\{f_n\}\subset \Hcal_0(K)\) of continuous functions such that \(\|f_n - f\|_K \to 0\).
-But then \(f_n \to f\) uniformly by \((ii)\), and therefore \(f\) is continuous as the uniform limit of continuous functions.
+But then \(f_n \to f\) uniformly by \textit{(ii)}, and therefore \(f\) is continuous as the uniform limit of continuous functions.
 Note that this statement remains valid, with a slightly different proof, even when the underlying domain is not compact.
@@
  Note that, as expected, the covariance function of \(X\) plays a crucial role in characterizing the associated RKHS.
-An integral operator closely related to the kernel is the so-called covariance operator, namely
+An integral operator closely related to this covariance function is the so-called covariance operator, namely
 \[
@@
 In connection with our RKHS theory, it can be shown that the set \(\{\sqrt{\lambda_j}\phi_j\}\) constitutes an orthonormal basis of \(\Hcal(K)\) \citep[see e.g.][Sec.~4.4]{cucker2007learning}.
-Furthermore, a similar orthogonal decomposition holds for our stochastic process \(X\), an expansion which lies at the base of many \(L^2\) functional regression models.
+Furthermore, a similar orthogonal decomposition holds for the stochastic process \(X\), an expansion that lies at the base of many \(L^2\) functional regression models.
 The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Sec.~3.2]{berlinet2004reproducing}.
@@
 
+\enlargethispage{1\baselineskip}
+
 In addition to these results, it is worth mentioning that the covariance operator \(\K\) provides several alternative definitions of \(\Hcal(K)\).
@@
 Nevertheless, we will now hopefully clarify that this space can indeed be seen as \textit{the} Hilbert space inherently associated with the process \(X\) itself.
-Let us shift the perspective momentarily and consider the linear span of \(X\) in the space of all zero-mean random variables with finite second moment, i.e.:
+Let us shift the perspective momentarily and consider the space \(L^2(\Omega)\) of all zero-mean random variables with finite second moment, endowed with the usual norm \(\|U\|^2=\E[U]^2\).
+Then, the linear span of \(X\) in \(L^2(\Omega)\) is given by
 \[
-\Lcal_0(X) = \left\{\sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
+\Lcal_0(X) = \left\{U \in L^2(\Omega): U = \sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
 \]
@@
 It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
-For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere.
-An heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
+For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the associated RKHS are differentiable almost everywhere.
+A heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
 \[
@@
 \end{proposition}
-Note that the statement \((iii)\) above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.
+Note that the statement \textit{(iii)} above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.
 
@@
 Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning.
-For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS.
+For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property.
+First, a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS.
 Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
@@
 \]
-The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex feature space (e.g.
-infinite-dimensional), such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).
+The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex (e.g.
+infinite-dimensional) feature space, such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).
 
@@
 \begin{theorem}[Generalized representer theorem]
-  Consider a RKHS of functions \(\Hcal(K)\) defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
+  Consider a RKHS of functions \(\Hcal(K)\) defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(\mathcal R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\).
 Then, any minimizer \(f^\ast\) of the regularized empirical risk
   \[
-    R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
+    \mathcal R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
   \]
@@
 In these cases, we could work with a function \(f(x)\) proportional to our target density \(\pi(x)\).
-More specifically, in a Bayesian framework the posterior distribution is often intractable due to the normalizing integral constant, but we do know that \(posterior \propto prior \times likelihood\).
+More to the point, in a Bayesian framework the posterior distribution is often intractable due to the normalizing integral constant, but we do know that \textit{posterior} \(\propto\) \textit{prior} \(\times\) \textit{likelihood}.
 
@@
 This is where Markov chain Monte Carlo (MCMC) methods come into play.
+
+\vspace{1em}
 
@@
 In this way, we can take the samples of a sufficiently advanced chain as approximate realizations of \(X\sim \pi(x)\).
-When it comes to building the chains, transition probabilities are assigned in such a way that regions with higher density with respect to \(\pi(x)\) are favored (by means of the known proportional function \(f(x)\)).
+When it comes to building the chains, the transition probabilities are assigned in such a way that regions with higher density with respect to \(\pi(x)\) are favored (by means of the known proportional function \(f(x)\)).
 The dimensionality issues are somewhat mitigated, and on top of that there are specific tuning techniques to tackle them.
-However, the samples obtained from this procedure are not independent, though they can be made so by \textit{thinning} the chain and considering only one every few samples.
+However, the samples obtained from this procedure are not independent, though they can be made more or less so by \textit{thinning} the chain and considering only one every few samples.
 
 Another advantage of these methods is that they recover the marginal distributions directly.
-Suppose that after a successful MCMC run we have \(M\) multivariate samples \(\{\theta^{(m)}=(\theta_1^{(m)},\dots, \theta_p^{(m)})\}\) of a joint \(p\)-dimensional distribution.
+Suppose that after a successful MCMC run we have \(M\) multivariate approximate samples \(\{\theta^{(m)*}=(\theta_1^{(m)*},\dots, \theta_p^{(m)*})\}\) of a joint \(p\)-dimensional distribution.
 The marginal distribution of each variable, which would be theoretically computed as
@@
 \]
-can be approximated by just retaining the samples \(\{\theta_i^{(m)}\}\) corresponding to \(\theta_i\).
+can be approximated by just retaining the samples \(\{\theta_i^{(m)*}\}\) corresponding to \(\theta_i\).
 
@@
 \end{enumerate}
-
  \begin{figure}[ht!]
@@
  \end{figure}
-
 Figure~\ref{fig:mh} shows a schematic representation of the process.
-Note that \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\).
-This specific acceptance rate is also chosen so that the detailed balance equations of the chain are satisfied, i.e., \(\pi(x)P(x'|x)=\pi(x')P(x|x')\), where \(P(x|y)\) represents the transition probability from state \(x\) to state \(y\).
+Note that the acceptance ratio \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\).
+This specific acceptance value is also chosen so that the detailed balance equations of the chain are satisfied, i.e., \(\pi(x)P(x'|x)=\pi(x')P(x|x')\), where \(P(x|y)\) is the transition probability from state \(x\) to state \(y\).
 These conditions essentially guarantee that \(\pi\) is the stationary distribution of the chain; see \citet{robert1999monte} for additional details.
@@
 via rejection sampling) of each variable conditional on the rest of them, using always the more up-to-date values available:
-
 \[
@@
 
-Generally speaking, a MCMC algorithm can be described at time \(t+1\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi(x)\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d.
+Generally speaking, a MCMC algorithm can be described through a function \(R\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d.
 random variables that represent the random behavior of the chain.
-With this notation, the property of affine-invariant can be characterized as \citep{goodman2010ensemble}:
+With this notation, the  affine-invariance property can be characterized as \citep{goodman2010ensemble}:
  \[
@@
 
-From a computational perspective, the Python library \textit{emcee}\footnote{\url{https://emcee.readthedocs.io/en/stable/}} \citep{foreman2013emcee} provides a parallel implementation of this algorithm.
+From a computational perspective, the Python library \textit{emcee}\footnote{\url{https://emcee.readthedocs.io}} \citep{foreman2013emcee} provides a parallel implementation of this algorithm.
 The idea is to divide the ensemble \(\Lambda\) into two equally-sized subsets \(\Lambda^{(0)}\) and \(\Lambda^{(1)}\), and then proceed on each iteration in the following alternate fashion:

### 2022-09-05 chapters/bayesian-functional-regression.tex
Finish document

@@
 
-However, after some intial experimentation we found that, for practical and computational reasons, the value of \(p\), the dimensionality of the model, is best fixed beforehand in a suitable way (see Chapter~\ref{ch:model-choice} for details).
+However, after some initial experimentation we found that, for practical and computational reasons, the value of \(p\), the dimensionality of the model, is best fixed beforehand in a suitable way (see Chapter~\ref{ch:model-choice} for details).
 Thus, we will regard only the \(\beta_j\) and \(t_j\) as free parameters, and search for our functional parameter in the space
@@
 
-In the case of functional linear regression, the simplified RKHS model considered in this work is given by
+In the case of functional linear regression, the simplified RKHS model considered is
 \begin{equation}\label{eq:rkhs-model-linear}
-  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
+  Y = \alpha_0 + \dotprod{X}{\alpha}_K + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
 \end{equation}
@@
 When \(p\) is fixed, the parameter space becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this \((2p+2)\)-dimensional space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\).
-Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\D_n\).
+Before proceeding further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\D_n\).
 For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
@@
 
-It is worth mentioning that the model remains linear in the sense that it fundamentally involves random variables belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
+It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable \(\dotprod{X}{\alpha}_K = \Psi_X(\alpha)\) belonging to the linear span of the process \(X\) in \(L^2(\Omega)\).
 Also, note that given the time instants \(t_j\), the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates.
-As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
-In addition, this model could be easily extended to the case of several covariates via an expression of type \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\).
+As a matter of fact, this model is particularly suited as a basis for variable selection methods, and furthermore the general RKHS model entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}.
+In addition, this model could be easily extended to the case of several covariates via an expression of type \(Y=\alpha_0 + \dotprod{X^1}{\alpha_1}_K + \cdots + \dotprod{X^q}{\alpha_q}_K + \epsilon\).
 In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.
+
+\enlargethispage{1\baselineskip}
 
@@
 
-  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) that acts as a summary of the marginal posterior distributions, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\), and then predict the responses in the usual way following model~\eqref{eq:rkhs-model-linear}, i.e.:
+  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) that acts as a summary of the marginal posterior distributions, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\), and then we can predict the responses in the usual way following model~\eqref{eq:rkhs-model-linear}, i.e.:
   \begin{equation}\label{eq:summarize-predict-linear}
@@
 
-  \paragraph{Predict-then-summarize.} Alternatively, we can  look at the approximate posterior distribution as a whole and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
+  \paragraph{Predict-then-summarize.} Alternatively, we can  look at the approximate posterior distribution as a whole, and compute the predictive distribution of the simulated responses at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
   \begin{equation}\label{eq:sampled-response-vector}
@@
 Moreover, one could even contemplate an \textit{ensemble model} in which some kind of aggregation of several of the available prediction methods is performed to produce a final result.
-Also, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
-In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(L^1\)-type loss function while the former implicitly optimizes an \(L^2\) loss.
+
+Besides, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters.
+In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(l^1\)-type loss function while the former implicitly optimizes an \(l^2\) loss.
 On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
@@
 
+\newpage
+
 \section{Functional logistic regression}\label{sec:rkhs-logistic-model}
 
-In the case of functional logistic regression, we can regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and as suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model is given  by the  equation
+In the case of functional logistic regression, we can regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given \(X=x \in L^2[0, 1]\), and as usual suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
+Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model might be given, in terms of the isometric transformation \(X \mapsto \dotprod{X}{\alpha}_K = \Psi_X(\alpha)\), by the equation
 \begin{equation}\label{eq:rkhs-model-logistic}
-  \P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
+  \P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \dotprod{X}{\alpha}_K\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
 \end{equation}
@@
 \end{equation}
-where in turn \(\beta_j\in\R\) and \(t_j\in[0, 1]\).
+where in turn \(\alpha_0, \beta_j\in\R\) and \(t_j\in[0, 1]\).
 
-In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model.
+In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model~\eqref{eq:l2-logistic-model}.
 First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS logistic functional model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model.
+Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in this case it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model.
 Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
@@
 \begin{align*}
-  \log \pi(\theta \mid \D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
+  \log \pi(\theta \mid \D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \dotprod{X_i}{\alpha}_K\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \dotprod{X_i}{\alpha}_K\right\}\right)\right]\\
   \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
 \end{align*}
-Remember that \(\Psi_{X_i}(\alpha) = \sum_{j=1}^p \beta_j X_i(t_j)\).
+Remember that \(\dotprod{X_i}{\alpha}_K = \sum_{j=1}^p \beta_j X_i(t_j)\).
 \end{proposition}
@@
 The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
-
 On the other hand, the prediction method that takes into account the entire posterior approximation (that is, the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
@@
   \]
-  In this case, each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
-Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) from the majority vote of all the \(Y_i^{(m)*}\).\\
+  In this case, each \(Y_i^{(m)*}\) is a random variable that follows a Bernoulli distribution with parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
+Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) according to the majority vote of all the \(Y_i^{(m)*}\).\\
 

### 2022-09-05 chapters/conclusions.tex
Finish document

@@
 In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios.
-Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective.
+Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations but still with a functional perspective.
 This leads to a conceptually simpler approach that facilitates the interpretation and subsequent analysis, which we believe is a compelling feature for most practitioners.
@@
 We implemented the whole inference and prediction pipeline in the Python programming language, and then used it to carry out a large empirical study to assess the performance of our models against other usual alternatives, functional and otherwise.
-The results were generally satisfactory, proving that we have competitive methods that achieve good results in different scenarios, often with low dimensionality and thus increased tractability and interpretability.
-Lastly, we believe that with more computational resources we could consider a larger pool of hyperparameters in the model selection phase, which would most likely contribute to better results.
+The results were generally satisfactory, proving that we have competitive methods that achieve good performance in different scenarios, often with low dimensionality and thus increased tractability and interpretability.
+Lastly, we believe that with more computational resources we could consider a larger pool of hyperparameters in the model selection phase, and also increase the number of folds and random train/test splits, which would most likely contribute to better results.
 
@@
 
-Some lines of further research suggested by our work are the following.
+Some lines of further research suggested by our work are the following:
 

### 2022-09-05 chapters/experiments.tex
Finish document

@@
 
-In this chapter we present the results of the experiments carried out to test the performance of our models in different scenarios.
+In this chapter we present the results of the experiments carried out to test the performance of our models in different scenarios, both simulated and with real data.
 The structure of the code and other minor computational details are discussed in Appendix~\ref{ch:code}.
@@
 We use 64 chains and run them for 900 iterations in total, discarding the first 500 iterations as burn-in.
-Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains at each iteration, selecting the stretch move (the default) with probability \(0.7\), or the walk move with probability \(0.3\).
+Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains at each iteration, selecting the stretch move (the default) with probability 0.7 or the walk move with probability 0.3.
 
@@
 
-We consider a set of functional regressors common to linear and logistic regression problems. They are 4 Gaussian processes (GP's), each with a different covariance function:
+We consider a set of functional regressors common to linear and logistic regression problems. They are four Gaussian processes (GP's), each with a different covariance function:
 \begin{enumerate}
@@
 
-\paragraph{Linear regression data sets.} We consider three different types of data sets, all with a common value of \(\alpha_0=5\).
+\paragraph{Linear regression data sets.} We consider three different types of simulated data sets, all with a common value of \(\alpha_0=5\).
 \begin{itemize}
@@
 The first consists of near-infrared spectra of 100 wheat samples and the objective is to predict the samples' moisture content, whereas the second contains 268 samples of sugar fluorescence data in order to predict ash content.
-The three data sets are measured on a grid of 100, 101 and 115 equispaced points in \([0, 1]\), respectively.
+The three data sets are measured on a grid of 100, 101 and 115 equispaced points on \([0, 1]\), respectively.
 
-\paragraph{Logistic regression data sets.} Again we consider three different types of data sets,  with a common value of \(\alpha_0=-0.5\).
+\paragraph{Logistic regression data sets.} Again we consider three different types of simulated data sets,  with a common value of \(\alpha_0=-0.5\).
 In this case we randomly permute 10\% of the labels to introduce some noise in the simulations.
@@
   \[
-  \P(Y=1\mid X) = \frac{1}{1 + e^{0.5 +5X(0.1) - X(0.4) - 10X(0.8)}}.
+  \P(Y=1\mid X) = \frac{1}{1 + \exp\left\{0.5 +5X(0.1) - X(0.4) - 10X(0.8)\right\}}.
   \]
-  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case.
+  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case, i.e.:
+  \[
+  \P(Y=1\mid X) = \frac{1}{\displaystyle 1 + \exp\left\{0.5 - \int_0^1 \log(1+4t) X(t)\, dt\right\}}.
+  \]
   \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's (with equal probability \(p=1/2\)) and label them according to their origin.
 Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\).
-Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance \(2\), that is, with kernel \(K(t,s)=2\min\{t,s\}\).
+Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance 2, that is, with kernel \(K(t,s)=2\min\{t,s\}\).
 \end{itemize}
-In this case there are again three real data sets.
+In this case we use three real data sets well known in the literature.
 The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, trying to predict if their longevity is high or low.
@@
 We have included a fairly comprehensive suite of comparison algorithms, chosen among the most common methods used in machine learning and FDA.
-There are finite-dimensional models that work on the discretized data, variable selection/dimension reduction procedures and purely functional methods.
+There are purely functional methods, finite-dimensional models that work on the discretized data, and variable selection/dimension reduction procedures.
 
-\paragraph{Linear regression comparison algorithms.} We consider the following methods.
+\paragraph{Linear regression comparison algorithms.} We consider the following regression methods:
 \begin{itemize}
@@
 
-\paragraph{Logistic regression comparison algorithms.} All the variable selection and dimension reduction algorithms from above are also considered in this case.
+\paragraph{Logistic regression comparison algorithms.} All the variable selection and dimension reduction algorithms from above are also considered in this case, plus the following classification methods:
 \begin{itemize}
@@
   \item RKVS.\hspace{.3em} RKHS-based variable selection method proposed in \citet{berrendero2018use}.
-  \item APLS.\hspace{.3em} Functional PLS used as a dimension reduction method, as proposed in \citet{delaigle2012achieving} in combination with the nearest centroid algorithm.
+  \item APLS.\hspace{.3em} Functional PLS used as a dimension reduction method, as proposed in \citet{delaigle2012achieving} in combination with the nearest centroid (NC) algorithm.
   \item FLog.\hspace{.3em} Functional RKHS-based logistic regression algorithm proposed in \citet{berrendero2018functional}.
-  \item FLDA.\hspace{.3em} Implementation of the so-called functional linear discriminant analysis in \citet{preda2007pls}.
+  \item FLDA.\hspace{.3em} Implementation of the functional version of linear discriminant analysis proposed in \citet{preda2007pls}.
   \item MDC.\hspace{.3em} Maximum depth classifier \citep[e.g.][]{ghosh2005maximum}.
@@
 Indeed, we can see that in most cases our algorithms are the ones with lower RMSE, save for a few exceptions, notably with the Gaussian kernel.
-A subsequent analysis showed that this particular data set is especially sensitive to the value of the hyperparameter \(\eta\) in our prior distribution, so a more tailored approach would be needed to obtain better results.
+A subsequent analysis showed that this particular data set is especially sensitive to the value of the hyperparameter \(\eta\) in our prior distribution, so a more customized approach would be needed to obtain better results.
+
+\vspace{.3em}
 
@@
 Moreover, whenever one of our models has a higher RMSE, the difference is pretty small in comparison.
+Note that some of our Bayesian models have a higher standard deviation, partly because there is an intrinsic randomness in the methods (apart from the train/test splits), and it can be the cause of the occasional worse performance.
+In relation to this, we observe that the methods that use the mean as a summary statistic tend to perform much worse.
+This is because the mean is very sensitive to outliers, so if at some point there is a chain that randomly deviates from the rest, the mean of the posterior distribution will be greatly impacted.
+In particular, very high values of the parameter \(\sigma^2\) were reported sometimes, producing misleading results when averaged with the rest of the values.
+
+\enlargethispage{1\baselineskip}
+
+Lastly, the results for the non-GP case can be seen in Figure~\ref{fig:reg_emcee_nongp}, in which the regressors are realizations of a GBM.
+In this case we still get better results under the RKHS model, while the results under the \(L^2\)-model are slightly worse.
+However, as before, the difference is small (except for the \textit{emcee\_mean} methods).
+
+\vspace{.2em}
+
 \begin{figure}[ht!]
@@
 
-Note that some of our Bayesian models have a higher standard deviation, partly because there is an intrinsic randomness in the methods (apart from the train/test splits), and it can be the cause of the occasional worse performance.
-In relation to this, we observe that the methods that use the mean as a summary statistic tend to perform much worse.
-This is because the mean is very sensitive to outliers, so if at some point there is a chain that randomly deviates from the rest, the mean of the posterior distribution will be greatly impacted.
-In particular, very high values of the parameter \(\sigma^2\) were reported sometimes, producing misleading results when averaged with the rest of the values.
-
-Lastly, the results for the non-GP case can be seen in Figure~\ref{fig:reg_emcee_nongp}, in which the regressors are realizations of a GBM.
-In this case we still get better results under the RKHS model, while the results under the \(L^2\)-model are slightly worse.
-However, as before, the difference is small (except for the \textit{emcee\_mean} methods).
-
 \begin{figure}[ht!]
   \centering
-  \includegraphics[width=.7\textwidth]{reg_emcee_nongp}
+  \includegraphics[width=.65\textwidth]{reg_emcee_nongp}
   \caption{RMSE of predictors for simulated data with GBM regressors (lower is better).}\label{fig:reg_emcee_nongp}
@@
 However, the predict-then-summarize approach (represented as \textit{posterior\_mean}) seems to work quite well, always scoring near the mean RMSE of all the comparison algorithms.
-Moreover, our two-stage methods seem to perform better than our other methods in these cases, scoring again very close to the mean of the reference methods.
+Moreover, our two-stage methods seem to outperform the summarize-then-predict methods in these cases, scoring again very close to the mean of the reference methods.
 
-We have to bear in mind that real data is more complex and noisy than simulated data, and it is possible that after a suitable pre-preprocessing we could obtain better results with our methods.
-However, our goal was to perform a general comparison without exploring too much to the specifics of any particular data set.
+We have to bear in mind that real data is more complex and noisy than simulated data, and it is possible that after a suitable pre-preprocessing we would obtain better results with our methods.
+However, our goal was to perform a general comparison without focusing too much on the specifics of any particular data set.
 
@@
 
-\begin{figure}[ht!]
+Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case.
+Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e.
+with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}.
+Our results are in line with the empirical studies of said article, where the authors conclude that even though the asymptotic theoretical error is zero, most classification methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.
+
+\begin{figure}[htbp!]
   \centering
@@
 
-Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case.
-Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e.
-with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}.
-Our results are in line with the empirical studies of said article, where the authors conclude that even though the asymptotic theoretical error is zero, most methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.
-
-\begin{figure}[ht!]
+\begin{figure}[htbp!]
   \centering
-  \includegraphics[width=.7\textwidth]{clf_emcee_nongp}
+  \includegraphics[width=.65\textwidth]{clf_emcee_nongp}
   \caption{Accuracy of classifiers for simulated data coming from two different GP's, labeled according to their origin (higher is better).}\label{fig:clf_emcee_nongp}
@@
 As for the real data sets, in Figure~\ref{fig:clf_emcee_real} we see positive results in general, obtaining in most cases accuracies well above the mean of the reference models, and sometimes above most of them.
-In particular, the predict-then-summarize tend to have a good performance and achieve a lower standard error across executions, which is a trend that we also saw in the simulated data sets.
+In particular, the predict-then-summarize methods tend to have a good performance and achieve a lower standard error across executions, which is a trend that we also saw in the simulated data sets.
 However, our variable selection methods seem to be a bit worse in this case.
@@
 
-One requirement that our model should satisfy is that it should be able to recover the true parameter function when the underlying model is a RKHS model.
+One requirement that our model should satisfy is that it should be able to recover the true parameter function when the underlying data generation model is a finite-dimensional RKHS model.
 This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?
 
-Take for example a data set generated according to the formula \(Y=-5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\).
+Take for example a RKHS data set with two components generated according to the formula \(Y=-5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\).
 Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(\tau_1, \tau_2, \tau_3)\) for a model with 3 components.
@@
   \includegraphics[width=.8\textwidth]{beta_trace_3}
-  \caption{Estimated posterior distribution for \(\beta\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
+  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
 \end{figure}
@@
   \includegraphics[width=.8\textwidth]{beta_trace_4}
-  \caption{Estimated posterior distribution for \(\beta\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
+  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
 \end{figure}
@@
 Another thing we wanted to look at was the dependence of the final prediction result on the chosen value of \(p\), especially when there is no concept of ``components'' in the underlying model.
-We can take for example the homoscedastic mixture data set in the logistic regression problem, and fix the parameter \(\eta=0.01\).
+We can take for example the homoscedastic mixture data set described earlier for the logistic regression problem, and fix the parameter \(\eta=0.01\).
 Then the corresponding mean accuracies (in 10 repetitions) for RKHS models with \(p=1,2,\dots 10\) components are shown in Figure~\ref{fig:dependence_acc_p}, along with their standard errors (which are arguably not very informative).
@@
 
-It would appear that in the methods that use the whole posterior distribution are more stable and somewhat independent of the value of \(p>1\).
-The other methods show a slight downward trend as \(p\) increases (although not so much in the variable selection methods), and in general their best results are obtained at \(p=2\) and \(p=8\).
+It would appear that the methods that use the whole posterior distribution are more stable and somewhat independent of the value of \(p>1\).
+The other algorithms show a slight downward trend as \(p\) increases (although not so much in the variable selection methods), and in general their best results are obtained at \(p=2\) and \(p=8\).
 We expect that this effect or some small variation of it will remain valid in other situations; however, a profound study of this would be the subject of a different experiment altogether.
@@
 An individual MCMC run is not that expensive time-wise, but even for a modest 5-fold cross-validation loop the effects add up quickly to a considerable amount of time.
-In Figure~\ref{fig:dependence_time_n_train} we show the evolution of the execution time versus the number of training samples in a generic case, which not surprisingly has a more or less linear behavior.
+In Figure~\ref{fig:dependence_time_n_train} we show the evolution of the execution time (in seconds) versus the number of training samples in a generic case, which not surprisingly has a more or less linear behavior.
 This, however, is not an impediment for using our methods in practice.
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{l2_dependence_time_n_train}
-  \caption{Mean training times of a logistic RKHS model, averaged across 10 independent runs, as a function of the number of training samples.
-The values are surrounded by the corresponding standard deviations.
-}\label{fig:dependence_time_n_train}
+  \includegraphics[width=0.55\textwidth]{l2_dependence_time_n_train}
+  \caption{Mean training times of a logistic RKHS model with \(p=5\), as a function of the number of training samples (averaged across 10 independent runs).
+The values are surrounded by their corresponding standard deviations.}\label{fig:dependence_time_n_train}
 \end{figure}
@@
 Although these trials did not produce better results in general, we mention some of them here to show how the model could be easily extended.
-In fact, many of these modifications, if not all, are still available in the provided code and can be enabled through specific parameters.
+In fact, most of these modifications, if not all, are still available in the provided code and can be enabled through specific parameters.
 
@@
 The results were similar, but we ultimately discarded this approach because it introduced yet another hyperparameter into the model with no clear benefit.
-In addition, we tried to impose a \(\beta\) distribution on each of the time instants \(t_j\), selecting the hyperparameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable).
+In addition, we tried to impose a beta distribution on each of the time instants \(\{t_j\}\), selecting the corresponding shape parameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable).
 This approach does work, but in our estimation the computational overhead introduced by the beta density function (sometimes resulting in a 1.5x increase in execution time) is not worth the mild improvements, if any.

### 2022-09-05 chapters/introduction.tex
Finish document

@@
 Under a functional framework, the objects of interest are \textit{random functions} instead of random points in a finite-dimensional space.
-While in principle the functional data could be simply regarded as a discretized vector in a very high dimension (an indeed such a discretization is performed in practice), there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
-Thus, the general idea is to assume the existence of an underlying sufficiently smooth function that generates each (possibly noisy) functional observation, even though we only record it on a finite grid of points.
+While in principle the functional data could be simply regarded as a discretized vector in a very high dimension (and indeed such a discretization is performed in practice), there are often many advantages in taking into account the functional nature of the data, ranging from modeling the possibly high correlation among points that are close in the domain, to extracting information that may be hidden in the derivatives of the function in question.
+Thus, the general idea is to assume the existence of an underlying sufficiently smooth function that corresponds to each (possibly noisy) functional observation, even though we only record it on a finite grid of points.
 
@@
 In this case, after differentiating twice a clearer pattern emerges (see Figure~\ref{fig:tecator_derivatives}), one from which inference and prediction will surely be easier.
+
+\enlargethispage{1\baselineskip}
 
@@
 \begin{center}
+  \vspace{.25em}
 \color{teal}\FourStar
@@
 We will further suppose that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(x_i, y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\).
+In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\).
 Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).

### 2022-09-05 chapters/model-choice.tex
Finish document

--- 2022-09-02 chapters/model-choice.tex
@@
 
-In this chapter we gather together several remarks on the main choices made throughout the design of our model, as well as some examples of validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
+In this chapter we gather together several remarks on the main choices made during the design and implementation of our model, as well as some examples of validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.
 
@@
 The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein.
-A common approach that seems to work well is to simply impose an ordering in the offending parameters, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\).
+A common approach that seems to work well is to simply impose an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with \(\tau\).
 We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).
@@
 One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model.
-While theoretically we can impose a prior distribution on \(p\) as well (e.g.
+While theoretically we could impose a prior distribution on \(p\) as well (e.g.
 a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications.
@@
   \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator.
-That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter (which can be selected for example by cross-validation strategies).
+That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter (which can be selected for instance by cross-validation strategies).
 This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
@@
 The MCMC method chosen for approximating the posterior distribution in our models is the affine-invariant ensemble sampler described in Section~\ref{sec:mcmc}.
-As mentioned there, we use the implementation in the Python library \textit{emcee}, which is both reliable and easy to use; it aims to be a general-purpose package that performs well in a wide class of problems.
+As mentioned there, we utilize the computational implementation in the Python library \textit{emcee}, which is both reliable and easy to use; it aims to be a general-purpose package that performs well in a wide class of problems.
 One advantage of this method, apart from the property of affine-invariance, is that it only requires us to specify a few hyperparameters, irrespective of the underlying dimension.
@@
 To reduce the effects of randomness, we run the algorithm a few times and retain the point with the highest likelihood, and to avoid biasing the sampler too much towards the specific point selected, we let a fraction of the initial points be random (within reasonable bounds).
+This approximation is also used to specify the hyperparameter \(b_0\).
 
@@
   \fbox{\includegraphics[width=.9\textwidth]{credible_intervals}}
-  \caption{Example of 94\% credible intervals on the posterior distribution of the impact points}\label{fig:credible_intervals}
+  \caption{Example of 94\% credible intervals on the posterior distribution of the impact points.}\label{fig:credible_intervals}
 \end{figure}

### 2022-09-05 chapters/abstract.tex
Replace cover and minor adjustments

@@
 
-In this thesis we propose a novel Bayesian approach for functional linear and logistic regression models, based on the theory of reproducing kernel Hilbert spaces (RKHS's).
+In this thesis we propose a novel Bayesian approach for functional linear and logistic regression models based on the theory of reproducing kernel Hilbert spaces (RKHS's).
 These recently-proposed models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.

### 2022-09-05 chapters/background-related-work.tex
Replace cover and minor adjustments

@@
 In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{berrendero2018functional}.
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][]{berrendero2018functional}.
 

### 2022-09-05 chapters/bayesian-functional-regression.tex
Replace cover and minor adjustments

@@
 In the case of functional logistic regression, we can regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given \(X=x \in L^2[0, 1]\), and as usual suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model might be given, in terms of the isometric transformation \(X \mapsto \dotprod{X}{\alpha}_K = \Psi_X(\alpha)\), by the equation
+Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model might be given, in terms of the correspondence \(\dotprod{X}{\alpha}_K = \Psi_X(\alpha)\), by the equation
 \begin{equation}\label{eq:rkhs-model-logistic}

### 2022-09-05 chapters/experiments.tex
Minor addition

@@
 
-A lot of minor things were tried, such as standardizing the regressors and/or the responses, smoothing the functional data, or transforming the values of \(\tau\) during sampling so that the domain was unconstrained.
+A lot of minor things were tried, such as standardizing the regressors and/or the responses, smoothing the functional data, changing the proposal of moves in the MCMC sampler, or transforming the values of \(\tau\) during sampling so that the domain was unconstrained.
 It is likely that some of these strategies would lead to better results in specific cases, but overall they did not improve the general performance of the prediction procedure.

### 2022-09-17 slides/defense.tex
Modify slides

%
% Copyright (c) 2022 Antonio Coín Castro
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%

\RequirePackage{fix-cm}
\documentclass[9pt, english, professionalfonts]{beamer}

% BEAMER OPTIONS

\definecolor{Maroon}{cmyk}{0, 0.87, 0.88, 0.1}
\definecolor{teal}{rgb}{0.0, 0.45, 0.45}

\usetheme[block=fill, subsectionpage=progressbar, titleformat section=smallcaps]{metropolis}
\setbeamertemplate{frametitle continuation}[roman]
\setbeamerfont{section title}{size=\huge}
\setbeamertemplate{section in toc}[sections numbered]
%\setbeamertemplate{subsection in toc}[subsections unnumbered]
%\setsansfont[BviejoFont={Fira Sans SemiBold}]{Fira Sans Book}  % Increase font weigth
\widowpenalties 1 10000
\raggedbottom

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

% COLORS
\setbeamercolor{palette primary}{bg=teal}
\setbeamercolor{progress bar}{use=Maroon, fg=Maroon}

% PACKAGES

\usepackage{fontspec}
\usepackage[autostyle=true]{csquotes}
\usepackage[absolute,overlay]{textpos}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{subcaption}
\usepackage{epigraph}
\usepackage[export]{adjustbox}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{amssymb, amsmath, amsthm, amsfonts, amscd, dsfont}
\usepackage{listings}

\usepackage[backend=biber,style=numeric-comp,sorting=none]{biblatex}
\addbibresource{bibliography.bib}
\setbeamertemplate{bibliography item}[article]

\graphicspath{%
  {../figures/}
}

% FONTS
%\usefonttheme{professionalfonts}
%\usepackage{mathpazo}
%\usepackage{eulervm}

\definecolor{backg}{HTML}{F2F2F2} % Fondo
\definecolor{comments}{HTML}{a8a8a8} % Comentarios
\definecolor{keywords}{HTML}{08388c} % Palabras clave
\definecolor{strings}{HTML}{0489B1}  % Strings

% CUSTOM COMMANDS

\renewcommand{\baselinestretch}{1.3}
\definecolor{mLightBrown}{HTML}{f97e0b}
% \definecolor{mLightBrown}{HTML}{990000}  % Darker accent color
\newcommand\maroon[1]{\color{mLightBrown}#1\color{mDarkTeal}}

\renewcommand{\epsilon}{\varepsilon}

%% Math operators
\DeclareMathOperator*\argmin{arg\,min}
\DeclareMathOperator\supp{supp}
\DeclareMathOperator\Cov{Cov}
\DeclareMathOperator\Var{Var}

%% Math symbols
\newcommand{\N} {\ensuremath{\mathds{N}}}
\newcommand{\R} {\ensuremath{\mathds{R}}}
\newcommand{\E} {\ensuremath{\mathds{E}}}
\newcommand{\I} {\ensuremath{\mathds{I}}}
\renewcommand{\P} {\ensuremath{\mathds{P}}}

\newcommand{\D} {\ensuremath{\mathcal{D}}}
\newcommand{\X} {\ensuremath{\mathcal{X}}}
\newcommand{\K} {\ensuremath{\mathcal{K}}}
\newcommand{\T} {\ensuremath{\mathcal{T}}}
\newcommand{\U} {\ensuremath{\mathcal{U}}}
\newcommand{\Lcal} {\ensuremath{\mathcal{L}}}
\newcommand{\Hcal} {\ensuremath\mathcal{H}}

%% Scalar product
\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}

% TITLE

\title{\huge{Bayesian RKHS-based methods in functional regression}}
\providecommand{\subtitle}[1]{}
\subtitle{\Large{Master's degree in Data Science}}
\date{\normalsize{September 20th, 2022}\vspace{1em}\\\normalsize{Master's thesis}\vspace{2em}}
\author{\normalsize{Antonio Coín Castro}}
\institute{\small{Supervisors:\\José Ramón Berrendero Díaz\\ Antonio Cuevas González}}

\titlegraphic{
  \begin{textblock*}{2cm}(6.7cm, 7.5cm)
    \includegraphics[width=5cm]{uam_eps_logo}
  \end{textblock*}
}

% DOCUMENTO

\begin{document}
\maketitle

% \begin{frame}{Table of contents}
%    \tableofcontents
% \end{frame}

\section{Introduction}

\begin{frame}{What is FDA?}
\begin{center}
  \Large{\maroon{Functional Data Analysis}}
\end{center}
\vspace{1em}
\begin{figure}[ht!]
  \centering
  \includegraphics[width=.3\textwidth, valign=c]{example_points}
  \hspace{.4cm}
  \Huge{\(\longrightarrow\)}
  \hspace{.2cm}
  \includegraphics[width=.35\textwidth, valign=c]{example_functions}
  % Captions
  \large
  \makebox[.3\textwidth]{\(\R^p\)}
  \hspace{.4cm}
  \phantom{\(\longrightarrow\)}
  \hspace{.2cm}
  \makebox[.35\textwidth]{\(L^2[0,1]\)}
\end{figure}
\vspace{1em}
\normalsize
\begin{center}
\parbox{9cm}{\centering \textit{Statistical techniques to process, model and make inference on data varying over a continuum}}
\end{center}
\end{frame}

\begin{frame}{Functional linear regression}
  \textbf{First problem}: linear regression with functional data.
  \vspace{2em}

\begin{figure}
    \includegraphics[width=\textwidth]{data_lin}
  \end{figure}
\end{frame}

\begin{frame}{Functional logistic regression}
  \textbf{Second problem}: logistic regression with functional data.
\vspace{2em}

\begin{figure}
    \includegraphics[width=\textwidth]{data_log}
  \end{figure}
\end{frame}

\begin{frame}{The usual \(\bm{L^2}\)-models}
  \(\bullet\) \(X=X(t)\) second order stochastic process with trajectories in \(L^2[0, 1]\).

  \(\bullet\) \(Y\) real variable (continuous or dichotomous).

  \vspace{1em}

\begin{block}{Linear \(\bm{L^2}\)-model}
  \[
    Y = \alpha_0 + \maroon{\langle X, \beta \rangle_2} + \varepsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \varepsilon.
  \]
\end{block}

\begin{block}{Logistic \(\bm{L^2}\)-model}
    \[
    \mathbb P (Y=1\mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \int_0^1 X(t)\beta(t)\, dt\}}.
  \]
\end{block}
  \maroon{\(\beta \in L^2[0,1]\)}, \(\alpha_0\in\mathbb R\), \ \(\varepsilon \sim \mathcal N(0, \sigma^2), \ X \bot \varepsilon\).

  \footnotetext{\includegraphics[scale=.7]{beamericonbook} \ Ramsay, J. and Silverman, B. (1997). \textit{Functional Data Analysis}.}
\end{frame}

\begin{frame}{Shortcomings of the \(\bm{L^2}\)-models}
  \begin{itemize}
    \item Direct optimization is hopeless: \(L^2[0,1]\) is a \maroon{extremely wide space}.
    \item The usual least-squares theory cannot be applied directly; some regularization or dimensionality reduction techniques are required.
    \item They do not include simple models based on linear combinations of the marginal of the process as particular cases, such as
    \[
      Y = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \varepsilon.
    \]
    \item In the logistic case, the maximum likelihood estimator does not exist with probability one under fairly general conditions.
  \end{itemize}
\end{frame}

\section{A RKHS model for functional regression}

\begin{frame}{Alternative model: RKHS}

We will exploit the theory of \textit{reproducing kernel Hilbert spaces} (RKHS's).

\vspace{1em}
\begin{alertblock}{Proposal\footnote[2]{Based on the work of Berrendero et al. (2019) and Berrendero et al. (2022).}}
  Change the habitat of the functional parameter \(\beta(\cdot)\).
\end{alertblock}

\vspace{1em}
  We consider a functional parameter \maroon{\(\alpha \in \mathcal H(K)\)}, where \(\mathcal H(K)\) is the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\). Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\alpha}_K\)}.
\end{frame}

\begin{frame}{Historical review of RKHS theory}
  \begin{itemize}
    \item Initial development \(\longrightarrow\) 1950's (Mercer, Aronszajn, Bergman).
    \item Uses in statistics \(\longrightarrow\) 1960's, 1970's (Parzen, Wahba, \ldots).
    \item Uses in Machine Learning \(\longrightarrow\) 1990's, 2000's (representer theorems).
    \item Uses in FDA \(\longrightarrow\) 2010's onwards.
  \end{itemize}
\end{frame}

\begin{frame}{RKHS's in a nutshell}
  \begin{definition}
    Start with the preliminary space
    \[
    \Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N,  a_i \in \R,  t_i \in [0, 1] \right\},
    \]
    which is endowed with the scalar product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\).

    Then, \(\Hcal(K)\) is the completion of \(\Hcal_0(K)\) under this scalar product.
  \end{definition}

  \vspace{1em}

  Note that \(\Hcal(K)\subset L^2[0, 1]\) as subsets, but the elements of \(\Hcal(K)\) are simpler and more manageable in general. Plus, it is a space of \maroon{functions}, and not of equivalence classes.
\end{frame}

\begin{frame}{RKHS's in a nutshell (cont.)}
  This space has some interesting properties:

  \vspace{1em}

  \begin{block}{Proposition}
    \begin{enumerate}
      \item All the evaluation operators \(\delta_t(f)=f(t)\) are continuous \maroon{(characterization)}.
      \item Norm convergence implies pointwise convergence.
      \item If \(K\) is continuous, then so is every function \(f\in\mathcal H(K)\).
    \end{enumerate}
  \end{block}
\end{frame}

% \begin{frame}{RKHS's in a nutshell (cont.)}
%   There are other equivalent definitions of \(\Hcal(K)\), mainly based on the \textit{covariance operator}
% \[
% \mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
% \]
%
% \begin{definition}
% \begin{enumerate}
%   \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
%   \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\). Here, \(\lambda_j\to 0\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\).
% \end{enumerate}
% \end{definition}
%
% The second definition emphasizes the fact that functions in \(\Hcal(K)\) are ``simple'', in the sense that their components in an orthonormal basis need to vanish quickly.
%
% \end{frame}

\begin{frame}{A small subtlety}

  Generally speaking, the concrete realizations \(x=X(\omega)\) \maroon{do not belong to \(\Hcal(K)\) with probability \(1\)}, so the expression \(\dotprod{x}{\alpha}_K\) lacks meaning. However, we can make use of the relationship between \(\Hcal(K)\) and \(\mathcal L(X)\), the linear span of the process \(X\) within \(L^2(\Omega)\).

  \vspace{1em}
  \begin{block}{Loève's isometry}
    It is the continuous extension of the correspondence
    \[
\sum_{j=1}^p a_j X(t_j) \longleftrightarrow \sum_{j=1}^p a_j K(t_j, \cdot).
    \]
    Formally, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) for \(U \in \mathcal L(X)\).
  \end{block}
  With a slight abuse of notation, we will write \maroon{\(\dotprod{x}{\alpha}_K \equiv [\Psi_X(\alpha)](\omega) := \Psi_x(\alpha)\)}\ for \(x=X(\omega)\) and \(\alpha \in \Hcal(K)\).
\end{frame}

\begin{frame}{A new functional parameter\ldots}
  \begin{itemize}
    \item The initial idea was to suppose \(\alpha \in \Hcal(K)\).
    \item To further simplify the model, we will assume \(\alpha \in \Hcal_0(K)\), which is dense in \(\Hcal(K)\) by construction. That is,
    \[
      \alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot), \quad \beta_j \in \R, \ t_j \in [0, 1], \ p \in \N.
    \]
    \item When evaluating \(\dotprod{x}{\alpha}_K\) for a trajectory \(x=X(\omega)\), we resort to Loève's isometry:
    \[
      \dotprod{x}{\alpha}_K = \Psi_x(\alpha) = \sum_{j=1}^p \beta_j X(t_j).
    \]
  \end{itemize}

\end{frame}

\begin{frame}{\ldots and two new models}
  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), for \(\alpha \in \Hcal_0(K)\) we consider the following models:

  \vspace{1em}

  \begin{block}{Linear RKHS model}
  \[
    Y_i = \alpha_0 + \maroon{\langle X_i, \alpha \rangle_K} + \varepsilon = \alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j) + \varepsilon.
  \]
\end{block}
\vspace{1em}

\begin{block}{Logistic RKHS model}
    \[
    \mathbb P (Y_i=1| X_i) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X_i, \alpha\rangle_K}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\}}.
  \]
\end{block}

\end{frame}

\begin{frame}{Advantages of RKHS models}
  \begin{itemize}
    \item They are finite-dimensional approximations with a functional interpretation, which simplifies their understanding and handling.
    \item Under mild conditions, the general RKHS models include the \(L^2\)-models as particular cases.
    \item They mitigate the problem of non-existence of the MLE.
    \item They are particularly suited as a basis for variable selection methods.
  \end{itemize}
\end{frame}

\section{A Bayesian approach to parameter estimation}

\begin{frame}{Parameter estimation overview}

  We consider a parameter vector
  \[
  \theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (p, b, \tau, \alpha_0, \sigma^2),
  \]
  on which we impose a prior distribution \(\pi(\theta)\). The posterior distribution is obtained through Bayes' theorem, and it is approximated using Markov chain Monte Carlo (MCMC) methods.

  \vspace{1em}

      \begin{block}{Bayes's formula}
      \[
      \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
      \]
  \end{block}

  \vspace{1em}

  In practice we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.

\end{frame}

\begin{frame}{The value of \(\bm p\)}
    For practical and computational reasons, \maroon{the value of \(p\) is fixed beforehand} \ (equivalent to a degenerate prior).

    \begin{itemize}
      \item A variable-sized parameter vector causes unwanted problems in the implementation.
      \item In practice, the value of \(p\) will be small.
      \item The model is not that sensitive to the choice of \(p\).
      \item The parameter \(p\) can be chosen in several meaningful ways.
    \end{itemize}
\end{frame}

\begin{frame}{Prior distributions}
  Same prior scenario for both linear and logistic models.
  \vspace{.5em}
  \begin{block}{Prior for the parameters involved}
    \vspace{-1em}
  \begin{align*}
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  b\mid \tau, \sigma^2 & \sim \mathcal N_p\big(b_0, g\sigma^2{\underbrace{\left[\mathcal X_\tau' \mathcal X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
\end{align*}

\vspace{-1em}
where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\) is the data matrix, and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters.
\end{block}

\metroset{block=transparent}

\vspace{1em}
\begin{alertblock}{Hyperparameter selection}
  \vspace{0.25em}
  For \(b_0\) we use a computational approximation of the MLE; the rest of hyperparameters are set via \textit{cross-validation}.
\end{alertblock}
\metroset{block=fill}
\end{frame}

\begin{frame}{Posterior distributions}
  \begin{block}{Linear log-posterior}
    \vspace{-1em}
    \begin{align*}
  \log \pi(\theta\mid \mathcal D_n) & \propto \frac{1}{2}\log |G_\tau| - (p+n+2)\log \sigma\\
  &-\frac{1}{2\sigma^2} \left(\|\bm{Y}-\alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right),
\end{align*}
where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
  \end{block}
  \vspace{1em}
  \begin{block}{Logistic log-posterior}
      \vspace{-1em}
    \begin{align*}
  \log \pi(\theta \mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \dotprod{X_i}{\alpha}_K\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \dotprod{X_i}{\alpha}_K\right\}\right)\right]\\
  \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
\end{align*}
Remember that \(\dotprod{X_i}{\alpha}_K = \sum_j \beta_j X_i(t_j)\).
  \end{block}
\end{frame}

\begin{frame}{Posterior approximation}
 MCMC algorithms build a Markov chain with our posterior as its stationary distribution. We iteratively get a set of approximate samples \(\{\theta^{(m)*}: m=1,\dots, M\}\) of the posterior distribution.

  \begin{figure}
    \centering
    \includegraphics[width=.85\textwidth]{posterior}
  \end{figure}

Many alternatives: Metropolis-Hastings, \maroon{emcee}, Hamiltonian Monte Carlo, \ldots.
\end{frame}

\begin{frame}{Label switching}
The model suffers from non-identifiability of the components caused by their interchangeability: \(\pi(Y|X,\theta)=\pi(Y|X, \nu(\theta))\) for any permutation \(\nu\) that rearranges the indices \(j=1,\dots, p\):
    \[
    (\beta_1, \beta_2, \tau_1, \tau_2, \alpha_0, \sigma^2) \leftrightsquigarrow (\beta_2, \beta_1, \tau_2, \tau_1, \alpha_0, \sigma^2).
  \]

Hence, the components may be inadvertently exchanged from one iteration to the next in any MCMC algorithm.

\metroset{block=transparent}
\vspace{1em}
\begin{alertblock}{Partial solution}
  \vspace{0.25em}
  Post-process the trace and impose an artificial identifiability constraint on the parameters, such as \(\beta_i < \beta_j\) if \(i<j\) (or the analogous with \(\tau\)).
\end{alertblock}
\metroset{block=fill}
\end{frame}

\begin{frame}{Predictions}
  \textbf{Summarize-then-predict}. A summary of the posterior distribution \(\theta| \mathcal D_n\) is obtained via a point-estimate statistic (mean, median, mode), and then predictions are made according to the assumed model, e.g.:
  \[
  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j X_i(\hat t_j), \quad i=1,\dots, n.
  \]

  \textbf{Predict-then-summarize}. We first construct predictions on each step of the MCMC algorithm following our model, i.e.:
  \[
    Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}, \quad m=1,\dots,M, \quad i=1,\dots, n.
  \]
  Then, the mean of all such intermediate predictions is used as a proxy for the response variable:
  \[
    \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots, n.
  \]
\end{frame}

\begin{frame}{Predictions (cont.)}
    We are essentially approximating the \textit{posterior predictive distribution}
\[
\pi(y\mid x, \mathcal D_n) = \int_{\Theta} \pi(y\mid x, \theta)\pi(\theta\mid \mathcal D_n)\, d\theta.
\]
  \begin{figure}
    \includegraphics[width=.675\textwidth]{ppc_linear}
  \end{figure}
\end{frame}

\begin{frame}{Predictions (cont.)}
    \textbf{Variable selection}. We select \(p\) variables using only a point estimator of \(\tau|\mathcal D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
    \vspace{1em}

    \begin{figure}
      \includegraphics[width=\textwidth]{credible_intervals}
      \caption{Approximate posterior distribution for the impact points.}
    \end{figure}

    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
\end{frame}

\section{Experiments}

\begin{frame}{Methodology}
  \begin{itemize}
    \item Experiments on simulated data, with \(250\) samples of different Gaussian and non-Gaussian processes on a regular grid of \(100\) points. Also experiments on real data sets.
    \item We perform 10 random train/test partitions (60\%/40\%) to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
    \item The hyperparameters of our models are \(p\in\{1,2,\dots,10\}\), \(g=5\) and \(\eta \in \{10^{-4}, 10^{-3}, \dots, 10^2\}\).
    \item Several regression methods are considered for comparison, along with variable selection techniques, functional or otherwise. The evaluation metrics are the RMSE and the accuracy.
  \end{itemize}
\end{frame}

\begin{frame}{Linear regression: simulated GP-RKHS data}
  \[
  Y=5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon, \quad \epsilon \sim \mathcal N(0, 0.5).
  \]
  \begin{figure}
    \includegraphics[width=.95\textwidth]{reg_emcee_rkhs}
  \end{figure}
\end{frame}

\begin{frame}{Linear regression: simulated GP-\(\bm{L^2}\) data}
  \[
  Y = 5 + \int_0^1 \log(1+4t)X(t)\, dt + \epsilon, \quad \epsilon \sim \mathcal N(0, 0.5).
  \]
  \begin{figure}
    \includegraphics[width=.95\textwidth]{reg_emcee_l2}
  \end{figure}
\end{frame}

\begin{frame}{Linear regression: simulated non-GP data}
  \[
  X(t) = e^{BM(t)}, \quad BM \equiv \text{standard Brownian motion}.
  \]
  \begin{figure}
    \includegraphics[width=.6\textwidth]{reg_emcee_nongp}
  \end{figure}
\end{frame}

\begin{frame}{Linear regression: real data}
  \begin{center}
    Three real data sets.
  \end{center}
    \vspace{.5em}
  \begin{figure}
    \includegraphics[width=.8\textwidth]{reg_emcee_real}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: simulated GP-RKHS data}
\[
  \P(Y=1\mid X) = \frac{1}{1 + \exp\left\{0.5 +5X(0.1) - X(0.4) - 10X(0.8)\right\}}.
\]
  \begin{figure}
    \includegraphics[width=.95\textwidth]{clf_emcee_rkhs}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: simulated GP-\(\bm{L^2}\) data}
  \[
  \P(Y=1\mid X) = \frac{1}{\displaystyle 1 + \exp\left\{0.5 - \int_0^1 \log(1+4t) X(t)\, dt\right\}}.
  \]
  \begin{figure}
    \includegraphics[width=.95\textwidth]{clf_emcee_l2}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: simulated non-GP data}
  \[
  X^{\text{hom}} \sim BM(0, 1) + BM(m(t), 1), \quad X^{\text{het}} \sim BM(0, 1) + BM(0, 2).
  \]
  \begin{figure}
    \includegraphics[width=.6\textwidth]{clf_emcee_nongp}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: real data}
  \begin{center}
    Three real data sets.
  \end{center}
    \vspace{.5em}
  \begin{figure}
    \includegraphics[width=.8\textwidth]{clf_emcee_real}
  \end{figure}
\end{frame}

\begin{frame}{Other experiments}
\begin{itemize}
  \item Recover the underlying data generation model when it is a finite-dimensional RKHS model.
  \item Incorporate \(p\) into the Bayesian scheme with a categorical distribution.
  \item Change some prior distributions, such as Cauchy for \(\sigma^2\) or Beta for the times \(\{t_j\}\).
  \item Try different ways to address the label switching problem.
\end{itemize}
\end{frame}

\section{Conclusions}

\begin{frame}{Conclusions and future work}

     \metroset{block=transparent}
 \begin{alertblock}{Our proposal}
   \begin{itemize}
     \item Simple and interpretable finite-dimensional approximation to functional regression.
     \item Computationally feasible Bayesian approach to parameter estimation within the proposed model.
     \item Python implementation competitive with usual alternatives in both simulated and real data sets.
   \end{itemize}
 \end{alertblock}

\pause

 \begin{exampleblock}{The road ahead}
   \begin{itemize}
     \item Experiment with different prior distributions.
     \item Derive theoretical properties of the predictors.
     \item Try different MCMC algorithms (e.g. reversible-jump MCMC).
   \end{itemize}
 \end{exampleblock}

\end{frame}

\begin{frame}
    \frametitle{References}
    \nocite{*}
    \printbibliography[heading=none]
\end{frame}

\appendix

\begin{frame}[standout]
  Thank you for your attention
\end{frame}

%%
% BACKUP SLIDES
%%

\begin{frame}{Dependence on \(\bm p\)}
Mean accuracy for our logistic RKHS methods as a function of \(p\), using \(\eta=0.01\) and the homoscedastic mixture data set.

\vspace{1em}

\begin{figure}
  \centering
  \includegraphics[width=.95\textwidth]{mixture_dependence_acc_p}
  \caption*{}
\end{figure}
\end{frame}

\begin{frame}{Recovering the underlying model}
  \vspace{1em}
  Underlying finite-dimensional RKHS data generation model with 2 components.

  \vspace{.5em}
\begin{figure}[b]

  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=.95\textwidth]{beta_3}
    \caption{RKHS model with \(p=3\).}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=.9\textwidth]{beta_4}
    \caption{RKHS model with \(p=4\).}
  \end{subfigure}
\end{figure}
\end{frame}

\begin{frame}{Metropolis-Hastings}
  Approximate a distribution \(\pi(x)\) such that \(\pi(x)\propto f(x)\):
  \begin{enumerate}
  \item Generate a proposal \(x' \sim g(x'|x_t)\).
  \item Accept the proposal with probability \(\alpha=\min\{1, f(x')/f(x_t)\}\). If accepted, set \(x_{t+1}=x'\); otherwise set \(x_{t+1}=x_t\).
\end{enumerate}

\vspace{1em}

 \begin{figure}
   \centering
   \includegraphics[width=.6\textwidth]{mh}
 \end{figure}
\end{frame}

\begin{frame}{Affine-invariant ensemble sampler (emcee)}
Consider an \maroon{ensemble} \ of walkers \(\Lambda(t) = (\Lambda_1(t), \dots, \Lambda_L(t))\) and the complementary ensemble
\[
  \Lambda_{-l}(t) = \{\Lambda_1(t+1), \dots, \Lambda_{l-1}(t+1), \Lambda_{l+1}(t), \dots, \Lambda_L(t)\}, \quad l=1,\dots, L.
\]
\vspace{-1em}
\begin{enumerate}
  \item For each walker \(1\leq l \leq L\) another walker \(\Lambda_j \in \Lambda_{-l}(t)\) is chosen at random. The proposal is constructed as
  \[
    \Lambda_l(t) \to \Gamma = \Lambda_j + Z(\Lambda_l(t) - \Lambda_j),
\]
where \(Z \stackrel{i.i.d.}{\sim} g(z)\) satisfying \(g(z^{-1})=zg(z)\).
\item  The proposal is accepted with probability
  \[
    \alpha = \min\left\{1, \ Z^{p-1}\frac{\pi(\Gamma)}{\pi(\Lambda_l(t))}\right\}.
  \]
\end{enumerate}
\end{frame}

\end{document}

### 2022-09-17 chapters/abstract.tex
Minor changes

--- 2022-09-01 chapters/abstract.tex
@@
 In this thesis we propose a novel Bayesian approach for functional linear and logistic regression models, based on the theory of reproducing kernel Hilbert spaces (RKHS's).
-These recently-proposed models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
-The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest, which has a tractable parametric form.
+These models build upon the RKHS associated with the covariance function of the underlying stochastic process, and can be viewed as a finite-dimensional approximation to the classical functional regression paradigm.
+The corresponding functional model (or the functional logistic equation in the case of binary response) is determined by a function living on a dense subspace of the RKHS of interest, which has a tractable parametric form based on linear combinations of the kernel.
 By imposing a suitable prior distribution on this space, we can perform data-driven inference via standard Bayes methodology.
@@
 We show through a comprehensive set of experiments that these methods are competitive against other usual alternatives in terms of predictive performance, both in simulated examples and real data sets.
-Overall, our proposed model is simple with regard to interpretation and feasible with regard to implementation, while enjoying the added flexibility of an ambient Bayesian framework.\\
+Overall, our proposed model is simple with regard to interpretation and feasible with regard to implementation, while also enjoying the added flexibility of an ambient Bayesian framework.\\
 

### 2022-09-17 chapters/background-related-work.tex
Minor changes

@@
 \[
-\argmin_{(\alpha_0, b)\in\R\times\R^p} \ \left\{ \sum_{i=1}^n \left( y_i - \alpha_0 - \int_0^1 x_i(t)[\phi(t)'b]\, dt\right)^2 + \lambda \Omega(b) \right\},
+\argmin_{(\alpha_0, b)\in\R\times\R^p} \ \ \ \left\{ \sum_{i=1}^n \left( y_i - \alpha_0 - \int_0^1 x_i(t)[\phi(t)'b]\, dt\right)^2 + \lambda \Omega(b) \right\},
 \]
@@
 
-In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
-It turns out that in both linear and logistic scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
+In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals of the process, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
+It turns out that in both linear and logistic scenarios a natural and more general alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
@@
 Although in this definition there is no explicit mention of the kernel, it can be recovered through the Riesz representation theorem.
-We summarize below some interesting topological properties of \(\Hcal(K)\).
+We summarize below some interesting properties of \(\Hcal(K)\).
 
@@
   \]
-  as we have just shown in \textit{(i)} that the evaluation operator is bounded.
+  as we have shown in \textit{(i)} that the evaluation operator is bounded.
 Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform.
@@
   \[
-    \Hcal(K_{\text{bm}}) = \{f: f \text{ is absolutely continuous, } f(0)=0 \text{ and } f'\in L^2[0, 1]\},
+    \Hcal(K_{\text{bm}}) = \left\{f: f \text{ is absolutely continuous, } f(0)=0 \text{ and } f'\in L^2[0, 1]\right\},
   \]
-  with inner product \(\dotprod{f}{g}_{K_{\text{bm}}} = \int f'g'\).
+  with inner product \(\dotprod{f}{g}_{K_{\text{bm}}} = \text{\large\(\int\)}f'g'\).
 \end{example}
@@
   \[
-    f(s) = \int_{-\infty}^\infty f(t)\delta(t, s)\, dt \quad \forall s \in \R.
+    f(s) = \int_{-\infty}^\infty \delta(t, s) f(t)\, dt, \quad s \in \R.
   \]
@@
  Note that, as expected, the covariance function of \(X\) plays a crucial role in characterizing the associated RKHS.
-An integral operator closely related to this covariance function is the so-called covariance operator, namely
+An integral operator closely related to this covariance function is the so-called covariance operator, namely \(\K: L^2[0,1] \to L^2[0,1]\) given by
 \[
-\K f(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
+\K f(\cdot) = \int_0^1 K(t, \cdot)f(t)\, dt, \quad f \in L^2[0, 1],
 \]
 which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}, so for the remaining of this work we will indeed suppose that \(K\) is continuous.
-As it turns out, the covariance function admits a spectral decomposition in terms of the eigenvalues and eigenfunctions of this integral operator, a sort of continuous generalization of the eigendecomposition of a symmetric positive semidefinite matrix.
+As it turns out, the covariance function admits a spectral decomposition in terms of the eigenvalues and eigenfunctions of this operator, a sort of continuous generalization of the eigendecomposition of a symmetric positive semidefinite matrix.
 
@@
     Let \(K\) be a continuous kernel on \([0,1]^2\).
-Then, there exists an orthonormal system \(\{\phi_j\}\) in \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive and form a non-increasing sequence \(\{\lambda_j\}\to 0\), and such that
+Then, there exists an orthonormal basis \(\{\phi_j\}\) of \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive\footnote{Even though the spectral theorem only guarantees that the eigenvalues of \(\K\) are non-negative, there is no loss of generality in assuming \(\lambda_j> 0\) for all \(j\geq 1\); see Remark 3 in \citet[Ch.~3]{cucker2001mathematical}.} and form a non-increasing sequence \(\{\lambda_j\}\to 0\), and such that
     \[
@@
 Furthermore, a similar orthogonal decomposition holds for the stochastic process \(X\), an expansion that lies at the base of many \(L^2\) functional regression models.
-The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Sec.~3.2]{berlinet2004reproducing}.
 
@@
 
-\enlargethispage{1\baselineskip}
-
-In addition to these results, it is worth mentioning that the covariance operator \(\K\) provides several alternative definitions of \(\Hcal(K)\).
+In addition to these theorems, it is worth mentioning that the covariance operator \(\K\) provides several alternative definitions of \(\Hcal(K)\).
 For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\K^{-1/2}(f)}{\K^{-1/2}(g)}\).
-Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can also be seen as the set \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
+Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can also be seen as the set \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\),
+with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\).
 Note that, since \(\{\lambda_j\}\) tends to zero, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, not only in that they inherit the regularity of the kernel \(K\), but also in the sense that their components in an orthonormal basis, namely \(\dotprod{f}{\phi_j}\), need to vanish quickly.
+A proof of the previous results can be consulted for example in \citet[][Sec.~3.2]{berlinet2004reproducing}.
 
@@
 Nevertheless, we will now hopefully clarify that this space can indeed be seen as \textit{the} Hilbert space inherently associated with the process \(X\) itself.
-Let us shift the perspective momentarily and consider the space \(L^2(\Omega)\) of all zero-mean random variables with finite second moment, endowed with the usual norm \(\|U\|^2=\E[U]^2\).
+Let us shift the perspective momentarily and consider the space \(L^2(\Omega)\) of all zero-mean random variables with finite second moment, endowed with the usual norm \(\|U\|^2=\E[U^2]\).
 Then, the linear span of \(X\) in \(L^2(\Omega)\) is given by
+
+\setlength{\headsep}{11pt}
+
 \[
@@
 As it happens, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and \(\Lcal(X)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}.
-This isometry is essentially the completion of the correspondence
+This isometry is essentially the continuous extension of the correspondence
   \begin{equation}\label{eq:loeves-isometry}
@@
   \end{equation}
-and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \Lcal(X)\).
+and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for all \(U \in \Lcal(X)\).
 Thus, \(\Hcal(K)\) can be regarded as an isometric copy of \(\Lcal(X)\), an approach that is often useful in statistics.
@@
 However, despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process.
-It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}.
+It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \parencites[e.g.][Cor.~7.1]{lukic2001stochastic}[][Th.~11]{pillai2007characterizing}.
 For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the associated RKHS are differentiable almost everywhere.
@@
 
+\setlength{\headsep}{21.75pt}
+
 \begin{proposition}
-  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations are satisfied:
+  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations hold:
 \begin{enumerate}
@@
 
-\vspace{1em}
+\vspace{.1em}
 
-\begin{figure}[ht!]
+\begin{figure}[htbp!]
   \centering
-  \includegraphics[width=.25\textwidth]{cube_sphere}
+  \includegraphics[width=.2\textwidth]{cube_sphere}
   \caption{To cover the volume of \(f(x)\) in higher dimensions we intuitively need a volume that grows exponentially with the dimension itself.}\label{fig:curse_dimensionality}
@@
    \centering
-   \includegraphics[width=.55\textwidth]{mh}
-   \caption{Representation of the Metropolis-Hastings algorithm (taken from \citet{jaewook2015metamodel}). The discarded proposals are represented by dashed circles.}\label{fig:mh}
+   \includegraphics[width=.5\textwidth]{mh}
+   \caption{Representation of the Metropolis-Hastings algorithm (taken from \citet{jaewook2015metamodel}). The discarded proposals are depicted as dashed circles.}\label{fig:mh}
  \end{figure}
@@
 
+  \setlength{\footskip}{60pt}
+
   \paragraph{Walk move.} For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is chosen, and the proposed move is
@@
 where \(W\) is a normal distribution with mean \(0\) and the same covariance as the sample covariance of all walkers in \(S_l\).
-The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).\\
+The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).
+
+\vspace*{.75em}
 
@@
 The idea is to divide the ensemble \(\Lambda\) into two equally-sized subsets \(\Lambda^{(0)}\) and \(\Lambda^{(1)}\), and then proceed on each iteration in the following alternate fashion:
+
+\enlargethispage{2\baselineskip}
 

### 2022-09-17 chapters/bayesian-functional-regression.tex
Minor changes

@@
 \let\epsilon\varepsilon
+\setlength{\footskip}{50.75pt}
 
@@
 Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(x=X(\omega)\) and \(\Psi_x\) is Loève's isometry.
-Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we can write \(\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j)\) when \(\alpha\) is as in~\eqref{eq:alpha_h0}.
+Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), when \(\alpha\) is as in~\eqref{eq:alpha_h0} we can write \(\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j)\).
 
@@
 The general idea will be to impose a prior distribution on the functional parameter to eventually derive a posterior model after incorporating the available sample information.
+
 In view of~\eqref{eq:alpha_h0}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_0(K)\)) it suffices to first consider a discrete distribution on \(p\), and then impose \(p\)-dimensional continuous prior distributions on the coefficients \(\beta_j\) and the times \(t_j\) given \(p\).
@@
 
-However, after some initial experimentation we found that, for practical and computational reasons, the value of \(p\), the dimensionality of the model, is best fixed beforehand in a suitable way (see Chapter~\ref{ch:model-choice} for details).
+However, after some initial experimentation we found that, for practical and computational reasons, the value of \(p\) (the dimensionality of the model) is best fixed beforehand in a suitable way; see Chapter~\ref{ch:model-choice} for details.
 Thus, we will regard only the \(\beta_j\) and \(t_j\) as free parameters, and search for our functional parameter in the space
@@
 
-In the case of functional logistic regression, we can regard the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given \(X=x \in L^2[0, 1]\), and as usual suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
+In the case of functional logistic regression, we regard the response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given \(X=x \in L^2[0, 1]\), and as usual suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
 Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model might be given, in terms of the correspondence \(\dotprod{X}{\alpha}_K = \Psi_X(\alpha)\), by the equation

### 2022-09-17 chapters/conclusions.tex
Minor changes

@@
 In this work we have addressed some inference and prediction issues involved in the task of learning from functional data.
-In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios.
-Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations but still with a functional perspective.
+In particular, we have tackled functional linear and logistic regression problems, presenting an alternative to the \(L^2\)-models predominant in these scenarios.
+Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations to the regression function, but still with a functional perspective.
 This leads to a conceptually simpler approach that facilitates the interpretation and subsequent analysis, which we believe is a compelling feature for most practitioners.
@@
 For example, consistency and/or robustness results regarding the posterior distribution would be an excellent complement to the practical side of this work.
-  \item To find other prior distributions for our parameters that perform better in general, or to eliminate the need of the tuning hyperparameter \(\eta\).
+  \item To find other prior distributions for our parameters that perform better in general, or to eliminate the need of the hyperparameters \(b_0\) or \(\eta\).
   \item To experiment with other MCMC algorithms for posterior approximation.

### 2022-09-17 chapters/experiments.tex
Minor changes

@@
 \begin{itemize}
-  \item A RKHS model where the response is generated as \(Y= 0.5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
+  \item A RKHS model where the response is generated as \(Y=5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
   \item A \(L^2\)-model with underlying coefficient function \(\beta(t)=\log(1+4t)\), again for the same four GP's.
@@
 
-Take for example a RKHS data set with two components generated according to the formula \(Y=-5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\).
+Take for example a RKHS data set with two components generated according to the formula \(Y=5 -5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\).
 Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(\tau_1, \tau_2, \tau_3)\) for a model with 3 components.

### 2022-09-19 chapters/appendix-code.tex
Small changes after revision

@@
 
-The Python source code developed for this work is available at the GitHub repository \url{https://github.com/antcc/rk-bfr}.
+The Python source code developed for this work is available at the GitHub repository \url{https://github.com/antcc/rk-bfr}, under a GPLv3 license.
 The code is adequately documented and is structured in several directories as follows:
@@
   \item The \texttt{reference\_methods} folder contains our implementation of the functional comparison algorithms that were not available through a standard Python library.
+  \item The \texttt{utils} folder contains some utility files for simulation, experimentation and visualization.
   \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g.
-number of iterations, type of data set, \ldots), along with some utility files.
-In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our models without a cross-validation loop.
+number of iterations, type of data set, \ldots).
+In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our Bayesian methods without a cross-validation loop.
 \end{itemize}
@@
 When possible, the code was implemented in a generic way that would allow for easy extensions or derivations.
-It was also developed with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries.
+It was also developed with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy}\footnote{\url{https://numpy.org}} and \textit{scipy} libraries.
 Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them.
@@
 We provide a script file \texttt{launch.sh} that illustrates a typical execution.
+
 Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way.

### 2022-09-19 chapters/background-related-work.tex
Small changes after revision

@@
 At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework.
-Examples of this endeavor include the definition of centrality measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others.
+Examples of this endeavor include the definition of centrality measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep[e.g.][]{berrendero2020mahalanobis, galeano2015mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others.
 Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.
@@
 In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals of the process, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this.
-It turns out that in both linear and logistic scenarios a natural and more general alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space.
+It turns out that in both linear and logistic scenarios a natural and more general alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space.
 As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions.
@@
 It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}.
-Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model.
+Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model in question.
 
@@
 At this point, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\), which informally amounts to adding all the limits of Cauchy sequences in \(\Hcal_0(K)\), turning it into a genuine Hilbert space with the inner product extended accordingly.
-Then it is immediate to see that the reproducing property is retained.
+Then, it is immediate to see that the reproducing property is retained.
 An important consequence is that \(\Hcal(K)\) is a space of actual functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.
 
-Indeed, another way of defining a RKHS is via the continuity of all the \textit{evaluation operators}, i.e., \(\delta_t(f) := f(t)\) for \(f\in\Hcal(K)\) and \(t\in[0,1]\).
+Indeed, another way of characterizing a RKHS is via the continuity of all the \textit{evaluation operators}, i.e., \(\delta_t(f) := f(t)\) for \(f\in\Hcal(K)\) and \(t\in[0,1]\).
 In this case, the interpretation is that if two functions are close in the RKHS norm, they are also pointwise close.
@@
 \end{proposition}
-Note that the statement \textit{(iii)} above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.
+Note that statement \textit{(iii)} above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.
 
@@
 Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning.
-For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property.
+For example, the well-known \textit{kernel trick} can be seen as a specific usage of the reproducing property.
 First, a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS.
@@
 Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high.
-If we know the conditional distribution of each variable on the rest of them, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension.
+If we know the distribution of each variable conditional on the rest of them, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension.
 Specifically, in each step we get unidimensional samples \textit{in turn} (e.g.
@@
  \]
- for all \(A,b\) and \(\lambda\), and for almost all \(\xi(t)\).
+ for all \(A,b\) and \(\lambda\), and almost all \(\xi(t)\).
 This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Lambda(t)+b\) for all \(t\).
@@
 
-  \paragraph{Walk move.} For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is chosen, and the proposed move is
+  \paragraph{Walk move.} For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is selected, and the proposed move is
 \[

### 2022-09-19 chapters/bayesian-functional-regression.tex
Small changes after revision

@@
 
-At this point we propose to follow a Bayesian approach to estimate the parameters of the model, which we believe is in line with the idea of simplicity we pursue, and also introduces an additional layer of flexibility in the model.
-In this way, we can include problem-specific information in the model through the use of prior distributions, and on top of that, this method works almost unaltered for both linear and logistic regression models.
-The general idea will be to impose a prior distribution on the functional parameter to eventually derive a posterior model after incorporating the available sample information.
+At this point we propose to follow a Bayesian approach to estimate the parameters of the model, which we believe is in line with the idea of simplicity we pursue, and also introduces an additional layer of flexibility into the model.
+In this way, we can include problem-specific information through the use of prior distributions, and on top of that, this method works almost unaltered for both linear and logistic regression models.
+The general idea will be to impose a prior distribution on the functional parameter to eventually derive a posterior distribution after incorporating the available sample information.
 
@@
 Moreover, note that starting from a probability distribution \(\P_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\P\) on \(\Hcal(K)\) merely by defining \(\P(B) = \P_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\).
-Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
+Consequently, our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).
 
@@
 Before proceeding further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\D_n\).
-For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
+Indeed, for \(\theta \in \Theta_p\) the reinterpreted model assumes the form
 \begin{equation}\label{eq:rkhs-model-linear-2}
@@
 In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(l^1\)-type loss function while the former implicitly optimizes an \(l^2\) loss.
-On the other hand, the mode is also a good candidate because it represents the point of highest probability density.
+On the other hand, the mode is also a good candidate because it represents the point of highest probability density, following a \textit{maximum a posteriori} (MAP) approach.
 At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
@@
 First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in this case it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model.
+Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in some cases it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model.
 Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.
@@
 However, in this case the nuisance parameter \(\sigma^2\) only appears as part of the hierarchical prior distribution, and not in the final model.
-The posterior distribution is then derived after a routine calculation.
+The posterior distribution is again derived after a routine calculation.
 
@@
 The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.
-On the other hand, the prediction method that takes into account the entire posterior approximation (that is, the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
+On the other hand, the prediction method that takes into account the entire posterior approximation (i.e.
+the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples.
 Hence, there are primarily two possible outcomes.
@@
   In this case, each \(Y_i^{(m)*}\) is a random variable that follows a Bernoulli distribution with parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
-Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) according to the majority vote of all the \(Y_i^{(m)*}\).\\
+Note that when \(\gamma=1/2\) this is equivalent to predicting \(\hat Y_i\) according to the majority vote of all the \(Y_i^{(m)*}\).\\
 
-Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.
+Lastly, the variable selection method is essentially the same as in the case of functional linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.

### 2022-09-19 chapters/conclusions.tex
Small changes after revision

@@
   \item To experiment with other MCMC algorithms for posterior approximation.
-Specifically, it would be very interesting to implement an efficient and reliable reversible-jump MCMC method in Python, which as we have already mentioned is a better fit for our particular Bayesian model.
+Specifically, it would be interesting to implement an efficient and reliable reversible-jump MCMC method in Python, which as we have already mentioned is a better fit for our particular Bayesian model.
 \end{itemize}

### 2022-09-19 chapters/experiments.tex
Small changes after revision

@@
 We use 64 chains and run them for 900 iterations in total, discarding the first 500 iterations as burn-in.
-Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains at each iteration, selecting the stretch move (the default) with probability 0.7 or the walk move with probability 0.3.
+Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains in each iteration, selecting the stretch move (the default) with probability 0.7 or the walk move with probability 0.3.
 
@@
 \begin{itemize}
-  \item A RKHS model where the response is generated as \(Y=5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
+  \item A finite-dimensional RKHS model where the response is generated as \(Y=5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
   \item A \(L^2\)-model with underlying coefficient function \(\beta(t)=\log(1+4t)\), again for the same four GP's.
@@
 Specifically, we use a geometric Brownian motion (GBM) defined as \(X(t)=e^{BM(t)}\), where \(BM(t)\) is a standard Brownian motion.
-In this case, we consider two data sets, one with a RKHS response and one with a \(L^2\) response, with the same parameters as above.
+In this case we consider two data sets, one with a RKHS response and one with a \(L^2\) response, with the same parameters as above.
 \end{itemize}
@@
 \begin{itemize}
-  \item Four logistic RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
+  \item Four logistic finite-dimensional RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
   \[
@@
 \end{itemize}
-In this case we use three real data sets well known in the literature.
-The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, trying to predict if their longevity is high or low.
+Additionally, we use three real data sets well known in the literature.
+The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, to predict if their longevity is high or low.
 The second one is the Berkeley Growth Study data set \citep{tuddenham1954physical}, which records the height of 54 girls and 39 boys over 31 different points in their lives.
-Finally, we have selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.
+Finally, we selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.
 
@@
   \item FLin.\hspace{.3em} Functional \(L^2\) linear regression model with fixed basis expansion.
+  \item FPCA.\hspace{.3em} Functional principal component analysis.
   \item FPLS1.\hspace{.3em} Functional PLS regression through basis expansion, implemented as in \citet{aguilera2010using}.
-  \item FPCA.\hspace{.3em} Functional principal component analysis.
 \end{itemize}
@@
 The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair.
-In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is one of \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,9,11\}\).
+In particular, regularization parameters are searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for cubic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is one of \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,9,11\}\).
 
-Most algorithms have been taken from the libraries \textit{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/}} and \textit{scikit-fda}\footnote{\url{https://fda.readthedocs.io/en/latest/}}, the first oriented to machine learning in general and the second to functional data analysis in particular.
+Most algorithms have been taken from the libraries \textit{scikit-learn}\footnote{\url{https://scikit-learn.org}} and \textit{scikit-fda}\footnote{\url{https://fda.readthedocs.io}}, the first oriented to machine learning in general and the second to functional data analysis in particular.
 However, some methods were not found in these packages and had to be implemented from scratch.
@@
 This is the most favorable case for us, as the underlying model coincides with our assumed model.
-Indeed, we can see that in most cases our algorithms are the ones with lower RMSE, save for a few exceptions, notably with the Gaussian kernel.
+Indeed, we can see that in most cases our algorithms are the ones with lower RMSE, save for a few exceptions, notably the Gaussian kernel.
 A subsequent analysis showed that this particular data set is especially sensitive to the value of the hyperparameter \(\eta\) in our prior distribution, so a more customized approach would be needed to obtain better results.
@@
 
-Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case.
+Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic Gaussian processes, but they struggle in the heteroscedastic case.
 Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e.
 with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}.
-Our results are in line with the empirical studies of said article, where the authors conclude that even though the asymptotic theoretical error is zero, most classification methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.
+Our results are in line with the empirical studies of this article, where the authors conclude that even though the asymptotic theoretical error is zero, most classification methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.
 
@@
 
-In this section we present some experiments that are not directly related to measuring the prediction performance of our models relative to other methods.
+In this section we present some experiments that are not directly related to measuring the predictive performance of our models relative to other methods.
 
@@
 
-One requirement that our model should satisfy is that it should be able to recover the true parameter function when the underlying data generation model is a finite-dimensional RKHS model.
+One requirement that our model should satisfy is that it ought to be able to recover the true parameter function when the underlying data generation model is a finite-dimensional RKHS model.
 This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?
@@
 Take for example a RKHS data set with two components generated according to the formula \(Y=5 -5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\).
-Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(\tau_1, \tau_2, \tau_3)\) for a model with 3 components.
+Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(t_1, t_2, t_3)\) for a model with 3 components.
 As we can see, one of the coefficients has gone to zero to account for the overspecification of the model, while the other two have stabilized very close to the true parameters.
@@
   \includegraphics[width=.8\textwidth]{beta_trace_3}
-  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
+  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 400 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
 \end{figure}
@@
   \includegraphics[width=.8\textwidth]{beta_trace_4}
-  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
+  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 400 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
 \end{figure}
@@
 We can take for example the homoscedastic mixture data set described earlier for the logistic regression problem, and fix the parameter \(\eta=0.01\).
-Then the corresponding mean accuracies (in 10 repetitions) for RKHS models with \(p=1,2,\dots 10\) components are shown in Figure~\ref{fig:dependence_acc_p}, along with their standard errors (which are arguably not very informative).
+The corresponding mean accuracies (in 10 repetitions) for RKHS models with \(p=1,2,\dots 10\) components are shown in Figure~\ref{fig:dependence_acc_p}, along with their standard errors (which are arguably not very informative).
 
@@
 The results were similar, but we ultimately discarded this approach because it introduced yet another hyperparameter into the model with no clear benefit.
-In addition, we tried to impose a beta distribution on each of the time instants \(\{t_j\}\), selecting the corresponding shape parameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable).
+In addition, we tried to impose a beta distribution on each of the time instants \(t_j\), selecting the corresponding shape parameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable).
 This approach does work, but in our estimation the computational overhead introduced by the beta density function (sometimes resulting in a 1.5x increase in execution time) is not worth the mild improvements, if any.

### 2022-09-19 chapters/introduction.tex
Small changes after revision

@@
 We will further suppose that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\).
-In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\).
+In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid on \([0, 1]\).
 Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\).

### 2022-09-19 chapters/model-choice.tex
Small changes after revision

@@
 The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein.
-A common approach that seems to work well is to simply impose an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with \(\tau\).
+A common approach that seems to work well is to simply enforce an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the times \(\tau\).
 We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).
@@
 In particular, we tested the pivot method proposed by \citet{marin2005bayesian}, in which all samples are aligned to minimize their distance to a reference element (the ``pivot''), which is chosen as the sample that maximizes the posterior density.
-However, the process of finding the appropriate permutation in each case was time-consuming, and the results were similar, if not worse, than the ones obtained with the simpler order constraints, so the latter were chosen as the default relabeling method of our algorithm (though the pivot method can still be enabled through a dedicated argument in the sampling procedure).
+However, the process of finding the appropriate permutation in each case was time-consuming, and the results were similar, if not worse, than the ones obtained with the simpler order constraints.
+For this reason, the latter were chosen as the default relabeling method of our algorithm, though the pivot method can still be enabled through a dedicated argument in the sampling procedure.
 
@@
 Another possibility would be to adapt a purely Bayesian model selection technique to our framework \citep[see][]{piironen2017comparison, gelman2013bayesian}, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models.
-These methods are usually based in computing a quantity known as the \textit{Bayes factor}, which in turn needs the normalizing integral constant we have been trying to avoid all along.
+These methods are usually based in computing a quantity known as the \textit{Bayes factor}, which in turn requires the specific value of the normalizing integral constant we have been trying to avoid all along.
 In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g.
 BIC, DIC, cross-validation, \ldots) to select its optimal value.
-As we will see shortly in Chapter~\ref{ch:experiments}, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios.
+As we will see shortly in Chapter~\ref{ch:experiments}, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios, and they suggest that our models are not very sensitive to small variations of this parameter.
+
+\enlargethispage{1\baselineskip}
 

### 2022-09-19 slides/defense.tex
Finish defense slides

@@
 \documentclass[9pt, english, professionalfonts]{beamer}
-
-% BEAMER OPTIONS
-
-\definecolor{Maroon}{cmyk}{0, 0.87, 0.88, 0.1}
-\definecolor{teal}{rgb}{0.0, 0.45, 0.45}
-
-\usetheme[block=fill, subsectionpage=progressbar, titleformat section=smallcaps]{metropolis}
-\setbeamertemplate{frametitle continuation}[roman]
-\setbeamerfont{section title}{size=\huge}
-\setbeamertemplate{section in toc}[sections numbered]
-%\setbeamertemplate{subsection in toc}[subsections unnumbered]
-%\setsansfont[BviejoFont={Fira Sans SemiBold}]{Fira Sans Book}  % Increase font weigth
-\widowpenalties 1 10000
-\raggedbottom
-
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
-
-% COLORS
-\setbeamercolor{palette primary}{bg=teal}
-\setbeamercolor{progress bar}{use=Maroon, fg=Maroon}
 
@@
 \usepackage{amssymb, amsmath, amsthm, amsfonts, amscd, dsfont}
-\usepackage{listings}
+
+% BEAMER OPTIONS
+
+\usetheme[block=fill, subsectionpage=progressbar, titleformat section=smallcaps]{metropolis}
+\setbeamertemplate{frametitle continuation}[roman]
+\setbeamerfont{section title}{size=\huge}
+\setbeamertemplate{section in toc}[sections numbered]
+%\setbeamertemplate{subsection in toc}[subsections unnumbered]
+\widowpenalties 1 10000
+\raggedbottom
+
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\renewcommand{\baselinestretch}{1.3}
+
+% COLORS
+
+\definecolor{Maroon}{cmyk}{0, 0.87, 0.88, 0.1}
+\definecolor{teal}{rgb}{0.0, 0.45, 0.45}
+\definecolor{mLightBrown}{HTML}{f97e0b}
+% \definecolor{mLightBrown}{HTML}{990000}  % Darker accent color
+% \definecolor{backg}{HTML}{F2F2F2} % Fondo
+% \definecolor{comments}{HTML}{a8a8a8} % Comentarios
+% \definecolor{keywords}{HTML}{08388c} % Palabras clave
+% \definecolor{strings}{HTML}{0489B1}  % Strings
+\setbeamercolor{palette primary}{bg=teal}
+\setbeamercolor{progress bar}{use=Maroon, fg=Maroon}
+
+% BIBLIOGRAPHY
 
@@
 
+% GRAPHICS
+
 \graphicspath{%
@@
 
-% FONTS
-%\usefonttheme{professionalfonts}
-%\usepackage{mathpazo}
-%\usepackage{eulervm}
-
-\definecolor{backg}{HTML}{F2F2F2} % Fondo
-\definecolor{comments}{HTML}{a8a8a8} % Comentarios
-\definecolor{keywords}{HTML}{08388c} % Palabras clave
-\definecolor{strings}{HTML}{0489B1}  % Strings
-
 % CUSTOM COMMANDS
 
-\renewcommand{\baselinestretch}{1.3}
-\definecolor{mLightBrown}{HTML}{f97e0b}
-% \definecolor{mLightBrown}{HTML}{990000}  % Darker accent color
 \newcommand\maroon[1]{\color{mLightBrown}#1\color{mDarkTeal}}
-
 \renewcommand{\epsilon}{\varepsilon}
@@
 
-% DOCUMENTO
+% DOCUMENT
 
@@
 \begin{frame}{Functional linear regression}
-  \textbf{First problem}: linear regression with functional data.
+  \maroon{First problem}: linear regression with functional data.
   \vspace{2em}
@@
 \begin{frame}{Functional logistic regression}
-  \textbf{Second problem}: logistic regression with functional data.
+  \maroon{Second problem}: logistic regression with functional data.
 \vspace{2em}
@@
     \]
-    \item In the logistic case, the maximum likelihood estimator does not exist with probability one under fairly general conditions.
+    \item In the logistic case, the maximum likelihood estimator does not exist with probability 1 under fairly general conditions.
   \end{itemize}
@@
 \vspace{1em}
-  We consider a functional parameter \maroon{\(\alpha \in \mathcal H(K)\)}, where \(\mathcal H(K)\) is the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\).
+  We consider a functional parameter \maroon{\(\alpha \in \Hcal(K)\)}, where \(\Hcal(K)\) is the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\).
 Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\alpha}_K\)}.
@@
       \item Norm convergence implies pointwise convergence.
-      \item If \(K\) is continuous, then so is every function \(f\in\mathcal H(K)\).
+      \item If \(K\) is continuous, then so is every function \(f\in\Hcal(K)\).
     \end{enumerate}
@@
 
-% \begin{frame}{RKHS's in a nutshell (cont.)}
-%   There are other equivalent definitions of \(\Hcal(K)\), mainly based on the \textit{covariance operator}
-% \[
-% \mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
-% \]
-%
-% \begin{definition}
-% \begin{enumerate}
-%   \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
-%   \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\).
-Here, \(\lambda_j\to 0\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\).
-% \end{enumerate}
-% \end{definition}
-%
-% The second definition emphasizes the fact that functions in \(\Hcal(K)\) are ``simple'', in the sense that their components in an orthonormal basis need to vanish quickly.
-%
-% \end{frame}
-
 \begin{frame}{A small subtlety}
 
-  Generally speaking, the concrete realizations \(x=X(\omega)\) \maroon{do not belong to \(\Hcal(K)\) with probability \(1\)}, so the expression \(\dotprod{x}{\alpha}_K\) lacks meaning.
-However, we can make use of the relationship between \(\Hcal(K)\) and \(\mathcal L(X)\), the linear span of the process \(X\) within \(L^2(\Omega)\).
+  Generally speaking, the concrete realizations \(x=X(\omega)\) \maroon{do not belong to \(\Hcal(K)\) with probability 1}, so the expression \(\dotprod{x}{\alpha}_K\) lacks meaning.
+However, we can make use of the relationship between \(\Hcal(K)\) and \(\Lcal(X)\), the linear span of the process \(X\) within \(L^2(\Omega)\).
 
@@
     \]
-    Formally, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) for \(U \in \mathcal L(X)\).
+    Formally, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) for \(U \in \Lcal(X)\).
   \end{block}
@@
 \begin{frame}{\ldots and two new models}
-  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), for \(\alpha \in \Hcal_0(K)\) we consider the following models:
+  Given a data set \(\D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), for \(\alpha \in \Hcal_0(K)\) we consider the following models:
 
@@
       \[
-      \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
+      \pi(\theta \mid \D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
       \]
@@
   \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
-  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
-  b\mid \tau, \sigma^2 & \sim \mathcal N_p\big(b_0, g\sigma^2{\underbrace{\left[\mathcal X_\tau' \mathcal X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
+  \tau                     & \sim \U([0, 1]^p),                                              \\
+  b\mid \tau, \sigma^2 & \sim \mathcal N_p\big(b_0, g\sigma^2{\underbrace{\left[\X_\tau' \X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
 \end{align*}
@@
 \vspace{-1em}
-where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\) is the data matrix, and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters.
+where \(I\) is the identity matrix, \(\X_\tau = (X_i(t_j))_{i,j}\) is the data matrix, and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters.
 \end{block}
@@
     \begin{align*}
-  \log \pi(\theta\mid \mathcal D_n) & \propto \frac{1}{2}\log |G_\tau| - (p+n+2)\log \sigma\\
-  &-\frac{1}{2\sigma^2} \left(\|\bm{Y}-\alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right),
+  \log \pi(\theta\mid \D_n) & \propto \frac{1}{2}\log |G_\tau| - (p+n+2)\log \sigma\\
+  &-\frac{1}{2\sigma^2} \left(\|\bm{Y}-\alpha_0\bm{1} - \X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right),
 \end{align*}
@@
     \begin{align*}
-  \log \pi(\theta \mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \dotprod{X_i}{\alpha}_K\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \dotprod{X_i}{\alpha}_K\right\}\right)\right]\\
+  \log \pi(\theta \mid \D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \dotprod{X_i}{\alpha}_K\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \dotprod{X_i}{\alpha}_K\right\}\right)\right]\\
   \quad &+ \frac{1}{2}\log |G_\tau| - (p+2)\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
@@
     \centering
-    \includegraphics[width=.85\textwidth]{posterior}
+    \includegraphics[width=.9\textwidth]{posterior}
   \end{figure}
@@
   \textbf{Summarize-then-predict}.
-A summary of the posterior distribution \(\theta| \mathcal D_n\) is obtained via a point-estimate statistic (mean, median, mode), and then predictions are made according to the assumed model, e.g.:
+A summary of the posterior distribution \(\theta| \D_n\) is obtained via a point-estimate statistic (mean, median, mode), and then predictions are made according to the assumed model, e.g.:
   \[
@@
 \[
-\pi(y\mid x, \mathcal D_n) = \int_{\Theta} \pi(y\mid x, \theta)\pi(\theta\mid \mathcal D_n)\, d\theta.
+\pi(y\mid x, \D_n) = \int_{\Theta} \pi(y\mid x, \theta)\pi(\theta\mid \D_n)\, d\theta.
 \]
@@
     \textbf{Variable selection}.
-We select \(p\) variables using only a point estimator of \(\tau|\mathcal D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
+We select \(p\) variables using only a point estimator of \(\tau|\D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
     \vspace{1em}
@@
       \includegraphics[width=\textwidth]{credible_intervals}
-      \caption{Approximate posterior distribution for the impact points.}
     \end{figure}
@@
 where \(Z \stackrel{i.i.d.}{\sim} g(z)\) satisfying \(g(z^{-1})=zg(z)\).
-\item  The proposal is accepted with probability
+
+\item If \(\R^p\) is the sample space, the proposal is accepted with probability
   \[

### 2022-09-19 chapters/bayesian-functional-regression.tex
Minor changes

@@
 In this way we get a simpler, finite-dimensional approximation of the functional RKHS model, which we argue reduces the overall complexity of the model while still capturing most of the relevant information.
-When it comes to parameter estimation, a direct optimization of some loss function would probably require a tailored algorithm that took into account the continuous nature of the times \(t_j\).
+When it comes to parameter estimation, a direct optimization of some loss function would probably require a tailored algorithm that took into account the whole functional trajectories \(x(t)\) to select the appropriate times \(t_j\).
 Indeed, such an idea is explored in \citet{berrendero2018functional} for the logistic regression case, where the authors propose a ``greedy max-max'' method reminiscent of the EM algorithm that alternates between estimating the coefficients and the time instants through a maximum likelihood approach.

### 2022-09-22 chapters/experiments.tex
Some corrections after public defense

@@
 However, the predict-then-summarize approach (represented as \textit{posterior\_mean}) seems to work quite well, always scoring near the mean RMSE of all the comparison algorithms.
-Moreover, our two-stage methods seem to outperform the summarize-then-predict methods in these cases, scoring again very close to the mean of the reference methods.
+Moreover, our two-stage methods seem to outperform the summarize-then-predict methods in Moisture and Sugar, scoring again very close to the mean of the reference models.
 
@@
 In Figure~\ref{fig:clf_emcee_rkhs} we see the results for the GP regressors in the logistic RKHS case.
-Our models perform fairly well in this advantageous case, although the two-stage methods fare somewhat poorly in these data sets.
+Our models perform fairly well in this advantageous case, although they are not always better than the comparison methods.
 However, in most cases the differences observed account for only one or two misclassified samples.
@@
 The predict-then-summarize approaches (\textit{emcee\_posterior\_mean} and \textit{emcee\_posterior\_vote}) are particularly good in this case, and in general have low standard errors.
-Moreover, the overall accuracy of all methods is low (below 60\%), so this is indeed a difficult problem in which even small increases in accuracy are relevant.
+Moreover, the overall accuracy of all methods is poor (below 60\%), so this is indeed a difficult problem in which even small increases in accuracy are relevant.
 
@@
 In particular, the predict-then-summarize methods tend to have a good performance and achieve a lower standard error across executions, which is a trend that we also saw in the simulated data sets.
-However, our variable selection methods seem to be a bit worse in this case.
-Again, the models that use \textit{emcee\_mean} are the exception, and in all these data sets they perform steadily worse than the rest of our Bayesian models.
+As we have been seeing almost invariably, the models that use \textit{emcee\_mean} are the exception, and in all these data sets they perform steadily worse than the rest of our Bayesian models.
 

### 2023-11-06 chapters/bayesian-functional-regression.tex
Fix errata

@@
 \begin{equation}\label{eq:rkhs-model-linear-2}
-  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
+  Y_i \mid X_i, \theta \ \stackrel{\text{ind.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
 \end{equation}
@@
 \begin{equation}\label{eq:rkhs-model-logistic-2}
-Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
+Y_i \mid X_i,\theta \ \stackrel{\text{ind.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
 \end{equation}

### 2023-11-21 chapters/background-related-work.tex
Update reference

@@
 In this situation, the most common way of estimating the slope function \(\beta\) is via its Maximum Likelihood Estimator (MLE).
-However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][]{berrendero2018functional}.
+However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][]{berrendero2023functional}.
 
@@
 In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity.
-These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional} for the case of functional logistic regression.
+These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2023functional} for the case of functional logistic regression.
 

### 2023-11-21 chapters/bayesian-functional-regression.tex
Update reference

@@
 In this chapter we present the precise functional models and Bayesian methodologies explored in this work.
-The RKHS-based functional models under consideration \citep[see][]{berrendero2018functional, berrendero2019rkhs} are those obtained by substituting the functional parameter \(\beta\in L^2[0,1]\) for \(\alpha \in \Hcal(K)\), replacing also the scalar product \(\dotprod{X}{\beta}\) for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
+The RKHS-based functional models under consideration \citep[see][]{berrendero2023functional, berrendero2019rkhs} are those obtained by substituting the functional parameter \(\beta\in L^2[0,1]\) for \(\alpha \in \Hcal(K)\), replacing also the scalar product \(\dotprod{X}{\beta}\) for \(\dotprod{X}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}.
 However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subspace \(\Hcal_0(K)\) defined in~\eqref{eq:h0}, i.e.:
@@
 When it comes to parameter estimation, a direct optimization of some loss function would probably require a tailored algorithm that took into account the whole functional trajectories \(x(t)\) to select the appropriate times \(t_j\).
-Indeed, such an idea is explored in \citet{berrendero2018functional} for the logistic regression case, where the authors propose a ``greedy max-max'' method reminiscent of the EM algorithm that alternates between estimating the coefficients and the time instants through a maximum likelihood approach.
+Indeed, such an idea is explored in \citet{berrendero2023functional} for the logistic regression case, where the authors propose a ``greedy max-max'' method reminiscent of the EM algorithm that alternates between estimating the coefficients and the time instants through a maximum likelihood approach.
 
@@
 In the case of functional logistic regression, we regard the response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given \(X=x \in L^2[0, 1]\), and as usual suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\P(Y=1| X=x)\).
-Then, following the approach  suggested by \citet{berrendero2018functional}, a RKHS model might be given, in terms of the correspondence \(\dotprod{X}{\alpha}_K = \Psi_X(\alpha)\), by the equation
+Then, following the approach  suggested by \citet{berrendero2023functional}, a RKHS model might be given, in terms of the correspondence \(\dotprod{X}{\alpha}_K = \Psi_X(\alpha)\), by the equation
 \begin{equation}\label{eq:rkhs-model-logistic}
@@
 First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below.
-Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in some cases it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model.
+Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in some cases it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2023functional}; this provides a sound theoretical motivation for the reduced model.
 Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well.

### 2023-12-13 chapters/appendix-code.tex
Minor correction

@@
 number of iterations, type of data set, \ldots).
-In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our Bayesian methods without a cross-validation loop.
+In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while the script \texttt{results\_all.py} allows the execution of our Bayesian methods without a cross-validation loop.
 \end{itemize}

### 2023-12-21 chapters/appendix-code.tex
Add experiments folder to appendix on source code

@@
 \begin{itemize}
-  \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our models, separated according to the functionality they provide.
+  \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our Bayesian models, separated according to the functionality they provide.
   \item The \texttt{reference\_methods} folder contains our implementation of the functional comparison algorithms that were not available through a standard Python library.
-  \item The \texttt{utils} folder contains some utility files for simulation, experimentation and visualization.
-  \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g.
-number of iterations, type of data set, \ldots).
-In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while the script \texttt{results\_all.py} allows the execution of our Bayesian methods without a cross-validation loop.
+  \item The \texttt{utils} folder has some utility files for simulation, experimentation and visualization.
+  \item The \texttt{experiments} folder contains plain text files with the numerical experimental results shown in Appendix~\ref{ch:tables}, as well as \texttt{.csv} and \texttt{.npz} files that facilitate working with them directly in Python.
+  \item At the root folder we have files for executing our experiments, which accept several user-specified parameters (such as the number of iterations or the type of data set).
+In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while the script \texttt{results\_all.py} executes our Bayesian methods without a cross-validation loop.
 \end{itemize}

### 2024-02-03 chapters/experiments.tex
Update reference
Minor change in naming

@@
   \item APLS.\hspace{.3em} Functional PLS used as a dimension reduction method, as proposed in \citet{delaigle2012achieving} in combination with the nearest centroid (NC) algorithm.
-  \item FLog.\hspace{.3em} Functional RKHS-based logistic regression algorithm proposed in \citet{berrendero2018functional}.
+  \item FLog.\hspace{.3em} Functional RKHS-based logistic regression algorithm proposed in \citet{berrendero2023functional}.
   \item FLDA.\hspace{.3em} Implementation of the functional version of linear discriminant analysis proposed in \citet{preda2007pls}.
@@
 
-Lastly, the results for the non-GP case can be seen in Figure~\ref{fig:reg_emcee_nongp}, in which the regressors are realizations of a GBM.
+Lastly, the results for the non-GP case can be seen in Figure~\ref{fig:reg_emcee_non_gp}, in which the regressors are realizations of a GBM.
 In this case we still get better results under the RKHS model, while the results under the \(L^2\)-model are slightly worse.
@@
   \centering
-  \includegraphics[width=.65\textwidth]{reg_emcee_nongp}
-  \caption{RMSE of predictors for simulated data with GBM regressors (lower is better).}\label{fig:reg_emcee_nongp}
+  \includegraphics[width=.65\textwidth]{reg_emcee_non_gp}
+  \caption{RMSE of predictors for simulated data with GBM regressors (lower is better).}\label{fig:reg_emcee_non_gp}
 \end{figure}
@@
 
-Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic Gaussian processes, but they struggle in the heteroscedastic case.
+Finally, Figure~\ref{fig:clf_emcee_non_gp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic Gaussian processes, but they struggle in the heteroscedastic case.
 Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e.
@@
   \centering
-  \includegraphics[width=.65\textwidth]{clf_emcee_nongp}
-  \caption{Accuracy of classifiers for simulated data coming from two different GP's, labeled according to their origin (higher is better).}\label{fig:clf_emcee_nongp}
+  \includegraphics[width=.65\textwidth]{clf_emcee_non_gp}
+  \caption{Accuracy of classifiers for simulated data coming from two different GP's, labeled according to their origin (higher is better).}\label{fig:clf_emcee_non_gp}
 \end{figure}
